### 2024-11-06 docs/Homeownership_BankOfMomAndDad.tex
First submission

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb}
\usepackage[right=1.25in,left=1.25in,top=1.15in,bottom=1.1in]{geometry}
\usepackage{setspace}
\usepackage{makecell}
\usepackage[backref=page]{hyperref}
\renewcommand*{\backref}[1]{}
\renewcommand*{\backrefalt}[4]{%
	\ifcase #1 (Not cited.)%
	\or        [Cited on page~#2.]%
	\else      [Cited on pages~#2.]%
	\fi}
\usepackage{xcolor}
\definecolor{mLightBrown}{rgb}{0.92,0.51,0.1}
\hypersetup{
    colorlinks,
    linkcolor={mLightBrown},
    citecolor={mLightBrown},
    urlcolor={mLightBrown}
}

\usepackage[longnamesfirst]{natbib}
\usepackage{adjustbox}
\usepackage{threeparttable}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{pdflscape} 
\usepackage{graphicx} 
\usepackage{tikzit}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows.meta}
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\usepackage{tikz}

\newcommand{\E}{\mathbb{E}}
\newcommand{\mc}{\mathcal}

\title{{\LARGE Illiquid Homeownership and the \\ Bank of Mom and Dad}\thanks{Federal Reserve Board of Governors. E-mail: eirik.e.brandsaas@frb.gov. I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts, Corina Boar, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobson, William Gamber, Claudia Sahm, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, Boston Fed, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, YES 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop. \textit{Disclaimer: }The views expressed in this paper are solely the responsibility of the author and should not be interpreted as reflecting the views of the Board of Governors of the Federal Reserve System or of anyone else associated with the Federal Reserve System.}}
\author{Eirik Eylands Brandsaas}
\date{{\today }}

\begin{document}

\maketitle
\begin{abstract}
{\begin{spacing}{1.2}
Housing is the largest asset in U.S. household portfolios, and first-time homebuyers increasingly rely on parental transfers. This paper quantifies the contribution of parental transfers to the homeownership rate of young households. I build and estimate a life-cycle overlapping generations model with housing, where adult children and parents interact without commitment. I find that parental transfers account for 14 percentage points (29\%) of young households' homeownership. Transfers from wealthy parents not only help households overcome borrowing constraints, but also help sustain homeownership, mitigating the drawbacks of illiquidity. Surprisingly, policies lowering entry barriers to homeownership generally increase the reliance on parental wealth, whereas increased liquidity reduces it. Finally, I show that children of wealthy parents strategically use the illiquidity of housing as a commitment device to encourage transfers, resulting in a preference for illiquidity.
\end{spacing}}
\vspace{0.5cm}
\noindent \textbf{JEL:} D14, D15, E21, G51, R21

\noindent \textbf{Keywords:} Homeownership, Parental transfers, Altruism, Life-cycle models

\end{abstract}
\newpage
\onehalfspacing

\section{Introduction}
Housing is the largest asset in U.S. households portfolios. While homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}, households face different barriers to ownership. Between 2009 and 2016, around 30\% of American first-time homebuyers received parental assistance for their down payment, with such assistance averaging \$48,000 per receiving household.\footnote{All data from the Survey of Household Economic Decisionmaking and the Panel Study of Income Dynamics (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.

This paper quantifies the importance of parental transfers for young adults' homeownership and investigates how policies lowering homeownership barriers interact with parental transfers. Previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}. Additionally, parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}. However, while these studies show that transfer receipt increases the probability of renters becoming homeowners, they do not address the overall importance of parental transfers. For example, households may anticipate future transfers and subsequently save less. In this case, even if transfers increase the likelihood of homeownership, all else being equal, they may not affect aggregate homeownership. Furthermore, this is the first paper quantifying the importance of parental transfers as a determinant of homeownership; the existing literature has primarily focused on family formation and credit constraints \citep[see e.g.,][]{Chang2024,Paz-Pardo2019,Mabille2020}.

I first present empirical evidence indicating that parental wealth helps not only in the transition to ownership, but also in maintaining ownership. I show that, controlling for a wide set of household and parental characteristics, households with wealthier parents are less likely to fall behind on mortgage payments after becoming homeowners, even though they buy more expensive homes. Finally, using a variation of the event study in \cite{Chetty2007}, I show that while most households downsize their housing during unemployment spells, those with wealthy parents retain their current homes.

To quantify the impact of parental transfers on homeownership, I then combine two distinct models: housing models, which ignore parental transfers, and models of parental transfers, which ignore housing. I build a life-cycle overlapping generations model with altruistic parents and a rent-or-own decision. Parents and children interact without commitment through both inter-vivos transfers and end-of-life bequests. In models of altruism without housing (e.g., \citealp{Altonji1997a,Barczyk2014}), the transfer motive is to increase the child's consumption when they are borrowing constrained, thus having high marginal utility of wealth. The inclusion of a frictional housing market generates two new transfer motives. First, transfers can directly alleviate credit constraints, enabling wealthier parents to help their children meet down-payment requirements. Second, future transfers provide partial insurance against future income shocks, reducing risks associated with the large illiquid investment of homeownership. Moreover, the illiquidity of housing serves as a commitment device enabling children to secure future transfers. As in all transfer models, the expectation of future transfers reduces children's saving motives, potentially lowering the homeownership rate.

Allowing for parental transfers in all periods highlights that transfers are critical not only for entering homeownership---the main focus of empirical literature---but also for maintaining homeownership. Specifically, the prospect of future transfers mitigates the downsides of homeownership for liquidity-constrained households. Ameliorating illiquidity turns out to be as important as relaxing borrowing constraints. Crucially, the model rationalizes the previously discussed empirical pattern that households with wealthy parents do not downsize during income losses, which would not be possible without allowing transfers in all periods.

I estimate the model by matching data on homeownership, wealth, and transfers from the Panel Study of Income Dynamics (PSID). To quantify the importance of transfers, I find the counterfactual homeownership rate without parental transfers. Without parental transfers, the model simplifies to a standard life-cycle model with housing. I find that parental transfers account for 14 percentage points (29\%) of the homeownership rate among young households.

I then use the model to evaluate policies aiming to increase homeownership. I find that such policies disproportionately benefit households with wealthy parents. Relaxing mortgage constraints---for example, by lowering minimum down payments---strengthens the link between parental wealth and children's housing outcomes, since more households with wealthy parents are on the margin to become homeowners. Nonmortgage policies that lower barriers to entry, such as reducing prices or purchase costs, increase homeownership for all households, though disproportionately among those with wealthier parents. However, increasing liquidity by reducing sales costs weakens the link between parental wealth and children's housing outcomes, as the partial insurance provided by parental wealth becomes less important.

I then study how illiquidity affects the strategic behavior arising in the absence of commitment. I find that 27\% of young households prefer illiquid housing---due to sales costs---to liquid housing. These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth.'' Lower liquid wealth implies higher marginal propensities to consume (MPC) and, hence, more parental transfers. As an illustrative example, consider a child who buys a house and brings no liquid wealth to the next period. If the parent does not transfer in the next period, the child will either face very low consumption (with a correspondingly high marginal value of wealth) or be forced to liquidate the house and incur sales costs. The parent dislikes both outcomes and may find it optimal to transfer enough to keep the child in the home. This driver of preference for illiquidity differs from those studied in the literature, such as temptation or time inconsistency \citep{attanasio2024temptation,laibson1997golden}. Furthermore, parental wealth generating preference for illiquidity provides a novel explanation for the existence of high-MPC households \cite[see e.g.,][]{kaplan2022marginal}. That children with wealthy parents find illiquidity less problematic---or even prefer it---is consistent with the empirical results in \cite{choukhmane2023benefits}. They find that households with wealthy parents are more likely to invest in illiquid retirement accounts and less likely to make costly early withdrawals, even after controlling for a wide set of characteristics.

Finally, I quantify the contribution of parental transfers to the Black-White homeownership gap of 50\% among young adults. I adapt the methodology of \cite{Ashman2020} and \cite{aliprantis2022dynamics}, who study the racial wealth gap, and recalibrate the income process to capture racial differences. I find that parental transfers account for 7 percentage points (14\%) of the Black-White homeownership gap. For context, the Black-White homeownership gap has remained roughly constant since the passing of the Fair Housing Act in 1968.

%%%%%
%% Start of pure literature review
%%%%%

My paper first contributes to the literature studying the determinants of homeownership over the life cycle. Recent papers focus on marriage and family formation \citep{Fisher2011,Chang2024,Khorunzhina2019}, housing demand in old age \citep{McGee2019,Barczyk2020a}, and changing borrowing constraints \citep{Paz-Pardo2019,Mabille2020}. These studies highlight the importance of credit constraints and minimum down payments in constraining demand for owner-occupied housing. By considering parent-child interactions, I show how parental wealth affects the relative importance of constraints. For households with poorer parents, the sales and purchase costs decrease ownership by making it riskier to own. This risk is smaller for households with richer parents, making mortgage credit constraints relatively more important. My results highlight that policies increasing ownership through subsidies to first-time buyers---a common approach in many cities---will lead to increased housing inequality. My model-based results also complement some of the empirical research on the effect of parental resources and housing outcomes \citep[see e.g.,][ for recent work]{wold2024housing,daysal2023intergenerational,benetton2022dynastic} by quantifying the contribution of transfers to homeownership.

My paper also contributes to the literature on altruistic households interacting without commitment by studying how illiquid housing affects the commitment problem. In these models, children have an incentive to undersave to increase parental transfers (e.g., \citealp{Altonji1997a,Boar2018,Barczyk2014,Chu2020}). Illiquid housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}, which children of wealthy parents strategically exploit to encourage future parental transfers. The paper most closely related to mine is \cite{Barczyk2020a}, who study how the homeownership of retired parents affects the economic behavior of the elderly and their adult children. In their model, the parents' home serves as a commitment device, encouraging wealth bequests from the parent and informal care from the adult offspring. My contribution is to study the complementary problem: how parental transfers affect the homeownership decisions of young adults. Finally, while \cite{kaplan2012moving} studies cohabitation with parents among individuals under the age of 25, my focus is on homeownership decisions for households from age 25 onwards.

Finally, the paper contributes to the literature on parental transfers and life-cycle outcomes by focusing on how parental transfers to young adult households influence their wealth accumulation through homeownership. My results show that households with wealthier parents are willing to buy sooner, take on higher leverage, and hold less liquid precautionary savings. The previous literature has generally focused on parental investment in children's human capital (e.g., \citealp{Lee2019,Daruich2018,gilraine2023public}), transfers from adult children to retired parents (e.g., \citealp{Mommaerts2016, Barczyk2018, Barczyk2020a}), or smoothing income shocks (e.g., \cite{Boar2021ReStud,fagereng2023insuring,andersen2020bailing}) instead of the effect of parental wealth on household portfolio choices.

The paper proceeds as follows. In Section \ref{sec:data}, I describe the data sources and summary statistics, and I document that parental wealth is associated with better housing outcomes. Section \ref{sec:model} describes the quantitative model, and Section \ref{sec:estimation} discusses the structural estimation. Section \ref{sec:quant} performs the main quantitative exercise and robustness tests. Finally, Section \ref{sec:pol} studies how policies intended to increase homeownership also affect the role of parental wealth and housing illiquidity.

\section{Data on Transfers, Family, and Housing}\label{sec:data}
I first present time trends in parental transfers for down payments, before describing the estimation sample taken from the PSID. Finally, I use the estimation sample to show that parental wealth is positively associated with better housing outcomes, both before and after purchase.

\subsection{Parental Transfers for Owners Over Time}\label{sec:overtime}
I use the Survey of Household Economics and Decisionmaking (SHED) and American Housing Survey (AHS) datasets to show the time trends in parental housing transfers.

\begin{figure}
	\caption{Increased Reliance on Parental Transfers for Down Payments}\label{fig:motivation}
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/SHED_gift_scatter_SE_paper}%
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/AHS_majorsourcedown_surveyyear_paper}%

	 {\begin{footnotesize} \textit{Notes:} The left panel uses data from the SHED with the sample restricted to current owners who report this being their first home. The right panel uses data from the AHS with the sample restricted to all owner-occupied units that did not indicate the sale of a previous home as the main source of funding. Note that the left figure has year of purchase on the horizontal axis while the right figure has survey year. Dashed lines denote 95\% confidence intervals.\end{footnotesize}}
\end{figure}

\textit{SHED:} I use SHED, an annual cross-sectional survey conducted by the Federal Reserve, to observe the share of households who funded the down payment with a loan or gift from family or friends, by year of purchase. The results are plotted in the left panel of Figure \ref{fig:motivation}. The main observation is the large increase in the role of inter-vivos transfers for homeowners since 2001. From 2001 to 2007, only 10\% to 18\% of first-time buyers received transfers, while 20\% to 40\% received transfers after 2009.\footnote{The question was only included in the 2015 and 2016 waves. In the 2015 wave, first-time ownership and down-payment source is only reported for those who bought in 2001 to 2015, while in the 2016 wave it is only reported for 2015 to 2016. }

\textit{AHS:} I use the AHS, which surveys occupants of owner-occupied units, to obtain a time series on whether the down payment was mostly funded by gifts or inheritances (right panel of Figure \ref{fig:motivation}). We observe a flat pattern in the 1990s, followed by a sharp decline from 2005 to 2009, coinciding with the prevalence of mortgages featuring very low down payments (e.g., \cite{Corbae2015}), and a stark increase from 2013 to today. Minor fluctuations in the percentage of homeowners relying on gifts or inheritances reflect significant shifts among new homeowners, who comprise a small fraction of all homeowners. This is because the horizontal axis denotes the survey year, not the year of purchase.

\subsection{Panel Study of Income Dynamics}\label{sec:PSID}
My main data source is the PSID, which follows a nationally representative sample of U.S. households and their descendants over time since 1968. The PSID is the only publicly available U.S. dataset that satisfies this paper's three requirements. First, it has detailed wealth, income, and housing data for both parents and adult children. Second, it has information about inter-vivos transfers from parents to children, unlike most register data. Third, it follows households over time, so we can observe the transitions from renting to owning and how these transitions relate to parental wealth.

I use data from 1999 to 2017. In 1999, the PSID started to collect detailed wealth data every other year. In most waves of the PSID, there is limited transfer data, and the main question is whether households received gifts or bequests over \$5,000 in the last two years. In 2013, the PSID collected more detailed transfer data in the Family Roster and Transfer Module. They asked parents how much they gave their children in the last calendar year and how much they had given over their lifetime for school, house purchases, or other purposes. Household characteristics such as age, gender, and education refer to the household head. I classify top-coded values as missing observations. All monetary variables are expressed in constant 2016 U.S. dollars (in thousands). Table \ref{tab:vardef} in the Appendix provides variable code of the main variables.

\textit{Sample Selection:} Throughout this paper, the sample includes all households aged 25 to 84 in the PSID, excluding the Survey of Economic Opportunity and the Latino subsamples, to obtain a representative sample. All summary statistics are calculated using the provided family weights. I drop all observations with missing housing values (which is set to zero for renters) or wealth.

\textit{Matching Parents with Children:} I use the Parent/Child file from the PSID's 2013 transfer supplement. I can observe each household's reported transfers to and from parents and children. This leads to discrepancies, where the child and parent do not agree on the amount given from the parent to the child. I only use the parent's reported transfer. First, there may be some stigma about receiving transfers, which may induce receiving children to underreport. Second, in the model, parents determine the size of the transfer they give to their child.

\textit{Definition of Transfers:} The transfer supplement asked all parent households whether they gave money, gifts, or loans of \$100 or more to their children in 2012. I follow the literature \citep[e.g.,][]{mcgarry2016dynamic} and treat all transfers as gifts. Since this paper focuses on transfers that (a) relate to housing and (b) are quantitatively meaningful, I set transfers below \$500 to \$0. About 25\% of transfers are below this threshold, and ignoring them increases the mean transfer from \$2,921 to \$3,960.

\subsubsection{Descriptive Statistics: Who Receives Transfers?}
I now discuss descriptive statistics from the PSID sample in 2013 for households aged 25 to 44 with an observed parent household. Table \ref{tab:descrstats} contains the means of variables, by age, wealth, marriage, homeownership, and transfer receipt.

There are several main takeaways from the subgroup analysis. In 2013, 22\% of young households received a transfer, and transfers averaged \$3,950. Receivers have significantly richer parents, have similar wealth and income as nonreceivers, and are less likely to own. Receivers are more likely to transition from renting to owning, especially in the age groups where households are most likely to buy. The key determinant of transfer receipt seems to be parental wealth.

In panel a), the first two columns compare the whole sample, by transfers receipt in the last calendar year. The mean transfer is relatively high at \$3,960, and 22\% of households received transfers in the last year. The largest difference is parental wealth: transferring parents are 2.5 times as wealthy as nontransferring parents (\$1,029,000 vs. \$410,000). Perhaps surprisingly, receivers are slightly richer than nonreceivers (\$108,000 vs. \$78,500), are more likely to be college-educated and white, and are one year younger. Receivers are slightly less likely to be homeowners (39\% vs. 43\%), reflecting the age difference and the tendency for college attendance to delay homeownership.

\begin{table}[htbp]
	\begin{adjustbox}{center}
		\small
		\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
		\begin{threeparttable}
			\singlespacing
			\caption{Descriptive Statistics (Means), Households Aged 20-44}\label{tab:descrstats}
			\begin{tabular}{l@{\hspace{.5\tabcolsep}} r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
				\toprule
				& \multicolumn{2}{c}{All}  & \multicolumn{2}{c}{Single}  & \multicolumn{2}{c}{Married}  & \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\
				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}  & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   \\  \midrule
				\input ../tabfig/descr/descr_trsnf_disc_mean.tex
				\midrule
				Age                    & \multicolumn{2}{c}{25-28} & \multicolumn{2}{c}{29-32}  & \multicolumn{2}{c}{33-36}  & \multicolumn{2}{c}{37-40}   & \multicolumn{2}{c}{41-44}   \\
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}  & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   \\ 
				 \midrule
				\input ../tabfig/descr/descr_trsnf_age_mean.tex
				\midrule
				Wealth Quintile                    & \multicolumn{2}{c}{Quintile 1}  & \multicolumn{2}{c}{Quintile 2}  & \multicolumn{2}{c}{Quintile 3}   & \multicolumn{2}{c}{Quintile 4}  & \multicolumn{2}{c}{Quintile 5}   \\
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}  & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   \\ 
		\midrule
				\input ../tabfig/descr/descr_trsnf_wlt_mean.tex
		\bottomrule
			\end{tabular}
			{\footnotesize
				\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD.}
		\end{threeparttable}

	\end{adjustbox}
	\end{table}

By breaking down the sample by marital status, we see that married or cohabiting households have more wealth, higher income, are slightly older, and have three times the ownership rate of singles. Married households receive more frequent and slightly larger transfers. Next, I break the sample down by ownership. Homeowners have 10 times the wealth of renters and also significantly wealthier parents. Renters and owners receive transfers at about the same rate and size. Receivers are more likely to switch from renting to owning: 21\% of receiving owners rented two years ago, compared to 14\% of nonreceiving owners.

Next, I break the sample into five age groups from 25 to 44. We see that households' wealth, income, and homeownership rates increase with age. Wealthier households are less likely to receive transfers. Notably, among 29-to-32-year-olds, homeownership is more common among transfer recipients (40\%) than nontransfer recipients (32\%). Furthermore, receivers are not only more likely to own, but also to be recent homeowners: 21\% of receiving owners are new homeowners versus 13\% of nonreceiving owners.

\subsection{Parental Resources and Housing Outcomes in the Data}\label{sec:dataregr}

I now utilize microdata from the PSID to examine the correlation between parental financial resources and housing outcomes.

I use the sum of parental net worth and income (denoted as wealth) as a statistic to measure parents' financial resources in all specifications for two reasons. First, this is consistent with models of altruism, as the one I use later. Second, since parents' wealth and income are highly correlated and I have few observations---under 1,000 for most specifications---the standard errors are large when both are included. The sample is limited to households aged 25 to 44.

\subsubsection{Households with Wealthier Parents Buy More Expensive Houses}
I first show that households with wealthier parents buy more expensive homes, controlling for household characteristics.

I run an ordinary least squares (OLS) regression on house purchase values for first-time owners and test whether lagged parental wealth (logged) is positively associated with larger house purchases $(\beta_1)$, controlling for lagged log income, lagged log net worth, education, age, log household size, regions, time, and race:
\begin{equation*}
\ln{HVal_{i}} = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
\end{equation*}
Parents' wealth, as well as household income and net worth, are log-transformed. Households are denoted by $i$ and their parents by $p(i)$. First-time buyers are households who have never been observed as homeowners before.

Column 1 of Table \ref{tab:hypo} reports the results. Households with wealthier parents buy larger houses: A 1\% increase in parental wealth is associated with a 0.049\% increase in the purchase value of the child's house, and the effect is significant at the 5\% level. The effect of parental wealth is about the same as the child's own net worth (0.065\%).

\begin{table}
	\center
	\begin{threeparttable}
		\caption{Housing Choices and Parental Wealth}
		\label{tab:hypo}
		\small 
				\input{../tabfig/regr/hypoII.tex}

		{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses. `Behind' refers to whether the household is behind on a mortgage. Wealth, income, parental wealth, mortgage, family size, and house values are logged. All regressions include year and state fixed effects and control for age and age-squared of both the child and parent. Specifications 1-3 use ordinary least squares while specifications 4 and 5 use random and fixed effects, respectively. \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
	\end{threeparttable}
\end{table}

\subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales. I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses. The PSID has collected this data since 2009, and 3.9\% of owners aged 25 to 44 are behind, on average, while 9.2\% have been behind at least once.

The sample is limited to the first time a household is observed as owners, and the control variables are the same as in the previous regression:
\begin{equation*}
Behind_i = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
\end{equation*}
In Column 2 of Table \ref{tab:hypo}, the outcome variable is whether households are behind on the mortgage payments the first time they are observed as owners. Parental wealth in the period before purchase decreases the probability that a household will ever be behind: A 1\% increase in parental wealth decreases the probability of being behind by 0.016 percentage points and is significant at the 1\% level. Parental wealth has a larger effect than the child's net worth. In Column 3, the outcome variable is whether first-time homeowners will ever be observed to be behind on their mortgage, and coefficients and significance levels are virtually unchanged. The sample size increases in these regressions, since households who bought before 2009 are now included.

I also report the results from two specifications where I follow households post-purchase, using random effects and household fixed effects regressions (Columns 4 and 5). Once we follow households over time, the effect of parental wealth decreases. With household fixed effects, the effect is no longer significant. However, the fixed effects regressions must be interpreted with caution, since there is little within-parent variation in wealth over time. Indeed, the random and fixed effects coefficients are almost the same (-0.008 vs -0.006). Overall, the results support the hypothesis that parental wealth decreases the probability of being behind on mortgage payments.

The probability of being behind on mortgages is decreasing in parental wealth, highlighting that parental resources appear to help new homeowners maintain ownership after purchase---in addition to relaxing borrowing constraints.

\subsubsection{Households with Wealthier Parents Are Less Likely to Downsize during Unemployment Spells}\label{sec:eventstudy}
Distressed and forced home sales are costly and frequently driven by unemployment \citep{kermani2021racial,hsu2018unemployment}. If parental wealth provides insurance, households with wealthier parents should be less likely to downsize after income losses. I perform a simple event study on the effect of unemployment on housing consumption to test this hypothesis.

The exercise follows \cite{Chetty2007} closely. The outcome of interest is the change in log housing consumption, which is set to zero for households who do not move. For households who do move, housing consumption is defined as yearly rent while renting and 5\% of the market value for owner-occupied housing \citep{Davis2008}. The sample is limited to household heads who are unemployed only once between ages 25 and 45. I divide the sample by whether a household's parents were in the top parental wealth quartile (of parents with children aged 25 to 44) at the time of unemployment and compare the two groups' housing consumption growth rates at unemployment. The results are displayed in Figure \ref{fig:housinggrowthrates}. Among households with nonwealthy parents, housing consumption decreases by about 5\% during unemployment (significant at the 95\% level). In contrast, among households with wealthy parents, housing consumption even grows slightly at unemployment. Moreover, I replicate this event study using simulated data from the quantitative model and show that the results are similar (see Section \ref{sec:externalval} and Figure \ref{fig:housinggrowthrates_model}).
\begin{figure}
	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates}
	\noindent\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both}\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both}

	{\begin{footnotesize} \textit{Notes:} Solid lines denote means, and bars denote the 95\% confidence interval. The sample consists of households aged 25-45 with exactly one unemployment spell and without changes in head and/or spouse in the four years before and after unemployment. \end{footnotesize}}
\end{figure}

\subsection{Taking Stock of the Data}
Taking stock, we have seen that parental transfers have become more important for young homeowners over the past two decades. About 30\% of young households receive transfers from their parents in a given year, and recipients' parents are wealthier. Recipients are also more likely to transition from renting to owning. Moreover, households with wealthier parents buy more expensive houses, are less likely to be behind on their mortgages, and are less likely to downsize during unemployment. These results suggest that homeowners with wealthy parents find illiquidity of homeownership less problematic.

\subsection{Transfers and Taxes}
In the United States, transfers and bequests are subject to taxation due to the estate tax. Taxes are paid by the giver, not the recipient. In 2024, an individual can give \$13.61 million before they start owing gift taxes. However, individuals have to file if they give transfers above \$18,000 (in 2024) to any one individual within a calendar year. Thus, the gift tax is irrelevant for the vast majority of households, as it only applies if lifetime transfers---including bequests---exceed \$13.61 million.

\section{A Quantitative Model of Parental Transfers and Homeownership}\label{sec:model}
%% There is a missing paragraph linking the previous results this this part!
This section describes my life-cycle model of housing choices with overlapping generations, idiosyncratic earnings risk, and altruistic inter-vivos transfers from parents to their offspring (``children''). To answer how important parental transfers are for housing outcomes, I combine two models: altruism without commitment and illiquid housing.

Altruism without commitment is a standard model of intergenerational family interactions (e.g., \cite{Altonji1997a,Barczyk2018}). Altruistic parents, deriving utility from their child's utility, influence their child's consumption and housing choices through nonnegative transfers. The lack of commitment aligns with empirical transfer patterns and generates strategic behavior, as both children and parents internalize the effect of their choices on each other's future choices. The central theoretical prediction is that wealthy parents transfer to children with high marginal utility of wealth, typically poor or borrowing-constrained households \citep{Chu2020,Barczyk2020}.

The second model component is illiquid homeownership. Without illiquidity, transfers mainly relax mortgage constraints, as has been the focus in empirical literature (e.g., \cite{Blickle2019,Engelhardt1998,Guiso2002,Lee2018}). With illiquidity, we get a two-asset model where portfolio composition matters for transfers. Specifically, homeowning children with low liquid wealth (``house rich but cash poor'') have high marginal utility of wealth, strengthening parental transfer motives and, thus, the child's homeownership motive. Finally, while illiquid housing makes selling costly, wealthy parents can offer partial insurance, thus reducing the risk of costly liquidation or downsizing.

These two housing frictions---borrowing constraints and illiquidity---are not only theoretically important, but also empirically relevant. In the SHED, 57\% of renters could not afford a down payment or did not qualify for a mortgage, while 26\% said that renting was more convenient and 23\% cited planning to move as reasons for renting.

\subsection{Demographics, Preferences, and Technologies}
Time is discrete and finite. Each period consists of two years. At the beginning of each period, a constant mass of new households enters and exits the economy. The only economic agents in the model are households.

\textit{Demographics:} Households are economically active from age $a\in \mathcal{A}=\{25,\allowbreak 27,\allowbreak\dots,\allowbreak 83\}$. The life cycle is illustrated in \ref{fig:overview}. A family consists of one adult child $c$ household (age 25 to 53) and a parent $p$ household (age 55 to 83) with an age gap of 30. Thus, children have new children at age 30, but new children are economically inactive until age 25. Each parent-child pair overlaps for 15 periods (30 years), and only two households are economically active in any dynasty at a time. Three events happen simultaneously when a child household becomes 55. First, the household's parent dies (at age 85). Second, the child transitions to become a parent household. Third, the child of the child becomes economically active as a child household.

\begin{figure}\begin{center}
	\caption{Life Cycle of Three Generations in a Dynasty }\label{fig:overview}
	\input{modeloverview.tex}
\end{center}
\end{figure}

\textit{Preferences and Altruism:} Parents and children have time-separable expected utility with a discount factor of $\beta$. Households maximize expected utility, and the per-period utility function for children is
\begin{equation}
U_c(c_c,h_c) = u(c_c,h_c) = \frac{\left(c^\xi s(h)^{1-\xi}\right)^{1-\gamma}-1}{1-\gamma},
\end{equation}
where $c$ is consumption, $h$ is housing, $\gamma$ is the coefficient of relative risk aversion, and $\xi$ measures the relative importance of consumption to housing services. The within-period Cobb-Douglas aggregator imposes a constant desired expenditure share on housing services, roughly consistent with empirical evidence (e.g., \cite{davis2011household}). The function $s(h)$ allows utility to depend on ownership:
\begin{equation}
s(h) = \begin{cases}
h & \text{ if renting}, \\
\chi h & \text{ if owning}.
\end{cases}
\end{equation}
The parameter $\chi$ measures the owner-occupied utility premium, reflecting any additional benefits derived from owner-occupation, such as stability or ownership rights.

The parent's utility also depends on the altruistic utility derived from the child:
\begin{equation}
U_p(c_p,h_p,c_c,h_c) = u(c_p,h_p) + \eta u(c_c,h_c),
\end{equation}
where $\eta$ measures the intensity of the parent's altruism towards the child.

\textit{Intergenerational Transfers:} In the last period before death, the parent can leave a bequest, which the child receives in the next period. In all other periods, the parent can give inter-vivos transfers of a nonnegative amount, $t_p$, that the child receives immediately.

\textit{Housing:} Households can obtain housing services by renting or owning. They can rent housing of size $h_r$ or own houses of size $h_o$. The unit price of housing is $p$, and $q$ denotes the rent-to-price ratio. I assume that prices are constant for tractability, but I model aggregate house price risk as in \cite{Corbae2015} in Appendix \ref{sec:robust_pricerisk} and find that the quantitative results are almost unchanged. Homeowners incur proportional maintenance and depreciation costs $\delta$. Illiquidity is captured by proportional adjustment and moving costs on owner-occupied housing, as in \cite{Yang2009}:
\begin{equation}\label{eq:transcost}
adj(h_{a+1},h_a) = \begin{cases}
m_b p h_{a+1} & \text{if } h_{a}=h_r\mathbin{\&} h_{a+1}= h_o, \\
m_s p h_a 	& \text{if } h_a = h_o\mathbin{\&}h_{a+1}=h_r, \\
0 & \text{if } h_{a+1} = h_a,
\end{cases}
\end{equation} where $m_s$ and $m_b$ denote selling and buying costs, respectively. A household enters the period house $h_a$, while $h_{a+1}$ is the house chosen in this period.

\textit{Financial Market:} Households can save in one-period bonds that pay the risk-free rate $r$. Households can also borrow in one-period risk-free bonds (``mortgage'') at an interest rate $r + r^m$, where $r^m$ is the mortgage premium. However, only owners may borrow, subject to a loan-to-value (LTV) constraint:
\begin{align*}
\begin{cases}
b\ge - LTV \times p h_{a+1} & \text{if } h_{a+1} = h_o, \\ 
b\ge 0 & \text{if } h_{a+1} = h_r.
\end{cases}
\end{align*}

In the U.S., borrowers making low down payments typically pay for private mortgage insurance (PMI) until their home equity surpasses a threshold \citep{goodman2017sixty}. I model this as an extra fee $r^{pmi}$ that households must pay until they reach the $\overline{PMI}$ threshold. Since the mortgage premium is positive, households never hold both a mortgage and savings in the bond, and the interest rate on the net bond $b$ is given by
\begin{equation}\label{eq:rb}
r(b) = \begin{cases}
r & \text{ if } b\ge0, \\
r +r^m & \text{ if } b<0, \\
r +r^m +r^{pmi}& \text{ if } b < \overline{PMI}\times p\times h_o.
\end{cases}
\end{equation}

\textit{Income Endowment:} Households are endowed with an income process with a deterministic life-cycle profile $l_a$. Children face persistent idiosyncratic age-dependent productivity shocks $y_{i,a}\in\mathcal{Y}_a=\{y_1,\dots,y_{N_y}\}$, which follow a Markov chain, where $\pi_a(y'|y)$ is the probability of switching from state $y$ to $y'$ at age $a$. Parents face no income uncertainty. Consequently, income of household $i$ at age $a$ is given by
\begin{align}
w_{i,a} &= l_ay_{i,a} \; \forall a\in\{25,27,\dots,53\}, \label{eq:wk} \\
w_{i,a} &= l_a \; \forall a\in\{55,57,\dots,83\}. \label{eq:wp}
\end{align}
I assume that parents face no income risk for simplicity, consistent with the decrease in income risk with age \citep{Sanchez2020}. Furthermore, this paper focuses on the role of parental transfers for children' housing choices, where parental income risk is not a first-order concern. However, in Appendix \ref{sec:robust_incomerisk}, I show that parental income and health expenditure risks do not meaningfully change my results.

\textit{Initial Conditions of the New Child:} Households' initial wealth and productivity levels are stochastic yet correlated with the parent's wealth and productivity at age 53: $x_{25},y_{25} \sim F(x_{53},y_{53})$. While this paper does not consider parental investment in their child's human capital, such investments are an important factor in intergenerational persistence of economic outcomes \citep{Daruich2018,Lee2019}. The function $F$ allows me to match the intergenerational correlation in initial wealth and income levels, so the transfer motives are not over- or understated. Finally, all households begin as renters.

\textit{Timing:} The child's productivity is realized first. If the dynasty has a newly economically active child, they also realize their initial wealth. Next, the decision problem is separated into two stages. First, the parent chooses consumption $c_p$, housing $h_p$, savings position $b_p$, and nonnegative inter-vivos transfers $t_p$. After observing the parent's choices, the child in the second stage chooses consumption $c_c$, housing $h_c$, and savings $b_c$. The parent acts first to align with U.S. mortgage regulations, which mandate that gifts be deposited before approving mortgages.

\subsection{Household Decision Problems}
I now present the recursive formulation and the decision problems. For clarity, I omit individual subscripts and apply a prime superscript to all $a+1$ variables.

\textit{State Variables:} A parent $p$'s state variables are child's wealth $x_c\in X=[0,\infty)$, parent's wealth $x_p\in X$, child's housing $h_c\in H = \{h_r,h_o\}$, parent's housing $h_p\in H$, child's productivity $y_c\in Y_a=\{y_1,\dots,y_{N_y}\}$, and age $a_c\in A_c\in\{25,27,\dots,53\}$. The parent's state variable vector is $\mathbf{s}_p\equiv\left(x_c,x_p,y_c,h_c,h_p,a_c\right)$.

The child's state space differs because the parent's choices influence the child's decisions. The child is indifferent to receiving an additional unit of wealth or transfers, focusing only on total wealth (including transfers), $\tilde x_c\equiv x_c+t_p$. Furthermore, the parent's wealth and housing at the beginning of the period is redundant; the child only cares about the parent's bond and housing choices. The child's state variable vector is $\mathbf{s}_c=\left(b_p',h'_p,x_c+t_p,y_c,h_c,a_c\right)$.

\textit{Decision Problems:} 
I now show the decision problems for a dynasty at age $a_c=25$. For simplicity, I assume the child is a renter choosing to become an owner, while the parent remains a renter. Appendix \ref{sec:decextra} describes the decision problems in the final period of the current parent's life.

\textit{Child - Second Stage:} The child, conditional on becoming a new owner ($h_c=h_r,h'_c=h_o$), chooses consumption and bonds to maximize expected utility:
\begin{equation}\label{eq:Vk}
\begin{split}
 V_c(\mathbf{s}_c)^{own} = \max_{c_c,b'_c,h'_c=h_o} & u(c_c,h'_c) + \beta \E\left[V_{c}(\mathbf{s}_c') \right] \\ 
 \text{s.t.}\quad & 	b'_c = x_c + t_p + w_c - c_c - p h'_c - m_b p h'_c \\
 & x_c' = b'_c(1+r(b'_c)) + p h'_c(1-\delta) \\
 & b_c' \ge - LTV p h'_c. 
\end{split}
\end{equation}
The first constraint is the budget constraint, which determines the net bond position as available resources (net worth, transfers, and income) minus expenditures (consumption, purchase price, and purchase cost). The second constraint is the law of motion for net worth, which is the sum of the bond position and housing wealth after depreciation. The third constraint is the LTV constraint. We see that a transfer $t_p$ can move the child away from the LTV constraint by increasing the bond position $b_c'$, holding consumption and housing fixed.

For the other three possible housing choices (remain renting, remain owning, and own to rent), the problem is similar, but with appropriate changes to the budget equation, law of motion, and borrowing constraint. Finally, the child must choose between the two housing options:
\begin{equation}
	V_c(\mathbf{s}_c) = \max_{h_c'} \left\{V_c(\mathbf{s}_c)^{rent},V_c(\mathbf{s}_c)^{own}\right\}.
\end{equation}

\textit{Parent - First Stage:} The parent, conditional on remaining a renter ($h_p=h'_p=h_r$), chooses consumption, transfers, and bonds to maximize expected utility:
\begin{equation}\label{eq:Vp}
\begin{split}
V_p(\mathbf{s}_p)^{rent} = \max_{c_p,b_p',t_p} & u(c_p,h'_p) + {\eta} u\left (c_c^*({\mathbf{s}_c}),h_c'^*({\mathbf{s}_c})\right ) + \beta \E \left[V_{p}({\mathbf{s}_p'})\right] \\
\text{s.t.}\quad & b'_p = x_p + w_p - c_p - t_p - q p h_p'\\
& x_p' = b'_p(1+r(b'_p)) \\
& t_p\ge0, \\
& b_p'\ge 0. \\
\end{split}
\end{equation}
The first constraint is the budget constraint, which determines the net bond position as available resources (net worth and income) minus expenditures (consumption, transfers, and rent). The law of motion of net worth is simply the bond position for renters. The third and fourth constraints impose that transfers be nonnegative and that renters may not borrow, respectively. Finally, the parent chooses whether to rent or own.

\textit{Next-Period States:}
When making decisions, households consider how their choices will impact future states, as we can see from the next-period state vectors. For example, the parent internalizes that their next-period state includes the child's next-period wealth and housing, which are functions of the child's current choices, which in turn depend on the parent's current choices:
\begin{equation}
	\mathbf{s}_p' = \left(x_c'^{*}({\mathbf{s}_c}),x_p',y_c',h'^*_c({\mathbf{s}_c}),h'_p,a_c'\right).
\end{equation}
Similarly, the child knows that their next-period state includes the parent's next-period bond, housing, and transfer choices, which depend on the parent's next-period state, which in turn depends on the child's current choices:
\begin{equation}
	\mathbf{s}_c' = (b_p''^*({\mathbf{s}_p'}),h_p''^*({\mathbf{s}_p'}),x'_c + t_p'^*({\mathbf{s}_p'}),y'_c,h_c',a_c').
\end{equation}

\subsection{Markov Perfect Equilibrium Definition}
A stationary equilibrium, which is Markov Perfect, consists of value functions $V_c(\mathbf{s}_c)$ and $V_p(\mathbf{s}_p)$; policy functions of the child $c_c'^*(\mathbf{s}_c)$, $b_c'^*(\mathbf{s}_c)$, $h_c'^*(\mathbf{s}_c)$; and the parent $c_p'^*(\mathbf{s}_p)$, $b_p'^*(\mathbf{s}_p)$, $h_p'^*(\mathbf{s}_p)$, $t_p'^*(\mathbf{s}_p)$; and a distribution of households $\psi(\mathbf{s}_p)$ such that i) in each repetition of the parent-child game, the parent's policy functions are optimal given the child's policy functions; ii) the child's policy functions are optimal given the parent's policy functions; and iii) the measure of households is invariant. I derive the measure $\psi$ of households in Appendix \ref{sec:hhmeasures}.

Finally, given that this is an infinitely repeated game, the equilibrium may not be unique. However, experiments with various starting positions have consistently led to the same equilibrium. I discuss the computational algorithm in Appendix \ref{sec:computational}.

\subsection{Intuition - Transfers and Homeownership}\label{sec:modelintuition}
Before turning to estimation, there are several points to highlight about how housing interacts with altruistic transfers. For a thorough discussion of how transfers can lead to overconsumption, see \cite{Altonji1997a} and \cite{Barczyk2014}, and for insights on how a lack of commitment limits risk sharing, see \cite{Attanasio2018} and \cite{Mommaerts2016}.

\subsubsection{When Do Parents Transfer?}
Due to nonconvexities introduced by housing, the policy and value functions are not differentiable everywhere. To focus on the underlying intuition, I will sidestep the technical issues and refer to \cite{Barczyk2020} and \cite{Chu2020} for a formal derivation.

Consider a a wealthy parent with a poor, borrowing-constrained child. This implies that the child does not save and must rent and, thus, consumes any additional wealth. By assumption, the child's policy functions are locally differentiable, with $\frac{\partial c'_c(\cdot)}{\partial t_p}=1$ and $\frac{\partial x'_c(\cdot)}{\partial t_p}=\frac{\partial h'_c(\cdot)}{\partial t_p}=0$. We can ignore the nonnegativity constraint on transfers, since the parent is wealthy. The optimality condition (from Eq. \ref{eq:Vp}) is
\begin{equation}\label{eq:FOC}
	u'(c_p,h_p) = \eta u'(c_c + t_p,h_c).
\end{equation}
This is the classic first-order condition from \cite{Altonji1997a} and says that the parent transfers until their own marginal utility equals the altruism-weighted marginal utility of the child. This is the fundamental driver of the transfer motives in the model and, as we will see, interacts with housing choices in interesting ways. As discussed extensively in \cite{Barczyk2020a} and \cite{Chu2020}, other transfer motives exist, such as ``pushes to autarky'', though these do not appear in my parameterization.

\subsubsection{Transfers and Homeownership}
In the model, homeownership is determined by a threshold rule: Households become homeowners once their wealth is sufficiently high. The presence of transfers both moves the threshold (direct effect) and shifts the wealth level (indirect effect). I now discuss each effect in isolation, though they interact in equilibrium.

First, for a child with a given state, a transfer increases wealth, potentially pushing the child above the threshold, increasing homeownership indirectly.

Second, altruism reduces the child's savings motive, and so fewer children accumulate enough wealth to cross the threshold, lowering homeownership indirectly.

Third, transfers alleviate the LTV constraint in two ways. First, the parent give transfers to alleviate the borrowing constraint for a child who would otherwise be unable to buy. Second, the parent give transfers to a borrowing-constrained child, who can only afford very low consumption, shifting the ownership threshold down. The first increases homeownership directly and the second indirectly.

Fourth, transfers mitigate the costs of illiquid housing. Consider an owning child who receives a series of bad income shocks. Without transfers, the child optimally downsizes and pays the sales cost to better align consumption and housing expenditures. This creates a strong transfer motive, as transfers can save the dynasty from paying the sales cost. This increases homeownership directly by shifting the ownership threshold down.

\subsubsection{Policy Functions and Strategic Behavior}\label{sec:policyfuncs}
Notably, both the child and the parent strategically internalize these effects. For example, a child can use illiquid housing to commit to having low liquid wealth, thereby increasing the marginal utility of wealth and, consequently, the transfer. Conversely, a parent can tie up wealth in illiquid housing, increasing the marginal utility of wealth and, consequently, decreasing transfers. These strategic considerations are illustrated in the child's policy functions for homeownership and savings (Fig. \ref{fig:pol}). The vertical dashed lines denote the homeownership thresholds and the solid lines the bond position.

First, altruism lowers the threshold for homeownership among households with some parental support: Wealthier parents make homeownership more attractive. An increase in parental wealth shifts the threshold further down.

Next, we observe the child's savings choices. Altruism reduces saving in two ways. First, the shift in the ownership threshold means that the child switches from saving to borrowing sooner. Second, even in regions where the child makes the same ownership choice, they save less. For example, on the far left where all households rent, the child saves less if the parent is richer. On the right, where all households own, the child takes out a larger mortgage, holding wealth fixed. Note that once the child becomes sufficiently wealthy, they behave as if there is no altruism. Comparing the orange and dark red lines, we find that an increase in parental wealth decreases saving at all wealth levels.

\begin{figure}\caption{Policy Functions for Homeownership and Savings}\label{fig:pol}
	\begin{center} 
	\includegraphics[width=0.6\textwidth]{../tabfig/model/pol_bond.pdf}
	\end{center}
	{\begin{footnotesize} \textit{Notes:} The dashed vertical lines indicate the threshold where the child switches from renting to owning. The horizontal dotted lines indicate the PMI and LTV thresholds. The child states $\mathbf{s}_c'$ are \input{../tabfig/model/pol_bond_state.tex}\unskip. The figure plots the policy functions in the estimated model. I discuss the estimation in Section \ref{sec:estimation}. The qualitative patterns remain consistent across parameterizations. \end{footnotesize}}
\end{figure}

More generally, the child's savings choices show that with altruism, households are more likely to choose to be liquidity constrained, often called hand-to-mouth, as they become homeowners with less wealth and are closer to the borrowing constraints. Altruism reduces the downside of liquidity constraints---namely, the inability to smooth income shocks through partial insurance---while creating the advantage of receiving larger transfers. Given the importance of hand-to-mouth households in aggregate consumption responses \citep[see, e.g.,][]{Kaplan2014a,Boar2020}, this is an important result.

\section{Parameter Selection}\label{sec:estimation}
To quantify the impact of parental transfers on homeownership and to perform policy evaluations, I estimate the model using data from the PSID and a standard two-step estimation procedure. First, parameters independent of the model structure are estimated from the data or sourced from the literature. Second, I estimate the remaining parameters using simulated method of moments (SMM). I validate the model using nontargeted moments from the PSID as well as experimental evidence from empirical studies.

For details on the PSID sample selection, see Section \ref{sec:PSID}. Income, wealth, and housing values are winsorized at the 1st and 99th percentiles to control for extreme values. All calculations use family weights.

\subsection{First Stage: External Parameters}
I report all externally calibrated parameters and their values in Table \ref{tab:calpar}.

\textit{Period Length:} Each period corresponds to two years, matching the PSID interview frequency. I report all parameters in biennial form (e.g., $\beta$ measures how much households discount two years).

\textit{Income Process:} The income process consists of a deterministic life-cycle component $l_a$ for all ages and a persistent shock $y_a$ for children.

To calibrate the deterministic component, I first find the average income for each age by year and then average over all years. I then regress average income on a fourth-order polynomial of age. The fitted values give the deterministic component for households aged 25 to 65. I find retirement income by dividing the average income for households aged 67 to 83 by the predicted income at age 66 (Figure \ref{fig:y_a}).

To calibrate the stochastic income process $y_{i,a}$, I first set $N_y=3$ to create a three-state Markov Chain. The sample is divided into age-specific income tertiles. Within each age, I set $v_{i,a}$ as the ratio of the median income for that tertile to the overall median income (Figure \ref{fig:y_ia}). The empirical age-dependent transition matrices provide the transition matrix (Figure \ref{fig:Piv}).

\textit{Housing Parameters:} I set the rental rate $q=0.10$ \citep{Davis2008}, housing depreciation $\delta=0.05$ \citep{Harding2007}, and sales cost $m_s=0.075$ and purchase cost $m_b=0.02$ \citep{Yang2009}. I set the price level to $p=\$232,918/h_o$ to match the average market value of owner-occupied houses for households aged 25 to 44 in my sample. The size $h_o$ is estimated internally.

\textit{Financial Parameters:} Based on \cite{goodman2017sixty}, I set $\overline{PMI}=0.8$ so that households with LTVs above 80\% must pay a PMI premium $r^{PMI}=0.03$. I set $LTV=0.90$, broadly in line with the literature, to account for the mass of loans with LTVs above 80\%. The interest rate on savings is set at $r = 0.04$ (2\% per annum), while the mortgage premium is set at $r^m=0.03$ (1.5\% per annum). Both parameters vary in the literature, but both are typically set around 1\% to 2\% per annum \citep[see e.g.,][]{Cocco2005b,Kaplan2020,Paz-Pardo2019}.
% Mortgage premium: 1 percent in Kaplan et al (0.33 wedge on top of 3% interest rate)
% Mortgage premium: 2 percent in Cocco (2005)
% Mortgage premium: 2 percent in Paz-Pardo

% Risk free rate: 2 percent in Cocco (2005)
% Paz Pardo: 1 + 1 percent in Paz-Pardo
% Kaplan et al 3%

\textit{Risk Aversion, Discount Factor, and Housing Expenditure Share:} I set risk aversion $\gamma=2.0$ and the discount factor $\beta=0.92$ (annualized to $\approx0.96$), standard values in life-cycle models without portfolio choice \citep[see e.g.,][]{Kaplan2020,Boar2018}. The parameter $\xi$ pins down the desired expenditure share of housing consumption and is set to 0.2. This parameter is typically set around 0.15 to 0.25 (e.g., \citealp{Kaplan2020,Chatterjee2015,Paz-Pardo2019,davis2011household}) based on the share of housing expenditures in personal consumption expenditures. Due to minimum house sizes, low-income households generally spend more on rent than implied by $\xi$.
%% ξ parameter:
% Chatterjee 2015: xi=0.15
% Kaplan 2020: xi = 0.16
% Paz Pardo> xi = 0.2
% Vestman (Sweden) xi = = 0.25
% Davis & Ortalo-Magne xi = = 0.24
% Fisher and Gervais (2011) beta = 0.951, gamma = 2.0,

%% beta:
% Boar 2021 (WP version:) 0.959 (WP version, calibrated to match wealth, risk aversion =2)
% Kaplan 2020: 0.964 (estimated internally, risk aversion =2)
% Paz-Pardo: 0.929 (but have stocks and risk aversion=4)

\textit{Initial Conditions of the Young $y_{c,25},x_{c,25}\sim F(x_{c,53},y_{c,53})$:} 
The initial distribution is set to match the intergenerational correlation in wealth and income. I estimate it nonparametrically for transparency. First, I limit the sample to households aged 24 to 26 with parents 16 to 40 years older. I then divide wealth and income into four and three quantiles, respectively, for both parents and children. The PDF $F(\cdot)$ is then given by the empirical probability of each parent-child combination. The wealth level of a quartile is defined by its median value.

\input{../tabfig/parameters.tex} 
\begin{table}
	\center
	\begin{threeparttable}
		\caption{Summary of Externally and Independently Estimated Parameter}\label{tab:calpar}
		\small
		\input{outside_parameters.tex}
		\footnotesize
		\textit{Notes:} Superscript $^\dagger$ denotes parameters that are bi-annualized.
	\end{threeparttable}
\end{table}

\subsection{Moments and Identification}
The remaining three parameters $\theta=\{\eta,\chi,h_o\}$ are estimated internally to minimize the distance between four simulated and empirical moments.

\textit{Identification and Moment Estimation:}
I derive all empirical moments from the PSID sample discussed in Section \ref{sec:PSID}. I calculate moments by aggregating over all years, assigning equal weight to each year.

The two most important moments to match are the homeownership and transfer receipt rates of young households, since this is a paper about ownership and transfers. Homeownership pins down the strength of the homeownership utility premium $\chi$ (with a higher premium, more households are homeowners, Fig. \ref{fig:chiid}), while the transfer rate pins down the strength of altruism $\eta$ (with stronger altruism, more households receive transfers, Fig. \ref{fig:etaid}). I exclude transfers below \$500, as they are unimportant for homeownership and including them would increase the transfer rate.

While the model and the PSID have the same two-year frequency, households were asked about transfers in the last calendar year. In my cleaned sample, the annual transfer rate is 24\%. Thus, the two-year transfer rate could range between 24\% and 48\%. To be very conservative, I assume that the two-year transfer rate equals the one-year transfer rate. My estimate of altruism $\eta$ should be viewed as a lower bound.\footnote{I have experimented with targeted higher transfer rates. This requires stronger altruism, which, in turn, increases homeownership motives, lowering the estimated utility premium $\chi$. Overall, the results are qualitatively similar to my conservative estimate, which understates the role of altruistic transfers.}

Finally, I target two moments that govern the selection and timing of ownership---the rent-to-income ratio of renters and wealth at first purchase---to ensure the appropriate selection into homeownership based on income and that marginal first-time owners have the correct wealth levels. The rent-to-income ratio is largely pinned down by the size of the owner-occupied house $h_o$ (larger owner-occupied houses mean that more high-income households rent, reducing the rent-to-income ratio). The wealth at purchase decreases with altruism $\eta$ and the utility premium $\chi$, as they lower the wealth threshold for ownership. For a further discussion of the identification, I refer to Appendix \ref{app:SMM} and Figure \ref{fig:idall}.

\subsection{Model Fit}
I estimate the remaining parameters $\theta=\{\eta,\chi,h_o\}$ as follows. First, let $\hat m$ denote the vector of targeted empirical moments. Given a parameter vector $\theta$, let $\hat m(\theta)$ denote the corresponding model-simulated moments. The estimated parameters minimize the distance between the empirical and simulated moments:
 \begin{equation}
\label{eq:SMM}
 \theta^\star(\Omega) = \arg\min_\theta \left\{[\hat m(\theta) - \hat m]'\Omega [\hat m(\theta) - \hat m]\right\},
\end{equation}
where $\Omega$ is a diagonal weighting matrix with elements equal to the inverse of the squared empirical moments, $1/\hat m^2$. This normalization ensures that no moment receives disproportionate weight due to its units. For details on the estimation procedure, see Appendix \ref{app:SMM}, where I also discuss how the global estimation procedure is used to verify identification and calculate standard errors.

The estimated parameters, reported in Table \ref{tab:esttable}, are consistent with previous literature. The altruism parameter $\eta$ is roughly in the middle of the range in the related literature \cite[see e.g.,][]{Boar2018,Barczyk2020a,Mommaerts2016,Lee2019}. The utility premium of ownership $\chi$ means that the consumption bundle feels about 30\% larger for owners than renters, roughly in line with previous estimates based on life-cycle models \cite[see e.g.,][]{McGee2019,Fisher2011}. The size of the owner-occupied unit relative to the rental is model-dependent, but my estimate falls near the middle of the range in \cite{Kaplan2020}, where the ratio varies from 0.78 to 4.4. The model matches all targeted moments well, though it is overidentified, with no moment being particularly poorly matched.

%% eta altruism
%Mommaerts 2015: (Limited Commitment) eta = 0.08
% Barczyk 2020a: (0.45)
% Lee & Seshadri: 0.32
% Boar WP: 0.201

%% Extra preference
% Kaplan 2020: 1.015
% Fisher and Gervais 2011: ψy=0.95, ψf=0.87 => \chi = 1/0.95= 1.05 to 1/0.87=1.15
% McGee 1.15

%% House size:
% Kaplan Rentals: {1.17, 1.50, 1.92} owners {1.50, 1.92, 2.46, 3.15, 4.03, 5.15}: Average is 1.98, min,max = (1.5/1.92,5.15/1.17) = (0.78,4.4)
%

\begin{table}
	\center 
	\begin{threeparttable}
	\caption{Model Estimation}\label{tab:esttable}
		\input{../tabfig/est/esttable.tex}

		\begin{tablenotes}[para,flushleft]
			\footnotesize
			\textit{Notes:} Standard errors calculated by estimating the model to 100 bootstrapped empirical samples.
		\end{tablenotes}
	\end{threeparttable}
\end{table}

\subsubsection{Nontargeted Moments}\label{sec:externalval}
To validate the model, I show that it also matches nontargeted moments.

First, with a calibrated discount factor of $\beta=0.92$, the model matches the wealth levels of both young and old households. Second, the model replicates the correlation between parental wealth and children's homeownership. In the PSID, young homeowners' parents are 2.5 times wealthier as young renters' parents. The model predicts a similar, though slightly lower, ratio of about 2. This parental wealth gradient of homeownership captures the correlation between parental wealth and children's homeownership.

Next, the model aligns with other moments related to homeownership and housing decisions. First, it matches the age of first homeownership well, overshooting it slightly. Second, it somewhat overpredicts the aggregate homeownership rate of 65\%, primarily due to an excess of homeowner parents, partly due to their lack of income risk (see Appendix \ref{sec:robust_incomerisk}). Third, households hold slightly smaller mortgages than in the data, as their flexibility allows households to pay down debt quickly and borrow later if necessary. Fourth, the model also matches the LTV ratio for buyers well.\footnote{Note that this is not the LTV at origination but the first time the household is observed as an owner. If a household bought just after the previous survey, the LTV would decrease due to price appreciation, improvements, and mortgage paydown.}

Finally, I report the share of households that received a transfer in the two years before, during, and after a house purchase. The model matches this number well, indicating that not only are the right households purchasing at the right time, but also that buyers are receiving transfers at the right time.

\begin{table}
	\center 
	\begin{threeparttable}
	\caption{Non-Targeted Moments}\label{tab:nontargeted}
	\input{../tabfig/est/extramoments_mostrecent.tex}
	\footnotesize 
	\textit{Notes:} Wealth is measured in 1000s of 2016 US dollars.
\end{threeparttable}
\end{table}

\subsubsection{Household-Level Responses to Income Shocks by Parental Wealth}
As a final external validation, I repeat the exercise in Section \ref{sec:eventstudy} (see Figure \ref{fig:housinggrowthrates}), where we saw that households with wealthy parents do not downsize during unemployment spells. Since the model lacks an explicit unemployment state, households are considered unemployed if their productivity drops to the lowest level for a single period. Otherwise, I replicate the exercise exactly. Figure \ref{fig:housinggrowthrates_model} shows that the model matches both the qualitative and quantitative patterns from the empirical data: Housing consumption falls for households without wealthy parents during ``unemployment,'' while it grows for those with wealthy parents.

\begin{figure}
	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates_model}
	\includegraphics[width=0.5\textwidth]{../tabfig/model_housinggrowthpoor_both}%
	\includegraphics[width=0.5\textwidth]{../tabfig/model_housinggrowthrich_both}

	{\begin{footnotesize} \textit{Notes:} Lines denote means, and shaded areas indicate 95\% confidence intervals. The light orange area denotes the empirical estimates (Section \ref{sec:eventstudy}), and the darker grey areas denote the results from the simulated panel. The sample consists of households aged 25-45 with exactly one unemployment spell and no changes in head or spouse in the four years before and after unemployment. \end{footnotesize}}
\end{figure}

Taking stock, this section shows that the model can match targeted moments regarding homeownership, transfers, and the transition to homeownership with realistic preference parameter values. The model also matches a set of untargeted moments. Perhaps more importantly, the model replicates the household-level associations between parental wealth and downsizing during income losses.

\section{Homeownership Rates Without Transfers?}\label{sec:quant}
How important are parental transfers for homeownership? I simulate the model without altruism ($\eta=0$), holding all other parameters constant---the standard model without parent-child interactions. In the main exercise, I assume a perfectly elastic housing supply, meaning that changes in homeownership have no effect on rental or house prices. In Section \ref{sec:endoprice}, I show that the results remain robust even with endogenous housing supply and varying elasticities.

\subsection{The Effect of Parental Transfers on Homeownership}
Table \ref{tab:modelfitnoa} reports how the moments change without altruism and transfers. Homeownership rates decline by 14 percentage points from 49\%, a decrease of 29\% for households aged 25 to 44. This is a large decrease, since households endogenously choose to save more without transfers. By comparison, homeownership only decreased by about 8 percentage points during the financial crisis.

The reduction in homeownership cannot be attributed to an inability to afford buying: Without altruism, median wealth of the young increases (increasing homeownership, in isolation) to the average wealth level at purchase with altruism. The main driver is that the ownership threshold shifts out in wealth without altruism, as we saw in the policy functions (Section \ref{sec:policyfuncs}). Without transfers then, the average wealth at purchase increases by about \$12,500 (about 33\%), the age of first ownership is delayed by about 2.5 years, and the LTV at purchase falls by 13 percentage points. Although homeownership of young households declines sharply, the overall homeownership rate of all households (ages 25 to 74) only falls by 6 percentage points (8\%). Hence, the main aggregate effect of parental transfers is that they induce earlier homeownership. Finally, without altruism, the parental wealth gradient in homeownership decreases from 1.94 to 1.14, with the remaining effect driven by persistent productivity through initial conditions $F()$.

\begin{table}
	\center 
	\begin{threeparttable}
		\caption{Homeownership Decreases while Wealth Increases Without Altruism}\label{tab:modelfitnoa}
		\input{../tabfig/postest/quant_mostrecent.tex}
		\footnotesize
		\textit{Notes:} Wealth is measured in 1000s of 2016 US dollars.
	\end{threeparttable}
\end{table}

\subsection{Endogenous House Prices}\label{sec:endoprice}
I previously assumed that the supply of homes would fully adjust without affecting prices or rents, reflecting perfectly elastic supply. I now show that the results are similar even when housing supply is endogenous and under different elasticities.

I assume that housing supply ($H^S$) is log-linear in the house price:
\begin{equation}
\label{eq:hsupply}
\ln H^S = \alpha_0 + \alpha_1 \ln p,
\end{equation}
where $\alpha_1$ is the aggregate elasticity of supply to prices. I use a closed-form supply function because the model is too complex to introduce a construction sector. The rent-to-price ratio is unchanged, as in \cite{Kaplan2020}, where rental units are supplied by deep-pocketed landlords who convert rentals to owner-occupied units if the price deviates from the present value of rents. Letting $\alpha_1\to\infty$ yields a perfectly elastic supply function, leaving prices constant, as in the previous section. I first set the elasticity to 3, a typical estimated value for the United States \cite[see e.g.,][]{saiz2010geographic,aastveit2023changing}. Finally, I set the elasticity to 1 as a reasonable lower bound for the long-run aggregate elasticity.

\begin{table}
	\center 
	\begin{threeparttable}
		\caption{Housing Supply Elasticities Not Quantitatively Important}\label{tab:quant_endogenprices}

		\begin{tabular}{@{}llll@{}}
			\input{../tabfig/postest/quant_endogenprice_mostrecent.tex}
		\end{tabular}

		\footnotesize
				\textit{Notes:} I calibrate the housing supply elasticity from equation \eqref{eq:hsupply} as follows. For any elasticity $\alpha_1$, I set $\alpha_0$ to be such that housing supply would equal housing demand in the benchmark model with altruism: $\alpha_0(\alpha_1) = \ln(H^d) - \alpha_1 \ln p$, where $p$ is the benchmark price level (Table \ref{tab:calpar}). The ``Elastic'' column is the same as in the benchmark specification (Table \ref{tab:modelfitnoa}).
	\end{threeparttable}
\end{table}

Table \ref{tab:quant_endogenprices} reports the results. Allowing house prices to adjust does not meaningfully decrease the contribution of transfers to homeownership or other aggregate outcomes. This is due to the small difference in homeownership rates (ages 25 to 74) with and without altruism. Prices decline by just 7\% even with a supply elasticity as low as 1. Thus, transfers account for between 10 and 13 percentage points of the homeownership rate, depending on the supply elasticity.

\section{Policy, Illiquidity, and Racial Gaps}\label{sec:pol}
I have three main results regarding parental wealth and children's homeownership: (1) its relationship with housing-related policies, (2) how housing illiquidity amplifies its impact, and (3) its role in exacerbating the Black-White homeownership gap.

\subsection{Policies and the Importance of Parental Transfers}
Many countries have policies aimed at increasing homeownership, and housing affordability and delayed homeownership is at the forefront of economic policy and research \citep[see e.g.,][]{Mabille2020}. I now evaluate how five different policy levers influence the role of parental wealth in housing outcomes. The results are reported in Table \ref{tab:policy}. For each policy, I change the relevant parameter and solve the model before simulating a new distribution.

\textit{Relaxing Mortgaging Regulation:} Down-payment requirements are widely seen as a barrier to homeownership \citep[e.g.,][]{Lee2018,Guiso2002}. I tighten the LTV requirement by 3 percentage points to explore how mortgage regulation affects the parental wealth-housing channel. The tighter LTV limit decreases homeownership among the young by 5 percentage points (10\%), as households must have larger down payments. Homeownership falls more among households with parents in the middle of the wealth distribution, reducing the parental wealth gradient. While at first counterintuitive, the borrowing constraint is more binding for these households, all else being equal, since homeownership is more attractive when you have access to some parental wealth and a small transfer pushes you into ownership. However, the parents aren't wealthy enough to render the borrowing constraint irrelevant. This result suggests that relaxing borrowing constraints increases not only homeownership for all households, but also the role of parental wealth. Indeed, \cite{wold2024housing} find that households with poorer parents have lower LTV ratios at origination, suggesting that the LTV constraint is less binding for households with poorer parents.

\textit{Lowering PMI:} Conventional mortgages require that mortgages with down payments below 20\% have PMI \cite{goodman2017sixty}. I relax the PMI limit by 7 percentage points to explore how PMI affects the parental wealth-housing channel. The relaxed PMI limit increases homeownership by 2 percentage points (4\%). Similar to the LTV regulation, homeownership among households with parents in the middle of the wealth distribution rises more, further widening the parental wealth gradient. This result suggests that relaxing the PMI limit increases not only homeownership for all households, but also the role of parental wealth.

\textit{Lowering House Prices:} Although policymakers cannot directly control house prices, they can implement policies to lower them, such as by incentivizing supply. For example, both presidential candidates in the 2024 U.S. election promised to increase new construction. I lower house prices by \$6,000 (7\%) to explore how the price level affects the parental wealth-housing channel. As prices decline, homeownership increases for all households, though more so for households with the richest parents, slightly increasing the role of parental wealth.

\textit{Reducing Purchase Costs:} Another way to lower entry barriers is to reduce the costs of buying a house, and policymakers have lowered government fees and closing costs for first-time buyers. For example, in 2017, the stamp duty was lowered for first-time buyers in England. I decrease the purchase cost ($m_b$) by 2 percentage points (equivalent to a grant of about \$6,250) to understand how purchase costs affect the parental wealth-housing channel. While lower purchase costs increase homeownership, the gains are concentrated among households with wealthy parents, amplifying the parental wealth gradient. For these households, purchase costs were a key barrier to homeownership, because they are less likely to downsize after income shocks (Figure \ref{fig:housinggrowthrates_model}). This suggests that grants or credits for first-time buyers primarily benefit households with the wealthiest parents.

\textit{Reducing Sales Costs:} The earlier empirical results have highlighted how parental wealth may ameliorate housing illiquidity. I lower the sales cost $m_s$ from \parms\% to 5.5\% to understand how sales costs affect the parental wealth-housing channel. Strikingly, the homeownership rate decreases by 2 percentage points (4\%). This counterintuitive result is explained by selection into ownership. Homeownership among children with the wealthiest and poorest parents increases just a touch. In contrast, homeownership declines by 9 percentage points for households with moderately wealthy parents, leading to a net decline in homeownership. Homeownership decreases for households with moderately wealthy parents, as it is harder to use housing to extract transfers. For households with the poorest parents, who rarely get transfers, and the wealthiest parents, who frequently get transfers, there is no need to use the sales cost to ``extract'' future transfers.

\begin{table} 
	\centering
	\resizebox{\textwidth}{!}{\begin{threeparttable}
		\caption{Homeownership Policies on the role of Parental Wealth}\label{tab:policy}

ratio
		\begin{tabular}{@{}llll@{}}
			\input{../tabfig/postest/policy_mostrecent.tex}
		\end{tabular}

		\footnotesize
			\textit{Notes:} The results are from stationary distributions and ignore transition dynamics. Wealth is measured in 1000s of 2016 US dollars.
	\end{threeparttable}}
\end{table}

These experiments highlight two main results. First, parental wealth is an important determinant of homeownership, and having richer parents makes households more likely to prefer owning to renting. Some common policies intended to increase homeownership rates change the role of parental wealth in housing outcomes. Relaxing credit constraints increases not only homeownership, but the role of parental wealth, since housing is more attractive to households with wealthier parents.

However, increasing liquidity, either through reducing purchase costs or sales costs, decreases the role of parental wealth. Increasing liquidity can lower homeownership, as children of wealthy parents can no longer use housing as a commitment device to receive larger transfers.

\subsection{Illiquidity Preference}\label{sec:adjcost}
The previous section established that higher sales costs can increase ownership among households with wealthy parents. Here, I present a more surprising result: Households with wealthy parents prefer illiquidity.

I solve the model both with and without sales costs ($m_s=0$). For each household in the stationary distribution with sales costs, I assess whether they would prefer that neither they nor future generations in their dynasty face sales costs. This hypothetical corresponds to each member in the dynasty deciding whether they would prefer that a benevolent agent pays the sales costs for current and future generations. The results are reported in Table \ref{tab:adjcost}.

Without altruism, all households prefer to remove sales costs. I find that while all parents prefer to remove sales costs, roughly one-quarter of children prefer them. The key determinant is parental wealth: Parents of children who prefer sales costs are approximately three times wealthier. Children who prefer sales costs are also more likely to receive transfers (48\% vs. 13\%) and, conditional on receiving transfers, receive three times larger transfers. Notably, households that own their homes are more likely to prefer sales costs on their housing.

\begin{table}
	\center
	\caption{Household Observables and Support for Keeping Adjustment Cost}\label{tab:adjcost}
	\begin{threeparttable}
		\input{../tabfig/preferliq/tab_mostrecent.tex}
		\footnotesize
		\textit{Notes:} Only includes dynasties where the child household is aged 25-44. Wealth is measured in 1000s of 2016 US dollars.
	\end{threeparttable}
\end{table}

To explore why some households prefer illiquidity, I break the sample down by liquidity preference and age in Figure \ref{fig:prefliq}. Figure \ref{fig:pref} plots the share of children who would prefer to keep housing illiquid. Illiquidity preference decreases with age as reliance on parental transfers decreases, and roughly 50\% of the youngest households prefer illiquidity. The next figures plot average homeownership, parental wealth, and child wealth by liquidity preference. To use illiquid housing to extract transfers, households must be homeowners, so households who prefer illiquidity have higher homeownership (Figure \ref{fig:pref_own}). To extract transfers, parents must have wealth, so households who prefer illiquidity have wealthier parents (Figure \ref{fig:pref_wealthp}). Finally, Figure \ref{fig:pref_wealthk} plots the child's wealth, and we see that households who prefer illiquidity are initially wealthier, but this relationship flips at age 33. The relationship flips because households only benefit from having their housing subject to sales costs if they are homeowners, which requires some wealth. To continue benefiting, they must remain liquidity-constrained (i.e., homeowners with limited wealth) to induce transfers.
\begin{figure}[tb]
	\captionsetup[subfigure]{aboveskip=-1pt}
	\caption{Liquidity Preference over Age}\label{fig:prefliq}
	\begin{subfigure}{0.5\textwidth}%
		\caption{Prefer Costs}\label{fig:pref}%
		\includegraphics[width=\textwidth]{../tabfig/preferliq/prefer.pdf}%
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.5\textwidth}%
		\caption{Child's Homeownership}\label{fig:pref_own}%
		\includegraphics[width=\textwidth]{../tabfig/preferliq/prefer_owner.pdf}%	
	\end{subfigure} 
	\begin{subfigure}{0.5\textwidth}%
		\caption{Parent Wealth}\label{fig:pref_wealthp}%
		\includegraphics[width=\textwidth]{../tabfig/preferliq/prefer_wealthp.pdf}%
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.5\textwidth}%
		\caption{Child Wealth}\label{fig:pref_wealthk}%
		\includegraphics[width=\textwidth]{../tabfig/preferliq/prefer_wealthk.pdf}%	
	\end{subfigure}%
	\caption*{\footnotesize \textit{Notes:} These figures break down the sample of households by whether they prefer illiquidity.}
\end{figure}

To summarize, this section shows that altruism generates illiquidity preference. This is similar to models of behavioral biases, where households use the illiquidity of housing as a commitment device \cite{attanasio2024temptation}. In both settings, households use illiquidity to keep their marginal utility of wealth high in future periods. With altruism, this is done to induce future transfers. With time inconsistency, this is done to ensure that households will save in the future.

\subsection{Black-White Homeownership Gap}
Racial disparities in economic and financial outcomes are receiving increased attention. In the United States, young White households are nearly twice as likely to own homes as young Black households.

Several structural papers have used life-cycle models with overlapping generations to study differences in economic outcomes by race. \cite{Ashman2020} find that the Black-White income gap accounts for most of the 90\% wealth gap. Similarly, \cite{aliprantis2022dynamics} find that closing the income gap is the most effective mechanism to reduce the wealth gap. While these papers account for racial differences in inheritances, they do not include endogenous inter-vivos transfers and ignore housing.\footnote{\cite{aliprantis2022dynamics} include a form of inter-vivos transfer receipt: Households leave pooled bequests at death, primarily to middle-aged households. There are no strategic considerations, since the bequests are stochastically distributed and parents do not transfer directly to their children.} While the structural papers have omitted housing, many empirical papers have looked at racial differences in housing outcomes. For example, \cite{charles2002transition} find that differences in parental wealth account for 25\% of the Black-White mortgage application gap, and \cite{bond2021role} find that parental wealth accounts for 28\% of the gap in maintaining homeownership. I now use my model to bridge the gap between the structural and empirical papers and to quantify the contribution of parental wealth to the Black-White Homeownership gap.
\begin{table}
	\center
\begin{threeparttable}[tb]
				\singlespacing
		%			\footnotesize
		\caption{Black-White Homeownership Rate}\label{tab:BW_table}
		\input{../tabfig/postest/BW_mostrecent.tex}
		\footnotesize
		\textit{Notes:} Wealth is measured in 1000s of 2016 US dollars. See main text for details.
	\end{threeparttable}
\end{table}

I aim to bridge the gap between these structural approaches and empirical results by using my model to quantify the contribution of parental transfers to the Black-White homeownership gap. I follow the methodology in \cite{Ashman2020} and \cite{aliprantis2022dynamics} closely. I shift the income level ($l_a$) according to racial differences: In the matched parent-child household, White households have 8\% higher income and Black households 42\% lower than the average. I then solve the model for White households. For Black households, I reduce house size by 25\% to match observed homeownership rates. I then solve the models both with and without altruism. The results are reported in Table \ref{tab:BW_table}.

The observed Black-White homeownership gap is $\frac{0.54 - 0.27}{0.54}=50\%$ and 52\% in the model with altruism. Without altruism, the Black-White homeownership gap decreases to 45\%. Transfers account for 7 percentage points (14\%) of the Black-White gap. Though Black and White households receive transfers at similar rates in the model, transfers are less important for Black households' homeownership, as they receive smaller amounts due to lower parental wealth. My results show that racial differences in parental income, which affect parental wealth and transfers, are significant contributors to racial gaps in housing outcomes. Combined with the previous policy analysis, these results suggest that policies such as tax rebates or down-payment assistance for low-income or first-time buyers may widen the Black-White homeownership gap rather than narrow it. However, the model omits many other factors beyond income, such as discriminatory lending and appraisal bias, that affect the Black-White homeownership gap (e.g., \cite{shapiro2004hidden}).

\section{Conclusion}
In this paper, I demonstrate that parental transfers play a large role in households' homeownership decisions. I build and estimate a dynamic life-cycle model of homeownership where altruistic parents may transfer to their adult children in every period. I use the model to quantify the importance of parental transfers for young adults' homeownership and to understand how policies that lower barriers to homeownership interact with these transfers.

Using a counterfactual experiment without altruism (a standard life-cycle model), I find that transfers account for 14 percentage points (29\%) of the homeownership rate. I show that policies that lower entry barriers to homeownership amplify the importance of parental wealth on housing outcomes, while those that lower the downsides of illiquid homeownership reduce the importance of parental wealth. Finally, I study the role of housing illiquidity and parental transfers. My paper thus contributes to a better understanding of the determinants of homeownership and to the growing literature on the role of family economics in shaping macroeconomic outcomes (e.g., \cite{Doepke2016a,Daruich2018}).

The second contribution of the paper is to include an illiquid asset in the altruism framework. With altruistic parents, transfers generally flow to borrowing-constrained households, since they have a large marginal utility of wealth \citep{Barczyk2020,Chu2020}. With a single asset, this framework implies that households who receive transfers are poor. In the data, around 20\% of all households are ``wealthy hand-to-mouth'' and have positive wealth but no liquid wealth \citep{Kaplan2014a,Attanasio2018}. My paper bridges these two strands of the literature. First, I show that some households with wealthy parents choose to be liquidity-constrained to receive larger transfers. Second, I show that the lack of commitment generates preferences for illiquidity, even when households are rational and without behavioral biases \citep[see e.g.,][]{attanasio2024temptation}.

I made several simplifying assumptions to keep the model tractable. Adding richer housing dynamics and equilibrium effects would allow one to study other interesting questions, such as the extent to which transfers contribute to house price growth. My model omits a feedback loop, in which transfers raise demand and, thus, house prices, boosting the wealth of homeowning parents and, thus, higher transfers, and so on. Another potential extension could examine whether illiquidity may reduce the commitment problem in the family. Additionally, my preferences-for-illiquidity results have implications for the literature on high-MPC households \citep[see e.g.,][]{kaplan2022marginal}. In particular, we are likely overestimating the average MPC of liquidity-constrained households with wealthy parents, since these households use parental transfers to smooth shocks. For example, following a negative shock, consumption may fall less due to transfers, while for positive shocks, consumption may rise less as parents reduce transfers.

\newpage
\begingroup\singlespacing
\bibliographystyle{ecta}
\bibliography{bibliography}
\endgroup
\newpage

\appendix

\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\section{Quantitative Model Details}

\subsection{Decision Problems at Age 53 and 83}\label{sec:decextra}
As depicted in Figure \ref{fig:overview}, several transitions occur between age 53 and 55: The `old' parent dies, the `old' child becomes a `new' parent, and the `new' child becomes economically active.

These transitions mean that the child's decision problem at age 53 is very different from earlier ages (Eq. \ref{eq:Vk}). First, the continuation value is now the parent's value function at age 25 of the new child ($V_{p}({\mathbf{s}'_{p,25}})$). Second, the expectation is over the new child's initial productivity and net worth ($x_{c,25},v_{c,25}\sim F(x_{c,53},y_{c,53}$)), instead of the old child's productivity. Third, the law of motion now includes bequests from the old parent instead of a transfer. Fourth, it is necessary to map next-period variables for the old child into variables for the new parent. The problem, for a child aged 53 who choose to own, becomes:
\begin{equation*}
\begin{split}
V_c(\mathbf{s}_{c,53})^{own} = &\max_{c_{c,53},b'_{c,53},h'_{c,53}=h_o} u(c_{c,53},h'_{c,53}) + \beta \E\left[V_{p}({\mathbf{s}'_{p,25}}) \right] \\
\text{s.t.}\quad & 	b'_{c,53} = x_{c,53} + w_{c,53} - c_{c,53} - p h'_{c,53} - adj(h_{c,53},h'_{c,53}) \\
& x'_{p,25} = b'_{c,53}(1+r(b'_{c,53})) + x'_{p,53} \\
& b'_{c,53} \ge -LTV p h'_{c,53}, \\
& h_{p,25} = h'_{c,53},
\end{split}
\end{equation*} 
where $c,53$ and $p,53$ denote the variables associated with old child and old parent in this period, and $c,25$ and $p,25$ denote the variables associated with the new child and new parent in the next period. The last equation maps the old child's housing choice into the new parent's---the same household one period later---housing state. The same mapping applies to the law of motion (third equation).

The dying parent's decision problem undergoes similar changes. First, the continuation value is now the new parent's value function at age 25 of the new child, weighted by altruism ($\eta V_p(\mathbf{s}'_{p,25})$). The remaining changes (expectation, law of motion, and mapping of next-period variables) are the same as for the child and hence the decision problem is changed in the same way as for the child.

\subsection{Measures of Households}\label{sec:hhmeasures}
The state-space of a parent is $S_p = X_c \times X_p \times Y_c \times H_c \times H_p \times A_c$, with $\mathbf{s}_p$ denoting generic elements therein and $\mc{S}_p$ the associated Borel-$\sigma$ algebra. The state space of a child is $S_c= B_p \times H_p \times X_c \times Y_c \times H_c \times A_c$, where $B_p=\mathbb{R}$. For conciseness, I omit further definitions for the child. Let $\psi_p(\mathbf s_p)$ be a probability measure over $(S_p,\mc{S}_p)$ so that $\psi(\mathbf s_p)$ denotes the measure of households with state $\mathbf s_p$ (i.e., after the shock is realized but before choices are made). Finally, $\Psi_p$ denotes the corresponding cumulative distribution function. The mass of households for each age is normalized to 1/15.

\textit{Law of Motion for Dynasties with Children Aged 25-51:} The mass of households in state $\mathbf{s}_p$ is the mass of families adopting housing and savings policies that lead them to this state, adjusted by the probability of experiencing a given income shock:
\begin{equation}\label{eq:LoM}
\begin{split}
\psi(\mathbf{s}'_p) = \int_{\mathbf{s}_p\in \mathcal{S}_p} & 
\mathbf{1}_{\left\{ x'_p = x'^*_p(\mathbf s_p) \right\} }
\mathbf{1}_{\left\{ h'_p = h'^*_p(\mathbf s_p) \right\} } 
\mathbf{1}_{\left\{ x'_c = x'^*_c(\mathbf s_c(\mathbf s_p))\right\}} 
\mathbf{1}_{\left\{ h'_c = h'^*_c(\mathbf s_c(\mathbf s_p))\right\}} \times \\
&\pi(y'_c|y_c)\diff \psi(\mathbf s_p).
\end{split}
\end{equation}
Note that the child's state $\mathbf{s}_c(\mathbf{s}_p)$ in these expressions depend on the choices of the parent which depend on their state. For example, 
\begin{equation}
h'^*_c\left(\mathbf s_c(\mathbf s_p)\right) =h'^*_c\left(b_p'^*(\mathbf{s}_p),h_p'^*(\mathbf{s}_p'),x'_c + t_p'^*(\mathbf{s}_p'),y'_c,h_c',a_c+2 \right).
\end{equation}

\textit{Law of Motion for Children Aged 53:} In this special case, the distribution will depend on the choices of the new parent (old child), the now deceased parent and the stochastic initial conditions of the new child. 
\begin{equation}\label{eq:LoM55}
\begin{split}
\psi(\mathbf{s}'_p;a_c=25) = \int_{\mathbf{s}_p\in \mathcal{S}_p} & 
\mathbf{1}_{\left\{ x'_p = x^*_p(\mathbf s_p) + x^*_c(\mathbf s_c(\mathbf s_p)) \right\} }
\mathbf{1}_{\left\{ h'_p = h^*_p(\mathbf s_p) \right\} } 
\mathbf{1}_{\left\{ h_c' = h_r\right\}} \times \\
&F(x_c',y_c'|x_c,y_c) \diff \psi(\mathbf s_p;a_c=53),
\end{split}
\end{equation}
where we limit $s_p$ to only include the subset of the state-space where $a_c=53$. The initial wealth and productivity of the child depends on the wealth $x_c$ and productivity $y_c$ of the new parent at age 53. Further, all children start out as renters, and the next-period wealth of the new parent is savings plus bequests.

Finally, the function $\mathcal{H}$ operates on the distribution $\psi(\mathbf{s}_p)$ and policy functions %$g^*(\mathbf{s}_p):\mathbf{S}_p \rightarrow (X\times B \times T \times H)\times (X\times B \times H) \times (A_c)$ 
and maps them into a new distribution in accordance with equations (\ref{eq:LoM}, \ref{eq:LoM55}):
\begin{equation}\label{eq:DisIter}
\psi_{n+1} = \mathcal{H}(\psi_n,g^*),
\end{equation}
where the subscript denotes the iteration of the distribution. A stationary distribution is then a fixed point of equation \ref{eq:DisIter}.

\section{Estimation}\label{app:SMM}
In the internal estimation, I first draw $N=\parNest$ 3-dimensional candidate parameter vectors $\theta$, using Sobol sequencing. I then rank the vectors by the objective function, and use the top 5\% to set the bounds of a new Sobol sequence, from which I draw another {\parNest} candidate parameter vectors. I shrink the search space {\parNshrinks} times.

The global optimization procedure facilitates verifying identification. First, pick a parameter, say $h_o$, and divide it into 20 quantiles. The remaining parameters are uniformly distributed within each quantile. Next, find the 25th, 50th, and 75th percentiles within each quantile for a moment. We can then plot how the moment depends on the parameter by plotting the percentiles over quantiles. A moment is informative for a parameter if the percentiles shift as we move across the parameter's quantiles. A steeper slope indicates a more informative moment for the parameter. A parameter is relatively more important when the spread between the 25th and 75th percentiles is smaller.

The results of this exercise are plotted in Figure \ref{fig:idall}. For example, take the house size $h_o$ (Fig. \ref{fig:hoid}), where we see that the rent-to-income ratio decreases with the size of owner-occupied housing. The tightness of the 25th and 75th percentiles indicates that the other parameters have little effect on the rent-to-income ratio. The other moments are not sensitive to the house size. Finally, I also plot the SMM objective function (Eq. \ref{eq:SMM}) over $\eta$, showing that the global search space is wide enough. The results for the remaining parameters are shown in the other panels of Figure \ref{fig:idall}.

\begin{figure}\caption{Identification of $\eta,h_0,\chi$.}\label{fig:idall}
	\begin{subfigure}{\textwidth}\caption{Identification of $h_o$}\label{fig:hoid}
		\includegraphics[width=\textwidth]{../tabfig/est/identification/ho.pdf}
	\end{subfigure}
	\begin{subfigure}{\textwidth}\caption{Identification of $\eta$}\label{fig:etaid}
		\includegraphics[width=\textwidth]{../tabfig/est/identification/eta.pdf}
	\end{subfigure}
	\begin{subfigure}{\textwidth}\caption{Identification of $\chi$}\label{fig:chiid}
		\includegraphics[width=\textwidth]{../tabfig/est/identification/chi.pdf}
	\end{subfigure}
	{\begin{footnotesize} \textit{Notes:} Dashed grey lines denote the 25th and 75th percentiles, the solid black line the median, the solid red line the minimum (only used for the SMM objective function value), and the dashed orange horizontal line the empirical moment (not applicable for the SMM objective function).\end{footnotesize}}
\end{figure}

\section{Robustness Exercises}
\subsection{Parental Income Risk}\label{sec:robust_incomerisk}
In the benchmark model, households face no income risk after age 55. I now show that the results are robust to this assumption. To do so, I perform the following modifications. First, the labor income of parents is now assumed to be the product of a transitory productivity shock:
\begin{equation*}
w_{i,a} = l_a \nu_{i,a} \; \forall a\in\{55,57,\dots,83\}. \label{eq:wp2}
\end{equation*}
The process is calibrated as follows. First, I keep households aged 55-85. I subtract healthcare expenditures from household income as healthcare expenditure risk is a significant risk for older households \citep{denardi2024}. Next, I divide the sample into year-age specific income tertiles, find the median income within each age, and average over years to find the parent productivity shifter $\nu_{i,a}$. The results, fitted to a cubic trend, are plotted in Figure \ref{fig:nu}. The shock is transitory, and the PDF $\Pi(\nu)$ takes the value 1/3 for any outcome.

The results are reported in Table \ref{tab:robust}. We see that the introduction of income risk for the old has neglible effects and leaves the main findings intact. Income risk for parents lower their homeownership rate a touch while increasing their wealth a wee bit. Parental transfers still account for 14 percentange points of the homeownership rate of young households.

\subsection{Aggregate Price Risk}\label{sec:robust_pricerisk}
In the benchmark model house prices are certain. I now show that the results are robust to introducing price risk in the manner of \cite{Corbae2015}. This introduces a new state variable $z$ that denotes the aggregate price level, and the house price is either low, normal, or high depending on the value of $z$. The transition matrix for $z$ follows 
\begin{equation}
\Pi(z'|z) = \begin{bmatrix}
0.90 & 0.10 & 0.00 \\
0.02 & 0.96 & 0.02 \\
0.00 & 0.25 & 0.75 \\
\end{bmatrix}.
\end{equation}
The aggregate house prices are set to be $(p_l,p_n,p_h)=(0.7,1.0,1.3)p_{bench}$, where $p_{bench}$ is set to be the estimated value from Table \ref{tab:esttable}. I then solve the model with price uncertainty, but simulate the economy when house prices are at the normal level.

The results are reported in Table \ref{tab:robust}. We see that the introduction of aggregate price uncertainty increases savings for old households, with and without altruism. It also induces households to buy housing slightly later, with lower LTV's and higher wealth at purchase. However, the main quantitative finding remains: Parental transfers now account for 11pp (25\%) of the homeownership rate. I have also perform robustness checks where I allow for transitory idiosyncratic dividend/return shocks as in \cite{Chang2024} for homeowners, and find that, as expected, idiosyncratic price risk have even smaller quantitative effects.

\begin{figure}
    \caption{Calibration of Parental Income Shocks}\label{fig:nu}
	{\centering
    \includegraphics[width=0.5\textwidth]{../tabfig/empirical/lifecycleproductivity_old_3}
    \par}    {\begin{footnotesize} \textit{Notes:}
        Dashed lines are the empirical age-medians, and solid lines are fitted second-order polynomials used in model calibration. The lines denote the value of $\nu_{i,a}$ in the first tertile (black), middle (gray), and top (blue) by age.
    \end{footnotesize}}
\end{figure}

\begin{table}
	\center 
	\resizebox{\textwidth}{!}{
	\begin{threeparttable}
		\caption{Results Robust to Old-Age Risk and Uncertain House Prices}\label{tab:robust}
		\input{../tabfig/postest/robust_mostrecent.tex}
		\footnotesize
		%		\textit{Notes:} 
	\end{threeparttable}}
\end{table}

\section{Numerical Details}\label{sec:computational}
I now briefly discuss some details in the numerical solution of the model. Due to the nonconvex nature of the decision problems---due to discrete housing choices, noncontinuous transfer choices, and occasionally binding constraints---I use grid search. That is, I define a grid of possible choices for each choice variable, and for each state I loop over all possible choices, and the optimal policy is the one that maximizes the utility. The value and policy functions are linearly interpolated over child and parental wealth, conditional on the discrete choices. This solution algorithm satisfies both criteria established by \cite[p.30]{Barczyk2020}: ``algorithms should be able to deal with locally convex and even discontinuous value functions''. Naturally, the solution accuracy improves with denser grids. I use {\parnstate} nodes in the state vectors and {\parnchoice} nodes in choice variable grids.

\textit{Solving Decisions Problems:} 
The parent's choice can be written as a three-stage problem, which increases computational speed dramatically. First, the parent makes their housing choice. Second, the parent then chooses how much to consume. Third, the parent allocates the remainder between transfers and savings. \cite{Chu2020} shows how the consumption-transfer-savings choice can be separated into a two-stage problem, which is straightforward to extend to include housing.

\textit{Value Function Iteration Procedure:}
I use the following procedure for value function iteration. The initial guess is set to zero for all functions, except for consumption, set to 1.0, and housing, set to renting.
\begin{enumerate}
    \item Next, I solve the decision problem of the child at age 53, that is the final period before he becomes a parent. This yields $V_c(\cdot;a_c=53)$ and $g_c(\cdot;a_c=53)$. The solution to this problem depends on the value function of a new parent ($V_p(\cdot;a_p=55)$, which is set to 0), the intentional bequest left by the parent ($x_p^*(\cdot;a_p=83)$, aslo set to 0), but not the next-period transfer since then this child has become a parent and thus moves first and decides the transfer. The initial state of the new child is given by the joint distribution of initial productivity and wealth.
    \item Next, I solve the parent's decision problem in the terminal age of 83, when the child is 53. This yields $V_p(\cdot;a_c=53)$ and $g_p(\cdot;a_c=53)$, which depend on the policy functions for the child at age 53 ($g_c(\cdot;a_c=53)$)---found in the previous step---and the new parent's value function $V_p(\cdot;a_c=25)$---which we still haven't found and is set to 0.
    \item Next, I solve the child's decision problem at age 51. This yields $V_c(\cdot;a_c=51)$ and $g_c(\cdot;a_c=51)$, and depends on the value function we found in the first step ($V_c(\cdot;a_c=53)$) and the parent's policy function we found in the previous step ($g_p(\cdot;a_c=53)$).
    \item Next, I solve, the parent's decision problem at age 81. This yields $V_p(\cdot,a_c=51)$ and $g_p(\cdot,a_c=51)$, and depends on the policy function for the child we found in the previous step ($g_c(\cdot;a_c=51)$) and the value function we found in the second step ($V_p(\cdot;a_p=83)$).
    \item This is repeated backwards in age until the child is 25.
    \item Repeat these steps until the value functions converge.
\end{enumerate}

\textit{Fixed Point Convergence:} 
The parent-child interaction forms an infinitely repeated game within a dynasty, meaning the equilibrium may not be unique. In practice, value function iterations generally converge to a unique fixed poin, but when the altruism parameter is high, the equilibrium can become cyclical, with the solution alternating between iterations. For example, in iteration $n$ one solution appears, in $n+1$ another, and in $n+2$ the cycle returns. These behavioral differences are minor: the child homeownership rate varies by less than 0.3 percentage points, and average child wealth differs by less than 0.1 percent. While this cyclical behavior does not occur in the reported results, it sometimes occurs during the structural estimation, when the model is solved with higher altruism parameters $\eta$. In such cases, I simulate each dynasty once for every candidate solution.

Intuitively, this cyclical behavior can arise from the interaction between the discrete rent-or-own choice and strategic interactions. For example, a renting child may decide to buy given the parent's policy functions at a certain state. In the next iteration, the parent observes this and adjusts savings to discourage the child from buying, prompting the child to reoptimize. The cycle then repeats, with the child buying in one iteration and not buying in the next. When I solve the model without homeownership (reducing it to a standard altruistic consumption-savings model as in \cite{Barczyk2020a} and \cite{Chu2020}), the value function iterations consistently converge to a unique fixed point. Similarly, removing transaction costs, making housing more liquid and no longer a state variable, reduces the likelihood of cyclical fixed points. Denser grids also decreases the likelihood of cyclical behavior. In practice, the largest differences in policy functions within cycles are at the upper edges of the parental wealth grid. Since the upper edge is set so high that no households are close to it, this has little impact on the solution.

\textit{Interpolation Details:} The policy functions for consumption and savings are noncontinuous in both child and parental wealth which makes interpolation unattractive. However, the policy functions, conditional on choosing a specific house quality are smoother, and do not display large jumps in the quantitative model. Therefore, I use the following procedure. I first find the interpolated policies for each housing choice, and then find the probability that the parent/child makes this choice. This probability comes from the interpolated discrete choice function, which is binary ${0,1}$, while the interpolated policy itself (with flat extrapolation) is continuous over the interval $[0,1]$. The probability is between 0 and 1 only when the decision changes in one of the four nodes used in the interpolation. If the interpolated consumption choice of the child $c_c$ leads to a net bond position $b_c$ that violates the borrowing constraint I set $b_c$ equal to the constraint and reduce $c_c$ to make it hold.

\textit{Simulation of Households:} 
I simulate N={\parNdyn} dynasties. The initial states of each dynasty (initial child and initial old) $(x_p,h_p,x_c,v_c,h_c,a_c=25)$ is drawn from a five-dimensional joint uniform distribution. I then simulate all dynasties for {\parNdyn} generations, since the distribution, as measured by average homeownership, wealth, and, productivity levels stabilizes after four generations. I discard observations from the first four children and parents in each dynasty.

\textit{Computational Packages:}
The program to solve the model is written in Julia v1.10.4. I rely on the \texttt{interpolations.jl} v0.15.1 package for numerical interpolation routines. The Stata code used for empirical data and Julia code for the model is available through the author's website and Github.

\section{Supplementary Figures and Tables}
\begin{table}
	\small	
	\caption{Variable Definitions in the PSID}\label{tab:vardef}
	\begin{threeparttable}
	\begin{tabular}{@{}llll@{}}
		\toprule
		Variable& PSID code & Description & Note \\ \midrule
		\textit{Transfer Related} \\ 
		Received Transfer & ER67962 & 2/5years, gift/inherit \$10,000+ & Changing def. \\
		Gave Transfer & RT13V125 & Loans/gifts to child in 2012 & 2013
		prnt/chld file \\ 
		Transfer Amount & RT13V125 & Amount given in 2012 & 2013 prnt/chld file \\
		\textit{Other} \\ 
		Behind Mortgage & ER66062 & Behind on mortgage payments\\
		Income & ER65349 & Total Household Income & \\
		Employment & ER66164 & Working, Unemployed etc. \\
		House Value & ER60031 & Reported Market Value & \\
		Dollars Rent &	ER66090 & Monthly Rent\\
		Family Weight &		ER71570 & Weight of family unit & \\
		\bottomrule
	\end{tabular}
	{\textit{Notes:} This table lilsts the main variables I use from the PSID, as well as their variable code in the 2017 Main family level data set or the 2013 Transfer supplement.}
	\end{threeparttable}

\end{table}

\begin{figure}\caption{Calibrated Income Process}\label{fig:calincome}
	\begin{subfigure}{0.49\textwidth}\caption{Deterministic Income Profile $y_a$}\label{fig:y_a}
		\includegraphics[width=\textwidth]{../tabfig/empirical/lifecycleincome.pdf}%
	\end{subfigure} %
	%
	\begin{subfigure}{0.49\textwidth}\caption{Productivity Shifter $y_{i,a}$}\label{fig:y_ia}
		\includegraphics[width=\textwidth]{../tabfig/empirical/lifecycleproductivity_3.pdf}
	\end{subfigure}

	\begin{subfigure}{\textwidth}\caption{Age-State Dependent Transition Probabilities $\Pi(y_{i,a+2}|y_{i,a})$}\label{fig:Piv}
		\includegraphics[width=0.99\textwidth]{../tabfig/empirical/Pi_vk_3.pdf}

		{\begin{footnotesize}\textit{Notes:}
			Dashed lines are the empirical age-means and solid lines are the fitted third order polynomials used in the model calibration. Each line gives the probability of moving to the bottom tertile (black), middle (gray), and top (blue) income tertiles by the income tertile the child is currently in, by age. \end{footnotesize}}
	\end{subfigure}

	\begin{subfigure}{\textwidth}
		\caption{Initial Distribution $F(x_{53},y_{53})$ by wealth $x_{53}$ and productivity $y_{53}$}\label{fig:inidistr}
		\includegraphics[width=1\textwidth]{../tabfig/empirical/inidistr} 
		{\begin{footnotesize}\textit{Notes:} The vertial lines denote the first, second, and third productivty level for the child. Within each interval each point denotes a wealth quartile. For example, the first point on the black line in the left-most panel gives the probability that a new parent who are in the first wealth quartile (black line), in the first income tertile (left panel), has a child in the lowest income tertile and lowest wealth tertile. The second point gives the probability that the child is in the second wealth quartile but first income tertile. \end{footnotesize}}
		\end{subfigure}
\end{figure}

\end{document}

### 2024-11-06 docs/outside_parameters.tex
First submission

\begin{tabular}{@{}llll@{}}
    \toprule
    Parameter 			& 		 			& Value  	& Source \\	\midrule
    Period Length		& --				& 2 years   & PSID Frequency \\
    Discount Factor$^\dagger$		& $\beta$			& \parbeta		& Literature (see text)  \\
    Risk Aversion		& $\gamma$			& \pargamma		& Literature (see text)  \\
    Expenditure Share Housing & $\xi$		& \parxi        & Literature (see text) \\
    Rental Price$^\dagger$ 		& $q$				& \parq 		& \cite{Davis2008}  \\
    Deprecation$^\dagger$ 		& $\delta$			& \pardelta		& \cite{Harding2007} \\			
    Risk-free Rate$^\dagger$		& $r^f$ 			& \parrf 	    & Literature (see text) \\
    Mortgage premium$^\dagger$	& $r^m$ 			& \parrm 	    & Literature (see text) \\
    PMI premium$^\dagger$         & $r^{pmi}$         & \parrpmi      & \cite{goodman2017sixty} \\
    PMI limit           & $\overline{PMI}$   & \parPMIlim     & \cite{goodman2017sixty} \\ 
    House Price          & $p$		& \parprice      & Data (see text) \\
    Max Loan-to-Value   & LTV 			& \parLTV  & Literature/data (see text) \\
    Selling \& Buying Cost		& ($m_s,m_b$) 			& (\parms,\parmb) &  \cite{Yang2009} \\ 
    Rental Size & $h_r$ & \parhr & Normalization \\
    Deterministic Income & $ l_a$				& Fig. \ref{fig:y_a}		& PSID \\
    Productivity Shocks for Kids & $y,\Pi(y'|y)$ & Fig. \ref{fig:y_ia},\ref{fig:Piv}      & PSID \\
    Initial Distribution & $F(x_{53},v_{53}$)	& Fig. \ref{fig:inidistr}		& PSID \\			
    \bottomrule			
\end{tabular}

### 2024-11-06 tabfig/postest/BW_mostrecent.tex
First submission

\begin{tabular}{l rrr rrr}
\toprule & \multicolumn{3}{c}{White} & \multicolumn{3}{c}{Black} \\ \cmidrule(lr){2-4}\cmidrule(lr){5-7}
\multicolumn{1}{c}{Moment} & \multicolumn{1}{c}{Data} & \multicolumn{1}{c}{Altr.} & \multicolumn{1}{c}{No Altr.} & \multicolumn{1}{c}{Data} & \multicolumn{1}{c}{Altr.} & \multicolumn{1}{c}{No Altr.}\\
\midrule
\;Owner (25-44) & 0.54 & 0.54 & 0.40 & 0.27 & 0.26 & 0.22\\
\;Rent / Income (25-44) & 0.22 & 0.22 & 0.21 & 0.24 & 0.27 & 0.26\\
\;Wealth at Purchase (25-44) & 36.72 & 33.07 & 43.17 & 16.53 & 47.76 & 57.65\\
\;Transfer Rate (55-74) & 0.27 & 0.25 & 0.00 & 0.14 & 0.26 & 0.00\\
\bottomrule
\end{tabular}

### 2024-11-06 src/Julia/export_estimation.jl

@@ -1,0 +1,391 @@
+function varnames()
+    varnames = Dict(:owneryoung =>"Owner (25-44)",
+    :medwealthyoung =>"Median Wealth (25-44)",
+    :medwealthold =>"Median Wealth (55-74)",
+    :medwealthfyoung =>"Median Family Wealth (25-44)",
+    :wealthatpurchase => "Wealth at Purchase (25-44)",
+    :transfrateyoung => "Transfer Rate (55-74)",
+    :h2wyoung =>"House Value / Wealth (25-44)",
+    :r2incyoung =>"Rent / Income (25-44)",
+    :wealthatpurchaseyoung => "Wealth at Purchase (25-44)",
+    :first_ownyoung => "Age First Own (25-44)",
+    :LTVatpurchaseyoung => "LTV at Purchase (25-44)",
+    :tp2wealthpyoung => "Transfers / Parental Wealth (55-74)",
+    :transferbuyersyoung => "Transfers Around Purchase (25-44)",
+    :hand2mouthyoung  => "Hand to Mouth (25-44)",
+    :mortg2incyoung => "Mortgage / Income (25-44)",
+    :transfsizeyoung => "Transfer Size (55-74)",
+    :ownerall => "Owner (25-73)",
+    :ownerold => "Owner (55-73)",
+    :price => "House Price",
+    :α1 => "Supply Elasticity (\$\\alpha_1\$)",
+    :α0 => "Supply El. Intercept",
+    :supply => "Housing Supply",
+    :wealthp_owneryoung => "Own: Parent Wealth",
+    :wealthp_renteryoung => "Rent: Parent Wealth",
+    :wealthpgradyoung => "Parent Wealth Gradient (avg)",
+    :medwealthpgradyoung => "Parent Wealth Gradient (med)",
+    :medwealthp_owneryoung => "Own: Parent Wealth",
+    :medwealthp_renteryoung => "Rent: Parent Wealth",
+    :medconsp2conskyoung => "Consumption Ratio (P/K)",
+    :utilk => "Lifetime Utils Kid",
+    :utilp => "Lifetime Utils Parent",
+    :mortgyoung => "Mortgage (25-44)",
+    :owneryoung_prich => "Owner (25-44), Top 33\\%",
+    :owneryoung_pmiddle => "Owner (25-44), Middle 33\\%",
+    :owneryoung_ppoor => "Owner (25-44), Bottom 33\\%",
+    :transfrateyoung_prich => "Transfers (25-44), Parent Top 50\\%",
[…]

### 2024-11-06 src/Julia/fundamentals.jl

@@ -1,0 +1,123 @@
+function isowner(ih,np::NumPar)
+  @assert ih>0
+  @assert ih<=np.nh
+
+  owner = true
+  if ih <= np.nhr
+    owner = false
+  end
+  return owner
+
+end
+
+function BorrCons(h,own::Bool,price,ia::Int,mp::ModPar)
+  LTV = - mp.LTV[ia]*price*h*own
+
+  BorrCons = LTV
+end
+
+
+function hexp_func(h::Float64,own::Bool,ia::Int64,io::Int64,price::Float64,np::NumPar,mp::ModPar)
+  if own == false
+    hexp = mp.κ*price*h
+  else
+    hexp = price*h
+  end
+
+  if own == true && io <= np.nhr # Buyers must pay extra..
+    hexp = hexp + mp.mb[ia]*price*h
+  elseif own == false && io > np.nhr # Sellers must pay
+    hexp = hexp + mp.ms[ia]*price*np.h_grd[io]
+  end
+
+  return hexp
+end
+
+function BCp_netshock(x::Float64,c::Float64,t::Float64,h::Float64,own::Bool,ia::Int64,io::Int64,price::Float64,np::NumPar,mp::ModPar)
+  @assert ia>np.nyoung
+  hexp = hexp_func(h,own,ia,io,price,np,mp)
+
+  b = x - c - t - hexp + wp_netshock(ia,np)
+  
+  return b
+end
+
+function BCk(x::Float64,c::Float64,t::Float64,h::Float64,own::Bool,ia::Int64,io::Int64,price::Float64,np::NumPar,mp::ModPar,iinc)
+  @assert ia≤np.nyoung
+
+  hexp = hexp_func(h,own,ia,io,price,np,mp)
+  b = x - c - t - hexp + wk(ia,iinc,np)
+    
+  return b
+end
+
+function xk′_scalar(b,h,own::Bool,iak,np::NumPar,mp::ModPar,is,iδ)
+    @assert iak < np.nyoung && iak > 0
+    xkn = b*(1+rr(b,iak,h,own,np.p_grd[is],mp)) + h*own*(1-mp.δ-np.δ_grd[iδ])*np.p_grd[is]
+end
+
[…]

### 2024-11-06 src/Julia/policyplots.jl

@@ -1,0 +1,51 @@
+function plot_policyfuncs(Mb,Mn;store=false)
+    mp = Mb.mp
+    np = Mb.np
+
+    ## Setup for which state to plot
+    bph = 400
+    ibph = argmin(abs.(np.bp_grd.-bph))
+
+    bpl = 0
+    ibpl = argmin(abs.(np.bp_grd.-bpl))
+
+    ihp = 1
+    ivk = 3
+    iok = 1
+    is = 1
+    iak = 1
+    ihk = 2
+    _h2size = Dict(1=>L"h_r",2=>L"h_o")
+
+    plot(title="Child's Bond and Housing Choice $(L"b{\prime}_c,h{\prime}_c")",xlabel="Child Wealth $(L"x_c")")
+
+    iown = findfirst(x->x==true,Mn.g.gk.disc[:,ibpl,ihp,ivk,iok,is,iak,ihk].==1)
+    plot!(np.xk_grd[1:iown],Mn.g.gk.b[1:iown,ibpl,ihp,ivk,iok,is,iak,1] ,label="No Altruism",marker=:circle,color=:black,linewidth=2)
+    plot!(np.xk_grd[iown:end],Mn.g.gk.b[iown:end,ibpl,ihp,ivk,iok,is,iak,2] ,label="",marker=:circle,color=:black,linewidth=2)
+    vline!([np.xk_grd[iown]],line=(:black,:dash),label="",linewidth=2)
+
+    iown = findfirst(x->x==true,Mb.g.gk.disc[:,ibpl,ihp,ivk,iok,is,iak,ihk].==1)
+    plot!(np.xk_grd[1:iown],Mb.g.gk.b[1:iown,ibpl,ihp,ivk,iok,is,iak,1] ,label="Altruism, Non-Rich Parent",marker=:xcross,color=:brown,linewidth=2)
+    plot!(np.xk_grd[iown:end],Mb.g.gk.b[iown:end,ibpl,ihp,ivk,iok,is,iak,2] ,label="",marker=:xcross,color=:brown,linewidth=2)
+    vline!([np.xk_grd[iown]],line=(:brown,:dash),label="",linewidth=2)
+
+    iown = findfirst(x->x==true,Mb.g.gk.disc[:,ibph,ihp,ivk,iok,is,iak,ihk].==1)
[…]

### 2024-11-06 src/Julia/quant_blackwhite.jl

@@ -1,0 +1,104 @@
+
+function blackwhitegap(Mbench::Model)
+    gaps = DataFrame(CSV.File("data/model/calibration/racialincomegaps.csv"))
+    racdiff = ( # element of all racial differenctes
+        bincome = gaps[gaps.race.=="Black",:gap][1],
+        wincome = gaps[gaps.race.=="White",:gap][1],
+        bhsize = 0.75,
+        whsize = 1.0
+    )    
+    display("Racial income gap: White=$(round(racdiff.wincome;digits=3)). Black=$(round(racdiff.bincome;digits=3)). ")
+    
+    ## initialize
+    moms = deepcopy(Mbench.moms)
+    moms[!,:race] = ["bench", "bench"]
+    
+    
+    # Black
+    mp = deepcopy(Mbench.mp)
+    np = deepcopy(Mbench.np)
+    Mblack = Model(mp,np)
+    np.h_grd .*=  racdiff.bhsize
+    np.y_grd .*=  racdiff.bincome
+    display(np.h_grd)
+    display(np.y_grd)
+    solvemodel2nd!(Mblack,np,mp)
+    Mblack.moms[!, :race] .= ["Black w altr"]
+
+    Mblack_n  = altrcontrhousing(mp,np)
+    Mblack_n.moms[!, :race] .= ["Black no altr"]
+
+    # White    
+    mp = deepcopy(Mbench.mp)
+    np = deepcopy(Mbench.np)
+    Mwhite = Model(mp,np)
+    np.h_grd .*= racdiff.whsize
+    np.y_grd .*= racdiff.wincome
+    display(np.h_grd)
+    display(np.y_grd)
+    solvemodel2nd!(Mwhite,np,mp)
+    Mwhite.moms[!, :race] .= ["White w altr"]
+
+    Mwhite_n  = altrcontrhousing(mp,np)
+    Mwhite_n.moms[!, :race] .= ["White no altr"]
+
+    Racemoms = vcat(moms,Mwhite.moms,Mwhite_n.moms,Mblack.moms,Mblack_n.moms)
+    Racemoms = Racemoms[(Racemoms.type .== "Model") .| (Racemoms.race .== "bench") ,:]
+
+    return Racemoms
+
+
+end
+
+function output_BW(BW::DataFrame,targ_moms::Vector{Symbol},nontarg_moms::Vector;store = false)
[…]

### 2024-11-06 src/Julia/quant_endoprice.jl

@@ -1,0 +1,77 @@
+function SolveModelDifferentPrices(Min,pgrd)
+    mp = Min.mp
+    np = Min.np
+    df = DataFrame()
+    for price in pgrd
+        update_prices!(price,np)
+        _M = altrcontrhousing(mp,np)
+        demand = _M.moms[_M.moms.type.=="Model",:].ownerall[1]
+        append!(df, [_M.moms[_M.moms.type .== "Model",:] DataFrame(price=price,demand = demand)])
+        @assert np.p_grd[1]==price # Check that the correct prices are used
+    end
+
+    return df
+end
+
+
+
+function output_endoprice(dfp,Mb,moms,agg_moms,α;store=false)
+    keep = union(agg_moms,moms)
+    Bmoms= Mb.moms[Mb.moms.type.=="Model",:]
+
+    # Add variables to model
+    Bmoms[!,:price] .= Mb.mp.price
+    Bmoms[!,:α1] .= NaN
+    Bmoms[!,:supply] .= Mb.moms[Mb.moms.type.=="Model",:ownerall]
+    df= copy(Bmoms)
+    df = vcat(df,dfp,cols=:union)
+    # df[!,:α1] = [NaN;α]
+
+    df = df[!,keep]
+    df = select(df,keep)
+    @assert issorted(α) "Supply elasticities not sorted. Next block assumes that they are sorted"
+    out = DataFrame(Moment = names(df),
+        Benchmark = Matrix(df[isnan.(df.α1),:])[1,:],
+        Elastic = Matrix(df[df.α1.==α[3],:])[1,:],
+        Middle = Matrix(df[df.α1.==α[2],:])[1,:],
+        Low = Matrix(df[df.α1.==α[1],:])[1,:],
+        )
+    out[!,:Moment] = ["\\;"*varnames()[Symbol(var)] for var in out[!,:Moment]]
+    insert!.(eachcol(out), 1, ["\\textit{Aggregate Moments}"; fill(NaN,length(α)+1)])
+    insert!.(eachcol(out), length(aggmoms)+2, ["\\textit{Targeted Moments}"; fill(NaN,length(α)+1)])
+    display(out)
+    df_latex = latexify(out; env=:table,latex=false,fmt="%1.2f",escape_underscore=true,booktabs = true,convert_unicode=true)
+
[…]

### 2024-11-06 src/Julia/quant_preferadjustment.jl

@@ -1,0 +1,241 @@
+function prefer_adjustment(dfout::DataFrame,Vbk_itp,Vnk_itp,Vbp_itp,Vnp_itp,gnp_t_itp,gnp_b_itp,gnp_disc_itp,np::NumPar)
+    # To find whether kids prefer adjustment costs you must also update the parent's gift giving choices, which is why this is not straightforward.
+    # This also requires some interpolation and so on.
+    @assert maximum(dfout.ihk) == 2
+    preferk = fill(false,size(dfout)[1])
+    preferp = fill(false,size(dfout)[1])
+    wtpk = fill( -1.0,size(dfout)[1])
+    wtpp = fill( -1.0,size(dfout)[1])
+    tpncol = fill(0.0,size(dfout)[1])
+    for ind = 1:size(dfout)[1]
+        df = dfout[ind,:]
+        xk = df.xk
+        xp = df.xp
+        bp = df.bp
+        ivk = df.ivk
+        iok = df.iok
+        iop = df.iop
+        is = df.is
+        ihpn = df.ihp
+        iak = df.iak
+        tp = df.tp
+
+        ## Choices if illiquid
+        # Here the choices in the simulated panel is already optimal 
+        Vbk = Vbk_itp[ihpn,ivk,iok,is,iak](xk + tp,bp)
+        Vbp = Vbp_itp[ivk,iok,iop,is,iak](xk,xp)
+
+        ## Chocies if liquid
+        # Here you have to find the new choices, since the simulated choices are suboptimal with liquid housing
+        probp = [gnp_disc_itp[ihpnn,ivk,iok,iop,is,iak](xk,xp) for ihpnn = 1:np.nh]
+        ihpnn = argmax(probp)
+        tpn = gnp_t_itp[ihpnn,ivk,iok,iop,is,iak](xk,xp)
+        bpn = gnp_b_itp[ihpnn,ivk,iok,iop,is,iak](xk,xp)
+        Vnk = Vnk_itp[ihpnn,ivk,iok,is,iak](xk + tpn,bpn)
+        Vnp = Vnp_itp[ivk,iok,iop,is,iak](xk,xp)
+        tpncol[ind] = tpn
+
+        if Vbk > Vnk
+            preferk[ind] = true
+            wtpk[ind] = (Vbk - Vnk)
+        end
+        if Vbp > Vnp
[…]

### 2024-11-06 src/Julia/simpanel.jl

@@ -1,0 +1,421 @@
+function itp_gpage(np::NumPar,gtp,gcp,gdiscp)
+    itptp = Array{Interpolations.Extrapolation{Float64,2,Interpolations.GriddedInterpolation{Float64,2,Matrix{Float64},Gridded{Linear{Throw{OnGrid}}},Tuple{Array{Float64,1},Array{Float64,1}}},Gridded{Linear{Throw{OnGrid}}},Flat{Nothing}},6}
+    @views gtp_itp = [linear_interpolation((np.xk_grd,np.xp_grd),            gtp[:,:,ivk,ihk,iop,is,iak,ihp],extrapolation_bc=Flat()) for ivk=1:np.nv,ihk=1:np.no,iop=1:np.no,is=1:np.ns,iak=1:np.nyoung,ihp=1:np.nh]
+    @views gbp_itp  = [linear_interpolation((np.xk_grd,np.xp_grd),           gcp[:,:,ivk,ihk,iop,is,iak,ihp],extrapolation_bc=Flat()) for ivk=1:np.nv,ihk=1:np.no,iop=1:np.no,is=1:np.ns,iak=1:np.nyoung,ihp=1:np.nh]
+    @views gdiscp_itp = [linear_interpolation((np.xk_grd,np.xp_grd),      gdiscp[:,:,ivk,ihk,iop,is,iak,ihp],extrapolation_bc=Flat()) for ivk=1:np.nv,ihk=1:np.no,iop=1:np.no,is=1:np.ns,iak=1:np.nyoung,ihp=1:np.nh]
+  
+    return gtp_itp,gbp_itp,gdiscp_itp
+end
+  
+  
+function itp_gkage(np::NumPar,gck,gdisck)
+    t1 = Array{Array{Interpolations.Extrapolation{Float64,2,Interpolations.GriddedInterpolation{Float64,2,Matrix{Float64},Gridded{Linear{Throw{OnGrid}}},Tuple{Array{Float64,1},Array{Float64,1}}},Gridded{Linear{Throw{OnGrid}}},Flat{Nothing}},5},1}
+    t2 = Array{Array{Interpolations.Extrapolation{Float64,2,Interpolations.GriddedInterpolation{Float64,2,Matrix{Int64},Gridded{Linear{Throw{OnGrid}}},Tuple{Array{Float64,1},Array{Float64,1}}},Gridded{Linear{Throw{OnGrid}}},Flat{Nothing}},5},1}
[…]

### 2024-11-06 docs/outside_parameters.tex
Added docs

@@
     Deprecation$^\dagger$ 		& $\delta$			& \pardelta		& \cite{Harding2007} \\			
-    Risk-free Rate$^\dagger$		& $r^f$ 			& \parrf 	    & Literature (see text) \\
+    Risk-free Rate$^\dagger$		& $r^f$ 			& \parrf 	    & Literautre (see text) \\
     Mortgage premium$^\dagger$	& $r^m$ 			& \parrm 	    & Literature (see text) \\

### 2024-11-06 docs/paper.tex
Added docs

--- 2024-11-06 docs/Homeownership_BankOfMomAndDad.tex
@@
 E-mail: eirik.e.brandsaas@frb.gov.
-I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts, Corina Boar, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobson, William Gamber, Claudia Sahm, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, Boston Fed, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, YES 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop.
+I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts,  Corina Boar, Tim Kehoe, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobsen, William Gamber, Claudia Sahm, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, Young Economist Symposium 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop.
 \textit{Disclaimer: }The views expressed in this paper are solely the responsibility of the author and should not be interpreted as reflecting the views of the Board of Governors of the Federal Reserve System or of anyone else associated with the Federal Reserve System.}}
@@
 Housing is the largest asset in U.S.
-household portfolios, and first-time homebuyers increasingly rely on parental transfers.
+household portfolios, and a first-time homebuyers increasingly rely on parental transfers.
 This paper quantifies the contribution of parental transfers to the homeownership rate of young households.
@@
 I find that parental transfers account for 14 percentage points (29\%) of young households' homeownership.
-Transfers from wealthy parents not only help households overcome borrowing constraints, but also help sustain homeownership, mitigating the drawbacks of illiquidity.
-Surprisingly, policies lowering entry barriers to homeownership generally increase the reliance on parental wealth, whereas increased liquidity reduces it.
-Finally, I show that children of wealthy parents strategically use the illiquidity of housing as a commitment device to encourage transfers, resulting in a preference for illiquidity.
+Transfers from wealthy parents not only help households overcome borrowing constraints but also help them maintain homeownership, reducing the downsides of illiquidity.
+Policies lowering entry barriers to homeownership generally increase the reliance on parental wealth, while increasing liquidity lowers the role of parental wealth.
+Finally, I study how illiquidity affects the commitment problem  between parents and children.
+Wealthy parents' children increase investment in illiquid housing to extract larger transfers and exhibit a preference for illiquidity.
 \end{spacing}}
-\vspace{0.5cm}
-\noindent \textbf{JEL:} D14, D15, E21, G51, R21
-
-\noindent \textbf{Keywords:} Homeownership, Parental transfers, Altruism, Life-cycle models
-
 \end{abstract}
@@
 Housing is the largest asset in U.S.
-households portfolios.
-While homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}, households face different barriers to ownership.
-Between 2009 and 2016, around 30\% of American first-time homebuyers received parental assistance for their down payment, with such assistance averaging \$48,000 per receiving household.\footnote{All data from the Survey of Household Economic Decisionmaking and the Panel Study of Income Dynamics (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
-
-This paper quantifies the importance of parental transfers for young adults' homeownership and investigates how policies lowering homeownership barriers interact with parental transfers.
-Previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}.
-Additionally, parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}.
-However, while these studies show that transfer receipt increases the probability of renters becoming homeowners, they do not address the overall importance of parental transfers.
-For example, households may anticipate future transfers and subsequently save less.
-In this case, even if transfers increase the likelihood of homeownership, all else being equal, they may not affect aggregate homeownership.
-Furthermore, this is the first paper quantifying the importance of parental transfers as a determinant of homeownership; the existing literature has primarily focused on family formation and credit constraints \citep[see e.g.,][]{Chang2024,Paz-Pardo2019,Mabille2020}.
-
-I first present empirical evidence indicating that parental wealth helps not only in the transition to ownership, but also in maintaining ownership.
-I show that, controlling for a wide set of household and parental characteristics, households with wealthier parents are less likely to fall behind on mortgage payments after becoming homeowners, even though they buy more expensive homes.
-Finally, using a variation of the event study in \cite{Chetty2007}, I show that while most households downsize their housing during unemployment spells, those with wealthy parents retain their current homes.
-
-To quantify the impact of parental transfers on homeownership, I then combine two distinct models: housing models, which ignore parental transfers, and models of parental transfers, which ignore housing.
+households' portfolios and homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}.
+Yet access to homeownership appears to be increasingly reliant on parental transfers.
+Around 30\% of American first-time homebuyers received downpayment assistance from their parents between 2009 and 2016.
+Parental downpayment assistance averaged \$48,000 among receiving households.\footnote{Data from the Survey of Household Economic Decisionmaking and the Transfer Supplement of the 2013  Panel Study of Income Dynamics (see Section \ref{sec:data}).}  It is important to understand how transfers shape households' ability to access and maintain ownership, given  housing’s importance to household portfolios, economic outcomes \citep{sodini2023identifying}, and business cycles \citep{Kaplan2020,favilukis2017macroeconomic}.
+
+The main objective of this paper is to quantify the importance of parental transfers for young adults' homeownership.
+The primary mechanism by which transfers increase homeownership, and the focus of most empirical studies, is through relaxing borrowing constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}.
+A secondary mechanism is that parental transfers help households maintain homeownership, a mechanism that has received less attention  \citep[see e.g.,][]{bond2021role}.
+However, while these studies show that receiving transfers causally increases homeownership, they do not address the overall importance of parental transfers.
+For example, households might anticipate future transfers and save less.
+In this case, even if transfers increase the likelihood of homeownership all else equal, they may not affect aggregate homeownership.
+The secondary objective is to study how policies lowering homeownership barriers interact with parental transfers.
+
+I first present empirical results suggesting that parental wealth not only helps in the transition to ownership but also in maintaining ownership.
+I show that, controlling for a wide set of household and parental characteristics, households with wealthier parents not only purchase pricier homes but are also less likely to fall behind on mortgage payments after becoming homeowners.
+Finally, in a variation of the event study in \cite{Chetty2007}, I show that while most households downsize their housing during unemployment spells, those with wealthy parents retain their current homes.
+
+To answer the research questions, I combine two distinct models: housing models, which ignore parental transfers, and models of parental transfers, which ignore housing.
 I build a life-cycle overlapping generations model with altruistic parents and a rent-or-own decision.
-Parents and children interact without commitment through both inter-vivos transfers and end-of-life bequests.
-In models of altruism without housing (e.g., \citealp{Altonji1997a,Barczyk2014}), the transfer motive is to increase the child's consumption when they are borrowing constrained, thus having high marginal utility of wealth.
+Parents and children  interact without commitment through both inter-vivos transfers and bequests.
+In models of altruism without housing (e.g., \citealp{Altonji1997a,Barczyk2014}), the transfer motive is to increase the child's consumption when they are borrowing constrained and thus have high marginal utility of wealth.
 The inclusion of a frictional housing market generates two new transfer motives.
-First, transfers can directly alleviate credit constraints, enabling wealthier parents to help their children meet down-payment requirements.
+First, transfers directly alleviate credit constraints, and so wealthy parents may transfer to allow their children to meet the downpayment requirement.
 Second, future transfers provide partial insurance against future income shocks, reducing risks associated with the large illiquid investment of homeownership.
 Moreover, the illiquidity of housing serves as a commitment device enabling children to secure future transfers.
-As in all transfer models, the expectation of future transfers reduces children's saving motives, potentially lowering the homeownership rate.
-
-Allowing for parental transfers in all periods highlights that transfers are critical not only for entering homeownership---the main focus of empirical literature---but also for maintaining homeownership.
-Specifically, the prospect of future transfers mitigates the downsides of homeownership for liquidity-constrained households.
-Ameliorating illiquidity turns out to be as important as relaxing borrowing constraints.
-Crucially, the model rationalizes the previously discussed empirical pattern that households with wealthy parents do not downsize during income losses, which would not be possible without allowing transfers in all periods.
-
-I estimate the model by matching data on homeownership, wealth, and transfers from the Panel Study of Income Dynamics (PSID).
+As in all transfer models, the expectation of future transfers reduces children’s savings motives, potentially lowering the homeownership rate.
+
+Beyond computational challenges, what insights do we gain by allowing parental transfers in all periods?
+After all, previous studies have used alternative assumptions, such as a one-time transfer after which the child is independent at age 24 \citep[e.g.,][]{Lee2019} or warm-glow bequests that are distributed to younger households \citep[e.g.,][]{aliprantis2022dynamics}.
+The key insight of my modeling approach is that transfers are critical not only for entering homeownership---the main focus of empirical literature---but also for maintaining homeownership.
+Specifically, the possibility of future transfers mitigates the downsides of homeownership for liquidity-constrained households.
+This turns out to be as important as relaxing borrowing constraints.
+Moreover, the model replicates the empirical pattern that households with wealthy parents do not downsize during income losses, which would not be possible without allowing transfers in all periods.
+
+I estimate the model by matching data on homeownership, wealth, and transfers from the Panel Survey of Income Dynamics (PSID).
 To quantify the importance of  transfers, I find the counterfactual homeownership rate without parental transfers.
 Without parental transfers, the model simplifies to a standard life-cycle model with housing.
-I find that parental transfers account for 14 percentage points (29\%) of the homeownership rate among young households.
-
-I then use the model to evaluate policies aiming to increase homeownership.
-I find that such policies disproportionately benefit households with wealthy parents.
-Relaxing mortgage constraints---for example, by lowering minimum down payments---strengthens the link between parental wealth and children's housing outcomes, since more households with wealthy parents are on the margin to become homeowners.
-Nonmortgage policies that lower barriers to entry, such as reducing prices or purchase costs, increase homeownership for all households, though disproportionately among those with wealthier parents.
-However, increasing liquidity by reducing sales costs weakens the link between parental wealth and children's housing outcomes, as the partial insurance provided by parental wealth becomes less important.
-
-I then study how illiquidity affects the strategic behavior arising in the absence of commitment.
-I find that 27\% of young households prefer illiquid housing---due to sales costs---to liquid housing.
-These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth.'' Lower liquid wealth implies higher marginal propensities to consume (MPC) and, hence, more parental transfers.
-As an illustrative example, consider a child who buys a house and brings no liquid wealth to the next period.
-If the parent does not transfer in the next period, the child will either face very low consumption (with a correspondingly high marginal value of wealth) or be forced to liquidate the house and incur sales costs.
-The parent dislikes both outcomes and may find it optimal to transfer enough to keep the child in the home.
+I find that parental transfers account for 14 percentage points (29\%) of the young homeownership rate.
+In the benchmark model, the supply of owner-occupied housing is perfectly elastic; however, my results are robust to alternative supply elasticities.
+
+Next, I evaluate policies to increase homeownership and their effect on the link between parental wealth and children's housing outcomes.
+The results show that many policies mainly benefit households with wealthy parents.
+Relaxing mortgage constraints, for example by lowering minimum downpayments, strengthens the link between parental wealth and children's housing outcomes since households with wealthier parents find homeownership more attractive, all else equal.
+Other policies that lower barriers to entry, such as reducing prices or purchase costs, increase homeownership for all households, though mainly among those with wealthier parents.
+However, increasing liquidity by reducing sales costs weakens the link between parental wealth and children’s housing outcomes, as the partial insurance provided by parental wealth becomes less important.
+
+Next, I study how housing illiquidity affects the strategic behavior arising in the absence of commitment.
+I find that 27\% of young households prefer to face sales costs on housing.
+These households, who have wealthy parents, use sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth''.
+Lower liquid wealth implies higher marginal propensities to consume (MPC), and hence more parental transfers.
+As an illustrative example, imagine that a child buys a house and brings no liquid wealth to the next period.
+If the parent transfer in the next period, the child will either face very low consumption (with a correspondingly high marginal value of wealth) or be forced to liquidate the house and incur sales costs.
+The parent dislikes both outcomes, and may find it optimal to transfer enough to keep the child in the home.
 This driver of preference for illiquidity differs from those studied in the literature, such as temptation or time inconsistency \citep{attanasio2024temptation,laibson1997golden}.
-Furthermore, parental wealth generating preference for illiquidity provides a novel explanation for the existence of high-MPC households \cite[see e.g.,][]{kaplan2022marginal}.
-That children with wealthy parents find illiquidity less problematic---or even prefer it---is consistent with the empirical results in \cite{choukhmane2023benefits}.
-They find that households with wealthy parents are more likely to invest in illiquid retirement accounts and less likely to make costly early withdrawals, even after controlling for a wide set of characteristics.
-
-Finally, I quantify the contribution of parental transfers to the Black-White homeownership gap of 50\% among young adults.
+Furthermore, this preference for illiquidity provides a novel explanation for the existence of high-MPC households \cite[see e.g.,][]{kaplan2022marginal}.
+
+Finally, I quantify the contribution of parental transfers to the Black-White homeownership gap: in the U.S., young White households are twice as likely to be homeowners as Black households.
 I adapt the methodology of \cite{Ashman2020} and \cite{aliprantis2022dynamics}, who study the racial wealth gap, and recalibrate the income process to capture racial differences.
 I find that parental transfers account for 7 percentage points (14\%) of the Black-White homeownership gap.
-For context, the Black-White homeownership gap has remained roughly constant since the passing of the Fair Housing Act in 1968.
 
@@
 
-My paper first contributes to the literature studying the determinants of homeownership over the life cycle.
+By introducing parental transfers to models of housing, my paper contributes to the literature studying the determinants of homeownership over the life cycle.
 Recent papers focus on marriage and family formation \citep{Fisher2011,Chang2024,Khorunzhina2019}, housing demand in old age \citep{McGee2019,Barczyk2020a}, and changing borrowing constraints \citep{Paz-Pardo2019,Mabille2020}.
 These studies highlight the importance of credit constraints and minimum down payments in constraining demand for owner-occupied housing.
-By considering parent-child interactions, I show how parental wealth affects the relative importance of constraints.
+However, by considering parent-child interactions, I show a dichotomy based on parent wealth in the importance of constraints.
 For households with poorer parents, the sales and purchase costs decrease ownership by making it riskier to own.
-This risk is smaller for households with richer parents, making mortgage credit constraints relatively more important.
-My results highlight that policies increasing ownership through subsidies to first-time buyers---a common approach in many cities---will lead to increased housing inequality.
+This risk is less important for households with richer parents.
+making mortgage credit constraints more important.
+My results highlight  that policies increasing ownership through subsidies to first-time buyers will lead to increased housing inequality.
 My model-based results also complement some of the empirical research on the effect of parental resources and housing outcomes \citep[see e.g.,][ for recent work]{wold2024housing,daysal2023intergenerational,benetton2022dynastic} by quantifying the contribution of transfers to homeownership.
 
-My paper also contributes to the literature on altruistic households interacting without commitment by studying how illiquid housing affects the commitment problem.
-In these models, children have an incentive to undersave to increase parental transfers (e.g., \citealp{Altonji1997a,Boar2018,Barczyk2014,Chu2020}).
-Illiquid housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}, which children of wealthy parents strategically exploit to encourage future parental transfers.
-The paper most closely related to mine is \cite{Barczyk2020a}, who study how the homeownership of retired parents affects the economic behavior of the elderly and their adult children.
-In their model, the parents' home serves as a commitment device, encouraging wealth bequests from the parent and informal care from the adult offspring.
-My contribution is to study the complementary problem: how parental transfers affect the homeownership decisions of young adults.
-Finally, while \cite{kaplan2012moving} studies cohabitation with parents among individuals under the age of 25, my focus is on homeownership decisions for households from age 25 onwards.
-
-Finally, the paper contributes to the literature on parental transfers and life-cycle outcomes by focusing on how parental transfers to young adult households influence their wealth accumulation through homeownership.
+I contribute to the literature on altruistic households who interact without commitment by studying how illiquid housing affects the commitment problem.
+In these models children undersave to increase parental transfers (e.g., \citealp{Altonji1997a,Boar2018,Barczyk2014,Chu2020}).
+However, adjustment costs on housing impose future expenditure commitments \citep{Chetty2007,Shore2010}.
+The illiquidity of housing allows households to commit to having low liquid wealth, which increases future transfers.
+The paper most closely related to mine is \cite{Barczyk2020a}, which studies the mirror problem: how the homeownership of retired parents affects the economic behavior of the elderly and their adult children.
+In their model, the parent’s home serves as a commitment device, encouraging wealth bequests from the parent and informal care from the adult offspring.
+
+Finally, I contribute to the literature on parental transfers and life-cycle outcomes by focusing on how parental transfers to young adult households influence their wealth accumulation through homeownership.
 My results show that households with wealthier parents are willing to buy sooner, take on higher leverage, and hold less liquid precautionary savings.
-The previous literature has generally focused on parental investment in children's human capital (e.g., \citealp{Lee2019,Daruich2018,gilraine2023public}), transfers from adult children to retired parents (e.g., \citealp{Mommaerts2016, Barczyk2018, Barczyk2020a}), or smoothing income shocks (e.g., \cite{Boar2021ReStud,fagereng2023insuring,andersen2020bailing}) instead of the effect of parental wealth on household portfolio choices.
+The previous literature has generally focused on parental investment in children's human capital (e.g., \citealp{Lee2019,Daruich2018,gilraine2023public}) or transfers from adult children to retired parents (e.g., \citealp{Mommaerts2016, Barczyk2018, Barczyk2020a}), instead of the effect of parental wealth on household portfolio choices.
 
 The paper proceeds as follows.
-In Section \ref{sec:data}, I describe the data sources and summary statistics, and I document that parental wealth is associated with better housing outcomes.
-Section \ref{sec:model} describes the quantitative model, and Section \ref{sec:estimation} discusses the structural estimation.
+In Section \ref{sec:data}, I describe the data sources, summary statistics, and document that parental wealth is associated with better housing outcomes.
+Section \ref{sec:model} describes the quantitative model and Section \ref{sec:estimation} discusses the structural estimation.
 Section \ref{sec:quant} performs the main quantitative exercise and robustness tests.
@@
 
-\textit{SHED:} I use SHED, an annual cross-sectional survey conducted by the Federal Reserve, to observe the share of households who funded the down payment with a loan or gift from family or friends, by year of purchase.
+\textit{SHED:} I use SHED, an annual cross-sectional survey conducted by the Federal Reserve,  to observe the share of households who funded the downpayment with a loan or gift from family or friends, by year of purchase.
 The results are plotted in the left panel of Figure \ref{fig:motivation}.
 The main observation is the large increase in the role of inter-vivos transfers for homeowners since 2001.
-From 2001 to 2007, only 10\% to 18\% of first-time buyers received transfers, while 20\% to 40\% received transfers after 2009.\footnote{The question was only included in the 2015 and 2016 waves.
-In the 2015 wave, first-time ownership and down-payment source is only reported for those who bought in 2001 to 2015, while in the 2016 wave it is only reported for 2015 to 2016.
+In 2001-2007, only 10-18\% of first-time buyers received transfers, while 20-40\% received transfers after 2009.\footnote{The question was only included in the 2015 and 2016 waves.
+In the 2015 wave first-time ownership and downpayment source is only reported for those who bought in 2001-2015, while in the 2016 wave it is only reported for 2015-2016.
 }
 
-\textit{AHS:} I use the AHS, which surveys occupants of owner-occupied units, to obtain a time series on whether the down payment was mostly funded by gifts or inheritances (right panel of Figure \ref{fig:motivation}).
+\textit{AHS:} I use the AHS, which surveys occupants of owner-occupied units, to obtain a time series on whether the downpayment was mostly funded by gifts or inheritances (right panel of  Figure \ref{fig:motivation}).
 We observe a flat pattern in the 1990s, followed by a sharp decline from 2005 to 2009, coinciding with the prevalence of mortgages featuring very low down payments (e.g., \cite{Corbae2015}), and a stark increase from 2013 to today.
@@
 \subsection{Panel Study of Income Dynamics}\label{sec:PSID}
-My main data source is the PSID, which follows a nationally representative sample of U.S.
-households and their descendants over time since 1968.
-The PSID is the only publicly available U.S.
-dataset that satisfies this paper's three requirements.
+My main data source is the Panel Study of Income Dynamics (PSID), which follows a nationally representative sample of US households and their descendants over time since 1968.
+The PSID is the only publicly available US dataset that satisfies this paper's three requirements.
 First, it has detailed wealth, income, and housing data for both parents and adult children.
@@
 I use data from 1999 to 2017.
-In 1999, the PSID started to collect detailed wealth data every other year.
+In 1999 the PSID started to collect detailed wealth data every other year.
 In most waves of the PSID, there is limited transfer data, and the main question is whether households received gifts or bequests over \$5,000  in the last two years.
@@
 I classify top-coded values as missing observations.
-All monetary variables are expressed in constant 2016 U.S.
-dollars (in thousands).
+All monetary variables are expressed in constant 2016 US dollars (in thousands).
 Table \ref{tab:vardef} in the Appendix provides variable code of the main variables.
 
-\textit{Sample Selection:} Throughout this paper, the sample includes all households aged 25 to 84 in the PSID, excluding the Survey of Economic Opportunity and the Latino subsamples, to obtain a representative sample.
+\textit{Sample Selection:} Throughout this paper, the sample includes all households aged 25-84 in the PSID, excluding the Survey of Economic Opportunity and the Latino subsamples to obtain a representative sample.
 All summary statistics are calculated using the provided family weights.
@@
 \subsubsection{Descriptive Statistics: Who Receives Transfers?}
-I now discuss descriptive statistics from the PSID sample in 2013 for households aged 25 to 44 with an observed parent household.
-Table \ref{tab:descrstats} contains the means of variables, by age, wealth, marriage, homeownership, and transfer receipt.
-
-There are several main takeaways from the subgroup analysis.
-In 2013, 22\% of young households received a transfer, and transfers averaged \$3,950.
-Receivers have significantly richer parents, have similar wealth and income as nonreceivers, and are less likely to own.
+I now discuss descriptive statistics from the PSID sample in 2013 for households aged 25-44 with an observed parent household.
+Table \ref{tab:descrstats} contains the means of variables, by age, wealth, marriage, homeownership, and and transfer receipt.
+
+The main takeaways from the subgroup analysis are: in 2013, 22\% of young households received a transfer, and transfers averaged \$3,950.
+Receivers have significantly richer parents, have similar wealth and income as non-receivers, and are less likely to own.
 Receivers are more likely to transition from renting to owning, especially in the age groups where households are most likely to buy.
@@
 The mean transfer is relatively high at \$3,960, and  22\% of households received transfers in the last year.
-The largest difference is parental wealth: transferring parents are 2.5 times as wealthy as nontransferring parents (\$1,029,000 vs.
-\$410,000).
-Perhaps surprisingly, receivers are slightly richer than nonreceivers (\$108,000 vs.
-\$78,500), are more likely to be college-educated and white, and are one year younger.
+The largest difference is parental wealth: transferring parents are 2.5 times as wealthy as those who don't (\$1,029,000 vs \$410,000).
+Perhaps surprisingly, receivers are slightly richer (\$108,000 vs.
+\$78,500), more likely to be college-educated, white, and one year younger.
 Receivers are slightly less likely to be homeowners (39\% vs.
-43\%), reflecting the age difference and the tendency for college attendance to delay homeownership.
+43\%), likely reflecting the age difference and the tendendency for college attendance to delay homeownership.
 
@@
 By breaking down the sample by marital status, we see that married or cohabiting households have more wealth, higher income, are slightly older, and have three times the ownership rate of singles.
-Married households receive more frequent and slightly larger transfers.
+Married households are receive more frequent and slightly larger transfers.
 Next, I break the sample down by ownership.
-Homeowners have 10 times the wealth of renters and also significantly wealthier parents.
+Homeowners have ten times the wealth of renters and also significantly wealthier parents.
 Renters and owners receive transfers at about the same rate and size.
-Receivers are more likely to switch from renting to owning: 21\% of receiving owners rented two years ago, compared to 14\% of nonreceiving owners.
+Receivers are more likely to switch from renting to owning: 21\% of receiving owners rented two years ago, compared to 14\% of non-receiving owners.
 
@@
 Wealthier households are less likely to receive transfers.
-Notably, among 29-to-32-year-olds, homeownership is more common among transfer recipients (40\%) than nontransfer recipients (32\%).
-Furthermore, receivers are not only more likely to own, but also to be recent homeowners: 21\% of receiving owners are new homeowners versus 13\% of nonreceiving owners.
+Notably, among 29-32-year-olds, homeownership is more common among transfer recipients (40\%) than non-recipients (32\%).
+Furthermore, receivers are not only more likely to own but also to be recent homeowners: 21\% of receiving owners are new homeowners, versus 13\% of non-receiving owners.
 
@@
 
-I now utilize microdata from the PSID to examine the correlation between parental financial resources and housing outcomes.
-
-I use the sum of parental net worth and income (denoted as wealth) as a statistic to measure parents' financial resources in all specifications for two reasons.
+I now utilize microdata from the Panel Study of Income Dynamics (PSID) to examine the correlation between parental financial resources and housing outcomes.
+
+I use the sum of parental net worth and income (denoted as wealth) as a statistic to measure parents' financial resources in all specifications, for two reasons.
 First, this is consistent with models of altruism, as the one I use later.
 Second, since parents' wealth and income are highly correlated and I have few observations---under 1,000 for most specifications---the standard errors are large when both are included.
-The sample is limited to households aged 25 to 44.
-
-\subsubsection{Households with Wealthier Parents Buy More Expensive Houses}
+The sample is limited to households aged 25-44.
+
+\subsubsection{Households With Wealthier Parents Buy More Expensive Houses}
 I first show that households with wealthier parents buy more expensive homes, controlling for household characteristics.
@@
 \begin{equation*}
-\ln{HVal_{i}} = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
+\ln{HVal_{i}} = \beta_1 \ln(Wealth)_{p(i),t-2}  + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i},
 \end{equation*}
-Parents' wealth, as well as household income and net worth, are log-transformed.
+Parents' wealth, as well as household income and net-worth, are log-transformed.
 Households are denoted by $i$ and their parents by $p(i)$.
@@
 Column 1 of Table \ref{tab:hypo} reports the results.
-Households with wealthier parents buy larger houses: A 1\% increase in parental wealth is associated with a 0.049\% increase in the purchase value of the child's house, and the effect is significant at the 5\% level.
-The effect of parental wealth is about the same as the child's own net worth (0.065\%).
+Households with wealthier parents buy larger houses: a 1\% increase in parental wealth is associated with a 0.049\% increase in the purchase value of the child's house, and the effect is significant at the 5\% level.
+The effect of parental wealth is about the same the as child's own net worth (0.065\%).
 
@@
 
-\subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
+\subsubsection{Households With Wealthier Parents Are Less Likely to be Behind on Mortgages} 
 To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales.
 I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses.
-The PSID has collected this data since 2009, and 3.9\% of owners aged 25 to 44 are behind, on average, while 9.2\% have been behind at least once.
+The PSID has collected this data since 2009, and 3.9\% of owners aged 25-44 are behind on average while 9.2\% have been behind at least once.
 
@@
 \begin{equation*}
-Behind_i = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
+Behind_i = \beta_1 \ln(Wealth)_{p(i),t-2}  + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i},
 \end{equation*}
 In Column 2 of Table \ref{tab:hypo}, the outcome variable is whether households are behind on the mortgage payments the first time they are observed as owners.
-Parental wealth in the period before purchase decreases the probability that a household will ever be behind: A 1\% increase in parental wealth decreases the probability of being behind by 0.016 percentage points and is significant at the 1\% level.
+Parental wealth in the period before purchase decreases the probability that a household will ever be behind: a 1\% increase in parent wealth decreases the probability of being behind by 0.016 percentage points, and is significant at the 1\% level.
 Parental wealth has a larger effect than the child's net worth.
@@
 
-I also report the results from two specifications where I follow households post-purchase, using random effects and household fixed effects regressions (Columns 4 and 5).
+I also report the results from two specifications where I follow households after post-purchase, using random effects and household fixed effects regressions (Columns 4 and 5).
 Once we follow households over time, the effect of parental wealth decreases.
 With household fixed effects, the effect is no longer significant.
-However, the fixed effects regressions must be interpreted with caution, since there is little within-parent variation in wealth over time.
+However, the fixed effects regressions must be interpreted with caution since there is little within-parent variation in wealth over time.
 Indeed, the random and fixed effects coefficients are almost the same (-0.008 vs -0.006).
@@
 
-\subsubsection{Households with Wealthier Parents Are Less Likely to Downsize during Unemployment Spells}\label{sec:eventstudy}
-Distressed and forced home sales are costly and frequently driven by unemployment \citep{kermani2021racial,hsu2018unemployment}.
-If parental wealth provides insurance, households with wealthier parents should be less likely to downsize after income losses.
+\subsubsection{Households With Wealthier Parents Are Less Likely To Downsize During Unemployment Spells}\label{sec:eventstudy}
+Distressed and forced home sales are costly, and frequently driven by unemployment \citep{kermani2021racial,hsu2018unemployment}.
+If parent wealth provides insurance, households with wealthier parents should be less likely to downsize after income losses.
 I perform a simple event study on the effect of unemployment on housing consumption to test this hypothesis.
@@
 The sample is limited to household heads who are unemployed only once between ages 25 and 45.
-I divide the sample by whether a household's parents were in the top parental wealth quartile (of parents with children aged 25 to 44) at the time of unemployment and compare the two groups' housing consumption growth rates at unemployment.
+I divide the sample by whether a household's parents were in the top parental wealth quartile (of parents with children aged 25-44) at the time of unemployment, and compare the two groups' housing consumption growth rates at unemployment.
 The results are displayed in Figure \ref{fig:housinggrowthrates}.
-Among households with nonwealthy parents, housing consumption decreases by about 5\% during unemployment (significant at the 95\% level).
+Among households with non-wealthy parents housing consumption decreases by about 5\% during unemployment (significant at the 95\% level).
 In contrast, among households with wealthy parents, housing consumption even grows slightly at unemployment.
@@
 Moreover, households with wealthier parents buy more expensive houses, are less likely to be behind on their mortgages, and are less likely to downsize during unemployment.
-These results suggest that homeowners with wealthy parents find illiquidity of homeownership less problematic.
 
@@
 However, individuals have to file if they give transfers above \$18,000 (in 2024) to any one individual within a calendar year.
-Thus, the gift tax is irrelevant for the vast majority of households, as it only applies if lifetime transfers---including bequests---exceed \$13.61 million.
+Thus, the gift tax is irrelevant for the vast majority of households, as it applies if life-time  transfers---including bequests---exceed \$13.61 million.
 
@@
 %% There is a missing paragraph linking the previous results this this part!
-This section describes my life-cycle model of housing choices with overlapping generations, idiosyncratic earnings risk, and altruistic inter-vivos transfers from parents to their offspring (``children'').
-To answer how important parental transfers are for housing outcomes, I combine two models: altruism without commitment and illiquid housing.
+This section describes my life-cycle model of housing choices with overlapping generations, idiosyncratic earnings risk, and altruistic inter-vivos transfers from parents to their offspring (``kids'').
+To answer how important parental transfers are for housing outcomes, I combine two models, altruism without commitment and illiquid housing.
 
 Altruism without commitment is a standard model of intergenerational family interactions (e.g., \cite{Altonji1997a,Barczyk2018}).
-Altruistic parents, deriving utility from their child's utility, influence their child's consumption and housing choices through nonnegative transfers.
-The lack of commitment aligns with empirical transfer patterns and generates strategic behavior, as both children and parents internalize the effect of their choices on each other's future choices.
-The central theoretical prediction is that wealthy parents transfer to children with high marginal utility of wealth, typically poor or borrowing-constrained households \citep{Chu2020,Barczyk2020}.
+Altruistic parents, deriving utility from their kid's utility, influence their kid's consumption and housing choices through non-negative transfers.
+The lack of commitment aligns with empirical transfer patterns and generates strategic behavior as both kids and parents internalize the effect of their choices on each other's future choices.
+The central theoretical prediction is that wealthy parents transfer to kids with high marginal utility of wealth, typically poor or borrowing-constrained households \citep{Chu2020,Barczyk2020}.
 
 The second model component is illiquid homeownership.
-Without illiquidity, transfers mainly relax mortgage constraints, as has been the focus in empirical literature (e.g., \cite{Blickle2019,Engelhardt1998,Guiso2002,Lee2018}).
+Without illiquidity, transfers mainly relax mortgage constraints, as has been the focus in empirical literature (see e.g., \cite{Blickle2019,Engelhardt1998,Guiso2002,Lee2018}).
 With illiquidity, we get a two-asset model where portfolio composition matters for transfers.
-Specifically, homeowning children with low liquid wealth (``house rich but cash poor'') have high marginal utility of wealth, strengthening parental transfer motives and, thus, the child's homeownership motive.
+Specifically, homeowning kids with low liquid wealth (`house-rich but cash-poor') have high marginal utility of wealth, strengthening parental transfer motives and thus also the kid's homeownership motive.
 Finally, while illiquid housing makes selling costly, wealthy parents can offer partial insurance, thus reducing the risk of costly liquidation or downsizing.
 
-These two housing frictions---borrowing constraints and illiquidity---are not only theoretically important, but also empirically relevant.
-In the SHED, 57\% of renters could not afford a down payment or did not qualify for a mortgage, while 26\% said that renting was more convenient and 23\% cited planning to move as reasons for renting.
-
-\subsection{Demographics, Preferences, and Technologies}
+These two housing frictions---borrowing constraints and illiquidity---are not only theoretically important but also empirically relevant.
+In the SHED, 57\% of renters could not afford a down-payment or did not qualify for a mortgage, while 26\% said that renting was more convenient and 23\% cited planning to move as reasons for renting.
+
+\subsection{Demographics, Preferences and Technologies}
 Time is discrete and finite.
@@
 The life cycle is illustrated in \ref{fig:overview}.
-A family consists of one adult child $c$ household (age 25 to 53) and a parent $p$ household (age 55 to 83) with an age gap of 30.
-Thus, children have new children at age 30, but new children are economically inactive until age 25.
-Each parent-child pair overlaps for 15 periods (30 years), and only two households are economically active in any dynasty at a time.
-Three events happen simultaneously when a child household becomes 55.
+A family consists of one adult kid $k$ household (age 25-53) and a parent $p$ household (age 55-83) with an age gap of 30.
+Thus, kids have new kids at age 30, but new kids are economically inactive until age 25.
+Each parent-kid pair overlaps for 15 periods (30 years) and only two households are economically active in any dynasty at a time.
+Three events happen simultaneously when a kid household becomes 55.
 First, the household's parent dies (at age 85).
-Second, the child transitions to become a parent household.
-Third, the child of the child becomes economically active as a child household.
+Second, the kid transitions to become a parent household.
+Third, the child of the kid becomes economically active as kid household.
 
@@
 \textit{Preferences and Altruism:} Parents and children have time-separable expected utility with a discount factor of $\beta$.
-Households maximize expected utility, and the per-period utility function for children is
+Households maximize expected utility, and the per-period utility function for kids is
 \begin{equation}
-U_c(c_c,h_c) = u(c_c,h_c) = \frac{\left(c^\xi s(h)^{1-\xi}\right)^{1-\gamma}-1}{1-\gamma},
+U_k(c_k,h_k) = u(c_k,h_k) = \frac{\left(c^\xi s(h)^{1-\xi}\right)^{1-\gamma}-1}{1-\gamma},
 \end{equation}
@@
 
-The parent's utility also depends on the altruistic utility derived from the child:
+The parent's utility also depends on the altruistic utility derived from the kid:
 \begin{equation}
-U_p(c_p,h_p,c_c,h_c) = u(c_p,h_p) + \eta u(c_c,h_c),
-\end{equation}
-where $\eta$ measures the intensity of the parent's altruism towards the child.
-
-\textit{Intergenerational Transfers:} In the last period before death, the parent can leave a bequest, which the child receives in the next period.
-In all other periods, the parent can give inter-vivos transfers of a nonnegative amount, $t_p$, that the child receives immediately.
+U_p(c_p,h_p,c_k,h_k) = u(c_p,h_p) + \eta u(c_k,h_k),
+\end{equation}
+where $\eta$ measures the intensity of the parent's altruism towards the kid.
+
+\textit{Inter-Generational Transfers:} In the last period before death, the parent can leave a bequest, which the child receives in the next period.
+In all other periods, the parent can give inter-vivos transfers of a non-negative amount, $t_p$, that the child receives immediately.
 
@@
 They can rent housing of size $h_r$ or own houses of size $h_o$.
-The unit price of housing is $p$, and $q$ denotes the rent-to-price ratio.
+The unit price of housing is $p$ and $q$ denotes the rent-to-price ratio.
 I assume that prices are constant for tractability, but I model aggregate house price risk as in \cite{Corbae2015} in Appendix \ref{sec:robust_pricerisk} and find that the quantitative results are almost unchanged.
@@
 \end{cases}
-\end{equation} where $m_s$ and $m_b$ denote selling and buying costs, respectively. A household enters the period house $h_a$, while $h_{a+1}$ is the house chosen in this period.
+\end{equation} where $m_s$ and $m_b$ denote selling and buying costs. A household enters the period house $h_a$ while $h_{a+1}$ is the house chosen in this period.
 
@@
 
-In the U.S., borrowers making low down payments typically pay for private mortgage insurance (PMI) until their home equity surpasses a threshold \citep{goodman2017sixty}.
+In the US, borrowers making low down payments typically pay for private mortgage insurance (PMI)  until their home equity surpasses a threshold \citep{goodman2017sixty}.
 I model this as an extra fee $r^{pmi}$ that households must pay until they reach the $\overline{PMI}$ threshold.
@@
 r +r^m & \text{ if } b<0, \\
-r +r^m +r^{pmi}& \text{ if } b < \overline{PMI}\times p\times h_o.
+r +r^m +r^{pmi}& \text{ if } b < \overline{PMI}\times p\times h_o
 \end{cases}
@@
 \textit{Income Endowment:} Households are endowed with an income process with a deterministic life-cycle profile $l_a$.
-Children face persistent idiosyncratic age-dependent productivity shocks $y_{i,a}\in\mathcal{Y}_a=\{y_1,\dots,y_{N_y}\}$, which follow a Markov chain, where $\pi_a(y'|y)$ is the probability of switching from state $y$ to $y'$ at age $a$.
+Kids face persistent idiosyncratic age-dependent productivity shocks $y_{i,a}\in\mathcal{Y}_a=\{y_1,\dots,y_{N_y}\}$, which follow a Markov chain, where $\pi_a(y'|y)$ is the probability of switching from state $y$ to $y'$ at age $a$.
 Parents face no income uncertainty.
@@
 I assume that parents face no income risk for simplicity, consistent with the decrease in income risk with age \citep{Sanchez2020}.
-Furthermore, this paper focuses on the role of parental transfers for children' housing choices, where parental income risk is not a first-order concern.
+Furthermore, this paper focuses on the role of parental transfers for kids' housing choices, where parental income risk is not a first-order concern.
 However, in Appendix \ref{sec:robust_incomerisk}, I show that parental income and health expenditure risks do not meaningfully change my results.
 
-\textit{Initial Conditions of the New Child:} Households' initial wealth and productivity levels are stochastic yet correlated with the parent's wealth and productivity at age 53: $x_{25},y_{25} \sim F(x_{53},y_{53})$.
+\textit{Initial Conditions of the New Kid:} Households' initial wealth and productivity levels are stochastic yet correlated with the parent's wealth and productivity at age 53: $x_{25},y_{25} \sim F(x_{53},y_{53})$.
 While this paper does not consider parental investment in their child's human capital, such investments are an important factor in intergenerational persistence of economic outcomes \citep{Daruich2018,Lee2019}.
@@
 
-\textit{Timing:} The child's productivity is realized first.
-If the dynasty has a newly economically active child, they also realize their initial wealth.
+\textit{Timing:} The kid's productivity is realized first.
+If the dynasty has a newly economically active kid, they also  realize their initial wealth.
 Next, the decision problem is separated into two stages.
-First, the parent chooses consumption $c_p$, housing $h_p$, savings position $b_p$, and nonnegative inter-vivos transfers $t_p$.
-After observing the parent's choices, the child in the second stage chooses consumption $c_c$, housing $h_c$, and savings $b_c$.
-The parent acts first to align with U.S.
-mortgage regulations, which mandate that gifts be deposited before approving mortgages.
+First, the parent chooses  consumption $c_p$, housing $h_p$, savings position $b_p$, and non-negative inter-vivos transfers $t_p$.
+After observing the parent's choices, the kid in the second stage chooses consumption $c_k$, housing $h_k$, and savings $b_k$.
+The parent acts first to align with US mortgage regulations, which mandate that gifts be deposited before approving mortgages.
 
@@
 
-\textit{State Variables:} A parent $p$'s state variables are child's wealth $x_c\in X=[0,\infty)$, parent's wealth $x_p\in X$, child's housing $h_c\in H = \{h_r,h_o\}$, parent's housing $h_p\in H$, child's productivity $y_c\in Y_a=\{y_1,\dots,y_{N_y}\}$, and age $a_c\in A_c\in\{25,27,\dots,53\}$.
-The parent's state variable vector is $\mathbf{s}_p\equiv\left(x_c,x_p,y_c,h_c,h_p,a_c\right)$.
+\textit{State Variables:}  A parent $p$'s state variables are  kid's wealth $x_k\in X=[0,\infty)$, parent's wealth $x_p\in X$, kid's housing $h_k\in H = \{h_r,h_o\}$, parent's housing $h_p\in H$, kid's productivity $y_k\in Y_a=\{y_1,\dots,y_{N_y}\}$, and age $a_k\in A_k\in\{25,27,\dots,53\}$.
+The parent's state variable vector is  $\mathbf{s}_p\equiv\left(x_k,x_p,y_k,h_k,h_p,a_k\right)$.
 
 The child's state space differs because the parent's choices influence the child's decisions.
-The child is indifferent to receiving an additional unit of wealth or transfers, focusing only on total wealth (including transfers), $\tilde x_c\equiv x_c+t_p$.
-Furthermore, the parent's wealth and housing at the beginning of the period is redundant; the child only cares about the parent's bond and housing choices.
-The child's state variable vector is $\mathbf{s}_c=\left(b_p',h'_p,x_c+t_p,y_c,h_c,a_c\right)$.
+The kid is indifferent to receiving an additional unit of wealth or transfers, focusing only on total wealth (including transfers), $\tilde x_k\equiv x_k+t_p$.
+Furthermore, the parent's wealth and housing at the beginning of the period is redundant; the kid only cares about the parent's bond and housing choices.
+The kid's state variable vector is $\mathbf{s}_k=\left(b_p',h'_p,x_k+t_p,y_k,h_k,a_k\right)$.
 
 \textit{Decision Problems:} 
-I now show the decision problems for a dynasty at age $a_c=25$.
+I now show the decision problems for a dynasty at age $a_k=25$.
 For simplicity, I assume the child is a renter choosing to become an owner, while the parent remains a renter.
@@
 
-\textit{Child - Second Stage:} The child, conditional on becoming a new owner ($h_c=h_r,h'_c=h_o$), chooses consumption and bonds to maximize expected utility:
+\textit{Kids - Second Stage:} The kid, conditional on becoming a new owner ($h_k=h_r,h'_k=h_o$), chooses consumption and bonds to maximize expected utility:
 \begin{equation}\label{eq:Vk}
 \begin{split}
- V_c(\mathbf{s}_c)^{own} = \max_{c_c,b'_c,h'_c=h_o} & u(c_c,h'_c) + \beta \E\left[V_{c}(\mathbf{s}_c') \right] \\ 
- \text{s.t.}\quad & 	b'_c = x_c + t_p + w_c - c_c - p h'_c - m_b p h'_c \\
- & x_c' = b'_c(1+r(b'_c)) + p h'_c(1-\delta) \\
- & b_c' \ge - LTV p h'_c. 
+ V_k(\mathbf{s}_k)^{own} = \max_{c_k,b'_k,h'_k=h_o} & u(c_k,h'_k)  + \beta \E\left[V_{k}(\mathbf{s}_k') \right]  \\ 
+ \text{s.t.}\quad & 	b'_k = x_k + t_p + w_k - c_k - p h'_k - m_b p h'_k \\
+ & x_k' = b'_k(1+r(b'_k)) + p h'_k(1-\delta) \\
+ & b_k' \ge - LTV  p h'_k. 
 \end{split}
@@
 The third constraint is the LTV constraint.
-We see that a transfer $t_p$ can move the child away from the LTV constraint by increasing the bond position $b_c'$, holding consumption and housing fixed.
-
-For the other three possible housing choices (remain renting, remain owning, and own to rent), the problem is similar, but with appropriate changes to the budget equation, law of motion, and borrowing constraint.
-Finally, the child must choose between the two housing options:
+We see that a transfer $t_p$ can move the kid away from the LTV constraint by increasing the bond position $b_k'$, holding consumption and housing fixed.
+
+For the other three possible housing choices (remain renting, remain owning, and own to rent), the problem is similar, but with appropriate changes to the budget equation, law of motion, and the borrowing constraint.
+Finally, the kid must choose between the two housing options:
 \begin{equation}
-	V_c(\mathbf{s}_c) = \max_{h_c'} \left\{V_c(\mathbf{s}_c)^{rent},V_c(\mathbf{s}_c)^{own}\right\}.
+	V_k(\mathbf{s}_k) = \max_{h_k'} \left\{V_k(\mathbf{s}_k)^{rent},V_k(\mathbf{s}_k)^{own}\right\}.
 \end{equation}
@@
 \begin{split}
-V_p(\mathbf{s}_p)^{rent} = \max_{c_p,b_p',t_p} & u(c_p,h'_p) + {\eta} u\left (c_c^*({\mathbf{s}_c}),h_c'^*({\mathbf{s}_c})\right ) + \beta \E \left[V_{p}({\mathbf{s}_p'})\right] \\
+V_p(\mathbf{s}_p)^{rent} = \max_{c_p,b_p',t_p} & u(c_p,h'_p) + {\eta} u\left (c_k^*({\mathbf{s}_k}),h_k'^*({\mathbf{s}_k})\right )  +  \beta \E \left[V_{p}({\mathbf{s}_p'})\right] \\
 \text{s.t.}\quad  & b'_p = x_p + w_p - c_p - t_p - q p h_p'\\
@@
 The law of motion of net worth is simply the bond position for renters.
-The third and fourth constraints impose that transfers be nonnegative and that renters may not borrow, respectively.
+The third and fourth constraints impose that transfers be non-negative and that renters may not borrow.
 Finally, the parent chooses whether to rent or own.
@@
 When making decisions, households consider how their choices will impact future states, as we can see from the next-period state vectors.
-For example, the parent internalizes that their next-period state includes the child's next-period wealth and housing, which are functions of the child's current choices, which in turn depend on the parent's current choices:
+For example, the parent internalizes that their next-period state includes the kid's next-period wealth and housing, which are functions of the kids's current choices, which in turn depend on the parent's current choices:
 \begin{equation}
-	\mathbf{s}_p' = \left(x_c'^{*}({\mathbf{s}_c}),x_p',y_c',h'^*_c({\mathbf{s}_c}),h'_p,a_c'\right).
-\end{equation}
-Similarly, the child knows that their next-period state includes the parent's next-period bond, housing, and transfer choices, which depend on the parent's next-period state, which in turn depends on the child's current choices:
+	\mathbf{s}_p' = \left(x_k'^{*}({\mathbf{s}_k}),x_p',y_k',h'^*_k({\mathbf{s}_k}),h'_p,a_k'\right).
+\end{equation}
+Similarly, the kid knows that their next-period state includes the parent's next-period bond, housing, and transfer choices, which depend on the parent's next-period state, which in turn depends on the kid's current choices:
 \begin{equation}
-	\mathbf{s}_c' = (b_p''^*({\mathbf{s}_p'}),h_p''^*({\mathbf{s}_p'}),x'_c + t_p'^*({\mathbf{s}_p'}),y'_c,h_c',a_c').
+	\mathbf{s}_k' = (b_p''^*({\mathbf{s}_p'}),h_p''^*({\mathbf{s}_p'}),x'_k + t_p'^*({\mathbf{s}_p'}),y'_k,h_k',a_k').
 \end{equation}
@@
 \subsection{Markov Perfect Equilibrium Definition}
-A stationary equilibrium, which is Markov Perfect, consists of value functions $V_c(\mathbf{s}_c)$ and $V_p(\mathbf{s}_p)$; policy functions of the child $c_c'^*(\mathbf{s}_c)$, $b_c'^*(\mathbf{s}_c)$, $h_c'^*(\mathbf{s}_c)$; and the parent $c_p'^*(\mathbf{s}_p)$, $b_p'^*(\mathbf{s}_p)$, $h_p'^*(\mathbf{s}_p)$, $t_p'^*(\mathbf{s}_p)$; and a distribution of households $\psi(\mathbf{s}_p)$ such that i) in each repetition of the parent-child game, the parent's policy functions are optimal given the child's policy functions; ii) the child's policy functions are optimal given the parent's policy functions; and iii) the measure of households is invariant.
+A stationary equilibrium, which is Markov Perfect, consists of value functions $V_k(\mathbf{s}_k)$ and $V_p(\mathbf{s}_p)$;  policy functions of the kid $c_k'^*(\mathbf{s}_k)$, $b_k'^*(\mathbf{s}_k)$, $h_k'^*(\mathbf{s}_k)$; and the parent $c_p'^*(\mathbf{s}_p)$, $b_p'^*(\mathbf{s}_p)$, $h_p'^*(\mathbf{s}_p)$, $t_p'^*(\mathbf{s}_p)$; and a distribution of households $\psi(\mathbf{s}_p)$ such that: i) In each repetition of the parent-kid game, the parent's policy functions are optimal given the kid's policy functions, ii) the kid's policy functions are optimal given the parent's policy functions, and iii) the measure of households is invariant.
 I derive the measure $\psi$ of households in Appendix \ref{sec:hhmeasures}.
@@
 
-\subsubsection{When Do Parents Transfer?}
-Due to nonconvexities introduced by housing, the policy and value functions are not differentiable everywhere.
+\subsubsection{When do Parents Transfer?}
+Due to non-convexities introduced by housing, the policy and value functions are not differentiable everywhere.
 To focus on the underlying intuition, I will sidestep the technical issues and refer to \cite{Barczyk2020} and \cite{Chu2020} for a formal derivation.
 
-Consider a a wealthy parent with a poor, borrowing-constrained child.
-This implies that the child does not save and must rent and, thus, consumes any additional wealth.
-By assumption, the child's policy functions are locally differentiable, with $\frac{\partial c'_c(\cdot)}{\partial t_p}=1$ and $\frac{\partial x'_c(\cdot)}{\partial t_p}=\frac{\partial h'_c(\cdot)}{\partial t_p}=0$.
-We can ignore the nonnegativity constraint on transfers, since the parent is wealthy.
+Consider a a wealthy parent and with a poor, borrowing-constrained kid.
+This implies that the kid does not save and must rent, and thus consumes any additional wealth.
+By assumption, the kid's policy functions are locally differentiable, with $\frac{\partial c'_k(\cdot)}{\partial t_p}=1$ and $\frac{\partial x'_k(\cdot)}{\partial t_p}=\frac{\partial h'_k(\cdot)}{\partial t_p}=0$.
+We can ignore the non-negativity constraint on transfers, since the parent is wealthy.
 The optimality condition (from Eq.
-\ref{eq:Vp}) is
+\ref{eq:Vp}) is:
 \begin{equation}\label{eq:FOC}
-	u'(c_p,h_p) = \eta u'(c_c + t_p,h_c).
-\end{equation}
-This is the classic first-order condition from \cite{Altonji1997a} and says that the parent transfers until their own marginal utility equals the altruism-weighted marginal utility of the child.
-This is the fundamental driver of the transfer motives in the model and, as we will see, interacts with housing choices in interesting ways.
+	u'(c_p,h_p) = \eta u'(c_k + t_p,h_k)
+\end{equation}
+This is the classic first-order condition from \cite{Altonji1997a}, and says that the parent transfers until their own marginal utility equals the altruism-weighted marginal utility of the kid.
+This is the fundamental driver of the transfer motives in the model, and as we  will see interacts with housing choices in interesting ways.
 As discussed extensively in \cite{Barczyk2020a} and \cite{Chu2020}, other transfer motives exist, such as ``pushes to autarky'', though these do not appear in my parameterization.
@@
 \subsubsection{Transfers and Homeownership}
-In the model, homeownership is determined by a threshold rule: Households become homeowners once their wealth is sufficiently high.
+In the model, homeownership is determined by a threshold rule; households become homeowners once their wealth is sufficiently high.
 The presence of transfers both moves the threshold (direct effect) and shifts the wealth level (indirect effect).
@@
 
-First, for a child with a given state, a transfer increases wealth, potentially pushing the child above the threshold, increasing homeownership indirectly.
-
-Second, altruism reduces the child's savings motive, and so fewer children accumulate enough wealth to cross the threshold, lowering homeownership indirectly.
+First, for a kid with a given state, a transfer increases  wealth, potentially pushing the child above the threshold, increasing homeownership indirectly.
+
+Second, altruism reduces the kid's savings motive, and so fewer kids accumulate enough wealth to cross the threshold, lowering homeownership indirectly.
 
 Third, transfers alleviate the LTV constraint in two ways.
-First, the parent give transfers to alleviate the borrowing constraint for a child who would otherwise be unable to buy.
-Second, the parent give transfers to a borrowing-constrained child, who can only afford very low consumption, shifting the ownership threshold down.
+First, the parent give transfers to alleviate the borrowing constraint for a kid who would otherwise be unable to buy.
+Second, the parents give transfers to a borrowing-constrained buyer, who can only afford very low consumption, shifting the ownership threshold down.
 The first increases homeownership directly and the second indirectly.
@@
 Fourth, transfers mitigate the costs of illiquid housing.
-Consider an owning child who receives a series of bad income shocks.
-Without transfers, the child optimally downsizes and pays the sales cost to better align consumption and housing expenditures.
+Consider an owning kid who receives a sequence of bad income shocks.
+Without transfers, the kid optimally downsizes and pays the sales cost, to better align consumption and housing expenditures.
 This creates a strong transfer motive, as transfers can save the dynasty from paying the sales cost.
-This increases homeownership directly by shifting the ownership threshold down.
+This increases homeownership directly, by shifting the ownership threshold down.
 
@@
 Notably, both the child and the parent strategically internalize these effects.
-For example, a child can use illiquid housing to commit to having low liquid wealth, thereby increasing the marginal utility of wealth and, consequently, the transfer.
+For example, a kid can use illiquid housing to commit to having low liquid wealth, thereby increasing the marginal utility of wealth and, consequently, the transfer.
 Conversely, a parent can tie up wealth in illiquid housing, increasing the marginal utility of wealth and, consequently, decreasing transfers.
-These strategic considerations are illustrated in the child's policy functions for homeownership and savings (Fig.
+These strategic considerations are illustrated in the kid's policy functions for homeownership and savings (Fig.
 \ref{fig:pol}).
@@
 
-First, altruism lowers the threshold for homeownership among households with some parental support: Wealthier parents make homeownership more attractive.
+First, altruism lowers the threshold for homeownership among households with some parental support: wealthier parents make homeownership more attractive.
 An increase in parental wealth shifts the threshold further down.
 
-Next, we observe the child's savings choices.
+Next, we observe the kid's savings choices.
 Altruism reduces saving in two ways.
-First, the shift in the ownership threshold means that the child switches from saving to borrowing sooner.
-Second, even in regions where the child makes the same ownership choice, they save less.
+First, the shift in the ownership threshold means that the kid switches from saving to borrowing sooner.
+Second, even in regions where the kid makes the same ownership choice, they save less.
 For example, on the far left where all households rent, the child saves less if the parent is richer.
-On the right, where all households own, the child takes out a larger mortgage, holding wealth fixed.
-Note that once the child becomes sufficiently wealthy, they behave as if there is no altruism.
+On the right, where all households own, the kid takes out a larger mortgage, holding wealth fixed.
+Note that once the kid becomes sufficiently wealthy, they behave as if there is no altruism.
 Comparing the orange and dark red lines, we find that an increase in parental wealth decreases saving at all wealth levels.
@@
 	\end{center}
-	{\begin{footnotesize} \textit{Notes:} The dashed vertical lines indicate the threshold where the child switches from renting to owning.
+	{\begin{footnotesize} \textit{Notes:} The dashed vertical lines indicate the threshold where the kid switches from renting to owning.
 The horizontal dotted lines indicate the PMI and LTV thresholds.
-The child states $\mathbf{s}_c'$ are \input{../tabfig/model/pol_bond_state.tex}\unskip.
+The child states $\mathbf{s}_k'$ are \input{../tabfig/model/pol_bond_state.tex}\unskip.
 The figure plots the policy functions in the estimated model.
@@
 
-More generally, the child's savings choices show that with altruism, households are more likely to choose to be liquidity constrained, often called hand-to-mouth, as they become homeowners with less wealth and are closer to the borrowing constraints.
+More generally, the kid's savings choices show that with altruism, households are more likely to choose to be liquidity constrained, often called hand-to-mouth, as they become homeowners with less wealth and are closer to the borrowing constraints.
 Altruism reduces the downside of liquidity constraints---namely, the inability to smooth income shocks through partial insurance---while creating the advantage of receiving larger transfers.
@@
 \section{Parameter Selection}\label{sec:estimation}
-To quantify the impact of parental transfers on homeownership and to perform policy evaluations, I estimate the model using data from the PSID and a standard two-step estimation procedure.
+To quantify the impact of parental transfers on homeownership and perform policy evaluations, I estimate the model using data from the PSID and a standard two-step estimation procedure.
 First, parameters independent of the model structure are estimated from the data or sourced from the literature.
 Second, I estimate the remaining parameters using simulated method of moments (SMM).
-I validate the model using nontargeted moments from the PSID as well as experimental evidence from empirical studies.
+I validate the model using non-targeted moments from the PSID as well as experimental evidence from empirical studies.
 
@@
 
-\textit{Income Process:} The income process consists of a deterministic life-cycle component $l_a$ for all ages and a persistent shock $y_a$ for children.
-
-To calibrate the deterministic component, I first find the average income for each age by year and then average over all years.
+\textit{Income Process:} The income process consists of a deterministic life-cycle component $l_a$, for all ages, and a persistent shock $y_a$, for kids.
+
+To calibrate the deterministic component, I first find the average income for each age by year, and then average over all years.
 I then regress average income on a fourth-order polynomial of age.
-The fitted values give the deterministic component for households aged 25 to 65.
-I find retirement income by dividing the average income for households aged 67 to 83 by the predicted income at age 66 (Figure \ref{fig:y_a}).
+The fitted values give the deterministic component for households aged 25-65.
+I find retirement income by dividing the average income for households aged 67-83 by the predicted income at age 66 (Figure \ref{fig:y_a}).
 
@@
 \textit{Housing Parameters:} I set the rental rate $q=0.10$ \citep{Davis2008}, housing depreciation $\delta=0.05$ \citep{Harding2007}, and sales cost $m_s=0.075$ and purchase cost $m_b=0.02$ \citep{Yang2009}.
-I set the price level to $p=\$232,918/h_o$ to match the average market value of owner-occupied houses for households aged 25 to 44 in my sample.
+I set the price level to $p=\$232,918/h_o$, to match the average market value of owner-occupied houses for households aged 25-44 in my sample.
 The size $h_o$ is estimated internally.
@@
 I set $LTV=0.90$, broadly in line with the literature, to account for the mass of loans with LTVs above 80\%.
-The interest rate on savings is set at $r = 0.04$ (2\% per annum), while the mortgage premium is set at $r^m=0.03$ (1.5\% per annum).
-Both parameters vary in the literature, but both are typically set around 1\% to 2\% per annum \citep[see e.g.,][]{Cocco2005b,Kaplan2020,Paz-Pardo2019}.
+The interest rate on savings is set at $r = 0.04$ (2\% per annum) while the mortgage premium $r^m=0.03$ (1.5\% per annum).
+Both parameters varies in the literature, but both are typically set around 1\%-2\% per annum \citep[see e.g.,][]{Cocco2005b,Kaplan2020,Paz-Pardo2019}.
 % Mortgage premium: 1 percent in Kaplan et al (0.33 wedge on top of 3% interest rate)
@@
 The parameter $\xi$ pins down the desired expenditure share of housing consumption and is set to 0.2.
-This parameter is typically set around 0.15 to 0.25 (e.g., \citealp{Kaplan2020,Chatterjee2015,Paz-Pardo2019,davis2011household}) based on the share of housing expenditures in personal consumption expenditures.
+This parameter is typically set around 0.15-0.25 (e.g., \citealp{Kaplan2020,Chatterjee2015,Paz-Pardo2019,davis2011household}) based on the share of housing expenditures in personal consumption expenditures.
 Due to minimum house sizes, low-income households generally spend more on rent than implied by $\xi$.
@@
 
-\textit{Initial Conditions of the Young $y_{c,25},x_{c,25}\sim F(x_{c,53},y_{c,53})$:} 
+\textit{Initial Conditions of the Young $y_{k,25},x_{k,25}\sim F(x_{k,53},y_{k,53})$:} 
 The initial distribution is set to match the  intergenerational correlation in wealth and income.
 I estimate it nonparametrically for transparency.
-First, I limit the sample to households aged 24 to 26 with parents 16 to 40 years older.
-I then divide wealth and income into four and three quantiles, respectively, for both parents and children.
-The PDF $F(\cdot)$ is then given by the empirical probability of each parent-child combination.
-The wealth level of a quartile is defined by its median value.
+First, I limit the sample to households aged 24-26 with parents 16-40 years older.
+I then divide wealth and income into 4 and 3 quantiles, respectively, for both parents and kids.
+The PDF $F(\cdot)$ is then given by the empirical probability of each parent-kid combination.
+The wealth level for each kid wealth quartile is the empirical median within that quartile.
 
@@
 \ref{fig:etaid}).
-I exclude transfers below \$500, as they are unimportant for homeownership and including them would increase the transfer rate.
+I exclude transfers below \$500 as they are unimportant for homeownership and including them would increase the transfer rate.
 
@@
 My estimate of altruism $\eta$ should be viewed as a lower bound.\footnote{I have experimented with targeted higher transfer rates.
-This requires stronger altruism, which, in turn, increases homeownership motives, lowering the estimated utility premium $\chi$.
-Overall, the results are qualitatively similar to my conservative estimate, which understates the role of altruistic transfers.}
+This requires stronger altruism, which in turn increases homeownership motives, lowering the estimated utility premium $\chi$.
+Overall, the results are qualitatively similar to my conservative estimate, which understates the role of the Bank of Mom and Dad.}
 
@@
 \subsection{Model Fit}
-I estimate the remaining parameters $\theta=\{\eta,\chi,h_o\}$ as follows.
+I use the simulated method of moments (SMM) to estimate the remaining parameters $\theta=\{\eta,\chi,h_o\}$ as follows.
 First, let $\hat m$ denote the vector of targeted empirical moments.
@@
 	\begin{threeparttable}
-	\caption{Model Estimation}\label{tab:esttable}
+	\caption{Model Estimation Moments}\label{tab:esttable}
 		\input{../tabfig/est/esttable.tex}
@@
 
-\subsubsection{Nontargeted Moments}\label{sec:externalval}
-To validate the model, I show that it also matches nontargeted moments.
+\subsubsection{Non-Targeted Moments}\label{sec:externalval}
+To validate the model, I show that it also matches non-targeted moments.
 
@@
 Third, households hold slightly smaller mortgages than in the data, as their flexibility allows households to pay down debt quickly and borrow later if necessary.
-Fourth, the model also matches the LTV ratio for buyers well.\footnote{Note that this is not the LTV at origination but the first time the household is observed as an owner.
+Fourth, the model also matches the LTV ratio for buyers well.\footnote{Note that this is not the LTV at origination, but the first time the household is observed as an owner.
 If a household bought just after the previous survey, the LTV would decrease due to price appreciation, improvements, and mortgage paydown.}
@@
 \subsubsection{Household-Level Responses to Income Shocks by Parental Wealth}
-As a final external validation, I repeat the exercise in Section \ref{sec:eventstudy} (see Figure \ref{fig:housinggrowthrates}), where we saw that households with wealthy parents do not downsize during unemployment spells.
-Since the model lacks an explicit unemployment state, households are considered unemployed if their productivity drops to the lowest level for a single period.
-Otherwise, I replicate the exercise exactly.
-Figure \ref{fig:housinggrowthrates_model} shows that the model matches both the qualitative and quantitative patterns from the empirical data: Housing consumption falls for households without wealthy parents during ``unemployment,'' while it grows for those with wealthy parents.
+As a final external validation, I repeat the exercise in Section \ref{sec:eventstudy} (see Fig.
+\ref{fig:housinggrowthrates}), where we saw that households with wealthy parents do not downsize during unemployment spells.
+Since the model lacks an explicit unemployment state, households are considered unemployed if their productivity drops to the lowest level for a single period.Otherwise, I replicate the exercise exactly.
+Figure \ref{fig:housinggrowthrates_model} shows that the model matches both the qualitative and quantitative patterns from the empirical data: housing consumption falls for households without wealthy parents during ‘unemployment,’ while it grows for those with wealthy parents.
 
@@
 Table \ref{tab:modelfitnoa} reports how the moments change without altruism and transfers.
-Homeownership rates decline by 14 percentage points from 49\%, a decrease of 29\% for households aged 25 to 44.
-This is a large decrease, since households endogenously choose to save more without transfers.
-By comparison, homeownership only decreased by about 8 percentage points during the financial crisis.
-
-The reduction in homeownership cannot be attributed to an inability to afford buying: Without altruism, median wealth of the young increases (increasing homeownership, in isolation) to the average wealth level at purchase with altruism.
-The main driver is that the ownership threshold shifts out in wealth without altruism, as we saw in the policy functions (Section \ref{sec:policyfuncs}).
+Homeownership rates decline by 14 pp.
+from 49\%, a decrease of 29\% for households aged 25-44.
+This is a large decrease since households endogenously choose to save more without transfers.
+By comparison, homeownership only decreased by about 8 percentage points during the financial crisis
+
+The reduction in homeownership cannot be attributed to an inability to afford buying: without altruism, median wealth of the young increases (increasing homeownership, in isolation) to the average wealth level at purchase with altruism.
+The main driver is that the ownership threshold shifts out in wealth without altruism, as we saw in the policy functions (Sec.
+\ref{sec:policyfuncs}).
 Without transfers then, the average wealth at purchase increases by about \$12,500 (about 33\%), the age of first ownership is delayed by about 2.5 years, and the LTV at purchase falls by 13 percentage points.
-Although homeownership of young households declines sharply, the overall homeownership rate of all households (ages 25 to 74) only falls by 6 percentage points (8\%).
-Hence, the main aggregate effect of parental transfers is that they induce earlier homeownership.
+Although homeownership of young households declines sharply, the overall homeownership rate of all households (25-74) only falls by 6 percentage points (8\%).
+Hence the main aggregate effect of parental transfers is that they induce earlier homeownership.
 Finally, without altruism, the parental wealth gradient in homeownership decreases from 1.94 to 1.14, with the remaining effect driven by persistent productivity through initial conditions $F()$.
@@
 I use a closed-form supply function because the model is too complex to introduce a construction sector.
-The rent-to-price ratio is unchanged, as in \cite{Kaplan2020}, where rental units are supplied by deep-pocketed landlords who convert rentals to owner-occupied units if the price deviates from the present value of rents.
+The rent-to-price ratio is unchanged as in \cite{Kaplan2020}, where rental units are supplied by deep-pocketed landlords who convert rentals to owner-occupied units if the price deviates from the present value of rents.
 Letting $\alpha_1\to\infty$ yields a perfectly elastic supply function, leaving prices constant, as in the previous section.
@@
 Allowing house prices to adjust does not meaningfully decrease the contribution of transfers to homeownership or other aggregate outcomes.
-This is due to the small difference in homeownership rates (ages 25 to 74) with and without altruism.
+This is due to the small difference in homeownership rates (ages 25-74) with and without altruism.
 Prices decline by just 7\% even with a supply elasticity as low as 1.
@@
 
-\textit{Relaxing Mortgaging Regulation:} Down-payment requirements are widely seen as a barrier to homeownership \citep[e.g.,][]{Lee2018,Guiso2002}.
+\textit{Relaxing Mortgaging Regulation:} Downpayment requirements are widely seen as a barrier to homeownership \citep[see e.g.,][]{Lee2018,Guiso2002}.
 I tighten the LTV requirement by 3 percentage points to explore how mortgage regulation affects the parental wealth-housing channel.
-The tighter LTV limit decreases homeownership among the young by 5 percentage points (10\%), as households must have larger down payments.
+The tighter LTV limit decreases homeownership among the young by 5pp (10\%), as households must have larger downpayments.
 Homeownership falls more among households with parents in the middle of the wealth distribution, reducing the parental wealth gradient.
-While at first counterintuitive, the borrowing constraint is more binding for these households, all else being equal, since homeownership is more attractive when you have access to some parental wealth and a small transfer pushes you into ownership.
-However, the parents aren't wealthy enough to render the borrowing constraint irrelevant.
-This result suggests that relaxing borrowing constraints increases not only homeownership for all households, but also the role of parental wealth.
-Indeed, \cite{wold2024housing} find that households with poorer parents have lower LTV ratios at origination, suggesting that the LTV constraint is less binding for households with poorer parents.
-
-\textit{Lowering PMI:} Conventional mortgages require that mortgages with down payments below 20\% have PMI \cite{goodman2017sixty}.
+While at first counterintuitive, the borrowing constraint is more binding for these households, all else equal, since homeownership is more attractive when you have access to some parental wealth and a small transfer pushes you into ownership.
+However, the parents arent wealthy enough to render the borrowing constraint irrelevant.
+This result suggests that relaxing borrowing constraints increases homeownership for all households, but also the role of parental wealth.
+
+\textit{Lowering PMI:} Conventional mortgages require that mortgages with downpayments below 20\% have PMI \cite{goodman2017sixty}.
 I relax the PMI limit by 7 percentage points to explore how PMI affects the parental wealth-housing channel.
-The relaxed PMI limit increases homeownership by 2 percentage points (4\%).
+The relaxed PMI limit increases homeownership by 2pp (4\%).
 Similar to the LTV regulation, homeownership among households with  parents in the middle of the wealth distribution rises more, further widening the parental wealth gradient.
-This result suggests that relaxing the PMI limit increases not only homeownership for all households, but also the role of parental wealth.
+This result suggests that relaxing the PMI limit increases homeownership for all households, but also the role of parental wealth.
 
@@
 I lower house prices by \$6,000 (7\%) to explore how the price level affects the parental wealth-housing channel.
-As prices decline, homeownership increases for all households, though more so for households with the richest parents, slightly increasing the role of parental wealth.
+As prices decline, homeownership increases for all households, though more so for households with the ricest parents, slightly increasing the role of parental wealth.
 
 \textit{Reducing Purchase Costs:} Another way to lower entry barriers is to reduce the costs of buying a house, and policymakers have lowered government fees and closing costs for first-time buyers.
-For example, in 2017, the stamp duty was lowered for first-time buyers in England.
+For example, in 2017 the stamp duty was lowered for first-time buyers in England and the 2024 Harris campaign has proposed a \$10,000 tax credit for first-time buyers.
 I decrease the purchase cost ($m_b$) by 2 percentage points (equivalent to a grant of about \$6,250) to understand how purchase costs affect the parental wealth-housing channel.
 While lower purchase costs increase homeownership, the gains are concentrated among households with wealthy parents, amplifying the parental wealth gradient.
-For these households, purchase costs were a key barrier to homeownership, because they are less likely to downsize after income shocks (Figure \ref{fig:housinggrowthrates_model}).
+For these households, purchase costs were a key barrier to homeownership, because they are less likely to downsize after income shocks (Fig.
+\ref{fig:housinggrowthrates_model}).
 This suggests that grants or credits for first-time buyers primarily benefit households with the wealthiest parents.
@@
 In contrast, homeownership declines by 9 percentage points for households with moderately wealthy parents, leading to a net decline in homeownership.
-Homeownership decreases for households with moderately wealthy parents, as it is harder to use housing to extract transfers.
-For households with the poorest parents, who rarely get transfers, and the wealthiest parents, who frequently get transfers, there is no need to use the sales cost to ``extract'' future transfers.
+Homeownership decreases for households with moderately wealthy parents as it is harder to use housing to extract transfers.
+Households with the poorest parents, who rarely get transfers, and the wealthiest parents, who frequently get transfers, there is no need to use the sales cost to ``extract'' future transfers.
 
@@
 These experiments highlight two main results.
-First, parental wealth is an important determinant of homeownership, and having richer parents makes households more likely to prefer owning to renting.
+First, parent wealth is an important determinant of homeownership, and having richer parents makes households more likely to prefer owning to renting.
 Some common policies intended to increase homeownership rates change the role of parental wealth in housing outcomes.
-Relaxing credit constraints increases not only homeownership, but the role of parental wealth, since housing is more attractive to households with wealthier parents.
-
-However, increasing liquidity, either through reducing purchase costs or sales costs, decreases the role of parental wealth.
+Relaxing credit constraints increases homeownership but also the role of parental wealth, since housing is more attractive to households with wealthier parents.
+
+However, increasing liquidity, either through reducing purchase costs or sales costs, decreases the role of parent wealth.
 Increasing liquidity can lower homeownership, as children of wealthy parents can no longer use housing as a commitment device to receive larger transfers.
@@
 The previous section established that higher sales costs can increase ownership among households with wealthy parents.
-Here, I present a more surprising result: Households with wealthy parents prefer illiquidity.
+Here, I present a more surprising result: households with wealthy parents prefer illiquidity.
 
@@
 Without altruism, all households prefer to remove sales costs.
-I find that while all parents prefer to remove sales costs, roughly one-quarter of children prefer them.
-The key determinant is parental wealth: Parents of children who prefer sales costs are approximately three times wealthier.
+I find that, while all parents prefer to remove sales costs, roughly one-quarter of children prefer them.
+The key determinant is parental wealth: parents of children who prefer sales costs are approximately three times wealthier.
 Children who prefer sales costs are also more likely to receive transfers (48\% vs.
-13\%) and, conditional on receiving transfers, receive three times larger transfers.
+13\%), and, conditional on receiving transfers, receive three times larger transfers.
 Notably, households that own their homes are more likely to prefer sales costs on their housing.
@@
 The next figures plot average homeownership, parental wealth, and child wealth by liquidity preference.
-To use illiquid housing to extract transfers, households must be homeowners, so households who prefer illiquidity have higher homeownership (Figure \ref{fig:pref_own}).
-To extract transfers, parents must have wealth, so households who prefer illiquidity have wealthier parents (Figure \ref{fig:pref_wealthp}).
+To use illiquid housing to extract transfers, households must be homeowners, so households who prefer illiquidity have higher homeownership (Fig.
+\ref{fig:pref_own}).
+To extract transfers, parents must have wealth, so households who prefer illiquidity have wealthier parents (Fig.
+\ref{fig:pref_wealthp}).
 Finally, Figure \ref{fig:pref_wealthk} plots the child's wealth, and we see that households who prefer illiquidity are initially wealthier, but this relationship flips at age 33.
@@
 Similarly, \cite{aliprantis2022dynamics} find that closing the income gap is the most effective mechanism to reduce the wealth gap.
-While these papers account for racial differences in inheritances, they do not include endogenous inter-vivos transfers and ignore housing.\footnote{\cite{aliprantis2022dynamics} include a form of inter-vivos transfer receipt: Households leave pooled bequests at death, primarily to middle-aged households.
-There are no strategic considerations, since the bequests are stochastically distributed and parents do not transfer directly to their children.} While the structural papers have omitted housing, many empirical papers have looked at racial differences in housing outcomes.
-For example, \cite{charles2002transition} find that differences in parental wealth account for 25\% of the Black-White mortgage application gap, and \cite{bond2021role} find that parental wealth accounts for 28\% of the gap in maintaining homeownership.
-I now use my model to bridge the gap between the structural and empirical papers and to quantify the contribution of parental wealth to the Black-White Homeownership gap.
+While these papers account for racial differences in inheritances, they do not include endogenous inter-vivos transfers and ignore housing.\footnote{\cite{aliprantis2022dynamics} include a form of inter-vivos transfer receipt: households leave pooled bequests at death, primarily to middle-aged households.
+There are no strategic considerations since the bequests are stochastically distributed and parents do not transfer directly to their children.} While the structural papers have omitted housing, many empirical papers have looked at racial differences in housing outcomes.
+For example, \cite{charles2002transition} find that differences in parental wealth account for 25\% of the Black-White mortgage application gap and \cite{bond2021role} find that parental wealth accounts for 28\% of the gap in maintaining homeownership.
+I now use my model to bridge the gap between the structural and empirical papers, and quantify the contribution of parental wealth to the Black-White Homeownership gap
 \begin{table}
@@
 I follow the methodology in \cite{Ashman2020} and \cite{aliprantis2022dynamics} closely.
-I shift the income level ($l_a$) according to racial differences: In the matched parent-child household, White households have 8\% higher income and Black households 42\% lower than the average.
+I shift the income level ($l_a$) according to racial differences: In the matched parent-child household, White households have 8\% higher income, and Black households 42\% lower than the average.
 I then solve the model for White households.
@@
 Transfers account for 7 percentage points (14\%) of the Black-White gap.
-Though Black and White households receive transfers at similar rates in the model, transfers are less important for Black households' homeownership, as they receive smaller amounts due to lower parental wealth.
+Though Black and White households receive transfers at similar rates in the model, transfers are less important for Black households’ homeownership, as they receive smaller amounts due to lower parental wealth.
 My results show that racial differences in parental income, which affect parental wealth and transfers, are significant contributors to racial gaps in housing outcomes.
-Combined with the previous policy analysis, these results suggest that policies such as tax rebates or down-payment assistance for low-income or first-time buyers may widen the Black-White homeownership gap rather than narrow it.
+Combined with the previous policy analysis, these results suggest that policies such as tax rebates or downpayment assistance for low-income or first-time buyers may widen the Black-White homeownership gap rather than narrowing it.
 However, the model omits many other factors beyond income, such as discriminatory lending and appraisal bias, that affect the Black-White homeownership gap  (e.g.,  \cite{shapiro2004hidden}).
@@
 I build and estimate a dynamic life-cycle model of homeownership where altruistic parents may transfer to their adult children in every period.
-I use the model to quantify the importance of parental transfers for young adults' homeownership and to understand how policies that lower barriers to homeownership interact with these transfers.
+I use the model to quantify the importance of parental transfers for young adults’ homeownership and understand how policies that lower barriers to homeownership interact with these transfers.
 
 Using a counterfactual experiment without altruism (a standard life-cycle model), I find that transfers account for 14 percentage points (29\%) of the homeownership rate.
-I show that policies that lower entry barriers to homeownership amplify the importance of parental wealth on housing outcomes, while those that lower the downsides of illiquid homeownership reduce the importance of parental wealth.
+I show that policies that lower entry barriers to homeownership amplify the importance of parental wealth on housing outcomes while those that lower the downsides of illiquid homeownership reduce the importance of parental wealth.
 Finally, I study the role of housing illiquidity and parental transfers.
@@
 
-The second contribution of the paper is to include an illiquid asset in the altruism framework.
-With altruistic parents, transfers generally flow to borrowing-constrained households, since they have a large marginal utility of wealth \citep{Barczyk2020,Chu2020}.
+The second contribution of the paper is to include an illiquid asset into the altruism framework.
+With altruistic parents, transfers generally flow to borrowing-constrained households since they have a large marginal utility of wealth \citep{Barczyk2020,Chu2020}.
 With a single asset, this framework implies that households who receive transfers are poor.
@@
 Adding richer housing dynamics and equilibrium effects would allow one to study other interesting questions, such as the extent to which transfers contribute to house price growth.
-My model omits a feedback loop, in which transfers raise demand and, thus, house prices, boosting the wealth of homeowning parents and, thus, higher transfers, and so on.
-Another potential extension could examine whether illiquidity may reduce the commitment problem in the family.
+My model omits a feedback loop, in which transfers raise demand and thus house prices, boosting the wealth of homeowning parents and thus higher transfers, and so on.
+Another potential extension would examine whether illiquidity may reduce the commitment problem in the family.
 Additionally, my preferences-for-illiquidity results have implications for the literature on high-MPC households \citep[see e.g.,][]{kaplan2022marginal}.
-In particular, we are likely overestimating the average MPC of liquidity-constrained households with wealthy parents, since these households use parental transfers to smooth shocks.
+In particular, we are likely overestimating the average MPC of liquidity-constrainted households with wealthy parents, since these households use parental transfers to smooth shocks.
 For example, following a negative shock, consumption may fall less due to transfers, while for positive shocks, consumption may rise less as parents reduce transfers.
@@
 \subsection{Decision Problems at Age 53 and 83}\label{sec:decextra}
-As depicted in Figure \ref{fig:overview}, several transitions occur between age 53 and 55: The `old' parent dies, the `old' child becomes a `new' parent, and the `new' child becomes economically active.
-
-These transitions mean that the child's decision problem at age 53 is very different from earlier ages (Eq.
+As depicted in Figure \ref{fig:overview}, several transitions occur between age 53 and 55: The `old' parent dies, the `old' kid becomes a `new' parent, and the `new' kid becomes economically active.
+
+These transitions mean that the kid's decision problem at age 53 is very different from earlier ages (Eq.
 \ref{eq:Vk}).
-First, the continuation value is now the parent's value function at age 25 of the new child ($V_{p}({\mathbf{s}'_{p,25}})$).
-Second, the expectation is over the new child's initial productivity and net worth ($x_{c,25},v_{c,25}\sim F(x_{c,53},y_{c,53}$)), instead of the old child's productivity.
+First, the continuation value is now the parent's value function at age 25 of the new kid ($V_{p}({\mathbf{s}'_{p,25}})$).
+Second, the expectation is over the new kid's initial productivity and net worth ($x_{k,25},v_{k,25}\sim F(x_{k,53},y_{k,53}$)), instead of the old kid's productivity.
 Third, the law of motion now includes bequests from the old parent instead of a transfer.
-Fourth, it is necessary to map next-period variables for the old child into variables for the new parent.
-The problem, for a child aged 53 who choose to own, becomes:
+Fourth, it is necessary to map next-period variables for the old kid into variables for the new parent.
+The problem, for a kid aged 53 who choose to own, becomes:
 \begin{equation*}
 \begin{split}
-V_c(\mathbf{s}_{c,53})^{own} = &\max_{c_{c,53},b'_{c,53},h'_{c,53}=h_o} u(c_{c,53},h'_{c,53}) + \beta \E\left[V_{p}({\mathbf{s}'_{p,25}}) \right] \\
-\text{s.t.}\quad & 	b'_{c,53} = x_{c,53} + w_{c,53} - c_{c,53} - p h'_{c,53} - adj(h_{c,53},h'_{c,53}) \\
-& x'_{p,25} = b'_{c,53}(1+r(b'_{c,53})) + x'_{p,53} \\
-& b'_{c,53} \ge -LTV p h'_{c,53}, \\
-& h_{p,25} = h'_{c,53},
+V_k(\mathbf{s}_{k,53})^{own} = &\max_{c_{k,53},b'_{k,53},h'_{k,53}=h_o}  u(c_{k,53},h'_{k,53}) + \beta \E\left[V_{p}({\mathbf{s}'_{p,25}}) \right] \\
+\text{s.t.}\quad & 	b'_{k,53} = x_{k,53} + w_{k,53} - c_{k,53} - p h'_{k,53} - adj(h_{k,53},h'_{k,53}) \\
+& x'_{p,25} = b'_{k,53}(1+r(b'_{k,53})) + x'_{p,53} \\
+& b'_{k,53} \ge -LTV p h'_{k,53}, \\
+& h_{p,25} = h'_{k,53},
 \end{split}
 \end{equation*} 
-where $c,53$ and $p,53$ denote the variables associated with old child and old parent in this period, and $c,25$ and $p,25$ denote the variables associated with the new child and new parent in the next period.
-The last equation maps the old child's housing choice into the new parent's---the same household one period later---housing state.
+where $k,53$ and $p,53$ denote the variables associated with old kid and old parent in this period, and $k,25$ and $p,25$ denote the variables associated with the new kid and new parent in the next period.
+The last equation maps the old kid's housing choice into the new parent's---the same household one period later---housing state.
 The same mapping applies to the law of motion (third equation).
@@
 The dying parent's decision problem undergoes similar changes.
-First, the continuation value is now the new parent's value function at age 25 of the new child, weighted by altruism ($\eta V_p(\mathbf{s}'_{p,25})$).
-The remaining changes (expectation, law of motion, and mapping of next-period variables) are the same as for the child and hence the decision problem is changed in the same way as for the child.
+First, the continuation value is now the new parent's value function at age 25 of the new kid, weighted by altruism ($\eta V_p(\mathbf{s}'_{p,25})$).
+The remaining changes (expectation, law of motion, and mapping of next-period variables) are the same as for the kid and hence the decision problem is changed in the same way as for the kid.
 
 \subsection{Measures of Households}\label{sec:hhmeasures}
-The state-space of a parent is $S_p = X_c \times X_p \times Y_c \times H_c \times H_p \times A_c$, with $\mathbf{s}_p$ denoting generic elements therein and $\mc{S}_p$ the associated Borel-$\sigma$ algebra.
-The state space of a child is $S_c= B_p \times H_p \times X_c \times Y_c \times H_c \times A_c$, where $B_p=\mathbb{R}$.
-For conciseness, I omit further definitions for the child.
+The state-space of a parent household is $S_p = X_k \times X_p \times Y_k \times H_k \times H_p  \times A_k$, with $\mathbf{s}_p$ denoting generic elements therein and $\mc{S}_p$ the associated Borel-$\sigma$ algebra.
+The state space of a kid household is $S_k= B_p \times H_p \times X_k \times Y_k \times H_k \times A_k$, where $B_p=\mathbb{R}$.
+For conciseness, I omit further definitions for the kid.
 Let $\psi_p(\mathbf s_p)$ be a probability measure over $(S_p,\mc{S}_p)$ so that $\psi(\mathbf s_p)$ denotes the measure of households with state $\mathbf s_p$ (i.e., after the shock is realized but before choices are made).
@@
 
-\textit{Law of Motion for Dynasties with Children Aged 25-51:} The mass of households in state $\mathbf{s}_p$ is the mass of families adopting housing and savings policies that lead them to this state, adjusted by the probability of experiencing a given income shock:
+\textit{Law of Motion for Dynasties with Kids Aged 25-51:} The mass of households in state $\mathbf{s}_p$ is the mass of families adopting housing and savings policies that lead them to this state, adjusted by the probability of experiencing a given income shock:
 \begin{equation}\label{eq:LoM}
@@
 \mathbf{1}_{\left\{ h'_p = h'^*_p(\mathbf s_p) \right\} } 
-\mathbf{1}_{\left\{ x'_c = x'^*_c(\mathbf s_c(\mathbf s_p))\right\}} 
-\mathbf{1}_{\left\{ h'_c = h'^*_c(\mathbf s_c(\mathbf s_p))\right\}} \times \\
-&\pi(y'_c|y_c)\diff \psi(\mathbf s_p).
+\mathbf{1}_{\left\{ x'_k = x'^*_k(\mathbf s_k(\mathbf s_p))\right\}}  
+\mathbf{1}_{\left\{ h'_k = h'^*_k(\mathbf s_k(\mathbf s_p))\right\}} \times \\
+&\pi(y'_k|y_k)\diff \psi(\mathbf s_p).
 \end{split}
 \end{equation}
-Note that the child's state $\mathbf{s}_c(\mathbf{s}_p)$ in these expressions depend on the choices of the parent which depend on their state. For example, 
+Note that the kid's state $\mathbf{s}_k(\mathbf{s}_p)$ in these expressions depend on the choices of the parent which depend on their state. For example, 
 \begin{equation}
-h'^*_c\left(\mathbf s_c(\mathbf s_p)\right) =h'^*_c\left(b_p'^*(\mathbf{s}_p),h_p'^*(\mathbf{s}_p'),x'_c + t_p'^*(\mathbf{s}_p'),y'_c,h_c',a_c+2 \right).
-\end{equation}
-
-\textit{Law of Motion for Children Aged 53:} In this special case, the distribution will depend on the choices of the new parent (old child), the now deceased parent and the stochastic initial conditions of the new child.
+h'^*_k\left(\mathbf s_k(\mathbf s_p)\right) =h'^*_k\left(b_p'^*(\mathbf{s}_p),h_p'^*(\mathbf{s}_p'),x'_k + t_p'^*(\mathbf{s}_p'),y'_k,h_k',a_k+2 \right).
+\end{equation}
+
+\textit{Law of Motion for Kids Aged 53:} In this special case, the distribution will depend on the choices of the new parent (old kid), the now deceased parent and the stochastic initial conditions of the new kid.
 \begin{equation}\label{eq:LoM55}
 \begin{split}
-\psi(\mathbf{s}'_p;a_c=25) = \int_{\mathbf{s}_p\in \mathcal{S}_p} & 
-\mathbf{1}_{\left\{ x'_p = x^*_p(\mathbf s_p) + x^*_c(\mathbf s_c(\mathbf s_p)) \right\} }
+\psi(\mathbf{s}'_p;a_k=25)  =  \int_{\mathbf{s}_p\in \mathcal{S}_p} & 
+\mathbf{1}_{\left\{ x'_p = x^*_p(\mathbf s_p) + x^*_k(\mathbf s_k(\mathbf s_p)) \right\} }
 \mathbf{1}_{\left\{ h'_p = h^*_p(\mathbf s_p) \right\} } 
-\mathbf{1}_{\left\{ h_c' = h_r\right\}} \times \\
-&F(x_c',y_c'|x_c,y_c) \diff \psi(\mathbf s_p;a_c=53),
+\mathbf{1}_{\left\{ h_k' = h_r\right\}} \times \\
+&F(x_k',y_k'|x_k,y_k)  \diff \psi(\mathbf s_p;a_k=53),
 \end{split}
 \end{equation}
-where we limit $s_p$ to only include the subset of the state-space where $a_c=53$.
-The initial wealth and productivity of the child depends on the wealth $x_c$ and productivity $y_c$ of the new parent at age 53.
-Further, all children start out as renters, and the next-period wealth of the new parent is savings plus bequests.
-
-Finally, the function $\mathcal{H}$ operates on the distribution $\psi(\mathbf{s}_p)$ and policy functions %$g^*(\mathbf{s}_p):\mathbf{S}_p \rightarrow (X\times B \times T \times H)\times (X\times B \times H) \times (A_c)$
+where we limit $s_p$ to only include the subset of the state-space where $a_k=53$.
+The initial wealth and productivity of the child depends on the wealth $x_k$ and productivity $y_k$ of the new parent at age 53.
+Further, all kids start out as renters, and the next-period wealth of the new parent is savings plus bequests.
+
+Finally, the function $\mathcal{H}$ operates on the distribution $\psi(\mathbf{s}_p)$ and policy functions %$g^*(\mathbf{s}_p):\mathbf{S}_p \rightarrow (X\times B \times T \times H)\times (X\times B \times H) \times (A_k)$
 and maps them into a new distribution in accordance with equations (\ref{eq:LoM}, \ref{eq:LoM55}):
@@
 I now briefly discuss some details in the numerical solution of the model.
-Due to the nonconvex nature of the decision problems---due to discrete housing choices, noncontinuous transfer choices, and occasionally binding constraints---I use grid search.
+Due to the non-convex nature of the decision problems---due to discrete housing choices, non-continuous transfer choices, and occasionally binding constraints---I use grid search.
 That is, I define a grid of possible choices for each choice variable, and for each state I loop over all possible choices, and the optimal policy is the one that maximizes the utility.
-The value and policy functions are linearly interpolated over child and parental wealth, conditional on the discrete choices.
+The value and policy functions are linearly interpolated over kid and parent wealth, conditional on the discrete choices.
 This solution algorithm satisfies both criteria established by \cite[p.30]{Barczyk2020}: ``algorithms should be able to deal with locally convex and even discontinuous value functions''.
@@
 \begin{enumerate}
-    \item Next, I solve the decision problem of the child at age 53, that is the final period before he becomes a parent.
-This yields $V_c(\cdot;a_c=53)$ and $g_c(\cdot;a_c=53)$.
-The solution to this problem depends on the value function of a new parent ($V_p(\cdot;a_p=55)$, which is set to 0), the intentional bequest left by the parent ($x_p^*(\cdot;a_p=83)$, aslo set to 0), but not the next-period transfer since then this child has become a parent and thus moves first and decides the transfer.
-The initial state of the new child is given by the joint distribution of initial productivity and wealth.
-    \item Next, I solve the parent's decision problem in the terminal age of 83, when the child is 53.
-This yields $V_p(\cdot;a_c=53)$ and $g_p(\cdot;a_c=53)$, which depend on the policy functions for the child at age 53 ($g_c(\cdot;a_c=53)$)---found in the previous step---and the new parent's value function $V_p(\cdot;a_c=25)$---which we still haven't found and is set to 0.
-    \item Next, I solve the child's decision problem at age 51.
-This yields $V_c(\cdot;a_c=51)$ and $g_c(\cdot;a_c=51)$, and depends on the value function we found in the first step ($V_c(\cdot;a_c=53)$) and the parent's policy function we found in the previous step ($g_p(\cdot;a_c=53)$).
+    \item Next, I solve the decision problem of the kid at age 53, that is the final period before he becomes a parent.
+This yields $V_k(\cdot;a_k=53)$ and $g_k(\cdot;a_k=53)$.
+The solution to this problem depends on the value function of a new parent ($V_p(\cdot;a_p=55)$, which is set to 0), the intentional bequest left by the parent ($x_p^*(\cdot;a_p=83)$, aslo set to 0), but not the next-period transfer since then this kid has become a parent and thus moves first and decides the transfer.
+The initial state of the new kid is given by the joint distribution of initial productivity and wealth.
+    \item Next, I solve the parent's decision problem in the terminal age of 83, when the kid is 53.
+This yields $V_p(\cdot;a_k=53)$ and $g_p(\cdot;a_k=53)$, which depend on the policy functions for the kid at age 53 ($g_k(\cdot;a_k=53)$)---found in the previous step---and the new parent's value function $V_p(\cdot;a_k=25)$---which we still haven't found and is set to 0.
+    \item Next, I solve the kid's decision problem at age 51.
+This yields $V_k(\cdot;a_k=51)$ and $g_k(\cdot;a_k=51)$, and depends on the value function we found in the first step ($V_k(\cdot;a_k=53)$) and the parent's policy function we found in the previous step ($g_p(\cdot;a_k=53)$).
     \item Next, I solve, the parent's decision problem at age 81.
-This yields $V_p(\cdot,a_c=51)$ and $g_p(\cdot,a_c=51)$, and depends on the policy function for the child we found in the previous step ($g_c(\cdot;a_c=51)$) and the value function we found in the second step ($V_p(\cdot;a_p=83)$).
-    \item This is repeated backwards in age until the child is 25.
+This yields $V_p(\cdot,a_k=51)$ and $g_p(\cdot,a_k=51)$, and depends on the policy function for the kid we found in the previous step ($g_k(\cdot;a_k=51)$) and the value function we found in the second step ($V_p(\cdot;a_p=83)$).
+    \item This is repeated backwards in age until the kid is 25.
     \item Repeat these steps until the value functions converge.
@@
 Intuitively, this cyclical behavior can arise from the interaction between the discrete rent-or-own choice and strategic interactions.
-For example, a renting child may decide to buy given the parent's policy functions at a certain state.
+For example, a renting child may decide to buy given the parent’s policy functions at a certain state.
 In the next iteration, the parent observes this and adjusts savings to discourage the child from buying, prompting the child to reoptimize.
@@
 
-\textit{Interpolation Details:} The policy functions for consumption and savings are noncontinuous in both child and parental wealth which makes interpolation unattractive.
+\textit{Interpolation Details:} The policy functions for consumption and savings are non-continuous in both kid and parental wealth which makes interpolation unattractive.
 However, the policy functions, conditional on choosing a specific house quality are smoother, and do not display large jumps in the quantitative model.
@@
 The probability is between 0 and 1 only when the decision changes in one of the four nodes used in the interpolation.
-If the interpolated consumption choice of the child $c_c$ leads to a net bond position $b_c$ that violates the borrowing constraint I set $b_c$ equal to the constraint and reduce $c_c$ to make it hold.
+If the interpolated consumption choice of the kid $c_k$ leads to a net bond position $b_k$ that violates the borrowing constraint I set $b_k$ equal to the constraint and reduce $c_k$ to make it hold.
 
@@
 I simulate N={\parNdyn} dynasties.
-The initial states of each dynasty (initial child and initial old) $(x_p,h_p,x_c,v_c,h_c,a_c=25)$ is drawn from a five-dimensional joint uniform distribution.
+The initial states of each dynasty (initial kid and initial old) $(x_p,h_p,x_k,v_k,h_k,a_k=25)$ is drawn from a five-dimensional joint uniform distribution.
 I then simulate all dynasties for {\parNdyn} generations, since the distribution, as measured by average homeownership, wealth, and, productivity levels stabilizes after four generations.
-I discard observations from the first four children and parents in each dynasty.
+I discard observations from the first four kids and parents in each dynasty.
 
@@
 			Dashed lines are the empirical age-means and solid lines are the fitted third order polynomials used in the model calibration.
-Each line gives the probability of moving to the bottom tertile (black), middle (gray), and top (blue) income tertiles by the income tertile the child is currently in, by age.
+Each line gives the probability of moving to the bottom tertile (black), middle (gray), and top (blue) income tertiles by the income tertile the kid is currently in, by age.
 \end{footnotesize}}
@@
 		\includegraphics[width=1\textwidth]{../tabfig/empirical/inidistr} 
-		{\begin{footnotesize}\textit{Notes:} The vertial lines denote the first, second, and third productivty level for the child.
+		{\begin{footnotesize}\textit{Notes:} The vertial lines denote the first, second, and third productivty level for the kids.
 Within each interval each point denotes a wealth quartile.

### 2024-11-07 docs/paper.tex
Half-done with Jo's edits

@@
 Policies lowering entry barriers to homeownership generally increase the reliance on parental wealth, while increasing liquidity lowers the role of parental wealth.
-Finally, I study how illiquidity affects the commitment problem  between parents and children.
-Wealthy parents' children increase investment in illiquid housing to extract larger transfers and exhibit a preference for illiquidity.
+Finally, I study how illiquidity affects the commitment problem and find that children with wealthy parents use the illiquidity of housing to encourage transfers and exhibit a preference for illiquidity.
 \end{spacing}}
@@
 Housing is the largest asset in U.S.
-households' portfolios and homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}.
-Yet access to homeownership appears to be increasingly reliant on parental transfers.
-Around 30\% of American first-time homebuyers received downpayment assistance from their parents between 2009 and 2016.
+households' portfolios.
+While homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}, households face different barriers to ownership.
+In particular, around 30\% of American first-time homebuyers received downpayment assistance from their parents between 2009 and 2016.
 Parental downpayment assistance averaged \$48,000 among receiving households.\footnote{Data from the Survey of Household Economic Decisionmaking and the Transfer Supplement of the 2013  Panel Study of Income Dynamics (see Section \ref{sec:data}).}  It is important to understand how transfers shape households' ability to access and maintain ownership, given  housing’s importance to household portfolios, economic outcomes \citep{sodini2023identifying}, and business cycles \citep{Kaplan2020,favilukis2017macroeconomic}.
 
-The main objective of this paper is to quantify the importance of parental transfers for young adults' homeownership.
-The primary mechanism by which transfers increase homeownership, and the focus of most empirical studies, is through relaxing borrowing constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}.
-A secondary mechanism is that parental transfers help households maintain homeownership, a mechanism that has received less attention  \citep[see e.g.,][]{bond2021role}.
+This paper quantifies the importance of parental transfers for young adults' homeownership and investigates how policies lowering homeownership barriers interact with parental transfers.
+Previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints.\citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}.
+Additionally.
+parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}.
 However, while these studies show that receiving transfers causally increases homeownership, they do not address the overall importance of parental transfers.
-For example, households might anticipate future transfers and save less.
+For example, households may anticipate future transfers and subsequently save less.
 In this case, even if transfers increase the likelihood of homeownership all else equal, they may not affect aggregate homeownership.
-The secondary objective is to study how policies lowering homeownership barriers interact with parental transfers.
-
-I first present empirical results suggesting that parental wealth not only helps in the transition to ownership but also in maintaining ownership.
+
+I first present empirical evidence indicating that parental wealth helps not only in the transition to ownership, but also in maintaining ownership.
 I show that, controlling for a wide set of household and parental characteristics, households with wealthier parents not only purchase pricier homes but are also less likely to fall behind on mortgage payments after becoming homeowners.
-Finally, in a variation of the event study in \cite{Chetty2007}, I show that while most households downsize their housing during unemployment spells, those with wealthy parents retain their current homes.
-
-To answer the research questions, I combine two distinct models: housing models, which ignore parental transfers, and models of parental transfers, which ignore housing.
+Finally, using a variation of the event study in \cite{Chetty2007}, I show that while most households downsize their housing during unemployment spells, those with wealthy parents retain their current homes.
+
+To quantify the impact of parental transfers on homeownership, I then combine two distinct models: housing models, which ignore parental transfers, and models of parental transfers, which ignore housing.
 I build a life-cycle overlapping generations model with altruistic parents and a rent-or-own decision.
-Parents and children  interact without commitment through both inter-vivos transfers and bequests.
-In models of altruism without housing (e.g., \citealp{Altonji1997a,Barczyk2014}), the transfer motive is to increase the child's consumption when they are borrowing constrained and thus have high marginal utility of wealth.
+Parents and children interact without commitment through both inter-vivos transfers and end-of-life bequests.
+In models of altruism without housing (e.g., \citealp{Altonji1997a,Barczyk2014}), the transfer motive is to increase the child's consumption when they are borrowing constrained, thus having high marginal utility of wealth.
 The inclusion of a frictional housing market generates two new transfer motives.
-First, transfers directly alleviate credit constraints, and so wealthy parents may transfer to allow their children to meet the downpayment requirement.
+First, transfers can directly alleviate credit constraints, enabling wealthier parents to help their children meet downpayment requirements.
 Second, future transfers provide partial insurance against future income shocks, reducing risks associated with the large illiquid investment of homeownership.
@@
 
-Beyond computational challenges, what insights do we gain by allowing parental transfers in all periods?
-After all, previous studies have used alternative assumptions, such as a one-time transfer after which the child is independent at age 24 \citep[e.g.,][]{Lee2019} or warm-glow bequests that are distributed to younger households \citep[e.g.,][]{aliprantis2022dynamics}.
-The key insight of my modeling approach is that transfers are critical not only for entering homeownership---the main focus of empirical literature---but also for maintaining homeownership.
-Specifically, the possibility of future transfers mitigates the downsides of homeownership for liquidity-constrained households.
-This turns out to be as important as relaxing borrowing constraints.
-Moreover, the model replicates the empirical pattern that households with wealthy parents do not downsize during income losses, which would not be possible without allowing transfers in all periods.
+Allowing for parental transfers in all periods highlights that transfers are critical not only for entering homeownership---the main focus of empirical literature---but also for maintaining homeownership.
+Specifically, the prospect of future transfers mitigates the downsides of homeownership for liquidity-constrained households.
+Ameliorating illiquidity turns out to be as important as relaxing borrowing constraints.
+Crucially, the model rationalizes the previously discussed empirical pattern that households with wealthy parents do not downsize during income losses, which would not be possible without allowing transfers in all periods.
 
@@
 I find that parental transfers account for 14 percentage points (29\%) of the young homeownership rate.
-In the benchmark model, the supply of owner-occupied housing is perfectly elastic; however, my results are robust to alternative supply elasticities.
 

### 2024-11-08 docs/paper.tex
Fixed Jo's comments and revised first paragraph

@@
 Housing is the largest asset in U.S.
-households' portfolios.
+households portfolios.
 While homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}, households face different barriers to ownership.
-In particular, around 30\% of American first-time homebuyers received downpayment assistance from their parents between 2009 and 2016.
-Parental downpayment assistance averaged \$48,000 among receiving households.\footnote{Data from the Survey of Household Economic Decisionmaking and the Transfer Supplement of the 2013  Panel Study of Income Dynamics (see Section \ref{sec:data}).}  It is important to understand how transfers shape households' ability to access and maintain ownership, given  housing’s importance to household portfolios, economic outcomes \citep{sodini2023identifying}, and business cycles \citep{Kaplan2020,favilukis2017macroeconomic}.
+Between 2009 and 2016, around 30\% of American first-time homebuyers received parental assistance for their down payment, with such assistance averaging \$48,000 per receiving household.\footnote{All data from the Survey of Household Economic Decisionmaking and the Panel Study of Income Dynamics (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
+Understanding how these transfers shape households' ability to access and maintain ownership, given  housing’s importance to household portfolios, economic outcomes \citep{sodini2023identifying}, and business cycles \citep{Kaplan2020,favilukis2017macroeconomic}.
 
 This paper quantifies the importance of parental transfers for young adults' homeownership and investigates how policies lowering homeownership barriers interact with parental transfers.
-Previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints.\citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}.
-Additionally.
-parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}.
+Previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}.
+Additionally, parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}.
 However, while these studies show that receiving transfers causally increases homeownership, they do not address the overall importance of parental transfers.
@@
 Moreover, the illiquidity of housing serves as a commitment device enabling children to secure future transfers.
-As in all transfer models, the expectation of future transfers reduces children’s savings motives, potentially lowering the homeownership rate.
+As in all transfer models, the expectation of future transfers reduces children’s saving motives, potentially lowering the homeownership rate.
 
@@
 
-I estimate the model by matching data on homeownership, wealth, and transfers from the Panel Survey of Income Dynamics (PSID).
+I estimate the model by matching data on homeownership, wealth, and transfers from the Panel Study of Income Dynamics (PSID).
 To quantify the importance of transfers, I find the counterfactual homeownership rate without parental transfers.
 Without parental transfers, the model simplifies to a standard life-cycle model with housing.
-I find that parental transfers account for 14 percentage points (29\%) of the young homeownership rate.
-
-Next, I evaluate policies to increase homeownership and their effect on the link between parental wealth and children's housing outcomes.
-The results show that many policies mainly benefit households with wealthy parents.
-Relaxing mortgage constraints, for example by lowering minimum downpayments, strengthens the link between parental wealth and children's housing outcomes since households with wealthier parents find homeownership more attractive, all else equal.
-Other policies that lower barriers to entry, such as reducing prices or purchase costs, increase homeownership for all households, though mainly among those with wealthier parents.
+I find that parental transfers account for 14 percentage points (29\%) of the homeownership rate among young households.
+
+I then use the model to evaluate policies aiming to increase homeownership.
+I find that such policies disproportionately benefit households with wealthy parents.
+Relaxing mortgage constraints, for example by lowering minimum downpayments, strengthens the link between parental wealth and children's housing outcomes since more households with wealthy parents are on the margin to become homeowners.
+Non-mortgage policies that lower barriers to entry, such as reducing prices or purchase costs, increase homeownership for all households, though disproportionately among those with wealthier parents.
 However, increasing liquidity by reducing sales costs weakens the link between parental wealth and children’s housing outcomes, as the partial insurance provided by parental wealth becomes less important.
@@
 Next, I study how housing illiquidity affects the strategic behavior arising in the absence of commitment.
-I find that 27\% of young households prefer to face sales costs on housing.
-These households, who have wealthy parents, use sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth''.
+I find that 27\% of young households prefer that housing is illiquid---due to sales costs---rather than liquid.
+These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth''.
 Lower liquid wealth implies higher marginal propensities to consume (MPC), and hence more parental transfers.
-As an illustrative example, imagine that a child buys a house and brings no liquid wealth to the next period.
-If the parent transfer in the next period, the child will either face very low consumption (with a correspondingly high marginal value of wealth) or be forced to liquidate the house and incur sales costs.
+As an illustrative example, consider a child who buys a house and brings no liquid wealth to the next period.
+If the parent does not transfer in the next period, the child will either face very low consumption (with a correspondingly high marginal value of wealth) or be forced to liquidate the house and incur sales costs.
 The parent dislikes both outcomes, and may find it optimal to transfer enough to keep the child in the home.
 This driver of preference for illiquidity differs from those studied in the literature, such as temptation or time inconsistency \citep{attanasio2024temptation,laibson1997golden}.
-Furthermore, this preference for illiquidity provides a novel explanation for the existence of high-MPC households \cite[see e.g.,][]{kaplan2022marginal}.
-
-Finally, I quantify the contribution of parental transfers to the Black-White homeownership gap: in the U.S., young White households are twice as likely to be homeowners as Black households.
+Furthermore, that parental wealth generates preference for illiquidity provides a novel explanation for the existence of high-MPC households \cite[see e.g.,][]{kaplan2022marginal}.
+
+Finally, I quantify the contribution of parental transfers to the Black-White homeownership gap of 50\% among young adults.
 I adapt the methodology of \cite{Ashman2020} and \cite{aliprantis2022dynamics}, who study the racial wealth gap, and recalibrate the income process to capture racial differences.
 I find that parental transfers account for 7 percentage points (14\%) of the Black-White homeownership gap.
+For context, the Black-White homeownership gap has remained roughly constant since the passing of the Fair Housing Act in 1968.
 
@@
 
-By introducing parental transfers to models of housing, my paper contributes to the literature studying the determinants of homeownership over the life cycle.
+My paper first contributes to the literature studying the determinants of homeownership over the life cycle.
 Recent papers focus on marriage and family formation \citep{Fisher2011,Chang2024,Khorunzhina2019}, housing demand in old age \citep{McGee2019,Barczyk2020a}, and changing borrowing constraints \citep{Paz-Pardo2019,Mabille2020}.
 These studies highlight the importance of credit constraints and minimum down payments in constraining demand for owner-occupied housing.
-However, by considering parent-child interactions, I show a dichotomy based on parent wealth in the importance of constraints.
+By considering parent-child interactions, I show how parental wealth affects the relative importance of constraints.
 For households with poorer parents, the sales and purchase costs decrease ownership by making it riskier to own.
-This risk is less important for households with richer parents.
-making mortgage credit constraints more important.
-My results highlight  that policies increasing ownership through subsidies to first-time buyers will lead to increased housing inequality.
+This risk is smaller  for households with richer parents, making mortgage credit constraints relatively more important.
+My results highlight that policies increasing ownership through subsidies to first-time buyers---a common approach in many cities---will lead to increased housing inequality.
 My model-based results also complement some of the empirical research on the effect of parental resources and housing outcomes \citep[see e.g.,][ for recent work]{wold2024housing,daysal2023intergenerational,benetton2022dynastic} by quantifying the contribution of transfers to homeownership.
 
-I contribute to the literature on altruistic households who interact without commitment by studying how illiquid housing affects the commitment problem.
-In these models children undersave to increase parental transfers (e.g., \citealp{Altonji1997a,Boar2018,Barczyk2014,Chu2020}).
-However, adjustment costs on housing impose future expenditure commitments \citep{Chetty2007,Shore2010}.
+My paper also contributes to the literature on altruistic households who interact without commitment by studying how illiquid housing affects the commitment problem.
+In these models children have an incentive to undersave to increase parental transfers (e.g., \citealp{Altonji1997a,Boar2018,Barczyk2014,Chu2020}).
+However, the illiquidity of housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}.
 The illiquidity of housing allows households to commit to having low liquid wealth, which increases future transfers.
@@
 
-Finally, I contribute to the literature on parental transfers and life-cycle outcomes by focusing on how parental transfers to young adult households influence their wealth accumulation through homeownership.
+Finally, the paper contributes to the literature on parental transfers and life-cycle outcomes by focusing on how parental transfers to young adult households influence their wealth accumulation through homeownership.
 My results show that households with wealthier parents are willing to buy sooner, take on higher leverage, and hold less liquid precautionary savings.

### 2024-11-08 docs/paper.tex
Removed a ton of double spaces
A ew minor tweaks

@@
 Housing is the largest asset in U.S.
-household portfolios, and a first-time homebuyers increasingly rely on parental transfers.
+household portfolios, and first-time homebuyers increasingly rely on parental transfers.
 This paper quantifies the contribution of parental transfers to the homeownership rate of young households.
@@
 I find that parental transfers account for 14 percentage points (29\%) of young households' homeownership.
-Transfers from wealthy parents not only help households overcome borrowing constraints but also help them maintain homeownership, reducing the downsides of illiquidity.
+Transfers from wealthy parents not only help households overcome borrowing constraints but also help sustain homeownership, reducing the downsides of illiquidity.
 Policies lowering entry barriers to homeownership generally increase the reliance on parental wealth, while increasing liquidity lowers the role of parental wealth.
-Finally, I study how illiquidity affects the commitment problem and find that children with wealthy parents use the illiquidity of housing to encourage transfers and exhibit a preference for illiquidity.
+Finally, I show that children of wealthy parents strategically use the illiquidity of housing as a commitment device to encourage transfers and exhibit a preference for illiquidity.
 \end{spacing}}
@@
 
-Next, I study how housing illiquidity affects the strategic behavior arising in the absence of commitment.
-I find that 27\% of young households prefer that housing is illiquid---due to sales costs---rather than liquid.
-These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth''.
-Lower liquid wealth implies higher marginal propensities to consume (MPC), and hence more parental transfers.
+I then study how housing illiquidity affects the strategic behavior arising in the absence of commitment.
+I find that 27\% of young households prefer illiquid housing---due to sales costs---to liquid housing.
+These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth.'' Lower liquid wealth implies higher marginal propensities to consume (MPC), and hence more parental transfers.
 As an illustrative example, consider a child who buys a house and brings no liquid wealth to the next period.
@@
 
-My paper also contributes to the literature on altruistic households who interact without commitment by studying how illiquid housing affects the commitment problem.
+My paper also contributes to the literature on altruistic households interacting without commitment by studying how illiquid housing affects the commitment problem.
 In these models children have an incentive to undersave to increase parental transfers (e.g., \citealp{Altonji1997a,Boar2018,Barczyk2014,Chu2020}).
-However, the illiquidity of housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}.
-The illiquidity of housing allows households to commit to having low liquid wealth, which increases future transfers.
+Illiquid housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}, which children of wealthy parents strategically exploit to encourage future parental transfers.
 The paper most closely related to mine is \cite{Barczyk2020a}, which studies the mirror problem: how the homeownership of retired parents affects the economic behavior of the elderly and their adult children.
-In their model, the parent’s home serves as a commitment device, encouraging wealth bequests from the parent and informal care from the adult offspring.
+In their model, the parents' home serves as a commitment device, encouraging wealth bequests from the parent and informal care from the adult offspring.
 

### 2024-11-12 docs/paper.tex
Took on Maggie comments

@@
 E-mail: eirik.e.brandsaas@frb.gov.
-I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts, Corina Boar, Tim Kehoe, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobsen, William Gamber, Claudia Sahm, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, Young Economist Symposium 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop.
+I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts, Corina Boar, Tim Kehoe, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobson, William Gamber, Claudia Sahm, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, Young Economist Symposium 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop.
 \textit{Disclaimer: }The views expressed in this paper are solely the responsibility of the author and should not be interpreted as reflecting the views of the Board of Governors of the Federal Reserve System or of anyone else associated with the Federal Reserve System.}}
@@
 I find that parental transfers account for 14 percentage points (29\%) of young households' homeownership.
-Transfers from wealthy parents not only help households overcome borrowing constraints but also help sustain homeownership, reducing the downsides of illiquidity.
-Policies lowering entry barriers to homeownership generally increase the reliance on parental wealth, while increasing liquidity lowers the role of parental wealth.
-Finally, I show that children of wealthy parents strategically use the illiquidity of housing as a commitment device to encourage transfers and exhibit a preference for illiquidity.
+Transfers from wealthy parents not only help households overcome borrowing constraints but also help sustain homeownership, mitigating the drawbacks of illiquidity.
+Surprisingly, policies lowering entry barriers to homeownership generally increase the reliance on parental wealth, whereas increased liquidity reduces it.
+Finally, I show that children of wealthy parents strategically use the illiquidity of housing as a commitment device to encourage transfers, resulting in a preference for illiquidity.
 \end{spacing}}
@@
 Between 2009 and 2016, around 30\% of American first-time homebuyers received parental assistance for their down payment, with such assistance averaging \$48,000 per receiving household.\footnote{All data from the Survey of Household Economic Decisionmaking and the Panel Study of Income Dynamics (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
-Understanding how these transfers shape households' ability to access and maintain ownership, given housing’s importance to household portfolios, economic outcomes \citep{sodini2023identifying}, and business cycles \citep{Kaplan2020,favilukis2017macroeconomic}.
 
@@
 Additionally, parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}.
-However, while these studies show that receiving transfers causally increases homeownership, they do not address the overall importance of parental transfers.
+However, while these studies show that  transfer receipt increases the probability of renters becoming homeowners, they do not address the overall importance of parental transfers.
 For example, households may anticipate future transfers and subsequently save less.
 In this case, even if transfers increase the likelihood of homeownership all else equal, they may not affect aggregate homeownership.
+Furthermore, this is the first paper quantifying the importance of parental transfers as a determinant of homeownership; the existing literature has primarily focused on family formation and credit constraints \citep[see e.g.,][]{Chang2024,Paz-Pardo2019,Mabille2020}.
 
@@
 Illiquid housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}, which children of wealthy parents strategically exploit to encourage future parental transfers.
-The paper most closely related to mine is \cite{Barczyk2020a}, which studies the mirror problem: how the homeownership of retired parents affects the economic behavior of the elderly and their adult children.
+The paper most closely related to mine is \cite{Barczyk2020a}, who studies how the homeownership of retired parents affects the economic behavior of the elderly and their adult children.
 In their model, the parents' home serves as a commitment device, encouraging wealth bequests from the parent and informal care from the adult offspring.
+My contribution is to study the complementary problem: how parental transfers affect the homeownership decisions of young adults.
+Finally, while \cite{kaplan2012moving} study cohabitation with parents among individuals under the age of 25, my focus is on homeownership decisions for households from age 25 onwards.
 

### 2024-12-13 docs/Homeownership_BankOfMomAndDad.tex
Added link to code

@@
 I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts, Corina Boar, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobson, William Gamber, Claudia Sahm, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, Boston Fed, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, YES 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop.
-\textit{Disclaimer: }The views expressed in this paper are solely the responsibility of the author and should not be interpreted as reflecting the views of the Board of Governors of the Federal Reserve System or of anyone else associated with the Federal Reserve System.}}
+Code and replication instructions are available at the \href{https://github.com/eirikbrandsaas/HomeownershipBankMomDad.jl}{author's  website and Github}.\newline \textit{Disclaimer: }The views expressed in this paper are solely the responsibility of the author and should not be interpreted as reflecting the views of the Board of Governors of the Federal Reserve System or of anyone else associated with the Federal Reserve System.}}
 \author{Eirik Eylands Brandsaas}
@@
 I rely on the \texttt{interpolations.jl} v0.15.1 package for numerical interpolation routines.
-The Stata code used for empirical data and Julia code for the model is available through the author's website and Github.
+The Stata code used for the empirical analysis and the Julia code for the model is available through the \href{https://github.com/eirikbrandsaas/HomeownershipBankMomDad.jl}{author's website and Github}.
 

### 2025-04-15 docs/Homeownership_BankOfMomAndDad.tex
Update data through 2021 (not, code not return, but model implications are tiny!)
Show less subgroups for the summary statistics table

@@
 
-I use data from 1999 to 2017.
+I use data from 1999 to 2021.
 In 1999, the PSID started to collect detailed wealth data every other year.
@@
 \subsubsection{Descriptive Statistics: Who Receives Transfers?}
-I now discuss descriptive statistics from the PSID sample in 2013 for households aged 25 to 44 with an observed parent household.
-Table \ref{tab:descrstats} contains the means of variables, by age, wealth, marriage, homeownership, and transfer receipt.
+\textcolor{red}{!! [[rewrite this]] !!}
+I now discuss descriptive statistics from the PSID sample in 2013 for households aged 25 to 45 with an observed parent household.
+Table \ref{tab:descrstats} contains the means of variables, by age, wealth, homeownership, and transfer receipt.
 
@@
 In panel a), the first two columns compare the whole sample, by transfers receipt in the last calendar year.
-The mean transfer is relatively high at \$3,960, and 22\% of households received transfers in the last year.
+The mean transfer is relatively high at \$3,950, and 22\% of households received transfers in the last year.
 The largest difference is parental wealth: transferring parents are 2.5 times as wealthy as nontransferring parents (\$1,029,000 vs.
@@
 			\singlespacing
-			\caption{Descriptive Statistics (Means), Households Aged 20-44}\label{tab:descrstats}
-			\begin{tabular}{l@{\hspace{.5\tabcolsep}} r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
+			\caption{Descriptive Statistics (Means), Households Aged 25-45}\label{tab:descrstats}
+			\begin{tabular}{l  r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
 				\toprule
-				& \multicolumn{2}{c}{All}  & \multicolumn{2}{c}{Single}  & \multicolumn{2}{c}{Married}  & \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\
-				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}  & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   \\  \midrule
+				& \multicolumn{2}{c}{All}  &  \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\
+				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\  \midrule
 				\input ../tabfig/descr/descr_trsnf_disc_mean.tex
 				\midrule
-				Age                    & \multicolumn{2}{c}{25-28} & \multicolumn{2}{c}{29-32}  & \multicolumn{2}{c}{33-36}  & \multicolumn{2}{c}{37-40}   & \multicolumn{2}{c}{41-44}   \\
-				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}  & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   \\
+				Age                    & \multicolumn{2}{c}{25-31} & \multicolumn{2}{c}{32-38}  & \multicolumn{2}{c}{39-45}   \\
+				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
 				 \midrule
@@
 				\midrule
-				Wealth Quintile                    & \multicolumn{2}{c}{Quintile 1}  & \multicolumn{2}{c}{Quintile 2}  & \multicolumn{2}{c}{Quintile 3}   & \multicolumn{2}{c}{Quintile 4}  & \multicolumn{2}{c}{Quintile 5}   \\
-				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}  & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}   \\
+				Wealth Tertile                    & \multicolumn{2}{c}{Tertile 1}  & \multicolumn{2}{c}{Tertile 2}  & \multicolumn{2}{c}{Tertile 3}    \\
+				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
 		\midrule
@@
 
-By breaking down the sample by marital status, we see that married or cohabiting households have more wealth, higher income, are slightly older, and have three times the ownership rate of singles.
-Married households receive more frequent and slightly larger transfers.
 Next, I break the sample down by ownership.

### 2025-04-15 docs/Homeownership_BankOfMomAndDad.tex
Descr stat drop age 45

@@
 \textcolor{red}{!! [[rewrite this]] !!}
-I now discuss descriptive statistics from the PSID sample in 2013 for households aged 25 to 45 with an observed parent household.
+I now discuss descriptive statistics from the PSID sample in 2013 for households aged 25 to 44 with an observed parent household.
 Table \ref{tab:descrstats} contains the means of variables, by age, wealth, homeownership, and transfer receipt.
@@
 				\toprule
-				& \multicolumn{2}{c}{All}  &  \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\
+				& \multicolumn{2}{c}{All}  &  \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
 				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\  \midrule
@@
 				\midrule
-				Age                    & \multicolumn{2}{c}{25-31} & \multicolumn{2}{c}{32-38}  & \multicolumn{2}{c}{39-45}   \\
+				Age                    & \multicolumn{2}{c}{25-31} & \multicolumn{2}{c}{32-38}  & \multicolumn{2}{c}{39-44}   \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
 				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
@@
 				\midrule
-				Wealth Tertile                    & \multicolumn{2}{c}{Tertile 1}  & \multicolumn{2}{c}{Tertile 2}  & \multicolumn{2}{c}{Tertile 3}    \\
+				Wealth Tertile                    & \multicolumn{2}{c}{Tertile 1}  & \multicolumn{2}{c}{Tertile 2}  & \multicolumn{2}{c}{Tertile 3}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
 				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
@@
 			{\footnotesize
-				\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD.}
+				\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. }
 		\end{threeparttable}
@@
 
-Next, I break the sample into five age groups from 25 to 44.
+Next, I break the sample into three age groups from 25 to 44.
 We see that households' wealth, income, and homeownership rates increase with age.
-Wealthier households are less likely to receive transfers.
 Notably, among 29-to-32-year-olds, homeownership is more common among transfer recipients (40\%) than nontransfer recipients (32\%).

### 2025-04-15 docs/Homeownership_BankOfMomAndDad.tex
Include marriage and non-winsorized data in regressions

@@
 
-I run an ordinary least squares (OLS) regression on house purchase values for first-time owners and test whether lagged parental wealth (logged) is positively associated with larger house purchases $(\beta_1)$, controlling for lagged log income, lagged log net worth, education, age, log household size, regions, time, and race:
+I run an ordinary least squares (OLS) regression on house purchase values for first-time owners and test whether lagged parental wealth (logged) is positively associated with larger house purchases $(\beta_1)$, controlling for lagged log income, lagged log net worth, education, age, log household size, marriage, state, time, and race:
 \begin{equation*}

### 2025-04-17 docs/Homeownership_BankOfMomAndDad.tex
Added preliminary results on effect of transfers

@@
 
+\textcolor{red}{In all other waves, there is a general question of gifts and inhertitances from people outside the hosuehold.
+We do not know the source of these transfers (e.g., they could be from friends or an inheritance from grandparents) and hence these do not know the extent to which these captures parental assistance.
+Yet, they, to some extent, capture the impact of transfer receipt on household decisions, and hence I use these variables to capture the role of transfers on housing outcomes, following \cite{Lee2018}.}
+
 \subsubsection{Descriptive Statistics: Who Receives Transfers?}
@@
 	\begin{adjustbox}{center}
-		\small
 		\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
@@
 			\singlespacing
-			\caption{Descriptive Statistics (Means), Households Aged 25-45}\label{tab:descrstats}
+			\caption{Descriptive Statistics (Means), Households Aged 25-44}\label{tab:descrstats}
 			\begin{tabular}{l  r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
@@
 
-I now utilize microdata from the PSID to examine the correlation between parental financial resources and housing outcomes.
+I now utilize microdata from the PSID to examine the empirical association between transfer receipt and parental financial resources on housing outcomes.
 
@@
 The sample is limited to households aged 25 to 44.
+
+\subsubsection{Transfers and the Rent-to-Own Transition}
+\textcolor{red}{!! [[rewrite this]] !!}
+As discussed in the introduction, several papers look at the role of parental transfers and resources for entering into the housing market \cite[see e.g.,][]{wold2024housing,Lee2019,Blickle2019}.
+The goal of this section is not to `re-invent' the wheel, but to simply show that these relationships look like in my data.
+I run regressions on renting households aged 25-44 who have not been observed as homeowners before.
+In the first set of regressions, the sample are all households aged 25-44 in 2011 who are renting with non-missing parental transfer data from the 2013 Transfer supplement.
+The outcome variable is whether the household is an owner in the next period, and the set of control variables are (log) parental cash-on-hand, (log) household wealth, (log) household income, household marital states, (log) household size, education, age, parental age, and fixed effects for race, year, and state.
+The results of the transfer coefficients are reported in Table \ref{tab:newowners_short} while the full regression output is in Appendix Table \ref{tab:newowners}
+
+Column 1 reports the effect of recieving a parental transfers (above \$500) (in year 2012) on the probability of becoming a homeowner in 2013, which increases by 1.6 pp.
+Next, I also include the inheritance/transfer dummy, (hence includes households who received transfers/inheritances above \$10,000 in 2011, 2012, or 2013).
+Column 2 also includes households who report having recieved transfers or inheritances above \$10,000 while column (3) only denotes those who have received transfers above \$10k (whether transfers from parents or any inheritance).
+Renters who receive transfers are much more likely to transition to homeowners, even after controlling for other observable characteristics.
+If we only include large transfers (above \$10,000) transfer receipt is associated with a 7.7 percentage point increase in the probability of transitioning into homeownership, more than half of the transition rate.
+The coefficents are not statistically significant, perhaps not surprising given the small sample size, the few households who receive transfers and the few households who become homeowners.
+
+To increase statistical power I run the same regression using the same specification, but over the entire sample period (1999-2021) using OLS and household fixed effects (FE).
+Using fixed effects also allows us to control for household fixed effects such as genetics, risk aversion, etc.
+The downside is that we only observe larger transfers (above 10k) and that we do not know the source of the transfer.
+If we focus on the FE estimator, we see that transfer receipt is associated with an 8.3 percentage point increase (statistically signifacant at the 10\% level) in the probability of becoming a homeowner, relative to a benchmark probability of 18.1\%.
+
+It is perhaps not so surprising that the effect of transfer receipt is so big.
+First, beyond the 2013 transfer supplement (which has the smallest effect, 1.6 pp) I only capture transfers above \$10,000.
+Standard models, including the quantiative model in this paper, predicts that renters follow a threshold rule: for any renter that chooses to rent, there exists a wealth level at which they prefer to own (Indeed, the regression results also show that a 10\% increase in net worth increases the probability of becoming a homeowner by 4 percentage points.).
+Hence, large transfers can shift households above the trehsold if they were `close'.
+Second, these transfers are not exogenous: indeed, it is likely that some of these transfers given exactly because the household is buying.
+
+\begin{table}
+	\center
+	\begin{threeparttable}
+		\caption{The Transition to Ownership}
+		\label{tab:newowners_short}
+		\small 
+				\input{../tabfig/regr/newowners_short.tex}
+
+		{\begin{footnotesize}\begin{flushleft}
+		\textit{Notes:} Standard errors in parentheses.
+\textcolor{red}{update footnote} \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
+		\end{flushleft}\end{footnotesize}}		
+	\end{threeparttable}
+\end{table}
+
+To illustrate the size issue and data limitations, and the need for a model, I plot the size of transfers to renters, by whether they become homeowners in $t+2$ in Figure \ref{fig:PSID_transfer_to_renters}.
+The left planel only includes transfers from parents to children (from the transfer supplement) in 2013 while the right panel plots the data from the transfer/inheritance above 10k figure.
+There are several things to note.
+First, we note that, among receivers, households who transition to ownership receive larger transfers on average, in both figures.
+Second, the transfers in the transfer supplement is much smaller than than those in the gift/inheritance question.
+Indeed, the median parental transfer is just about \$1k, the mean about \$3k, and the 95th percentile is at \$10k.
+Third, In all measures, no matter the measure, we see that receiving renters who become owners receive much larger transfers.
+\begin{figure}
+	\caption{Transfer Sizes to Renters by Transition to Ownership}\label{fig:PSID_transfer_to_renters}
+	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_transfers.pdf}% 
+	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_inheritance.pdf}%
+
+	 {\begin{footnotesize} \textit{Notes:} The left panel uses the transfers, gifts, and inheritance variable which is available in all waves of the PSID while the right panel uses the parental transfer data only available in the 2013 transfer supplement.
+See main text for a discussion.
+Vertical dotted and dashed lines denote the median and means, respectively.
+The figures show the kernel density estimates by group, capped at the the 99th percentile.
+\end{footnotesize}}
+\end{figure}
 
@@
 
+\begin{table}
+	\center
+	\begin{threeparttable}
+		\caption{The Transition to Ownership}
+		\label{tab:newowners}
+		\small 
+				\input{../tabfig/regr/newowners.tex}
+
+		{\begin{footnotesize}\begin{flushleft}
+		\textit{Notes:} Standard errors in parentheses.
+\textcolor{red}{update footnote} \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
+		\end{flushleft}\end{footnotesize}}		
+	\end{threeparttable}
+\end{table}
+
 \begin{figure}\caption{Calibrated Income Process}\label{fig:calincome}

### 2025-04-17 docs/Homeownership_BankOfMomAndDad.tex
Update behind regressions

@@
 	\begin{adjustbox}{center}
+		\small
 		\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
@@
 
+\textcolor{red}{Missing regression equation}
 It is perhaps not so surprising that the effect of transfer receipt is so big.
@@
 \begin{table}
-	\center
 	\begin{threeparttable}
@@
 
-\subsubsection{Households with Wealthier Parents Buy More Expensive Houses}
-I first show that households with wealthier parents buy more expensive homes, controlling for household characteristics.
-
-I run an ordinary least squares (OLS) regression on house purchase values for first-time owners and test whether lagged parental wealth (logged) is positively associated with larger house purchases $(\beta_1)$, controlling for lagged log income, lagged log net worth, education, age, log household size, marriage, state, time, and race:
-\begin{equation*}
-\ln{HVal_{i}} = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
-\end{equation*}
-Parents' wealth, as well as household income and net worth, are log-transformed.
-Households are denoted by $i$ and their parents by $p(i)$.
-First-time buyers are households who have never been observed as homeowners before.
-
-Column 1 of Table \ref{tab:hypo} reports the results.
-Households with wealthier parents buy larger houses: A 1\% increase in parental wealth is associated with a 0.049\% increase in the purchase value of the child's house, and the effect is significant at the 5\% level.
-The effect of parental wealth is about the same as the child's own net worth (0.065\%).
-
-\begin{table}
-	\center
-	\begin{threeparttable}
-		\caption{Housing Choices and Parental Wealth}
-		\label{tab:hypo}
-		\small 
-				\input{../tabfig/regr/hypoII.tex}
-
-		{\begin{footnotesize}\begin{flushleft}
-		\textit{Notes:} Standard errors in parentheses.
-`Behind' refers to whether the household is behind on a mortgage.
-Wealth, income, parental wealth, mortgage, family size, and house values are logged.
-All regressions include year and state fixed effects and control for age and age-squared of both the child and parent.
-Specifications 1-3 use ordinary least squares while specifications 4 and 5 use random and fixed effects, respectively.
-\textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
-		\end{flushleft}\end{footnotesize}}		
-	\end{threeparttable}
-\end{table}
-
 \subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
+\textcolor{red}{!! [[rewrite this]] !!}
 To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales.
@@
 The sample size increases in these regressions, since households who bought before 2009 are now included.
+\begin{table}
+	\begin{threeparttable}
+		\caption{Housing Choices and Parental Wealth}
+		\label{tab:hypo}
+		\small 
+				\input{../tabfig/regr/hypoII.tex}
+
+		{\begin{footnotesize}\begin{flushleft}
+		\textit{Notes:} Standard errors in parentheses.
+`Behind' refers to whether the household is behind on a mortgage.
+Wealth, income, parental wealth, mortgage, family size, and house values are logged.
+All regressions include year and state fixed effects and control for age and age-squared of both the child and parent.
+Specifications 1 and 2 use OLS while specifications 3 and 4 use RE and FE, respectively.
+\textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
+		\end{flushleft}\end{footnotesize}}		
+	\end{threeparttable}
+\end{table}
 

### 2025-04-17 docs/Homeownership_BankOfMomAndDad.tex
Updated behind regressions
Adding the prelim maintaing ownership res.

@@
 		\small 
-				\input{../tabfig/regr/hypoII.tex}
+				\input{../tabfig/regr/hypoII_short.tex}
 
@@
 The probability of being behind on mortgages is decreasing in parental wealth, highlighting that parental resources appear to help new homeowners maintain ownership after purchase---in addition to relaxing borrowing constraints.
+
+\subsubsection{Households with Wealthier Parents Are More Likely to Remain Homeowners}
+\textcolor{red}{!! [[rewrite this]] !!}
+Perhaps surprisingly, homeownerhsip is not a `permanent' state, and in fact a growing literature is focusing on the exit from homeownership... (Bond \& Eriksen, ... other papers)\dots
+
+First, I plot the share of homeowners who are homeowners in period $t+\Delta$.
+The results are striking, among first-time owners (i.e., households whose first observation was as renters and then the first time they were observed as an owner), about 15\% are renting two years later and 20\% are renting after 4 years.
+For all owners (i.e., the sample includes all households who are observed as owners, in any period), there is a marked decrease though it is more muted, after four years about 10\% of all owners will be renting.
+
+Which variable predict the transition out of ownership?
+I run similar regresions to the previous sections, for various lead lengths, and plot the coefficient for household net worth and parental wealth.
+While the coefficient intervals are overlapping, we see a strikin result: at every horizon, a 1\% increase in parental wealth is associated with a larger decrease in the probability of transitioning out of owning.
+
+\begin{figure}[tb]
+	\captionsetup[subfigure]{aboveskip=-1pt}
+	\caption{Mainting Homeownership}\label{fig:homeowner_exit}
+	\begin{subfigure}{0.33\textwidth}%
+		\caption{Exit Out of Ownership}\label{fig:owner_hazard}%
+		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_ownerexit.pdf}% 
+	\end{subfigure}%
+	\hfill
+	\begin{subfigure}{0.33\textwidth}%
+		\caption{All Owners}\label{fig:coef_allowner}%
+		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefowner.pdf}% 
+	\end{subfigure}%
+	\hfill
+	\begin{subfigure}{0.33\textwidth}%
+		\caption{First-Time Owners}\label{fig:coef_ftowner}%
+		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefftowner.pdf}% 
+	\end{subfigure}%
+	\hfill
+
+	\caption*{\footnotesize \textit{Notes:} \textcolor{red}{write this footnote.}}
+\end{figure}
 

### 2025-04-21 src/Julia/export_estimation.jl
Added non-targetted distributional moments

@@ -328,6 +328,22 @@ function output_extramoments(Mb::Model,mom_extra::Vector{Symbol},;store=false)
     out_latex = replace(out_latex,"-Inf" => "0.07")
     out_latex = replace(out_latex,"ccc" => "lrr")
 
+    ## Code that prints some distributional moments to show that the model matches these patterns okay
+    wa_label = "Wealth percentiles, Age 35 (10,25,50,75,90)" 
+    wa_dat = string(round.(Int,collect(Mb.moms[Mb.moms.type.=="Data","wealth_age3536_p" .* string.([10, 25, 50, 75, 90]) .* "young"][1,:])))
+    wa_mod = string(round.(Int,collect(Mb.moms[Mb.moms.type.=="Model","wealth_age3536_p" .* string.([10, 25, 50, 75, 90]) .* "young"][1,:])))
+    wa_row  = wa_label * " & " * wa_dat * " & "* wa_mod * "  \\\\ "
+
+    oi_label = "Ownership, by income tertile, Age 35"
+    oi_dat = string(round.(collect(Mb.moms[Mb.moms.type.=="Data","owner_age3536_inc" .* string.([1,2,3]) .* "young"][1,:]);digits=2))
+    oi_mod = string(round.(collect(Mb.moms[Mb.moms.type.=="Model","owner_age3536_inc" .* string.([1,2,3]) .* "young"][1,:]);digits=2))
+    oi_row  = oi_label * " & " * oi_dat * " & "* oi_mod * "  \\\\ "
+
+    out_latex = replace(out_latex,"\\bottomrule" => wa_row * "\n\\bottomrule \\\\")
+    out_latex = replace(out_latex,"\\bottomrule" => oi_row * "\n\\bottomrule \n")
+    out_latex = replace(out_latex,"[" => "")
+    out_latex = replace(out_latex,"]" => "")
+    
     println(out_latex)
     append_info = @sprintf("m%i_d%i",Dates.month(Dates.today()),Dates.day(Dates.today()))
     filename = "tabfig/est/extramoments_"*append_info*".tex"

### 2025-04-21 src/Julia/various_functions.jl

@@ -138,6 +138,19 @@ function find_moments(np::NumPar,mp::ModPar,pan,period::String)
     smom[!,:transfrateyoung_pmiddle] .= mean(dfy.tp[(dfy.middlep .== 1),:] .> 0)
     smom[!,:transfrateyoung_ppoor] .= mean(dfy.tp[(dfy.poorp .== 1),:] .> 0)
 
+    ## Calculate a wealth percentiles and ownership by income groups age 35-36
+    df35 = dfy[dfy.age.==35,:]
+    q = quantile(df35.xk, [0.10, 0.25, 0.5, 0.75, 0.90]) # Wealth percentiles
+    smom[!,:wealth_age3536_p10young] .= q[1]
+    smom[!,:wealth_age3536_p25young] .= q[2]
+    smom[!,:wealth_age3536_p50young] .= q[3]
+    smom[!,:wealth_age3536_p75young] .= q[4]
+    smom[!,:wealth_age3536_p90young] .= q[5]
+    # Ownership by income levels
+    smom[!,:owner_age3536_inc1young] .= mean(df35[df35.ivk.==1,:iok])-1  # subtract 1 since iok =1 if renting and =2 if owning
+    smom[!,:owner_age3536_inc2young] .= mean(df35[df35.ivk.==2,:iok])-1 
+    smom[!,:owner_age3536_inc3young] .= mean(df35[df35.ivk.==3,:iok])-1 
+
     df_owners = df_owners[!,[:iak, :id, :ownk, :first_own, :tp]]
     df_owners = df_lag(df_owners,:tp,:iak,:id,forward=true)
     df_owners = df_lag(df_owners,:tp,:iak,:id,forward=false)

### 2025-04-21 src/Julia/export_estimation.jl
Added more non-targetted moments

@@ -10,7 +10,7 @@ function varnames()
     :wealthatpurchaseyoung => "Wealth at Purchase (25-44)",
     :first_ownyoung => "Age First Own (25-44)",
     :LTVatpurchaseyoung => "LTV at Purchase (25-44)",
-    :tp2wealthpyoung => "Transfers / Parental Wealth (55-74)",
+    :tp2wealthpyoung => "Transfers / Parental Wealth (55-74) (\\%)",
     :transferbuyersyoung => "Transfers Around Purchase (25-44)",
     :hand2mouthyoung  => "Hand to Mouth (25-44)",
     :mortg2incyoung => "Mortgage / Income (25-44)",
@@ -36,6 +36,8 @@ function varnames()
     :owneryoung_ppoor => "Owner (25-44), Bottom 33\\%",
     :transfrateyoung_prich => "Transfers (25-44), Parent Top 50\\%",
     :transfrateyoung_ppoor => "Transfers (25-44), Parent Bot 50\\%",
+    :transfraterenteryoung => "Transfers (25-44), Renters",
+    :transfrateowneryoung => "Transfers (25-44), Owners",
     :smm_obj => "SMM Objective Function",
     )
 end
@@ -314,6 +316,11 @@ function output_extramoments(Mb::Model,mom_extra::Vector{Symbol},;store=false)
     
     dfe = Mb.moms[:,[mom_extra;:type]]
 
+    try
+        dfe.tp2wealthpyoung *= 100
+    catch
+    end
+
     out = DataFrame(Moment = names(dfe[!,mom_extra]),
             Data = Matrix(dfe[dfe.type .== "Data",mom_extra])[1,:],
             Model = Matrix(dfe[dfe.type .== "Model",mom_extra])[1,:]
@@ -339,8 +346,8 @@ function output_extramoments(Mb::Model,mom_extra::Vector{Symbol},;store=false)
     oi_mod = string(round.(collect(Mb.moms[Mb.moms.type.=="Model","owner_age3536_inc" .* string.([1,2,3]) .* "young"][1,:]);digits=2))
     oi_row  = oi_label * " & " * oi_dat * " & "* oi_mod * "  \\\\ "
 
-    out_latex = replace(out_latex,"\\bottomrule" => wa_row * "\n\\bottomrule \\\\")
-    out_latex = replace(out_latex,"\\bottomrule" => oi_row * "\n\\bottomrule \n")
[…]

### 2025-04-21 src/Julia/run_all.jl

@@ -8,7 +8,7 @@ include("../HomeownershipBankMomDad.jl")
 ## Set "parameters"
 pars = [:η, :χ, :ho] # Estimated parameters
 targ_moms =  [:owneryoung, :r2incyoung, :wealthatpurchaseyoung,:transfrateyoung]
-nontarg_moms = [:medwealthyoung, :medwealthold, :medwealthpgradyoung, :first_ownyoung, :ownerall, :mortgyoung,:LTVatpurchaseyoung,:transferbuyersyoung,]
+nontarg_moms = [:medwealthyoung, :medwealthold, :medwealthpgradyoung, :first_ownyoung, :ownerall, :mortgyoung,:LTVatpurchaseyoung, :tp2wealthpyoung, :transferbuyersyoung,:transfraterenteryoung,:transfrateowneryoung]
 nstate = 65 # Grid points in state
 nchoice = 145 # Grid points in choice
 Nest = 1500 # How many iterations in each step of the estimation

### 2025-04-21 src/Julia/various_functions.jl

@@ -138,7 +138,7 @@ function find_moments(np::NumPar,mp::ModPar,pan,period::String)
     smom[!,:transfrateyoung_pmiddle] .= mean(dfy.tp[(dfy.middlep .== 1),:] .> 0)
     smom[!,:transfrateyoung_ppoor] .= mean(dfy.tp[(dfy.poorp .== 1),:] .> 0)
 
-    ## Calculate a wealth percentiles and ownership by income groups age 35-36
+    ## Calculate a wealth percentiles, ownership by income groups age 35-36, and transferrate by ownership
     df35 = dfy[dfy.age.==35,:]
     q = quantile(df35.xk, [0.10, 0.25, 0.5, 0.75, 0.90]) # Wealth percentiles
     smom[!,:wealth_age3536_p10young] .= q[1]
@@ -150,6 +150,9 @@ function find_moments(np::NumPar,mp::ModPar,pan,period::String)
     smom[!,:owner_age3536_inc1young] .= mean(df35[df35.ivk.==1,:iok])-1  # subtract 1 since iok =1 if renting and =2 if owning
     smom[!,:owner_age3536_inc2young] .= mean(df35[df35.ivk.==2,:iok])-1 
     smom[!,:owner_age3536_inc3young] .= mean(df35[df35.ivk.==3,:iok])-1 
+    # Transfer rate by ownership
+    smom[!,:transfraterenteryoung] .= mean(dfy[dfy.iok.==1,:tp] .> 0)
+    smom[!,:transfrateowneryoung] .= mean(dfy[dfy.iok.==2,:tp] .> 0)
 
     df_owners = df_owners[!,[:iak, :id, :ownk, :first_own, :tp]]
     df_owners = df_lag(df_owners,:tp,:iak,:id,forward=true)

### 2025-04-23 docs/Homeownership_BankOfMomAndDad.tex
First pass at new empirical section (too long!)

--- 2025-04-17 docs/Homeownership_BankOfMomAndDad.tex
@@
 Moreover, the illiquidity of housing serves as a commitment device enabling children to secure future transfers.
+These mechanisms differ from those examined by \citet{Barczyk2020a}, who study how elderly parents’ homeownership decisions shape nursing‑home entry, care arrangements, and intergenerational transfers.
 As in all transfer models, the expectation of future transfers reduces children's saving motives, potentially lowering the homeownership rate.
@@
 	 {\begin{footnotesize} \textit{Notes:} The left panel uses data from the SHED with the sample restricted to current owners who report this being their first home.
+The question was only asked in the 2015 and 2016 waves.
 The right panel uses data from the AHS with the sample restricted to all owner-occupied units that did not indicate the sale of a previous home as the main source of funding.
+The question was discontinued after 2019.
 Note that the left figure has year of purchase on the horizontal axis while the right figure has survey year.
@@
 The main observation is the large increase in the role of inter-vivos transfers for homeowners since 2001.
-From 2001 to 2007, only 10\% to 18\% of first-time buyers received transfers, while 20\% to 40\% received transfers after 2009.\footnote{The question was only included in the 2015 and 2016 waves.
-In the 2015 wave, first-time ownership and down-payment source is only reported for those who bought in 2001 to 2015, while in the 2016 wave it is only reported for 2015 to 2016.
-}
-
-\textit{AHS:} I use the AHS, which surveys occupants of owner-occupied units, to obtain a time series on whether the down payment was mostly funded by gifts or inheritances (right panel of Figure \ref{fig:motivation}).
+From 2001 to 2007, only 10\% to 18\% of first-time buyers received transfers, while 20\% to 40\% received transfers after 2009.
+
+\textit{AHS:} I use the AHS, which surveys occupants of housing units, to obtain a time series on whether the down payment was mostly funded by gifts or inheritances (right panel of Figure \ref{fig:motivation}).
 We observe a flat pattern in the 1990s, followed by a sharp decline from 2005 to 2009, coinciding with the prevalence of mortgages featuring very low down payments (e.g., \cite{Corbae2015}), and a stark increase from 2013 to today.
@@
 
-\textcolor{red}{In all other waves, there is a general question of gifts and inhertitances from people outside the hosuehold.
+In all other waves, there is a general question of gifts and inhertitances from people outside the household.
 We do not know the source of these transfers (e.g., they could be from friends or an inheritance from grandparents) and hence these do not know the extent to which these captures parental assistance.
-Yet, they, to some extent, capture the impact of transfer receipt on household decisions, and hence I use these variables to capture the role of transfers on housing outcomes, following \cite{Lee2018}.}
+Yet, they, to some extent, capture the impact of transfer receipt on household decisions, and hence I use these variables in some of the empirical results, following \cite{Lee2018}.
 
@@
 I now utilize microdata from the PSID to examine the empirical association between transfer receipt and parental financial resources on housing outcomes.
-
-I use the sum of parental net worth and income (denoted as wealth) as a statistic to measure parents' financial resources in all specifications for two reasons.
-First, this is consistent with models of altruism, as the one I use later.
-Second, since parents' wealth and income are highly correlated and I have few observations---under 1,000 for most specifications---the standard errors are large when both are included.
-The sample is limited to households aged 25 to 44.
+The sample is limited to households aged 25 to 44 to focus on the housing choices of young adults.
+I use the sum of parental net worth and income (denoted as wealth) to measure parents' financial resources in all specifications for two reasons.
+First, this is consistent with models of altruism, as the one used later in the paper.
+Second, since parents' wealth and income are highly correlated and I have few observations---fewer than 1,000 for many specifications---the standard errors are large when both are included.
 
 \subsubsection{Transfers and the Rent-to-Own Transition}
-\textcolor{red}{!! [[rewrite this]] !!}
-As discussed in the introduction, several papers look at the role of parental transfers and resources for entering into the housing market \cite[see e.g.,][]{wold2024housing,Lee2019,Blickle2019}.
-The goal of this section is not to `re-invent' the wheel, but to simply show that these relationships look like in my data.
-I run regressions on renting households aged 25-44 who have not been observed as homeowners before.
-In the first set of regressions, the sample are all households aged 25-44 in 2011 who are renting with non-missing parental transfer data from the 2013 Transfer supplement.
-The outcome variable is whether the household is an owner in the next period, and the set of control variables are (log) parental cash-on-hand, (log) household wealth, (log) household income, household marital states, (log) household size, education, age, parental age, and fixed effects for race, year, and state.
-The results of the transfer coefficients are reported in Table \ref{tab:newowners_short} while the full regression output is in Appendix Table \ref{tab:newowners}
-
-Column 1 reports the effect of recieving a parental transfers (above \$500) (in year 2012) on the probability of becoming a homeowner in 2013, which increases by 1.6 pp.
-Next, I also include the inheritance/transfer dummy, (hence includes households who received transfers/inheritances above \$10,000 in 2011, 2012, or 2013).
-Column 2 also includes households who report having recieved transfers or inheritances above \$10,000 while column (3) only denotes those who have received transfers above \$10k (whether transfers from parents or any inheritance).
-Renters who receive transfers are much more likely to transition to homeowners, even after controlling for other observable characteristics.
-If we only include large transfers (above \$10,000) transfer receipt is associated with a 7.7 percentage point increase in the probability of transitioning into homeownership, more than half of the transition rate.
-The coefficents are not statistically significant, perhaps not surprising given the small sample size, the few households who receive transfers and the few households who become homeowners.
-
-To increase statistical power I run the same regression using the same specification, but over the entire sample period (1999-2021) using OLS and household fixed effects (FE).
-Using fixed effects also allows us to control for household fixed effects such as genetics, risk aversion, etc.
-The downside is that we only observe larger transfers (above 10k) and that we do not know the source of the transfer.
-If we focus on the FE estimator, we see that transfer receipt is associated with an 8.3 percentage point increase (statistically signifacant at the 10\% level) in the probability of becoming a homeowner, relative to a benchmark probability of 18.1\%.
-
-\textcolor{red}{Missing regression equation}
-It is perhaps not so surprising that the effect of transfer receipt is so big.
-First, beyond the 2013 transfer supplement (which has the smallest effect, 1.6 pp) I only capture transfers above \$10,000.
-Standard models, including the quantiative model in this paper, predicts that renters follow a threshold rule: for any renter that chooses to rent, there exists a wealth level at which they prefer to own (Indeed, the regression results also show that a 10\% increase in net worth increases the probability of becoming a homeowner by 4 percentage points.).
-Hence, large transfers can shift households above the trehsold if they were `close'.
-Second, these transfers are not exogenous: indeed, it is likely that some of these transfers given exactly because the household is buying.
-
-\begin{table}
+
+While a growing literature documents that parental transfers increase the probability that renters become homeowners (e.g.\ \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}), I replicate this result to highlight the importance of transfers for housing outcomes.
+I estimate
+\begin{equation}	\label{eq:transferrent}
+	\Pr\left(Own_{i,t+2}=1 \mid Own_{i,t}=0\right)
+	= \alpha + \beta\,Transfer_{i,t+2}
+	  + \gamma' X_{i,t} + \varepsilon_{i,t},
+\end{equation}
+where \(Own_{i,t}=0\) denotes a renter at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income and log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.
+The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners.
+Table \ref{tab:newowners_short} reports the estimated coefficient on transfer ($\beta$), with the full coefficient table relegated to Appendix Table \ref{tab:newowners}.
+These regressions are similar to those in \cite{Lee2018}, who also use the PSID.
+
+To measure transfer receipt, I compare two definitions.
+Definition 1 (narrow): Parental gifts exceeding \$500 in 2012, as reported in the PSID’s 2013 Transfer Module, limiting the sample to renters in 2011.
+Definition 2 (broad): Any gift or inheritance exceeding \$10,000 reported in all waves of the PSID, or a parental gift exceeding \$10,000 in 2012.
+This expands the sample include all waves of the PSID, and lets us include household fixed effects, though we now only capture large transfers and include some non-parental transfers.
+
+In a univariate regression, receiving a parental transfer is associated with a 2.9 percentage point higher probability of becoming a homeowner in 2013 among renters in 2011.
+With a full set of controls parental transfers are associated with a 1.2 percentage point higher probability of becoming a homeowner.
+This represents a substantial increase relative to the baseline rent-to-own transition rate of about 14\%, though the effect is not statistically significant.
+
+To increase statistical power, I now use the second transfer receipt definition, which also includes non-parental gifts and bequests.
+This leads to a ten-fold increase in sample size, but the transfer receipt rate also decrease to around 3\%, and the median transfer size jumps from \$1,500 to \$24,200.
+In addition, we can now follow renters over time, enabling the inclusion of household fixed effect.
+In a univariate regression, transfers increase the probability of becoming a homeowner by 21.9 percentage points.
+Even with the full set of controls, the effect of transfer receipt on the future homeownership is large and statisticaly significant: 13.5 percentage point in the OLS regression ($p<0.001$) and a 7.9 percentage point in the FE regression  ($p<0.1$).
+These effects are large, similar in magnitude of the baseline rent-to-own transition rate of 18 percent.
+A few caveats are in order: I only capture transfers above \$10,000, and the effect of transfer receipt is likely increasing with transfer size.
+Second, these transfers are not exogenous: indeed, it is likely that some of these transfers are given \emph{because} households are buying a house.
+
+\begin{table}[tb]
+	\centering
 	\begin{threeparttable}
@@
 
+	\end{threeparttable}
 	{\begin{footnotesize}\begin{flushleft}
-		\textit{Notes:} Standard errors in parentheses.
-\textcolor{red}{update footnote} \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
+		\textit{Notes:} Standard errors in parentheses, clusted at the household level.
+\textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
 		\end{flushleft}\end{footnotesize}}		
-	\end{threeparttable}
 \end{table}
 
-To illustrate the size issue and data limitations, and the need for a model, I plot the size of transfers to renters, by whether they become homeowners in $t+2$ in Figure \ref{fig:PSID_transfer_to_renters}.
-The left planel only includes transfers from parents to children (from the transfer supplement) in 2013 while the right panel plots the data from the transfer/inheritance above 10k figure.
-There are several things to note.
-First, we note that, among receivers, households who transition to ownership receive larger transfers on average, in both figures.
-Second, the transfers in the transfer supplement is much smaller than than those in the gift/inheritance question.
-Indeed, the median parental transfer is just about \$1k, the mean about \$3k, and the 95th percentile is at \$10k.
-Third, In all measures, no matter the measure, we see that receiving renters who become owners receive much larger transfers.
+\subsubsection{Transfer Sizes}
+For additional detail on the transfer variables, I plot kernel density estimates in Figure \ref{fig:PSID_transfer_to_renters}, for renters in $t$, separated by whether they own or rent in $t+2$.
+The left panel only includes parental transfers exceeding \$500 from the 2013 transfer supplement.
+The right panel only includes transfer/inheritance exceeding \$10,000 which is included in every wave.
+There are two takeaways.
+First, households who transition to ownership receive larger transfers on average, under both transfer measures.
+Second, the transfers in the transfer supplement are much smaller than than those in the gift/inheritance question.
+Indeed, the median parental transfer is about \$1k, the mean about \$3k, and the 95th percentile is at \$10k, compared to the median gift/inheritance of \$25k, mean of \$60k, and a 95th percentile of about \$130k.
+
+While these are the best data available to follow household over time---in particular their wealth, income, and housing choices---and transfers, the data have several some limitations.
+First, in the transfer supplement we only observe transfers in 2012, while the PSID is collected every other year.
+Hence we undercount both transfer rates and sizes.
+Second, in the gift/inheritance question, we do not know the source of the transfer.
+Third, we do not know the extent to which households report intra-family loans, nor the extent to which these are repaid.
 \begin{figure}
@@
 Vertical dotted and dashed lines denote the median and means, respectively.
-The figures show the kernel density estimates by group, capped at the the 99th percentile.
+All estimates are weighted and omit the top 5\% of transfer sizes.
 \end{footnotesize}}
@@
 
+\subsubsection{Households with Wealthier Parents Are More Likely to Remain Homeowners}
+While the previous sections have focused on the transition into ownership, I now focus on mainting homeownership.
+\textcolor{red}{rewrite this better: That said, \cite{bond2021role} is an exception and shows the positive association between parental wealth and children's housing outcomes in the PSID and the Health and Retirement Study.
+}
+
+First, the left panel of Figure \ref{fig:owner_hazard} plots the share of owners aged 25 to 44 in year $t$ who remain owners in period $t+\Delta$.
+Among first-time owners, only 80\% have retained ownership four years later.
+For non-first-time owners, there is still a marked decrease though it is more muted, after four years about 10\% of all owners will be renting.
+
+To understand which variables predicts retainment of ownership, I estimate
+\begin{equation}	\label{eq:regmaintain}
+	\Pr\left(Own_{i,t+\Delta}=1 \mid Own_{i,t}=1\right)
+	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
+	  + \gamma' X_{i,t} + \varepsilon_{i,t},
+\end{equation}
+
+where $Own_{i,t+\Delta}=1$ denotes an owner at time $t+\Delta$, $\beta$ and $\zeta$ measures the effect of log household net worth and parental wealth, respectively  and $X_{i,t}$ is a vector of control variables.
+The only changes relative to Eq.
+\ref{eq:transferrent} is that the outcome variable conditions on current ownership and measures future ownership, that there are no controls for transfer receipt, and that the household net worth and parental wealth are taken out of the vector of control variables.
+The sample is limited to households aged 25 to 44.
+These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Survey.
+Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plots the two coefficients of interest, over variying lead, times and their 90\% confidence intervals for all only and first-time owners only.
+
+Two results stand out: First, we see that parental wealth is at least as important as household net worth at all horizons, though the difference is never statistically significant.
+Second, we see that the impact of parental wealth is increasing over time.
+For example, we see that a 10\% increase in parental wealth increases the probability of retaining ownership after 8 years by 0.45 percentage points among first-time owners.
+These patterns will be serve as one part of model validation.
+\begin{figure}[tb]
+	\captionsetup[subfigure]{aboveskip=-1pt}
+	\caption{Mainting Homeownership}\label{fig:homeowner_exit}
+	\begin{subfigure}{0.33\textwidth}%
+		\caption{Exit Out of Ownership}\label{fig:owner_hazard}%
+		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_ownerexit.pdf}% 
+	\end{subfigure}%
+	\hfill
+	\begin{subfigure}{0.33\textwidth}%
+		\caption{All Owners}\label{fig:coef_allowner}%
+		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefowner.pdf}% 
+	\end{subfigure}%
+	\hfill
+	\begin{subfigure}{0.33\textwidth}%
+		\caption{First-Time Owners}\label{fig:coef_ftowner}%
+		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_coefftowner.pdf}% 
+	\end{subfigure}%
+	\hfill
+
+	\caption*{\footnotesize \textit{Notes:} \textcolor{red}{write this footnote.}}
+\end{figure}
+
 \subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
-\textcolor{red}{!! [[rewrite this]] !!}
-To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales.
+% To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales.
 I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses.
 The PSID has collected this data since 2009, and 3.9\% of owners aged 25 to 44 are behind, on average, while 9.2\% have been behind at least once.
-
-The sample is limited to the first time a household is observed as owners, and the control variables are the same as in the previous regression:
+One potential roadblock to maintaining ownership are mortgage delinquencies, as failure to meet mortgage payments can lead to distressed and forced sales.
+I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses.
+The PSID has collected this data since 2009, and about 2\% of owners aged 25 to 44 are behind.
+
+To understand which variables predicts whether households are behind on mortgage payments, I estimate
 \begin{equation*}
-Behind_i = \beta_1 \ln(Wealth)_{p(i),t-2} + \beta_2 \ln(Inc_{i,t-2}) + \beta_3 \ln(NetWorth_{i,t-2}) + \mathbf{\gamma}{X_{i,t}} + \varepsilon_{i}.
+\Pr(Behind_{i,t+2}) = \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t}) + \gamma' X_{i,t} + \varepsilon_{i,t},
 \end{equation*}
-In Column 2 of Table \ref{tab:hypo}, the outcome variable is whether households are behind on the mortgage payments the first time they are observed as owners.
-Parental wealth in the period before purchase decreases the probability that a household will ever be behind: A 1\% increase in parental wealth decreases the probability of being behind by 0.016 percentage points and is significant at the 1\% level.
-Parental wealth has a larger effect than the child's net worth.
-In Column 3, the outcome variable is whether first-time homeowners will ever be observed to be behind on their mortgage, and coefficients and significance levels are virtually unchanged.
-The sample size increases in these regressions, since households who bought before 2009 are now included.
+and the control variables are the same as in the previous regression.
+Table \ref{tab:hypo} reports the estimated coefficient on household net worth ($\beta$) and parental wealth ($\zeta$), with the full coefficient table relegated to Appendix Table \ref{tab:hypo_long}.
+
+I first limit the sample to first-time buyers to focus on distress among new owners.
+In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$) WWith the full set of controls, the effect increases to 0.1 percentage points, though it is no longer statistically significant.
+These are economically meaningful since 1.6\% of first-time owners are behind on their mortgage.
+Moreover, parental wealth is more predictive of future delinqucies than household net worth.
+In Columns 3 to 5, I instead focus on all owners (who are currently not behind), increasing the sample sizes and allowing us to include household fixed effects.
+In these specifications, a 10\% increase in parental wealth decreases the probability of being behind two years later by 0.09 percentage points in the univariate OLS regression ($p<0.001$), 0.02 percentage points in the OLS regression (not significant) and  by 0.07 percentage points in the FE regression ($p<0.1$).
+All told, the probability of being behind on mortgages is decreasing in parental wealth, highlighting that parental resources appear to help new homeowners maintain ownership after purchase---in addition to relaxing borrowing constraints.
 \begin{table}
+	\centering
 	\begin{threeparttable}
@@
 		\small 
-				\input{../tabfig/regr/hypoII.tex}
-
+				\input{../tabfig/regr/hypoII_short.tex}
+
+	\end{threeparttable}
 	{\begin{footnotesize}\begin{flushleft}
-		\textit{Notes:} Standard errors in parentheses.
-`Behind' refers to whether the household is behind on a mortgage.
-Wealth, income, parental wealth, mortgage, family size, and house values are logged.
-All regressions include year and state fixed effects and control for age and age-squared of both the child and parent.
-Specifications 1 and 2 use OLS while specifications 3 and 4 use RE and FE, respectively.
+		\textit{Notes:} Standard errors in parentheses, clustered at the household level.
 \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
 		\end{flushleft}\end{footnotesize}}	
-	\end{threeparttable}
 \end{table}
-
-I also report the results from two specifications where I follow households post-purchase, using random effects and household fixed effects regressions (Columns 4 and 5).
-Once we follow households over time, the effect of parental wealth decreases.
-With household fixed effects, the effect is no longer significant.
-However, the fixed effects regressions must be interpreted with caution, since there is little within-parent variation in wealth over time.
-Indeed, the random and fixed effects coefficients are almost the same (-0.008 vs -0.006).
-Overall, the results support the hypothesis that parental wealth decreases the probability of being behind on mortgage payments.
-
-The probability of being behind on mortgages is decreasing in parental wealth, highlighting that parental resources appear to help new homeowners maintain ownership after purchase---in addition to relaxing borrowing constraints.
 
@@
 The outcome of interest is the change in log housing consumption, which is set to zero for households who do not move.
-For households who do move, housing consumption is defined as yearly rent while renting and 5\% of the market value for owner-occupied housing \citep{Davis2008}.
-The sample is limited to household heads who are unemployed only once between ages 25 and 45.
+For households who do move, housing consumption is defined as yearly rent while renting and the imputed rent of owner-occupied housing is set to 5\% of the market value \citep{Davis2008}.
+The sample is limited to household heads who are unemployed only once between ages 25 and 44.
 I divide the sample by whether a household's parents were in the top parental wealth quartile (of parents with children aged 25 to 44) at the time of unemployment and compare the two groups' housing consumption growth rates at unemployment.
@@
 \begin{table}
+	\centering
+	\begin{threeparttable}
+		\caption{Housing Choices and Parental Wealth}
+		\label{tab:hypo_long}
+		\small 
+				\input{../tabfig/regr/hypoII.tex}
+
+	\end{threeparttable}
+	{\begin{footnotesize}\begin{flushleft}
+		\textit{Notes:} Standard errors in parentheses, clustered at the household level.
+`Behind' refers to whether the household is behind on a mortgage.
+Wealth, income, parental wealth, mortgage, family size, and house values are logged.
+All regressions include year and state fixed effects and control for age and age-squared of both the child and parent.
+Specifications 1 and 2 use OLS while specifications 3 and 4 use RE and FE, respectively.
+\textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
+		\end{flushleft}\end{footnotesize}}	
+\end{table}
+
+\begin{table}
 	\center
@@
 		{\begin{footnotesize}\begin{flushleft}
-		\textit{Notes:} Standard errors in parentheses.
-\textcolor{red}{update footnote} \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
+		\textit{Notes:} Standard errors in parentheses, clustered at the household level.
+\textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
 		\end{flushleft}\end{footnotesize}}		

### 2025-04-25 docs/Homeownership_BankOfMomAndDad.tex
Done with ref. #1

@@
 I first present empirical evidence indicating that parental wealth helps not only in the transition to ownership, but also in maintaining ownership.
+\textcolor{red}{bring in new empirical results here.
+Highlight that the results on behind and unemployment are new}.
 I show that, controlling for a wide set of household and parental characteristics, households with wealthier parents are less likely to fall behind on mortgage payments after becoming homeowners, even though they buy more expensive homes.
@@
 			\end{tabular}
-			{\footnotesize
-				\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. }
 		\end{threeparttable}
-
 	\end{adjustbox}
+	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}%
+		\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. 
+	\end{flushleft}\end{footnotesize}}		
 	\end{table}
@@
 	\end{threeparttable}
-	{\begin{footnotesize}\begin{flushleft}
+	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}%
 		\textit{Notes:} Standard errors in parentheses, clusted at the household level.
@@
 Moreover, I replicate this event study using simulated data from the quantitative model and show that the results are similar (see Section \ref{sec:externalval} and Figure \ref{fig:housinggrowthrates_model}).
+In appendix \ref{app:eventstudy_details} I, instead of using the means, use a formal event study, which shows that households with poor parents have a statsticaly significant decrease in housing consumption during unemployment spells.
 \begin{figure}
@@
 About 30\% of young households receive transfers from their parents in a given year, and recipients' parents are wealthier.
-Recipients are also more likely to transition from renting to owning.
-Moreover, households with wealthier parents buy more expensive houses, are less likely to be behind on their mortgages, and are less likely to downsize during unemployment.
+Recipients are also more likely to transition from renting to owning, also when we condition on a rich set of covariates.
+Parental wealth is also positively associated with maintaing homeownership.
+The probability of mainting homeownership is increasing in parental wealth, households with richer parents are less likely to behind on mortgages and less likely to downsize during unemployment sptells.
 These results suggest that homeowners with wealthy parents find illiquidity of homeownership less problematic.
@@
 \subsubsection{Policy Functions and Strategic Behavior}\label{sec:policyfuncs}
-Notably, both the child and the parent strategically internalize these effects.
-For example, a child can use illiquid housing to commit to having low liquid wealth, thereby increasing the marginal utility of wealth and, consequently, the transfer.
+The child and the parent strategically internalize these effects.
+For example, a child can use illiquid housing to commit to having low liquid wealth, increasing the marginal utility of wealth and, consequently, the transfer.
 Conversely, a parent can tie up wealth in illiquid housing, increasing the marginal utility of wealth and, consequently, decreasing transfers.
@@
 On the right, where all households own, the child takes out a larger mortgage, holding wealth fixed.
-Note that once the child becomes sufficiently wealthy, they behave as if there is no altruism.
-Comparing the orange and dark red lines, we find that an increase in parental wealth decreases saving at all wealth levels.
+One the far right, the child is wealthy and behave as if there is no altruism.
+Comparing the orange and red lines, we see that an increase in parental wealth decreases saving at all wealth levels.
 
@@
 	\end{center}
-	{\begin{footnotesize} \textit{Notes:} The dashed vertical lines indicate the threshold where the child switches from renting to owning.
+	{\vspace{-0.2in}\begin{footnotesize}\textit{Notes:} The dashed vertical lines indicate the threshold where the child switches from renting to owning.
 The horizontal dotted lines indicate the PMI and LTV thresholds.
@@
 I discuss the estimation in Section \ref{sec:estimation}.
-The qualitative patterns remain consistent across parameterizations.
 \end{footnotesize}}
@@
 \section{Parameter Selection}\label{sec:estimation}
-To quantify the impact of parental transfers on homeownership and to perform policy evaluations, I estimate the model using data from the PSID and a standard two-step estimation procedure.
+I estimate the model using data from the PSID and a standard two-step estimation procedure.
 First, parameters independent of the model structure are estimated from the data or sourced from the literature.
@@
 
-\begin{table}
+\begin{table}[tb]
 	\center 
@@
 		\input{../tabfig/est/esttable.tex}
-
-		\begin{tablenotes}[para,flushleft]
-			\footnotesize
+	\end{threeparttable}
+	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}
 		\textit{Notes:} Standard errors calculated by estimating the model to 100 bootstrapped empirical samples.
-		\end{tablenotes}
-	\end{threeparttable}
+		\end{flushleft}\end{footnotesize}}
 \end{table}
-
 \subsubsection{Nontargeted Moments}\label{sec:externalval}
@@
 	{\begin{footnotesize} \textit{Notes:} Lines denote means, and shaded areas indicate 95\% confidence intervals.
-The light orange area denotes the empirical estimates (Section \ref{sec:eventstudy}), and the darker grey areas denote the results from the simulated panel.
+The light orange area denotes the empirical estimates (Section \ref{sec:eventstudy}), and the darker grey area denotes the model simulation.
 The sample consists of households aged 25-45 with exactly one unemployment spell and no changes in head or spouse in the four years before and after unemployment.
@@
 
-		\footnotesize
+	\end{threeparttable}
+	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}%
 		\textit{Notes:} I calibrate the housing supply elasticity from equation \eqref{eq:hsupply} as follows.
@@
 The ``Elastic'' column is the same as in the benchmark specification (Table \ref{tab:modelfitnoa}).
-	\end{threeparttable}
+	\end{flushleft}\end{footnotesize}}		
 \end{table}
@@
 I now evaluate how five different policy levers influence the role of parental wealth in housing outcomes.
+For each policy, I change the relevant parameter slightly and solve the model before simulating a new distribution.
+The idea is not to see what the effect of various policy proposals would be in general equilibrium, but how different policy levers interact with parental transfers.
 The results are reported in Table \ref{tab:policy}.
-For each policy, I change the relevant parameter and solve the model before simulating a new distribution.
-
-\textit{Relaxing Mortgaging Regulation:} Down-payment requirements are widely seen as a barrier to homeownership \citep[e.g.,][]{Lee2018,Guiso2002}.
-I tighten the LTV requirement by 3 percentage points to explore how mortgage regulation affects the parental wealth-housing channel.
-The tighter LTV limit decreases homeownership among the young by 5 percentage points (10\%), as households must have larger down payments.
-Homeownership falls more among households with parents in the middle of the wealth distribution, reducing the parental wealth gradient.
-While at first counterintuitive, the borrowing constraint is more binding for these households, all else being equal, since homeownership is more attractive when you have access to some parental wealth and a small transfer pushes you into ownership.
-However, the parents aren't wealthy enough to render the borrowing constraint irrelevant.
-This result suggests that relaxing borrowing constraints increases not only homeownership for all households, but also the role of parental wealth.
-Indeed, \cite{wold2024housing} find that households with poorer parents have lower LTV ratios at origination, suggesting that the LTV constraint is less binding for households with poorer parents.
-
-\textit{Lowering PMI:} Conventional mortgages require that mortgages with down payments below 20\% have PMI \cite{goodman2017sixty}.
-I relax the PMI limit by 7 percentage points to explore how PMI affects the parental wealth-housing channel.
+
+\textit{Relaxing Mortgaging Regulation:} To study how parental transfers interact with downpayment requirements, I tighten the LTV limit by 3 percentage points.
+The tighter limit decreases homeownership among the young by 5 percentage points (10\%).
+Homeownership declines more among households with parents in the middle of the wealth distribution, reducing the parental wealth gradient.
+While initially counterintuitive, these middle-wealth households rely heavily on small transfers to overcome borrowing constraints; poorer households remain constrained, and richer ones are unaffected.
+Thus, looser borrowing constraints not only boost homeownership overall but also amplify the importance of parental wealth.
+Consistent with this mechanism, \citet{wold2024housing} find that owners with poorer parents have lower LTV ratios at origination.
+
+\textit{Lowering PMI:} To study how parental transfers interact with the PMI, which increases the interest rate on high LTV loans, I relax the PMI limit by 7 percentage points.
 The relaxed PMI limit increases homeownership by 2 percentage points (4\%).
@@
 
-\textit{Lowering House Prices:} Although policymakers cannot directly control house prices, they can implement policies to lower them, such as by incentivizing supply.
-For example, both presidential candidates in the 2024 U.S.
-election promised to increase new construction.
-I lower house prices by \$6,000 (7\%) to explore how the price level affects the parental wealth-housing channel.
+\textit{Lowering House Prices:} To study how parental transfers interact with housing affordability, I lower house prices by \$6,000 (7\%).
 As prices decline, homeownership increases for all households, though more so for households with the richest parents, slightly increasing the role of parental wealth.
 
-\textit{Reducing Purchase Costs:} Another way to lower entry barriers is to reduce the costs of buying a house, and policymakers have lowered government fees and closing costs for first-time buyers.
-For example, in 2017, the stamp duty was lowered for first-time buyers in England.
-I decrease the purchase cost ($m_b$) by 2 percentage points (equivalent to a grant of about \$6,250) to understand how purchase costs affect the parental wealth-housing channel.
-While lower purchase costs increase homeownership, the gains are concentrated among households with wealthy parents, amplifying the parental wealth gradient.
-For these households, purchase costs were a key barrier to homeownership, because they are less likely to downsize after income shocks (Figure \ref{fig:housinggrowthrates_model}).
-This suggests that grants or credits for first-time buyers primarily benefit households with the wealthiest parents.
-
-\textit{Reducing Sales Costs:} The earlier empirical results have highlighted how parental wealth may ameliorate housing illiquidity.
-I lower the sales cost $m_s$ from \parms\% to 5.5\% to understand how sales costs affect the parental wealth-housing channel.
+\textit{Reducing Purchase Costs:} To study how closing costs and first-time buyer assistance interact with parental transfers, I lower the purchase cost ($m_b$) by 2 percentage points---equivalent to a grant of about \$6,250.\footnote{Many U.S.
+states offer closing cost assistance to first-time buyers.
+In England, stamp duty was reduced for first-time buyers in 2017.} While homeownership rises, the gains are concentrated among households with wealthy parents, amplifying the parental wealth gradient.
+Lowering purchase costs has little effect for renters with non-wealthy parents, who remain constrained by LTV limits and sales costs.
+In contrast, for household with non-wealthy parents, lower purchase costs reduce the effective price—the main remaining barrier to ownership.
+This suggests that grants or credits for first-time buyers primarily benefit households of the wealthiest.
+
+\textit{Reducing Sales Costs:} 
+The earlier empirical results suggested that parental wealth helps offset the illiquidity of housing.
+To study this interaction, I lower the sales cost $m_s$ from \parms\% to 5.
 Strikingly, the homeownership rate decreases by 2 percentage points (4\%).
-This counterintuitive result is explained by selection into ownership.
-Homeownership among children with the wealthiest and poorest parents increases just a touch.
-In contrast, homeownership declines by 9 percentage points for households with moderately wealthy parents, leading to a net decline in homeownership.
-Homeownership decreases for households with moderately wealthy parents, as it is harder to use housing to extract transfers.
-For households with the poorest parents, who rarely get transfers, and the wealthiest parents, who frequently get transfers, there is no need to use the sales cost to ``extract'' future transfers.
-
-\begin{table} 
+This counterintuitive result reflects selection into ownership.
+homeownership rises slightly among households with the wealthiest and poorest parents but drops sharply---by 9 percentage points---among households with moderately wealthy parents.
+For these households, lower sales costs weaken the ability to use housing as a commitment device to secure future transfers.
+In contrast, children of the poorest parents (who rarely receive transfers) and the wealthiest parents (who expect transfers regardless) are less affected.
+
+\begin{table}[tb]
 	\centering
@@
 
-These experiments highlight two main results.
-First, parental wealth is an important determinant of homeownership, and having richer parents makes households more likely to prefer owning to renting.
-Some common policies intended to increase homeownership rates change the role of parental wealth in housing outcomes.
-Relaxing credit constraints increases not only homeownership, but the role of parental wealth, since housing is more attractive to households with wealthier parents.
-
-However, increasing liquidity, either through reducing purchase costs or sales costs, decreases the role of parental wealth.
-Increasing liquidity can lower homeownership, as children of wealthy parents can no longer use housing as a commitment device to receive larger transfers.
+These experiments highlight two main findings.
+First, common policies intended to raise homeownership rates can either strengthen or weaken the role of parental wealth.
+Relaxing credit constraints increases both homeownership and reliance on parental wealth, as homeownership becomes more attractive to households with wealthier parents.
+Second, and in contrast, increasing housing liquidity---by lowering purchase or sales costs---reduces the importance of parental wealth.
+Greater liquidity weakens the role of housing as a commitment device, and can even lower homeownership among children of wealthy parents.
 
@@
 For each household in the stationary distribution with sales costs, I assess whether they would prefer that neither they nor future generations in their dynasty face sales costs.
-This hypothetical corresponds to each member in the dynasty deciding whether they would prefer that a benevolent agent pays the sales costs for current and future generations.
+This hypothetical corresponds to a benevolent agent paying the sales costs for current and future generations in that dynasty only.
 The results are reported in Table \ref{tab:adjcost}.
@@
 		\footnotesize
+
+	\end{threeparttable}
+	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}%
 		\textit{Notes:} Only includes dynasties where the child household is aged 25-44. Wealth is measured in 1000s of 2016 US dollars.
-	\end{threeparttable}
+	\end{flushleft}\end{footnotesize}}		
 \end{table}
@@
 
-To summarize, this section shows that altruism generates illiquidity preference.
+To summarize, this section shows that altruism not only reduces the downside of illiquidity, but even generates illiquidity preference.
 This is similar to models of behavioral biases, where households use the illiquidity of housing as a commitment device \cite{attanasio2024temptation}.
@@
 
+\subsubsection{MPCs}
+If altruism makes illiquid assets more attractive, a natural question is how altruism affects MPCs, as liquidity constraints are an important driver of MPCs \citep[see e.g.,][]{aguiar2024hand,Kaplan2014,fagereng2021mpc}.
+We begin by computing counterfactual MPCs for households in the stationary distribution, assuming no altruistic transfers.
+The average MPC of young households is 0.35.
+The MPC of liquidity constrained households (0.44) is about three times that of unconstrained households (0.16).
+While these MPCs are lower than empirical estimates---typically ranging from 0.5 to 0.9---they are relatively standard in models not designed to generate high MPCs.
+
+Introducing altruistic transfers generates two offsetting effects.
+First, future transfers increase MPCs by decreasing precautionary savings motives.
+This effect is small, and only increases the average MPC by about 3 percentage points.
+The increase is almost entirely driven by unconstrained households.
+Second, current transfers provide partial insurance and decrease MPCs.
+This effect is large, reducing the MPC by 11 percentage points.
+The decrease is almost entirely driven by liquidity constrained households.
+Finally, altruism---including current and future transers---reduces the MPC from 0.35 to 0.25.
+Altruistic transfers lower the MPC of liquidity constrained households but increases it for unconstrained households---the ratio of MPC for liquidity constrained to unconstrained decreases from 3 to 2.
+
+\begin{table}[tb]
+	\center\singlespacing	
+	\begin{threeparttable}[tb]
+		\caption{Altruistic Transfers Decrease the MPC}
+		\input{../tabfig/postest/MPCs}
+	\end{threeparttable}
+	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}
+		\textit{Notes:} Sample includes all households aged 25-44 in the stationary distribution.
+Following \cite{Kaplan2014}, households are liquidity constrained if they have negative liquid wealth $(b_k)$ exceeding 16.5\% of annual earnings.
+The MPC is calculated from a \$5,000 positive shock.\end{flushleft}\end{footnotesize}}
+\end{table}
+
+\textcolor{red}{rewrite this paragraph}All told, these results have several implications for the MPC literature.
+Most model-implied MPCs are inflated by ignoring informal intra-family transfers.
+Second, one common strategy for raising model-implied MPCs is to introduce some households with low desired wealth holdings, for example through impatience \cite{aguiar2024hand}.
+However, this behavior is observationally similar to that of households receiving altruistic transfers from wealthy parents.
+Another typical method is to introduce an illiquid asset with a large return premium \cite{kaplan2022marginal}.
+However, as my results show, households with wealthier parents are not only more likely to be homeowners, but also have much lower MPCs than expected due to parental transfers.
+
 \subsection{Black-White Homeownership Gap}
-Racial disparities in economic and financial outcomes are receiving increased attention. In the United States, young White households are nearly twice as likely to own homes as young Black households.
+One natural application of this framework is to ask how much of the black-white homeownership gap---young White households are nearly twice as likely to own homes as young Black households---is due to different in parental wealth.
 
@@
 For example, \cite{charles2002transition} find that differences in parental wealth account for 25\% of the Black-White mortgage application gap, and \cite{bond2021role} find that parental wealth accounts for 28\% of the gap in maintaining homeownership.
-I now use my model to bridge the gap between the structural and empirical papers and to quantify the contribution of parental wealth to the Black-White Homeownership gap.
+
 \begin{table}
@@
 I aim to bridge the gap between these structural approaches and empirical results by using my model to quantify the contribution of parental transfers to the Black-White homeownership gap.
-I follow the methodology in \cite{Ashman2020} and \cite{aliprantis2022dynamics} closely.
+I repeat the main experiment, seperately for Black and White households seperately, following the methodology in \cite{Ashman2020} and \cite{aliprantis2022dynamics} closely.
 I shift the income level ($l_a$) according to racial differences: In the matched parent-child household, White households have 8\% higher income and Black households 42\% lower than the average.
@@
 Transfers account for 7 percentage points (14\%) of the Black-White gap.
-Though Black and White households receive transfers at similar rates in the model, transfers are less important for Black households' homeownership, as they receive smaller amounts due to lower parental wealth.
+Though Black and White households receive transfers at similar rates in the model, transfers are less important for Black households' homeownership, as they are further away from the ownership threshold and receive smaller transfers.
 My results show that racial differences in parental income, which affect parental wealth and transfers, are significant contributors to racial gaps in housing outcomes.
@@
 The Stata code used for the empirical analysis and the Julia code for the model is available through the \href{https://github.com/eirikbrandsaas/HomeownershipBankMomDad.jl}{author's website and Github}.
+
+\section{Event Study Details}\label{app:eventstudy_details}
+In the end, the sample consists of 3,553 households with non-missing log growth rates of housing consumption from 948 households with one observed unemployment spells.
+Means and standard errors are constructed using family weights.
+
+The main event study does not control for household characteristics for three reasons.
+First, this replicates the analysis in \cite{Chetty2007}.
+Second, this makes it straightforward to replicate the event study in the model.
+Third, for ease of interpretation.
+I now show that the results are robust to including household controls.
+
+I use the same estimation samples and variable definitions.
+I then run linear regressions where I interact years relative to unemployment with a dummy for having wealthy parents at unemployment.
+The set of controls include dummies for children's wealth quintiles, a full set of age, year, and state dummies, and dummy variables for college, high-school, marriage, race, log income and log family size.
+The standard errors are clustered at the household level.
+The results are plotted in figure \ref{fig:housinggrowthrates_controls}.
+The main result is that households with wealthier parents do not significantly decrease housing consumption at unemployment while those with poor parents, on average, decrease housing consumption by 8 percent ($p<0.05$).
+However, we cannot reject the hypothesis that the effect of unemployment is the same for both groups.
+\begin{figure}
+	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates_controls}
+	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both_controls}%
+	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both_controls}
+
+	{\begin{footnotesize} \textit{Notes:} Solid lines denote means and dashed lines denote the 95\% confidence interval.
+Sample consists of households aged 25-45 with exactly one observed unemployment spell and without changes in head and/or spouse in the four years before and after unemployment.
+The reference (omitted) category is four years before unemployment.
+\end{footnotesize}}
+\end{figure}
 

### 2025-04-25 src/HomeownershipBankMomDad.jl

@@ -33,5 +33,6 @@ include("Julia/quant_endoprice.jl")
 include("Julia/quant_blackwhite.jl")
 include("Julia/quant_preferadjustment.jl")
 include("Julia/quant_robustness.jl")
+include("Julia/quant_MPC.jl")
 
 include("Julia/postest_plots.jl")

### 2025-04-25 src/Julia/run_all.jl

@@ -130,6 +130,9 @@ out = policyeffects(Mb)
 bypwealth_moms = [:owneryoung_prich, :owneryoung_pmiddle,:owneryoung_ppoor]
 output_policy(out,targ_moms,bypwealth_moms;store=store)
 
+## MPC analysis
+create_MPC_table(Mb,Mn,store)
+
 ## Adjustment costs
 Mb,mp,np = findMb2nd(nstate,nchoice)
 Mn = solve_adjustmentcost(Mb)

### 2025-04-28 docs/Homeownership_BankOfMomAndDad.tex
Done with referee #2

@@
 While homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}, households face different barriers to ownership.
-Between 2009 and 2016, around 30\% of American first-time homebuyers received parental assistance for their down payment, with such assistance averaging \$48,000 per receiving household.\footnote{All data from the Survey of Household Economic Decisionmaking and the Panel Study of Income Dynamics (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
+Between 2009 and 2016, around 30\% of American first-time homebuyers received parental assistance for their down payment, with such assistance averaging \$48,000 per receiving household\textcolor{red}{Check numbers}.\footnote{All data from the Survey of Household Economic Decisionmaking and the Panel Study of Income Dynamics (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
 
 This paper quantifies the importance of parental transfers for young adults' homeownership and investigates how policies lowering homeownership barriers interact with parental transfers.
-Previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}.
+A growing literature emphasizes that families matter for economic outcomes \citep{Doepke2016a}.
+In the context of homeownership, previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}.
 Additionally, parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}.
@@
 
-Finally, the paper contributes to the literature on parental transfers and life-cycle outcomes by focusing on how parental transfers to young adult households influence their wealth accumulation through homeownership.
+Finally, the paper contributes to the literature on parental transfers and economic outcomes by focusing on how parental transfers to young adult households influence their wealth accumulation through homeownership.
 My results show that households with wealthier parents are willing to buy sooner, take on higher leverage, and hold less liquid precautionary savings.
@@
 
-\subsection{Parental Transfers for Owners Over Time}\label{sec:overtime}
+\subsection{Parental Transfers for Downpayments Over Time}\label{sec:overtime}
 I use the Survey of Household Economics and Decisionmaking (SHED) and American Housing Survey (AHS) datasets to show the time trends in parental housing transfers.
+A benefit of these surveys is that they ask questions directly relating transfers to downpayments, unlike the PSID.
 
@@
 \textit{SHED:} I use SHED, an annual cross-sectional survey conducted by the Federal Reserve, to observe the share of households who funded the down payment with a loan or gift from family or friends, by year of purchase.
-The results are plotted in the left panel of Figure \ref{fig:motivation}.
+The sample includes 772 households with non-missing down payment information.The results are plotted in the left panel of Figure \ref{fig:motivation}.
 The main observation is the large increase in the role of inter-vivos transfers for homeowners since 2001.
@@
 \textit{AHS:} I use the AHS, which surveys occupants of housing units, to obtain a time series on whether the down payment was mostly funded by gifts or inheritances (right panel of Figure \ref{fig:motivation}).
+The sample has 73,479 observations in 2019.
 We observe a flat pattern in the 1990s, followed by a sharp decline from 2005 to 2009, coinciding with the prevalence of mortgages featuring very low down payments (e.g., \cite{Corbae2015}), and a stark increase from 2013 to today.
@@
 Since this paper focuses on transfers that (a) relate to housing and (b) are quantitatively meaningful, I set transfers below \$500 to \$0.
-About 25\% of transfers are below this threshold, and ignoring them increases the mean transfer from \$2,921 to \$3,960.
+About 25\% of transfers are below this threshold, and ignoring them increases the conditional mean transfer from \$2,921 to \$3,960.
 
@@
 However, individuals have to file if they give transfers above \$18,000 (in 2024) to any one individual within a calendar year.
-Thus, the gift tax is irrelevant for the vast majority of households, as it only applies if lifetime transfers---including bequests---exceed \$13.61 million.
+Thus, the gift tax is irrelevant for the vast majority of households, as it applies only once lifetime transfers---including bequests---exceed \$13.61 million.
 
@@
 
-\textit{Demographics:} Households are economically active from age $a\in \mathcal{A}=\{25,\allowbreak 27,\allowbreak\dots,\allowbreak 83\}$.
+\textit{Demographics:} Households enter the economy at age $a=25$ and die at age 83.
 The life cycle is illustrated in \ref{fig:overview}.
@@
 \begin{equation}
-U_c(c_c,h_c) = u(c_c,h_c) = \frac{\left(c^\xi s(h)^{1-\xi}\right)^{1-\gamma}-1}{1-\gamma},
+U_c(c_c,h_c) = u(c_c,h_c) = \frac{\left(c_c^\xi s(h_c)^{1-\xi}\right)^{1-\gamma}-1}{1-\gamma},
 \end{equation}
@@
 \end{equation}
-where $\eta$ measures the intensity of the parent's altruism towards the child.
+where $\eta$ measures the intensity of the parent's altruism towards the child. As usual with altruism, the parent also derives altruistic utility after death.
 
@@
 \textit{Housing:} Households can obtain housing services by renting or owning.
-They can rent housing of size $h_r$ or own houses of size $h_o$.
+They can rent housing of size $h_r$ or own houses of size $h_o$, with one size available for each tenure choice for tractability.
 The unit price of housing is $p$, and $q$ denotes the rent-to-price ratio.
@@
 r(b) = \begin{cases}
-r & \text{ if } b\ge0, \\
-r +r^m & \text{ if } b<0, \\
-r +r^m +r^{pmi}& \text{ if } b < \overline{PMI}\times p\times h_o.
+r & \text{ if } b \ge 0, \\
+r +r^m & \text{ if } -\overline{PMI}\times p\times h_o < b \le 0, \\
+r +r^m +r^{pmi}& \text{ if } b \le -\overline{PMI}\times p\times h_o.
 \end{cases}
 \end{equation}
+
+\textit{Net Worth:} Let $x$ denote net worth, defined as the sum of the bond position and the value of owner-occupied housing.
 
@@
 
-\textit{Initial Conditions of the New Child:} Households' initial wealth and productivity levels are stochastic yet correlated with the parent's wealth and productivity at age 53: $x_{25},y_{25} \sim F(x_{53},y_{53})$.
-While this paper does not consider parental investment in their child's human capital, such investments are an important factor in intergenerational persistence of economic outcomes \citep{Daruich2018,Lee2019}.
-The function $F$ allows me to match the intergenerational correlation in initial wealth and income levels, so the transfer motives are not over- or understated.
-Finally, all households begin as renters.
+\textit{Initial Conditions of the New Child:} When children enter the economy at age $a_c=25$, their initial net worth $x_{25}$ and productivity $y_{25}$ are stochastic but correlated with the parent's wealth $x_{53}$ and productivity $y_{53}$ according to $x_{25}, y_{25} \sim F(x_{53}, y_{53})$.
+The initial net worth is not funded by parental transfers, though children can receive transfers in the first period.
+While I abstract from parental investment in human capital—an important driver of intergenerational persistence \citep[see e.g.,][]{Daruich2018,Lee2019}—the function $F$ matches observed correlations in wealth and income, ensuring realistic transfer motives.
+All households begin as renters.
 
 \textit{Timing:} The child's productivity is realized first.
-If the dynasty has a newly economically active child, they also realize their initial wealth.
-Next, the decision problem is separated into two stages.
+If the dynasty has a newly economically active child, they also realize their initial wealth and productivity.
+The decision problem then proceeds in two stages.
 First, the parent chooses consumption $c_p$, housing $h_p$, savings position $b_p$, and nonnegative inter-vivos transfers $t_p$.
@@
 
-\textit{State Variables:} A parent $p$'s state variables are child's wealth $x_c\in X=[0,\infty)$, parent's wealth $x_p\in X$, child's housing $h_c\in H = \{h_r,h_o\}$, parent's housing $h_p\in H$, child's productivity $y_c\in Y_a=\{y_1,\dots,y_{N_y}\}$, and age $a_c\in A_c\in\{25,27,\dots,53\}$.
+\textit{State Variables and Setup:} Let $\mathbf{g}_p$ and $\mathbf{g}_c$ denote the vectors of the parent's and child's policy functions, respectively.
+
+A parent $p$'s state variables are child's wealth $x_c\in X=[0,\infty)$, parent's wealth $x_p\in X$, child's housing $h_c\in H = \{h_r,h_o\}$, parent's housing $h_p\in H$, child's productivity $y_c\in Y_a=\{y_1,\dots,y_{N_y}\}$, and age $a_c\in A_c\in\{25,27,\dots,53\}$.
 The parent's state variable vector is $\mathbf{s}_p\equiv\left(x_c,x_p,y_c,h_c,h_p,a_c\right)$.
@@
 \textit{Decision Problems:} 
-I now show the decision problems for a dynasty at age $a_c=25$.
-For simplicity, I assume the child is a renter choosing to become an owner, while the parent remains a renter.
-Appendix \ref{sec:decextra} describes the decision problems in the final period of the current parent's life.
+I now show the decision problems for a dynasty at age $a_c < 53$. For simplicity, I assume the child is a renter choosing to become an owner, while the parent remains a renter.
 
@@
 \begin{split}
- V_c(\mathbf{s}_c)^{own} = \max_{c_c,b'_c,h'_c=h_o} & u(c_c,h'_c) + \beta \E\left[V_{c}(\mathbf{s}_c') \right] \\ 
- \text{s.t.}\quad & 	b'_c = x_c + t_p + w_c - c_c - p h'_c - m_b p h'_c \\
+ V_c(\mathbf{s}_c;\mathbf{g}_p)^{own} = \max_{c_c,b'_c,h'_c=h_o} & u(c_c,h'_c) + \beta \E_{y_c}\left[V_{c}(\mathbf{s}_c';\mathbf{g}_p) \right] \\ 
+ \text{s.t.}\quad & 	b'_c = \tilde x_c + w_c - c_c - p h'_c - m_b p h'_c \\
  & x_c' = b'_c(1+r(b'_c)) + p h'_c(1-\delta) \\
@@
 We see that a transfer $t_p$ can move the child away from the LTV constraint by increasing the bond position $b_c'$, holding consumption and housing fixed.
+The expectation is conditional on the child's current productivity.
+The child's value functions are conditional on the parent's policy functions because the child's choices depend on the parent's choices.
 
@@
 \begin{split}
-V_p(\mathbf{s}_p)^{rent} = \max_{c_p,b_p',t_p} & u(c_p,h'_p) + {\eta} u\left (c_c^*({\mathbf{s}_c}),h_c'^*({\mathbf{s}_c})\right ) + \beta \E \left[V_{p}({\mathbf{s}_p'})\right] \\
+V_p(\mathbf{s}_p;\mathbf{g}_c)^{rent} = \max_{c_p,b_p',t_p} & u(c_p,h'_p) + {\eta} u\left (c_c^*({\mathbf{s}_c}),h_c'^*({\mathbf{s}_c})\right ) + \beta \E_{y_c} \left[V_{p}({\mathbf{s}_p';\mathbf{g}_c})\right] \\
 \text{s.t.}\quad & b'_p = x_p + w_p - c_p - t_p - q p h_p'\\
@@
 The first constraint is the budget constraint, which determines the net bond position as available resources (net worth and income) minus expenditures (consumption, transfers, and rent).
-The law of motion of net worth is simply the bond position for renters.
+For renters, next-period wealth is simply given by the bond position.
 The third and fourth constraints impose that transfers be nonnegative and that renters may not borrow, respectively.
@@
 \end{equation}
+
+\textit{Decision Problems at Generational Transitions:}
+At age $a_c=53$, the final period of each generation, the decision problem changes (see Appendix \ref{sec:decextra}).
+Two key adjustments occur: First, the child's continuation value becomes the parent's value function at age 25 for the next generation, $V_p(\mathbf{s}_{p,25})$.
+Second, since the parent dies, their continuation value is the child's new value as a parent, weighted by altruism, $\eta V_p(\mathbf{s}_{p,25})$.
 
@@
 The remaining three parameters $\theta=\{\eta,\chi,h_o\}$ are estimated internally to minimize the distance between four simulated and empirical moments.
+
+I estimate the altruism parameter $\eta$ because, relative to previous research, the inclusion of housing introduces new transfer motives, requiring re-estimation to match observed intra-family transfers.
+Similarly, I estimate the homeownership utility $\chi$, since transfers generate new ownership motives.
+Finally, I estimate the minimum size of owner-occupied housing, $h_0$, a key parameter in models of homeownership, because housing choices interact with transfer behavior.
 
@@
 First, with a calibrated discount factor of $\beta=0.92$, the model matches the wealth levels of both young and old households.
-Second, the model replicates the correlation between parental wealth and children's homeownership.
-In the PSID, young homeowners' parents are 2.5 times wealthier as young renters' parents.
-The model predicts a similar, though slightly lower, ratio of about 2.
-This parental wealth gradient of homeownership captures the correlation between parental wealth and children's homeownership.
-
-Next, the model aligns with other moments related to homeownership and housing decisions.
-First, it matches the age of first homeownership well, overshooting it slightly.
-Second, it somewhat overpredicts the aggregate homeownership rate of 65\%, primarily due to an excess of homeowner parents, partly due to their lack of income risk (see Appendix \ref{sec:robust_incomerisk}).
-Third, households hold slightly smaller mortgages than in the data, as their flexibility allows households to pay down debt quickly and borrow later if necessary.
-Fourth, the model also matches the LTV ratio for buyers well.\footnote{Note that this is not the LTV at origination but the first time the household is observed as an owner.
-If a household bought just after the previous survey, the LTV would decrease due to price appreciation, improvements, and mortgage paydown.}
-
-Finally, I report the share of households that received a transfer in the two years before, during, and after a house purchase.
-The model matches this number well, indicating that not only are the right households purchasing at the right time, but also that buyers are receiving transfers at the right time.
+It also replicates the correlation between parental wealth and children's homeownership: in the PSID, young homeowners' parents are 2.5 times wealthier than young renters' parents, compared to about 2 times in the model.
+This parental wealth gradient effectively captures the intergenerational link in homeownership.
+
+Second, the model aligns well with other moments related to homeownership.
+It slightly overshoots the age of first homeownership and overpredicts the aggregate homeownership rate of 65\%, primarily due to an excess of homeowner parents stemming from their lack of income risk (see Appendix \ref{sec:robust_incomerisk}).
+Mortgages are slightly smaller than in the data, as households repay and re-borrow flexibly.
+The model also matches the LTV ratio for buyers well.\footnote{The empirical LTV is measured at the first observation of ownership, not necessarily at origination, so it can reflect price appreciation, improvements, and mortgage paydown.}
+
+Third, the model also matches transfer-related moments well.
+It predicts average parental transfers of 2.3\% of parental wealth, close to the 2.6\% estimated from the 2013 Transfer Module (excluding outliers below 0\% or above 100\%).
+It accurately captures the share of households receiving transfers within two years of house purchases, suggesting proper synchronization of housing transactions and transfers.
+Renters also receive transfers more often than owners in both the model and the data, though the model slightly overstates this gap.
+
+Fourth, the model fits key dimensions of heterogeneity.
+It matches the wealth distribution between the 10th and 75th percentiles well, although it slightly overpredicts wealth among the poorest households due to the non-negativity constraint.
+As with many models, it underpredicts wealth at the top.
+Finally, it reproduces the positive correlation between income and homeownership across income tertiles, though the model slightly overstates this gap.
 
@@
 Otherwise, I replicate the exercise exactly.
-Figure \ref{fig:housinggrowthrates_model} shows that the model matches both the qualitative and quantitative patterns from the empirical data: Housing consumption falls for households without wealthy parents during ``unemployment,'' while it grows for those with wealthy parents.
+Figure \ref{fig:housinggrowthrates_model} shows that the model matches both the qualitative and quantitative patterns from the empirical data relatively well: Housing consumption falls for households without wealthy parents during ``unemployment,'' while it does not decline for those with rich parents.
 
@@
 
-\textit{Relaxing Mortgaging Regulation:} To study how parental transfers interact with downpayment requirements, I tighten the LTV limit by 3 percentage points.
+\textit{Relaxing Mortgaging Regulation:} To study how parental transfers interact with down payment requirements, I tighten the LTV limit by 3 percentage points.
 The tighter limit decreases homeownership among the young by 5 percentage points (10\%).
@@
 Without altruism, all households prefer to remove sales costs.
-I find that while all parents prefer to remove sales costs, roughly one-quarter of children prefer them.
+With altruism (baseline calibration), I find that while all parents prefer to remove sales costs, roughly one-quarter of children prefer them.
 The key determinant is parental wealth: Parents of children who prefer sales costs are approximately three times wealthier.
@@
 I aim to bridge the gap between these structural approaches and empirical results by using my model to quantify the contribution of parental transfers to the Black-White homeownership gap.
-I repeat the main experiment, seperately for Black and White households seperately, following the methodology in \cite{Ashman2020} and \cite{aliprantis2022dynamics} closely.
+I repeat the main experiment, seperately for Black and White households, following the methodology in \cite{Ashman2020} and \cite{aliprantis2022dynamics} closely.
 I shift the income level ($l_a$) according to racial differences: In the matched parent-child household, White households have 8\% higher income and Black households 42\% lower than the average.
@@
 \begin{split}
-V_c(\mathbf{s}_{c,53})^{own} = &\max_{c_{c,53},b'_{c,53},h'_{c,53}=h_o} u(c_{c,53},h'_{c,53}) + \beta \E\left[V_{p}({\mathbf{s}'_{p,25}}) \right] \\
-\text{s.t.}\quad & 	b'_{c,53} = x_{c,53} + w_{c,53} - c_{c,53} - p h'_{c,53} - adj(h_{c,53},h'_{c,53}) \\
+V_c(\mathbf{s}_{c,53;\mathbf{g}_p})^{own} = &\max_{c_{c,53},b'_{c,53},h'_{c,53}=h_o} u(c_{c,53},h'_{c,53}) + \beta \E_{y_c}\left[V_{p}({\mathbf{s}'_{p,25}};\mathbf{g_c}) \right] \\
+\text{s.t.}\quad & 	b'_{c,53} = \tilde x_{c,53} - c_{c,53} - p h'_{c,53}  - m_b p h'_{c,53} \\
 & x'_{p,25} = b'_{c,53}(1+r(b'_{c,53})) + x'_{p,53} \\
@@
 
-The dying parent's decision problem undergoes similar changes.
+The dying parent's decision problem undergoes similar changes (Eq.
+\ref{eq:Vp}).
 First, the continuation value is now the new parent's value function at age 25 of the new child, weighted by altruism ($\eta V_p(\mathbf{s}'_{p,25})$).
-The remaining changes (expectation, law of motion, and mapping of next-period variables) are the same as for the child and hence the decision problem is changed in the same way as for the child.
+The remaining changes---namely the expectation, the law of motion, and the mapping of next-period variables---are the same as for the child, and thus the decision problem is modified in the same way.
+
+\begin{equation*}
+	\begin{split}
+	V_p(\mathbf{s}_{c,53};\mathbf{g}_c)^{own} = &\max_{c_{p,53},b'_{p,53},h'_{p,53}=h_r} u(c_{p,53},h'_{p,53}) + \eta u(c_{c,53},h'_{c,53}) + \beta \eta \E_{y_c}\left[V_{p}({\mathbf{s}'_{p,25}},\mathbf{g}_p) \right] \\
+	\text{s.t.}\quad & b'_{p,25} = x_{p,53} + w_{p,53} - c_{p,53} - t_{p,53} - q p h_{p,53}'\\
+	& x_{p,53}' = b'_{p,53}(1+r(b'_{p,53})) \\
+	& t_{p,53}\ge0, \\
+	& b'_{p,53}\ge 0. \\
+	\end{split}
+\end{equation*}
 
@@
 
-\textit{Interpolation Details:} The policy functions for consumption and savings are noncontinuous in both child and parental wealth which makes interpolation unattractive.
-However, the policy functions, conditional on choosing a specific house quality are smoother, and do not display large jumps in the quantitative model.
-Therefore, I use the following procedure.
-I first find the interpolated policies for each housing choice, and then find the probability that the parent/child makes this choice.
-This probability comes from the interpolated discrete choice function, which is binary ${0,1}$, while the interpolated policy itself (with flat extrapolation) is continuous over the interval $[0,1]$.
-The probability is between 0 and 1 only when the decision changes in one of the four nodes used in the interpolation.
-If the interpolated consumption choice of the child $c_c$ leads to a net bond position $b_c$ that violates the borrowing constraint I set $b_c$ equal to the constraint and reduce $c_c$ to make it hold.
+\textit{Interpolation Details:} Because the policy functions for consumption and savings are discontinuous in both child and parental wealth (due to discrete housing choices and altruistic transfers), direct interpolation is unattractive.
+However, conditional on discrete housing choices, the policy functions are smoother.
+I therefore use the following procedure:
+\begin{enumerate}
+	\item Linearly interpolate the parent's housing choice ($h'_p$)---which equals 0 if renting and 1 if owning---over child ($x_c$) and parent wealth ($x_p$), with flat extrapolation.
+The interpolant takes values on $[0,1]$.
+	\begin{enumerate}
+		\item Find the interpolated housing choice. If it equals 0 or 1, proceed.
+		\item If not, it takes a value on $(0,1)$.
+This happens only when the housing choice is not the same at all the four nearest nodes used to calculate the interpolation weights.
+Randomly assign the parent a housing choice.
+	\end{enumerate}
+	\item Linearly interpolate the parent's bond ($b'_p$) and transfer ($t'_p$) policies, conditional on the realized rent/own choice. (e.g., using the problem in eq.~\ref{eq:Vp} if renting.)
+	\item Recover parental consumption from the budget constraint.
+If the bond choice violates the borrowing limit, set the bond choice at the constraint and adjust consumption accordingly.
+This gives us the four parental choices $(h'_p,t_p,b'_p,c_p)$.
+
+	\item Linearly interpolate the child's housing choice ($h'_k$) over child cash-on-hand ($\tilde x_c$) and parent savings ($b'_p$), conditional on the parent's housing choice from the first step.
+Resolve the housing choice as for the parent.
+	\item Linearly interpolate the child's bond choice ($b'_k$) over child cash-on-hand and parent savings, conditional on both the child's and parent's housing choices.
+	\item Recover child consumption from the budget constraint, and ensure that the borrowing constraint is satisfied, as for the parent. This gives us the three child choices $(h'_k,b'_k,c_k)$. 
+\end{enumerate}
 

### 2025-04-28 src/Julia/policyplots.jl

@@ -17,7 +17,7 @@ function plot_policyfuncs(Mb,Mn;store=false)
     ihk = 2
     _h2size = Dict(1=>L"h_r",2=>L"h_o")
 
-    plot(title="Child's Bond and Housing Choice $(L"b{\prime}_c,h{\prime}_c")",xlabel="Child Wealth $(L"x_c")")
+    plot(title="Child's Bond and Housing Choice $(L"b{\prime}_c,h{\prime}_c")",xlabel="Child Wealth $(L"x_c")",ylabel="Child Bond Choice $(L"b'_c")")
 
     iown = findfirst(x->x==true,Mn.g.gk.disc[:,ibpl,ihp,ivk,iok,is,iak,ihk].==1)
     plot!(np.xk_grd[1:iown],Mn.g.gk.b[1:iown,ibpl,ihp,ivk,iok,is,iak,1] ,label="No Altruism",marker=:circle,color=:black,linewidth=2)

### 2025-04-29 docs/Homeownership_BankOfMomAndDad.tex
Done with most of ref #3 comments

@@
 Housing is the largest asset in U.S.
-households portfolios.
+household portfolios.
 While homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}, households face different barriers to ownership.
-Between 2009 and 2016, around 30\% of American first-time homebuyers received parental assistance for their down payment, with such assistance averaging \$48,000 per receiving household\textcolor{red}{Check numbers}.\footnote{All data from the Survey of Household Economic Decisionmaking and the Panel Study of Income Dynamics (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
+Between 2009 and 2016, around 30\% of American first-time homebuyers received direct parental assistance for their down payment, with such assistance averaging \$48,000 per receiving household\textcolor{red}{Check numbers}.\footnote{All data from the Survey of Household Economic Decisionmaking and the Panel Study of Income Dynamics (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
 
@@
 	 {\begin{footnotesize} \textit{Notes:} The left panel uses data from the SHED with the sample restricted to current owners who report this being their first home.
-The question was only asked in the 2015 and 2016 waves.
 The right panel uses data from the AHS with the sample restricted to all owner-occupied units that did not indicate the sale of a previous home as the main source of funding.
-The question was discontinued after 2019.
 Note that the left figure has year of purchase on the horizontal axis while the right figure has survey year.
@@
 
-\textit{SHED:} I use SHED, an annual cross-sectional survey conducted by the Federal Reserve, to observe the share of households who funded the down payment with a loan or gift from family or friends, by year of purchase.
-The sample includes 772 households with non-missing down payment information.The results are plotted in the left panel of Figure \ref{fig:motivation}.
+\textit{SHED:} I use the 2015 and 2016 SHED, an annual cross-sectional survey conducted by the Federal Reserve, to observe the share of first-time owners who funded the down payment with a loan or gift from family or friends, by year of purchase.
+The sample includes 772 households with non-missing down payment and first-time ownership information.
+The results are plotted in the left panel of Figure \ref{fig:motivation}.
 The main observation is the large increase in the role of inter-vivos transfers for homeowners since 2001.
@@
 
-\textit{AHS:} I use the AHS, which surveys occupants of housing units, to obtain a time series on whether the down payment was mostly funded by gifts or inheritances (right panel of Figure \ref{fig:motivation}).
-The sample has 73,479 observations in 2019.
-We observe a flat pattern in the 1990s, followed by a sharp decline from 2005 to 2009, coinciding with the prevalence of mortgages featuring very low down payments (e.g., \cite{Corbae2015}), and a stark increase from 2013 to today.
-Minor fluctuations in the percentage of homeowners relying on gifts or inheritances reflect significant shifts among new homeowners, who comprise a small fraction of all homeowners.
-This is because the horizontal axis denotes the survey year, not the year of purchase.
+\textit{AHS:} I use the 1991-2019 AHS, which surveys occupants of housing units, to construct a time series of whether homeowners primarily funded their down payment with gifts or inheritances (right panel of Figure~\ref{fig:motivation}).
+After excluding households who report sale of previous homes as a source, the sample includes about 50,000 observations each year.
+The series is relatively flat through the 1990s, followed by a sharp decline from 2005 to 2009---coinciding with the rise of low–down-payment mortgages---and a marked increase from 2013 onward.
+Because the horizontal axis reflects survey year rather than purchase year, minor fluctuations largely reflect variation among recent buyers, who make up a small share of all homeowners.
 
@@
 I classify top-coded values as missing observations.
-All monetary variables are expressed in constant 2016 U.S.
-dollars (in thousands).
+All monetary variables are expressed in thousands of 2016 U.S.
+dollars.
 Table \ref{tab:vardef} in the Appendix provides variable code of the main variables.
 
-\textit{Sample Selection:} Throughout this paper, the sample includes all households aged 25 to 84 in the PSID, excluding the Survey of Economic Opportunity and the Latino subsamples, to obtain a representative sample.
+\textit{Sample Selection:} Throughout this paper, the sample includes all households aged 25 to 84 in the PSID.
 All summary statistics are calculated using the provided family weights.
-I drop all observations with missing housing values (which is set to zero for renters) or wealth.
+I drop 1851 observations with missing housing values (which is set to zero for renters).
+The final sample has 18,075 households and 100,326 observations.
 
@@
 
-\textit{Definition of Transfers:} The transfer supplement asked all parent households whether they gave money, gifts, or loans of \$100 or more to their children in 2012.
+\textit{Definition of Transfers:} To measure the transfer receipt rate, I rely on the transfer supplement which asked all parent households whether they gave money, gifts, or loans of \$100 or more to each children in 2012.
 I follow the literature \citep[e.g.,][]{mcgarry2016dynamic} and treat all transfers as gifts.
 Since this paper focuses on transfers that (a) relate to housing and (b) are quantitatively meaningful, I set transfers below \$500 to \$0.
-About 25\% of transfers are below this threshold, and ignoring them increases the conditional mean transfer from \$2,921 to \$3,960.
+About 25\% of transfers are below this threshold, and ignoring them increases the conditional mean transfer from \$2,921 to \$3,960 and reduce the transfer receipt rate from 45\% to 22\%.\footnote{These transfer rates are relatively high and higher than those calculated by \citealt{feiveson2019lifecycle} using the Surey of Consumer Finances (SCF).
+First, it seems likely that the PSID transfer supplement was more successfull in counting transfers than the SCF.
+Second, the SCF asks about transfers and gifts going further back in time, likely increasing underreporting.
+\textcolor{red}{write more here} }
 
@@
 Yet, they, to some extent, capture the impact of transfer receipt on household decisions, and hence I use these variables in some of the empirical results, following \cite{Lee2018}.
+
+Finally, why focus on inter-vivos transfers when most intra-family wealth transfers occur through bequests \citep[see e.g.,][]{feiveson2019lifecycle}?
+First, parental bequests are typically received after age 45, too late to affect young adults' transition into homeownership.
+Second, some bequests are given early and recorded as inter-vivos transfers.
+Third, inter-vivos transfers are often targeted to children's immediate liquidity needs, such as financing a home purchase, whereas bequests are often passive outcomes of unspent wealth.
+Thus, inter-vivos transfers are more relevant for understanding early-life economic outcomes.
 
@@
 Table \ref{tab:descrstats} contains the means of variables, by age, wealth, homeownership, and transfer receipt.
+Though limited to households with observed parents, the sample appears reasonably representative.
+Average income by age closely matches estimates from the Survey of Consumer Finances \cite[see e.g.,][]{kuhn20162013}, though average wealth is slightly lower.
 
@@
 
-Consider a a wealthy parent with a poor, borrowing-constrained child.
+Consider a wealthy parent with a poor, borrowing-constrained child.
 This implies that the child does not save and must rent and, thus, consumes any additional wealth.
@@
 
-First, altruism lowers the threshold for homeownership among households with some parental support: Wealthier parents make homeownership more attractive.
+First, altruism lowers the wealth threshold for homeownership among households with some parental support: Wealthier parents make homeownership more attractive.
 An increase in parental wealth shifts the threshold further down.
@@
 
-\begin{figure}\caption{Policy Functions for Homeownership and Savings}\label{fig:pol}
+\begin{figure}
 	\begin{center} 
+	\caption{Child's Bond and Homeownership Policy Functions}\label{fig:pol}
 	\includegraphics[width=0.6\textwidth]{../tabfig/model/pol_bond.pdf}
 	\end{center}
-	{\vspace{-0.2in}\begin{footnotesize}\textit{Notes:} The dashed vertical lines indicate the threshold where the child switches from renting to owning.
-The horizontal dotted lines indicate the PMI and LTV thresholds.
+	{\vspace{-0.2in}\begin{footnotesize}\textit{Notes:} The solid lines plots the child’s bond choice ($b'_c$) and the dashed lines the housing choice ($h'_c$) as a function of child cash-on-hand $\tilde x_c$.
+The dashed vertical lines mark the wealth thresholds at which the child switches from renting to owning.
 The child states $\mathbf{s}_c'$ are \input{../tabfig/model/pol_bond_state.tex}\unskip.
-The figure plots the policy functions in the estimated model.
-I discuss the estimation in Section \ref{sec:estimation}.
 \end{footnotesize}}
@@
 The PDF $F(\cdot)$ is then given by the empirical probability of each parent-child combination.
-The wealth level of a quartile is defined by its median value.
+The wealth level of a quartile is defined by its median value (Fig.
+\ref{fig:inidistr})
 
@@
 \begin{figure}\caption{Calibrated Income Process}\label{fig:calincome}
-	\begin{subfigure}{0.49\textwidth}\caption{Deterministic Income Profile $y_a$}\label{fig:y_a}
+	\begin{subfigure}{0.49\textwidth}\caption{Deterministic Income Profile $l_a$}\label{fig:y_a}
 		\includegraphics[width=\textwidth]{../tabfig/empirical/lifecycleincome.pdf}%
@@
 
-	\begin{subfigure}{\textwidth}\caption{Age-State Dependent Transition Probabilities $\Pi(y_{i,a+2}|y_{i,a})$}\label{fig:Piv}
+	\begin{subfigure}{\textwidth}\caption{Age and State Dependent Productivity Transition Probabilities $\Pi_a(y_{i,a+2}|y_{i,a})$}\label{fig:Piv}
 		\includegraphics[width=0.99\textwidth]{../tabfig/empirical/Pi_vk_3.pdf}
@@
 		{\begin{footnotesize}\textit{Notes:}
-			Dashed lines are the empirical age-means and solid lines are the fitted third order polynomials used in the model calibration.
-Each line gives the probability of moving to the bottom tertile (black), middle (gray), and top (blue) income tertiles by the income tertile the child is currently in, by age.
-\end{footnotesize}}
+			Dashed lines are the empirical means and solid lines are the fitted polynomials.
+The top-left panel plots the income level, the top-right panel plots the productivity levels, and the bottom panel plots the transition probabilities (Eq.
+\ref{eq:wk}).
+Each line in the bottom panel gives the probability of moving to the bottom tertile (black), middle (gray), and top (blue) income tertiles by the income tertile the child is currently in, by age.
+\end{footnotesize}\vspace{0.05in}}
 	\end{subfigure}
-
-	\begin{subfigure}{\textwidth}
-		\caption{Initial Distribution $F(x_{53},y_{53})$ by wealth $x_{53}$ and productivity $y_{53}$}\label{fig:inidistr}
+\end{figure}
+
+\begin{figure}\caption{Initial Distribution $F(x_{53},y_{53})$ by wealth $x_{53}$ and productivity $y_{53}$}\label{fig:inidistr}
+
 		\includegraphics[width=1\textwidth]{../tabfig/empirical/inidistr} 
-		{\begin{footnotesize}\textit{Notes:} The vertial lines denote the first, second, and third productivty level for the child.
+		{\begin{footnotesize}\textit{Notes:} The vertial lines denote the first, second, and third productivity level for the child.
 Within each interval each point denotes a wealth quartile.
-For example, the first point on the black line in the left-most panel gives the probability that a new parent who are in the first wealth quartile (black line), in the first income tertile (left panel), has a child in the lowest income tertile and lowest wealth tertile.
+For example, the first point on the black line in the left-most panel gives the probability that a new parent in the first wealth quartile (black lines) and the first income tertile (left panel), has a child in the lowest income tertile and lowest wealth tertile.
 The second point gives the probability that the child is in the second wealth quartile but first income tertile.
 \end{footnotesize}}
-		\end{subfigure}
 \end{figure}

### 2025-04-29 src/Julia/policyplots.jl

@@ -17,7 +17,7 @@ function plot_policyfuncs(Mb,Mn;store=false)
     ihk = 2
     _h2size = Dict(1=>L"h_r",2=>L"h_o")
 
-    plot(title="Child's Bond and Housing Choice $(L"b{\prime}_c,h{\prime}_c")",xlabel="Child Wealth $(L"x_c")",ylabel="Child Bond Choice $(L"b'_c")")
+    plot(xlabel="Child Cash-on-Hand $(L"\tilde x_c")",ylabel="Child Bond Choice $(L"b'_c")")
 
     iown = findfirst(x->x==true,Mn.g.gk.disc[:,ibpl,ihp,ivk,iok,is,iak,ihk].==1)
     plot!(np.xk_grd[1:iown],Mn.g.gk.b[1:iown,ibpl,ihp,ivk,iok,is,iak,1] ,label="No Altruism",marker=:circle,color=:black,linewidth=2)
@@ -40,7 +40,7 @@ function plot_policyfuncs(Mb,Mn;store=false)
     if store == true
         Plots.pdf(p2,"tabfig/model/pol_bond")
         open("tabfig/model/pol_bond_state.tex", "w") do file
-            write(file, "$(L"h'_p")=$(_h2size[ihp]), $(L"y_c")=$(round(np.v_grd[iak,ivk];digits=1)), $(L"h_c")=$(_h2size[ihk]), and $(L"a_c")=$(round(Int,np.age_grd[iak])). The poor and rich parent have $(L"b_p")=$(round(Int,bpl)) and $(L"b_p")=$(round(Int,bph))" )
+            write(file, "$(L"h'_p")=$(_h2size[ihp]), $(L"y_c")=$(round(np.v_grd[iak,ivk];digits=1)), $(L"h_c")=$(_h2size[ihk]), and $(L"a_c")=$(round(Int,np.age_grd[iak])). The non-rich and rich parent have $(L"b'_p")=$(round(Int,bpl)) and $(L"b'_p")=$(round(Int,bph))" )
         end
 
     

### 2025-04-30 docs/Homeownership_BankOfMomAndDad.tex
More text changes

@@
 Third, inter-vivos transfers are often targeted to children's immediate liquidity needs, such as financing a home purchase, whereas bequests are often passive outcomes of unspent wealth.
+Fourth, while bequests dominate in aggregate terms, they are highly concentrated among wealthy households, whereas inter-vivos transfers are the main form of transfer for most parents \citep{yang2023financial}.
 Thus, inter-vivos transfers are more relevant for understanding early-life economic outcomes.
@@
 \end{align}
-I assume that parents face no income risk for simplicity, consistent with the decrease in income risk with age \citep{Sanchez2020}.
+I assume that parents face no income risk for simplicity, consistent with the decrease in income risk with age \citep{Sanchez2020} and the fact that much of retirement income---such as Social Security and defined benefit pensions---is relatively certain.
 Furthermore, this paper focuses on the role of parental transfers for children' housing choices, where parental income risk is not a first-order concern.
@@
 After observing the parent's choices, the child in the second stage chooses consumption $c_c$, housing $h_c$, and savings $b_c$.
-The parent acts first to align with U.S.
-mortgage regulations, which mandate that gifts be deposited before approving mortgages.
+The parent acts first, so that transfers can alleviate LTV constraints and to comply with U.S.
+mortgage regulations, which require gifts to be deposited before loan approval.
 
@@
 The child is indifferent to receiving an additional unit of wealth or transfers, focusing only on total wealth (including transfers), $\tilde x_c\equiv x_c+t_p$.
-Furthermore, the parent's wealth and housing at the beginning of the period is redundant; the child only cares about the parent's bond and housing choices.
+Furthermore, the parent's current wealth and housing are redundant; the child cares only about the parent's bond and housing choices, since future transfers depend on the parent’s next-period resources.
 The child's state variable vector is $\mathbf{s}_c=\left(b_p',h'_p,x_c+t_p,y_c,h_c,a_c\right)$.
@@
 The first constraint is the budget constraint, which determines the net bond position as available resources (net worth, transfers, and income) minus expenditures (consumption, purchase price, and purchase cost).
-The second constraint is the law of motion for net worth, which is the sum of the bond position and housing wealth after depreciation.
+The second constraint is the law of motion for net worth, which is the sum of the bond position and housing wealth after maintenace costs.
 The third constraint is the LTV constraint.
@@
 \begin{table}
-	\center 
+	\begin{adjustbox}{center}
 	\begin{threeparttable}
@@
 	\end{threeparttable}
+\end{adjustbox}
 	\end{table}
@@
 How important are parental transfers for homeownership?
-I simulate the model without altruism ($\eta=0$), holding all other parameters constant---the standard model without parent-child interactions.
-In the main exercise, I assume a perfectly elastic housing supply, meaning that changes in homeownership have no effect on rental or house prices.
+I simulate the model without altruism ($\eta=0$), holding all other parameters constant---the standard model without parent-child interactions.\footnote{Alternatively, I could remove transfers entirely without removing altruism.
+Since utility is additively separable, the child's choices do not affect the parent's decisions when transfers are absent, aside from a small increase in parental savings due to the intergenerational correlation between wealth and productivity.
+As this yields nearly identical results, I report only the simpler case of no altruism.} In the main exercise, I assume a perfectly elastic housing supply, meaning that changes in homeownership have no effect on rental or house prices.
 In Section \ref{sec:endoprice}, I show that the results remain robust even with endogenous housing supply and varying elasticities.
@@
 Hence, the main aggregate effect of parental transfers is that they induce earlier homeownership.
-Finally, without altruism, the parental wealth gradient in homeownership decreases from 1.94 to 1.14, with the remaining effect driven by persistent productivity through initial conditions $F()$.
-
-\begin{table}
+Finally, without altruism, the parental wealth gradient in homeownership decreases from 1.94 to 1.14, with the remaining effect driven by the integenerational correlation in initial wealth and persistent labor productivity.
+
+\begin{table}[tb]
 	\center 
@@
 
+\subsubsection{Black-White Homeownership Gap}\label{sec:bwgap}
+One natural application of this framework is to quantify how much of the Black-White homeownership gap---young White households are nearly twice as likely to own homes as young Black households---is driven by differences in parental wealth.
+While existing structural models of racial inequality focus primarily on income and bequests, they typically omit housing and inter-vivos transfers.
+In contrast, many empirical studies focus on racial difference in housing outcomes.\footnote{For example, \citet{Ashman2020} and \citet{aliprantis2022dynamics} use overlapping-generations models to study the racial wealth gap, attributing most of it to income differences.
+Though the latter include pooled bequests, they do not model strategic behavior or direct transfers from parents to children.
+Many empirical studies highlight the role of parental wealth in racial housing gaps: \citet{charles2002transition} find that it explains 25\% of the mortgage application gap,  \citet{bond2021role} attribute 28\% of the gap in maintaining homeownership to parental wealth, and \cite{kermani2021racial} find a large gap in realized housing returns due to distressed sales.}
+
+\begin{table}
+	\center
+\begin{threeparttable}[tb]
+				\singlespacing
+		%			\footnotesize
+		\caption{Black-White Homeownership Rate}\label{tab:BW_table}
+		\input{../tabfig/postest/BW_mostrecent.tex}
+		\footnotesize
+		\textit{Notes:} Wealth is measured in 1000s of 2016 US dollars. See main text for details.
+	\end{threeparttable}
+\end{table}
+
+I aim to bridge the gap between these structural approaches and empirical results by using my model to quantify the contribution of parental transfers to the Black-White homeownership gap.
+I repeat the main experiment, seperately for Black and White households, following the methodology in \cite{Ashman2020} and \cite{aliprantis2022dynamics} closely.
+I shift the income level ($l_a$) according to racial differences: In the matched parent-child household, White households have 8\% higher income and Black households 42\% lower than the average.
+I then solve the model for White households.
+For Black households, I reduce house size by 25\% to match observed homeownership rates.
+I then solve the models both with and without altruism.
+The results are reported in Table \ref{tab:BW_table}.
+
+The observed Black-White homeownership gap is $\frac{0.54 - 0.27}{0.54} = 50\%$, compared to 52\% in the model with altruism.
+Removing altruism reduces the gap to 45\%, implying that parental transfers account for 7 percentage points---or 14\%---of the gap.
+While Black and White households receive transfers at similar rates in the model, transfers are less influential for Black households, who are further from the ownership threshold and receive smaller amounts.
+These results highlight the importance of racial differences in parental income, which shape both wealth accumulation and transfer capacity.
+
 \subsection{Endogenous House Prices}\label{sec:endoprice}
@@
 
-\section{Policy, Illiquidity, and Racial Gaps}\label{sec:pol}
-I have three main results regarding parental wealth and children's homeownership: (1) its relationship with housing-related policies, (2) how housing illiquidity amplifies its impact, and (3) its role in exacerbating the Black-White homeownership gap.
+\section{Policy Levers, Illiquidity, and MPCs}\label{sec:pol}
+I have three main results regarding parental transfers and children's homeownership: (1) how they interact with housing-related policies, (2) how housing illiquidity amplifies their impact, and (3) how they affect marginal propensities to consume (MPCs).
 
@@
 
-These experiments highlight two main findings.
+These experiments highlight three main findings.
 First, common policies intended to raise homeownership rates can either strengthen or weaken the role of parental wealth.
 Relaxing credit constraints increases both homeownership and reliance on parental wealth, as homeownership becomes more attractive to households with wealthier parents.
-Second, and in contrast, increasing housing liquidity---by lowering purchase or sales costs---reduces the importance of parental wealth.
+Moreover, in light of the earlier results on the Black–White homeownership gap (Section \ref{sec:bwgap}), broad-based down payment subsidies or tax rebates may unintentionally widen this gap.
+Second, and in contrast, increasing housing liquidity---by lowering sales costs---reduces the importance of parental wealth.
 Greater liquidity weakens the role of housing as a commitment device, and can even lower homeownership among children of wealthy parents.
@@
 
-To explore why some households prefer illiquidity, I break the sample down by liquidity preference and age in Figure \ref{fig:prefliq}.
-Figure \ref{fig:pref} plots the share of children who would prefer to keep housing illiquid.
-Illiquidity preference decreases with age as reliance on parental transfers decreases, and roughly 50\% of the youngest households prefer illiquidity.
-The next figures plot average homeownership, parental wealth, and child wealth by liquidity preference.
-To use illiquid housing to extract transfers, households must be homeowners, so households who prefer illiquidity have higher homeownership (Figure \ref{fig:pref_own}).
-To extract transfers, parents must have wealth, so households who prefer illiquidity have wealthier parents (Figure \ref{fig:pref_wealthp}).
-Finally, Figure \ref{fig:pref_wealthk} plots the child's wealth, and we see that households who prefer illiquidity are initially wealthier, but this relationship flips at age 33.
-The relationship flips because households only benefit from having their housing subject to sales costs if they are homeowners, which requires some wealth.
-To continue benefiting, they must remain liquidity-constrained (i.e., homeowners with limited wealth) to induce transfers.
+To explore why some households prefer illiquidity, Figure~\ref{fig:prefliq} breaks the sample by liquidity preference and age.
+Figure~\ref{fig:pref} shows that illiquidity preference declines with age, consistent with reduced reliance on parental transfers; around 50\% of the youngest households prefer illiquidity.
+The next panels plot homeownership, parental wealth, and child wealth by liquidity preference.
+Since illiquidity can only be used to extract transfers when the household owns a home and the parent is wealthy, those who prefer it have higher homeownership rates and wealthier parents.
+Figure~\ref{fig:pref_wealthk} shows that illiquidity-preferring households are initially wealthier, but this reverses around age 33.
+This flip reflects a tension: to use housing as a commitment device, households must be wealthy enough to own, yet poor enough to remain liquidity-constrained and elicit transfers.
+
 \begin{figure}[tb]
@@
 
-To summarize, this section shows that altruism not only reduces the downside of illiquidity, but even generates illiquidity preference.
-This is similar to models of behavioral biases, where households use the illiquidity of housing as a commitment device \cite{attanasio2024temptation}.
-In both settings, households use illiquidity to keep their marginal utility of wealth high in future periods.
-With altruism, this is done to induce future transfers.
-With time inconsistency, this is done to ensure that households will save in the future.
-
-\subsubsection{MPCs}
+In summary, altruism not only reduces the downside of illiquidity but can generate a preference for it.
+As in models of behavioral biases, households use housing illiquidity as a commitment device to keep marginal utility high in the future.
+With altruism, this induces transfers; with time inconsistency, it promotes saving \citep{attanasio2024temptation}.
+
+\subsection{MPCs}
 If altruism makes illiquid assets more attractive, a natural question is how altruism affects MPCs, as liquidity constraints are an important driver of MPCs \citep[see e.g.,][]{aguiar2024hand,Kaplan2014,fagereng2021mpc}.
@@
 
-\textcolor{red}{rewrite this paragraph}All told, these results have several implications for the MPC literature.
-Most model-implied MPCs are inflated by ignoring informal intra-family transfers.
-Second, one common strategy for raising model-implied MPCs is to introduce some households with low desired wealth holdings, for example through impatience \cite{aguiar2024hand}.
-However, this behavior is observationally similar to that of households receiving altruistic transfers from wealthy parents.
-Another typical method is to introduce an illiquid asset with a large return premium \cite{kaplan2022marginal}.
-However, as my results show, households with wealthier parents are not only more likely to be homeowners, but also have much lower MPCs than expected due to parental transfers.
-
-\subsection{Black-White Homeownership Gap}
-One natural application of this framework is to ask how much of the black-white homeownership gap---young White households are nearly twice as likely to own homes as young Black households---is due to different in parental wealth.
-
-Several structural papers have used life-cycle models with overlapping generations to study differences in economic outcomes by race.
-\cite{Ashman2020} find that the Black-White income gap accounts for most of the 90\% wealth gap.
-Similarly, \cite{aliprantis2022dynamics} find that closing the income gap is the most effective mechanism to reduce the wealth gap.
-While these papers account for racial differences in inheritances, they do not include endogenous inter-vivos transfers and ignore housing.\footnote{\cite{aliprantis2022dynamics} include a form of inter-vivos transfer receipt: Households leave pooled bequests at death, primarily to middle-aged households.
-There are no strategic considerations, since the bequests are stochastically distributed and parents do not transfer directly to their children.} While the structural papers have omitted housing, many empirical papers have looked at racial differences in housing outcomes.
-For example, \cite{charles2002transition} find that differences in parental wealth account for 25\% of the Black-White mortgage application gap, and \cite{bond2021role} find that parental wealth accounts for 28\% of the gap in maintaining homeownership.
-
-\begin{table}
-	\center
-\begin{threeparttable}[tb]
-				\singlespacing
-		%			\footnotesize
-		\caption{Black-White Homeownership Rate}\label{tab:BW_table}
-		\input{../tabfig/postest/BW_mostrecent.tex}
-		\footnotesize
-		\textit{Notes:} Wealth is measured in 1000s of 2016 US dollars. See main text for details.
-	\end{threeparttable}
-\end{table}
-
-I aim to bridge the gap between these structural approaches and empirical results by using my model to quantify the contribution of parental transfers to the Black-White homeownership gap.
-I repeat the main experiment, seperately for Black and White households, following the methodology in \cite{Ashman2020} and \cite{aliprantis2022dynamics} closely.
-I shift the income level ($l_a$) according to racial differences: In the matched parent-child household, White households have 8\% higher income and Black households 42\% lower than the average.
-I then solve the model for White households.
-For Black households, I reduce house size by 25\% to match observed homeownership rates.
-I then solve the models both with and without altruism.
-The results are reported in Table \ref{tab:BW_table}.
-
-The observed Black-White homeownership gap is $\frac{0.54 - 0.27}{0.54}=50\%$ and 52\% in the model with altruism.
-Without altruism, the Black-White homeownership gap decreases to 45\%.
-Transfers account for 7 percentage points (14\%) of the Black-White gap.
-Though Black and White households receive transfers at similar rates in the model, transfers are less important for Black households' homeownership, as they are further away from the ownership threshold and receive smaller transfers.
-My results show that racial differences in parental income, which affect parental wealth and transfers, are significant contributors to racial gaps in housing outcomes.
-Combined with the previous policy analysis, these results suggest that policies such as tax rebates or down-payment assistance for low-income or first-time buyers may widen the Black-White homeownership gap rather than narrow it.
-However, the model omits many other factors beyond income, such as discriminatory lending and appraisal bias, that affect the Black-White homeownership gap (e.g., \cite{shapiro2004hidden}).
+These results have several implications for the MPC literature.
+First, model-implied MPCs may be overstated when informal intra-family transfers are omitted, particularly for households with wealthy parents.
+Second, a common strategy to match high MPCs is to introduce households with low target wealth—e.g., through impatience \citep{aguiar2024hand}—but such behavior is observationally similar to that of households expecting altruistic support.
+Third, models that generate high MPCs by introducing illiquid assets with large return premia \citep{kaplan2022marginal} may overlook the fact that homeowners with wealthy parents have access to informal insurance, which lowers their effective MPC.
 
@@
 
-The second contribution of the paper is to include an illiquid asset in the altruism framework.
+Another contribution of the paper is to include an illiquid asset in the altruism framework.
 With altruistic parents, transfers generally flow to borrowing-constrained households, since they have a large marginal utility of wealth \citep{Barczyk2020,Chu2020}.
@@
 First, I show that some households with wealthy parents choose to be liquidity-constrained to receive larger transfers.
+While it is difficult to test this prediction, my empirical results are consistent with this behavior: parental wealth improves housing outcomes after purchase, including ownership retention, lower delinquency, and less downsizing in unemployment.
 Second, I show that the lack of commitment generates preferences for illiquidity, even when households are rational and without behavioral biases \citep[see e.g.,][]{attanasio2024temptation}.

### 2025-04-30 src/Julia/quant_noalt.jl

@@ -50,7 +50,7 @@ function output_noalt(Mb::Model,Mn::Model,targ_moms::Vector{Symbol},nontarg_moms
     append_info = @sprintf("m%i_d%i",Dates.month(Dates.today()),Dates.day(Dates.today()))
     filename = "tabfig/postest/quant_pres_"*append_info*".tex"
     filename2 = "tabfig/postest/quant_mostrecent.tex"
-    display(df_latex)
+    println(df_latex)
     write(filename, df_latex)
     if store == true
         write(filename2, df_latex)

### 2025-04-30 src/Julia/run_all.jl

@@ -60,7 +60,8 @@ pChetty = ChettyEventStudy(Mb,0.75,store=store)
 #######################
 ## Contribution of Altruism to Homeownership
 Mn = altrcontrhousing(Mb.mp,Mb.np);
-output_noalt(Mb,Mn,targ_moms,nontarg_moms;store=store)
+nontarg_moms_notransf = [x for x in nontarg_moms if x ∉ [:tp2wealthpyoung, :transferbuyersyoung,:transfraterenteryoung,:transfrateowneryoung]] # No reason to report transfer moments in this table
+output_noalt(Mb,Mn,targ_moms,nontarg_moms_notransf;store=store)
 
 ## Policy function plts
 plot_policyfuncs(Mb,Mn,store=store)

### 2025-04-30 docs/Homeownership_BankOfMomAndDad.tex
Done with empirical section

@@
 
-\textit{AHS:} I use the 1991-2019 AHS, which surveys occupants of housing units, to construct a time series of whether homeowners primarily funded their down payment with gifts or inheritances (right panel of Figure~\ref{fig:motivation}).
-After excluding households who report sale of previous homes as a source, the sample includes about 50,000 observations each year.
+\textit{AHS:} I use the 1991-2019 AHS, which surveys occupants of housing units, to construct a time series of whether homeowners primarily funded their down payment with gifts or inheritances (right panel).
+After excluding households who report sale of previous homes as a funding source, the sample includes about 50,000 observations each year.
 The series is relatively flat through the 1990s, followed by a sharp decline from 2005 to 2009---coinciding with the rise of low–down-payment mortgages---and a marked increase from 2013 onward.
@@
 dollars.
-Table \ref{tab:vardef} in the Appendix provides variable code of the main variables.
 
@@
 This leads to discrepancies, where the child and parent do not agree on the amount given from the parent to the child.
-I only use the parent's reported transfer.
+For each child, I use the parent's reported transfer to that child.
 First, there may be some stigma about receiving transfers, which may induce receiving children to underreport.
@@
 
-\textit{Definition of Transfers:} To measure the transfer receipt rate, I rely on the transfer supplement which asked all parent households whether they gave money, gifts, or loans of \$100 or more to each children in 2012.
+\textit{Definition of Transfers:} To measure the transfer receipt rate, I rely on the transfer supplement which asked all parent households whether they gave money, gifts, or loans of \$100 or more to each child in 2012.
 I follow the literature \citep[e.g.,][]{mcgarry2016dynamic} and treat all transfers as gifts.
-Since this paper focuses on transfers that (a) relate to housing and (b) are quantitatively meaningful, I set transfers below \$500 to \$0.
-About 25\% of transfers are below this threshold, and ignoring them increases the conditional mean transfer from \$2,921 to \$3,960 and reduce the transfer receipt rate from 45\% to 22\%.\footnote{These transfer rates are relatively high and higher than those calculated by \citealt{feiveson2019lifecycle} using the Surey of Consumer Finances (SCF).
-First, it seems likely that the PSID transfer supplement was more successfull in counting transfers than the SCF.
-Second, the SCF asks about transfers and gifts going further back in time, likely increasing underreporting.
-\textcolor{red}{write more here} }
+Since this paper focuses on transfers that (a) relate to housing and (b) are quantitatively meaningful, I ignore transfers below \$500.
+About 25\% of transfers are below this threshold, and ignoring them increases the conditional mean transfer from \$2,921 to \$3,960 and reduces the transfer receipt rate from 45\% to 22\%.\footnote{These transfer rates are relatively high and exceed those reported by \citealt{feiveson2019lifecycle} using the Survey of Consumer Finances (SCF).
+The PSID transfer supplement likely captures a broader range of transfers than the SCF, which deliberately focuses on substantial transfers.
+In addition, the SCF asks about transfers further back in time, increasing the risk of recall error and underreporting.}
 
 In all other waves, there is a general question of gifts and inhertitances from people outside the household.
-We do not know the source of these transfers (e.g., they could be from friends or an inheritance from grandparents) and hence these do not know the extent to which these captures parental assistance.
-Yet, they, to some extent, capture the impact of transfer receipt on household decisions, and hence I use these variables in some of the empirical results, following \cite{Lee2018}.
+We do not know the source of these transfers (e.g., they could be from friends or grandparents) and hence these do not know the extent to which these captures parental assistance.
+Nevertheless, they still capture some of the impact of transfer receipt on household decisions.
+I therefore include these variables in some of the empirical results, following \citet{Lee2018}.
 
@@
 \subsubsection{Descriptive Statistics: Who Receives Transfers?}
-\textcolor{red}{!! [[rewrite this]] !!}
-I now discuss descriptive statistics from the PSID sample in 2013 for households aged 25 to 44 with an observed parent household.
+I now discuss descriptive statistics from the 2013 PSID sample for households aged 25 to 44 with an observed parent household.
 Table \ref{tab:descrstats} contains the means of variables, by age, wealth, homeownership, and transfer receipt.
 Though limited to households with observed parents, the sample appears reasonably representative.
-Average income by age closely matches estimates from the Survey of Consumer Finances \cite[see e.g.,][]{kuhn20162013}, though average wealth is slightly lower.
+Average income by age closely matches estimates from the SCF \cite[see e.g.,][]{kuhn20162013}, though average wealth is slightly lower.
 
 There are several main takeaways from the subgroup analysis.
-In 2013, 22\% of young households received a transfer, and transfers averaged \$3,950.
+In 2012, 21\% of young households received a transfer, and transfers averaged \$3,960.
 Receivers have significantly richer parents, have similar wealth and income as nonreceivers, and are less likely to own.
@@
 
-In panel a), the first two columns compare the whole sample, by transfers receipt in the last calendar year.
-The mean transfer is relatively high at \$3,950, and 22\% of households received transfers in the last year.
-The largest difference is parental wealth: transferring parents are 2.5 times as wealthy as nontransferring parents (\$1,029,000 vs.
-\$410,000).
+The first two columns of the top panel compare households by transfers receipt.
+The mean transfer is relatively high at \$3,960, and 21\% of households received transfers in the last year.
+The largest difference is parental wealth: transferring parents are 2.5 times as wealthy as nontransferring parents.
 Perhaps surprisingly, receivers are slightly richer than nonreceivers (\$108,000 vs.
@@
 
-\begin{table}[htbp]
+\begin{table}[tbp]
 	\begin{adjustbox}{center}
@@
 	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}%
-		\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. 
+		\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules.
+Weighted using family weights.
+Transfer, wealth, and income measured in 1000s of 2016 USD.
+Table \ref{tab:vardef} in the Appendix provides variable codes of the main variables.
 	\end{flushleft}\end{footnotesize}}		
@@
 Next, I break the sample down by ownership.
-Homeowners have 10 times the wealth of renters and also significantly wealthier parents.
+Homeowners are both wealthier and have wealthier parents than renters.
 Renters and owners receive transfers at about the same rate and size.
@@
 These regressions are similar to those in \cite{Lee2018}, who also use the PSID.
-
-To measure transfer receipt, I compare two definitions.
-Definition 1 (narrow): Parental gifts exceeding \$500 in 2012, as reported in the PSID’s 2013 Transfer Module, limiting the sample to renters in 2011.
-Definition 2 (broad): Any gift or inheritance exceeding \$10,000 reported in all waves of the PSID, or a parental gift exceeding \$10,000 in 2012.
-This expands the sample include all waves of the PSID, and lets us include household fixed effects, though we now only capture large transfers and include some non-parental transfers.
+It is important to note that transfer receipt is not exogenous---many transfers may be made precisely because the renter is buying a home.
+
+I use two definitions of transfer receipt.
+The narrow definition includes parental gifts over \$500 in 2012, limiting the sample to renters in 2011.
+The broad definition includes any gift or inheritance over \$10,000 reported in any wave, or a large parental gift in 2012.
+This allows for household fixed effects and a larger sample, but captures only large transfers and includes some non-parental ones.
 
@@
 
-To increase statistical power, I now use the second transfer receipt definition, which also includes non-parental gifts and bequests.
-This leads to a ten-fold increase in sample size, but the transfer receipt rate also decrease to around 3\%, and the median transfer size jumps from \$1,500 to \$24,200.
-In addition, we can now follow renters over time, enabling the inclusion of household fixed effect.
-In a univariate regression, transfers increase the probability of becoming a homeowner by 21.9 percentage points.
-Even with the full set of controls, the effect of transfer receipt on the future homeownership is large and statisticaly significant: 13.5 percentage point in the OLS regression ($p<0.001$) and a 7.9 percentage point in the FE regression  ($p<0.1$).
-These effects are large, similar in magnitude of the baseline rent-to-own transition rate of 18 percent.
-A few caveats are in order: I only capture transfers above \$10,000, and the effect of transfer receipt is likely increasing with transfer size.
-Second, these transfers are not exogenous: indeed, it is likely that some of these transfers are given \emph{because} households are buying a house.
+To increase statistical power, I now use the broader transfer definition, which includes all large transfers.
+This expands the sample size tenfold but reduces the transfer receipt rate to around 3\%, while the median transfer size rises from \$1,500 to \$24,200.
+The broader sample also allows me to follow renters over time and include household fixed effects.
+In a univariate regression, transfer receipt is associated with a 21.9 percentage point increase in the probability of becoming a homeowner.
+Even with a full set of controls, the effect remains large and statistically significant: 13.5 percentage points in the OLS specification ($p<0.001$) and 7.9 percentage points with household fixed effects ($p<0.1$).
+These magnitudes are substantial compared to the baseline rent-to-own transition rate of 18\%.
 
@@
 \subsubsection{Transfer Sizes}
-For additional detail on the transfer variables, I plot kernel density estimates in Figure \ref{fig:PSID_transfer_to_renters}, for renters in $t$, separated by whether they own or rent in $t+2$.
-The left panel only includes parental transfers exceeding \$500 from the 2013 transfer supplement.
-The right panel only includes transfer/inheritance exceeding \$10,000 which is included in every wave.
+For additional detail on the transfer variables, I plot kernel density estimates in Figure \ref{fig:PSID_transfer_to_renters}, for renters in $t$, separated by whether they own or rent in $t+2$, for the two transfer definitions.
 There are two takeaways.
@@
 
-While these are the best data available to follow household over time---in particular their wealth, income, and housing choices---and transfers, the data have several some limitations.
-First, in the transfer supplement we only observe transfers in 2012, while the PSID is collected every other year.
-Hence we undercount both transfer rates and sizes.
-Second, in the gift/inheritance question, we do not know the source of the transfer.
-Third, we do not know the extent to which households report intra-family loans, nor the extent to which these are repaid.
-\begin{figure}
+\begin{figure}[tb]
 	\caption{Transfer Sizes to Renters by Transition to Ownership}\label{fig:PSID_transfer_to_renters}
@@
 
-\subsubsection{Households with Wealthier Parents Are More Likely to Remain Homeowners}
-While the previous sections have focused on the transition into ownership, I now focus on mainting homeownership.
-\textcolor{red}{rewrite this better: That said, \cite{bond2021role} is an exception and shows the positive association between parental wealth and children's housing outcomes in the PSID and the Health and Retirement Study.
-}
-
-First, the left panel of Figure \ref{fig:owner_hazard} plots the share of owners aged 25 to 44 in year $t$ who remain owners in period $t+\Delta$.
-Among first-time owners, only 80\% have retained ownership four years later.
-For non-first-time owners, there is still a marked decrease though it is more muted, after four years about 10\% of all owners will be renting.
-
-To understand which variables predicts retainment of ownership, I estimate
-\begin{equation}	\label{eq:regmaintain}
-	\Pr\left(Own_{i,t+\Delta}=1 \mid Own_{i,t}=1\right)
-	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
-	  + \gamma' X_{i,t} + \varepsilon_{i,t},
-\end{equation}
-
-where $Own_{i,t+\Delta}=1$ denotes an owner at time $t+\Delta$, $\beta$ and $\zeta$ measures the effect of log household net worth and parental wealth, respectively  and $X_{i,t}$ is a vector of control variables.
-The only changes relative to Eq.
-\ref{eq:transferrent} is that the outcome variable conditions on current ownership and measures future ownership, that there are no controls for transfer receipt, and that the household net worth and parental wealth are taken out of the vector of control variables.
-The sample is limited to households aged 25 to 44.
-These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Survey.
-Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plots the two coefficients of interest, over variying lead, times and their 90\% confidence intervals for all only and first-time owners only.
-
-Two results stand out: First, we see that parental wealth is at least as important as household net worth at all horizons, though the difference is never statistically significant.
-Second, we see that the impact of parental wealth is increasing over time.
-For example, we see that a 10\% increase in parental wealth increases the probability of retaining ownership after 8 years by 0.45 percentage points among first-time owners.
-These patterns will be serve as one part of model validation.
+While the PSID allows us to observe transfers and changes in housing tenure over time, it has several limitations.
+First, the transfer supplement captures transfers only in 2012, while the PSID is collected biennially, leading to undercounts of both transfer rates and amounts.
+Second, in the gift/inheritance question, the source of the transfer is unknown.
+Third, it is unclear whether intra-family loans are consistently reported or repaid.
+
+\subsubsection{Households with Wealthier Parents Retain Ownership}
+While most research focuses on the transition into ownership, I now focus on the challenge of retaining it.
+
 \begin{figure}[tb]
 	\captionsetup[subfigure]{aboveskip=-1pt}
-	\caption{Mainting Homeownership}\label{fig:homeowner_exit}
+	\caption{Retaining Homeownership}\label{fig:homeowner_exit}
 	\begin{subfigure}{0.33\textwidth}%
-		\caption{Exit Out of Ownership}\label{fig:owner_hazard}%
+		\caption{Retaining Ownership}\label{fig:owner_hazard}%
 		\includegraphics[width=\textwidth]{../tabfig/descr/PSID_ownerexit.pdf}% 
@@
 	\end{subfigure}%
-	\hfill
-
-	\caption*{\footnotesize \textit{Notes:} \textcolor{red}{write this footnote.}}
+
+	\caption*{\footnotesize \textit{Notes:} The left panel shows the share of current owners who remain owners in future years.
+The center and right panels show the estimated coefficients from Eq.~\ref{eq:regmaintain}, with dashed lines indicating 90\% confidence intervals.
+See main text for details.}
 \end{figure}
 
-\subsubsection{Households with Wealthier Parents Are Less Likely to Be Behind on Mortgages} 
-% To realize the benefits of homeownership, households must be able to maintain their mortgage payments, since failure to do so can lead to distressed and forced sales.
-I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses.
-The PSID has collected this data since 2009, and 3.9\% of owners aged 25 to 44 are behind, on average, while 9.2\% have been behind at least once.
-One potential roadblock to maintaining ownership are mortgage delinquencies, as failure to meet mortgage payments can lead to distressed and forced sales.
-I now show that households with wealthier parents are less likely to be behind on their mortgages, even though they buy more expensive houses.
-The PSID has collected this data since 2009, and about 2\% of owners aged 25 to 44 are behind.
-
-To understand which variables predicts whether households are behind on mortgage payments, I estimate
+First, the left panel of Figure \ref{fig:owner_hazard} plots the share of owners aged 25 to 44 in year $t$ who remain owners in period $t+\Delta$.
+Among first-time owners, only 80\% have retained ownership four years later.
+For non-first-time owners, there is still a marked decrease though it is more muted, after four years about 10\% of all owners will be renting.
+
+To understand which variables predicts retainment of ownership, I estimate
+\begin{equation}	\label{eq:regmaintain}
+	\Pr\left(Own_{i,t+\Delta}=1|Own_{i,t}=1\right)
+	= \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t})
+	  + \gamma' X_{i,t} + \varepsilon_{i,t},
+\end{equation}
+where $Own_{i,t+\Delta}=1$ denotes an owner at time $t+\Delta$, $\beta$ and $\zeta$ measures the effect of log household net worth and parental wealth, respectively  and $X_{i,t}$ is a vector of control variables.
+The only changes relative to Eq.
+\ref{eq:transferrent} is that the outcome variable conditions on current ownership and measures future ownership, that there are no controls for transfer receipt, and that I have taken the two variables of interest out of the vector of control variables.
+The sample is limited to households aged 25 to 44.
+These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Study.
+Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plots the two coefficients of interest over different time horizons.
+
+Two results stand out.
+First, parental wealth is at least as predictive as household net worth at all horizons, although the difference is never statistically significant.
+Second, the effect of current parental wealth grows with the time horizon.
+For instance, a 10\% increase in parental wealth raises the probability of retaining ownership after eight years by 0.45 percentage points among first-time owners.
+I return to these results later as part of the model validation.
+
+\subsubsection{Households with Wealthier Parents Are Less Delinquent} 
+One potential roadblock to maintaining ownership is mortgage delinquency, as missed payments can lead to distressed or forced sales.
+I show that households with wealthier parents are less likely to fall behind on their mortgages, despite purchasing more expensive homes.
+The PSID has collected mortgage delinquency data since 2009; about 2\% of owners aged 25 to 44 report being behind.
+
+To examine what predicts mortgage delinquency, I estimate: 
 \begin{equation*}
@@
 \end{equation*}
-and the control variables are the same as in the previous regression.
-Table \ref{tab:hypo} reports the estimated coefficient on household net worth ($\beta$) and parental wealth ($\zeta$), with the full coefficient table relegated to Appendix Table \ref{tab:hypo_long}.
+using the same control variables as in the previous regression.
+Table~\ref{tab:hypo} reports the estimated coefficients on household net worth ($\beta$); full results appear in Appendix Table~\ref{tab:hypo_long}.
 
 I first limit the sample to first-time buyers to focus on distress among new owners.
-In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$) WWith the full set of controls, the effect increases to 0.1 percentage points, though it is no longer statistically significant.
-These are economically meaningful since 1.6\% of first-time owners are behind on their mortgage.
-Moreover, parental wealth is more predictive of future delinqucies than household net worth.
-In Columns 3 to 5, I instead focus on all owners (who are currently not behind), increasing the sample sizes and allowing us to include household fixed effects.
-In these specifications, a 10\% increase in parental wealth decreases the probability of being behind two years later by 0.09 percentage points in the univariate OLS regression ($p<0.001$), 0.02 percentage points in the OLS regression (not significant) and  by 0.07 percentage points in the FE regression ($p<0.1$).
-All told, the probability of being behind on mortgages is decreasing in parental wealth, highlighting that parental resources appear to help new homeowners maintain ownership after purchase---in addition to relaxing borrowing constraints.
+In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$) With the full set of controls, the effect increases slightly, though it is no longer statistically significant.
+These magnitudes are economically meaningful since 1.6\% of first-time owners are behind on their mortgage.In Columns 3 to 5, I expand the analysis to include all owners who are currently not behind, increasing the sample sizes and allowing us to include household fixed effects.
+In these specifications, a 10\% increase in parental wealth is associated with a 0.09 percentage point decrease in delinquency risk in the univariate OLS regression ($p<0.001$), a 0.02 percentage point decrease in the multivariate OLS regression (not significant), and a 0.07 percentage point decrease in the fixed effects regression ($p<0.1$).
+
 \begin{table}
@@
 
+Overall, the probability of falling behind on mortgage payments declines with parental wealth, even after controlling for a wide range of household characteristics.
+This highlights a novel channel through which parental resources help new homeowners retain ownership after purchase.
+
 \subsubsection{Households with Wealthier Parents Are Less Likely to Downsize during Unemployment Spells}\label{sec:eventstudy}
-Distressed and forced home sales are costly and frequently driven by unemployment \citep{kermani2021racial,hsu2018unemployment}.
-If parental wealth provides insurance, households with wealthier parents should be less likely to downsize after income losses.
-I perform a simple event study on the effect of unemployment on housing consumption to test this hypothesis.
+Distressed and forced home sales are costly and often triggered by unemployment \citep{kermani2021racial,hsu2018unemployment}.
+To test whether parental wealth mitigates this risk, I perform a simple event study of how unemployment affects the likelihood of downsizing.
 
@@
 The outcome of interest is the change in log housing consumption, which is set to zero for households who do not move.
-For households who do move, housing consumption is defined as yearly rent while renting and the imputed rent of owner-occupied housing is set to 5\% of the market value \citep{Davis2008}.
-The sample is limited to household heads who are unemployed only once between ages 25 and 44.
-I divide the sample by whether a household's parents were in the top parental wealth quartile (of parents with children aged 25 to 44) at the time of unemployment and compare the two groups' housing consumption growth rates at unemployment.
+For movers, housing consumption is defined as yearly rent while renting and as imputed rent for owners, set to 5\% of the market value \citep{Davis2008}.
+The sample is limited to household heads who experience only one unemployement spell between ages 25 and 44.
+I divide the sample by whether a household's parents were in the top parental wealth quartile at the time of unemployment and compare the two groups' housing consumption growth at unemployment.
 The results are displayed in Figure \ref{fig:housinggrowthrates}.
-Among households with nonwealthy parents, housing consumption decreases by about 5\% during unemployment (significant at the 95\% level).
-In contrast, among households with wealthy parents, housing consumption even grows slightly at unemployment.
-Moreover, I replicate this event study using simulated data from the quantitative model and show that the results are similar (see Section \ref{sec:externalval} and Figure \ref{fig:housinggrowthrates_model}).
-In appendix \ref{app:eventstudy_details} I, instead of using the means, use a formal event study, which shows that households with poor parents have a statsticaly significant decrease in housing consumption during unemployment spells.
+Among households with non-wealthy parents, housing consumption falls by about 5\% during unemployment ($p<0.05$).
+In contrast, households with wealthy parents see a slight increase (not significant)in housing consumption during unemployment.\footnote{In Appendix \ref{app:eventstudy_details}, I also report the results of a formal event study using the same set of controls as in the previous regressions.}  I return to these results later as part of the model validation.
+
 \begin{figure}
@@
 \section{A Quantitative Model of Parental Transfers and Homeownership}\label{sec:model}
-%% There is a missing paragraph linking the previous results this this part!
 This section describes my life-cycle model of housing choices with overlapping generations, idiosyncratic earnings risk, and altruistic inter-vivos transfers from parents to their offspring (``children'').

### 2025-05-01 docs/Homeownership_BankOfMomAndDad.tex
Updated empirics, moved rent-to-own reg to appendix

@@
 
-\subsection{Parental Resources and Housing Outcomes in the Data}\label{sec:dataregr}
-
-I now utilize microdata from the PSID to examine the empirical association between transfer receipt and parental financial resources on housing outcomes.
-The sample is limited to households aged 25 to 44 to focus on the housing choices of young adults.
-I use the sum of parental net worth and income (denoted as wealth) to measure parents' financial resources in all specifications for two reasons.
-First, this is consistent with models of altruism, as the one used later in the paper.
-Second, since parents' wealth and income are highly correlated and I have few observations---fewer than 1,000 for many specifications---the standard errors are large when both are included.
-
-\subsubsection{Transfers and the Rent-to-Own Transition}
-
-While a growing literature documents that parental transfers increase the probability that renters become homeowners (e.g.\ \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}), I replicate this result to highlight the importance of transfers for housing outcomes.
-I estimate
-\begin{equation}	\label{eq:transferrent}
-	\Pr\left(Own_{i,t+2}=1 \mid Own_{i,t}=0\right)
-	= \alpha + \beta\,Transfer_{i,t+2}
-	  + \gamma' X_{i,t} + \varepsilon_{i,t},
-\end{equation}
-where \(Own_{i,t}=0\) denotes a renter at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income and log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.
-The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners.
-Table \ref{tab:newowners_short} reports the estimated coefficient on transfer ($\beta$), with the full coefficient table relegated to Appendix Table \ref{tab:newowners}.
-These regressions are similar to those in \cite{Lee2018}, who also use the PSID.
-It is important to note that transfer receipt is not exogenous---many transfers may be made precisely because the renter is buying a home.
-
-I use two definitions of transfer receipt.
-The narrow definition includes parental gifts over \$500 in 2012, limiting the sample to renters in 2011.
-The broad definition includes any gift or inheritance over \$10,000 reported in any wave, or a large parental gift in 2012.
-This allows for household fixed effects and a larger sample, but captures only large transfers and includes some non-parental ones.
-
-In a univariate regression, receiving a parental transfer is associated with a 2.9 percentage point higher probability of becoming a homeowner in 2013 among renters in 2011.
-With a full set of controls parental transfers are associated with a 1.2 percentage point higher probability of becoming a homeowner.
-This represents a substantial increase relative to the baseline rent-to-own transition rate of about 14\%, though the effect is not statistically significant.
-
-To increase statistical power, I now use the broader transfer definition, which includes all large transfers.
-This expands the sample size tenfold but reduces the transfer receipt rate to around 3\%, while the median transfer size rises from \$1,500 to \$24,200.
-The broader sample also allows me to follow renters over time and include household fixed effects.
-In a univariate regression, transfer receipt is associated with a 21.9 percentage point increase in the probability of becoming a homeowner.
-Even with a full set of controls, the effect remains large and statistically significant: 13.5 percentage points in the OLS specification ($p<0.001$) and 7.9 percentage points with household fixed effects ($p<0.1$).
-These magnitudes are substantial compared to the baseline rent-to-own transition rate of 18\%.
-
-\begin{table}[tb]
-	\centering
-	\begin{threeparttable}
-		\caption{The Transition to Ownership}
-		\label{tab:newowners_short}
-		\small 
-				\input{../tabfig/regr/newowners_short.tex}
-
-	\end{threeparttable}
-	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}%
-		\textit{Notes:} Standard errors in parentheses, clusted at the household level.
-\textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
-	\end{flushleft}\end{footnotesize}}		
-\end{table}
-
 \subsubsection{Transfer Sizes}
@@
 Third, it is unclear whether intra-family loans are consistently reported or repaid.
+
+\subsection{Parental Resources and Housing Outcomes in the Data}\label{sec:dataregr}
+
+I now utilize microdata from the PSID to examine the empirical association between parental financial resources on housing outcomes.
+The sample is limited to households aged 25 to 44 to focus on the housing choices of young adults.
+I use the sum of parental net worth and income (denoted as wealth) to measure parents' financial resources in all specifications for two reasons.
+First, this is consistent with models of altruism, as the one used later in the paper.
+Second, since parents' wealth and income are highly correlated and I have few observations---fewer than 1,000 for many specifications---the standard errors are large when both are included.
+
+\subsubsection{Transfers and the Transition into Homeownership}
+Since many papers have already estimated the relationship between parental transfers and transitions into homeownership (e.g.\ \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Lee2018}), I replicate these regressions to highlight the association between transfer receipt and rent-to-own transitions, but relegate the full results to Appendix~\ref{app:rent_to_own}.
+Receiving a parental transfer above \$500 is associated with a 1.2 percentage point increase in the probability of becoming a homeowner (not statistically significant), after controlling for household characteristics.
+Receiving large transfers above \$10,000 is associated with a 7.9 percentage point increase in the probability of becoming a homeowner ($p<0.1$), even after including household fixed effects.
+These are sizable effects relative to the baseline rent-to-own transition probability of about 18\% during the sample period.
 
@@
 \end{equation}
-where $Own_{i,t+\Delta}=1$ denotes an owner at time $t+\Delta$, $\beta$ and $\zeta$ measures the effect of log household net worth and parental wealth, respectively  and $X_{i,t}$ is a vector of control variables.
-The only changes relative to Eq.
-\ref{eq:transferrent} is that the outcome variable conditions on current ownership and measures future ownership, that there are no controls for transfer receipt, and that I have taken the two variables of interest out of the vector of control variables.
+where $Own_{i,t+\Delta}=1$ denote ownership at time $t+\Delta$, $\beta$ and $\zeta$ measure the effects of log household net worth and parental wealth, respectively.
+The vector of control variables  $X_{i,t}$ consists of log household income, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.
 The sample is limited to households aged 25 to 44.
 These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Study.
-Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plots the two coefficients of interest over different time horizons.
+Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plot the two coefficients of interest over different time horizons.
 
@@
 To examine what predicts mortgage delinquency, I estimate: 
-\begin{equation*}
+\begin{equation}\label{eq:regbehind}
 \Pr(Behind_{i,t+2}) = \beta\ln(NetWorth_{i,t}) + \zeta\ln(ParWealth_{i,t}) + \gamma' X_{i,t} + \varepsilon_{i,t},
-\end{equation*}
+\end{equation}
 using the same control variables as in the previous regression.
-Table~\ref{tab:hypo} reports the estimated coefficients on household net worth ($\beta$); full results appear in Appendix Table~\ref{tab:hypo_long}.
+Table~\ref{tab:hypo} reports the estimated coefficients on parental wealth ($\zeta$); full results appear in Appendix Table~\ref{tab:hypo_long}.
 
@@
 	\begin{threeparttable}
-		\caption{Housing Choices and Parental Wealth}
+		\caption{Parental Wealth and Future Mortgage Delinquencies}
 		\label{tab:hypo}
@@
 	{\begin{footnotesize}\begin{flushleft}
-		\textit{Notes:} Standard errors in parentheses, clustered at the household level.
+		\textit{Notes:} The estimated association between parental wealth and the probability of behind on mortgage payments in the next period from regressions using specification \ref{eq:regbehind}.
+See Appendix Table~\ref{tab:hypo_long} for all results.
+Standard errors in parentheses, clustered at the household level.
 \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
@@
 
-\subsubsection{Households with Wealthier Parents Are Less Likely to Downsize during Unemployment Spells}\label{sec:eventstudy}
+\subsubsection{Owners with Wealthy Parents Don't Downsize in Unemployment}\label{sec:eventstudy}
 Distressed and forced home sales are costly and often triggered by unemployment \citep{kermani2021racial,hsu2018unemployment}.
@@
 Taking stock, we have seen that parental transfers have become more important for young homeowners over the past two decades.
-About 30\% of young households receive transfers from their parents in a given year, and recipients' parents are wealthier.
-Recipients are also more likely to transition from renting to owning, also when we condition on a rich set of covariates.
-Parental wealth is also positively associated with maintaing homeownership.
-The probability of mainting homeownership is increasing in parental wealth, households with richer parents are less likely to behind on mortgages and less likely to downsize during unemployment sptells.
+About 20\% of young households receive transfers from their parents in a given year, and recipients have wealthier parents.
+Recipients are more likely to transition from renting to owning, even after conditioning on a rich set of covariates.
+Parental wealth is also positively associated with maintaining homeownership.
+Households with richer parents are less likely to fall behind on mortgages and less likely to downsize during unemployment spells.
 These results suggest that homeowners with wealthy parents find illiquidity of homeownership less problematic.
@@
 \renewcommand{\thetable}{A\arabic{table}}
+
+\section{Transfers and the Rent-to-Own Transition}\label{app:rent_to_own}
+A large literature documents that parental transfers increase the probability that renters become homeowners (e.g., \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Guiso2002}; \citealp{Engelhardt1998}), I replicate this result to highlight the importance of transfers for housing outcomes.
+
+I estimate
+\begin{equation}	\label{eq:transferrent}
+	\Pr\left(Own_{i,t+2}=1 \mid Own_{i,t}=0\right)
+	= \alpha + \beta\,Transfer_{i,t+2}
+	  + \gamma' X_{i,t} + \varepsilon_{i,t},
+\end{equation}
+where \(Own_{i,t}=0\) denotes a renter at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income and log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.
+The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners.
+Table \ref{tab:newowners} reports the estimated coefficients.
+These regressions are similar to those in \cite{Lee2018}, who also use the PSID.
+It is important to note that transfer receipt is not exogenous---many transfers may be made precisely because the renter is buying a home.
+
+I use two definitions of transfer receipt.
+The narrow definition includes parental gifts over \$500 in 2012, limiting the sample to renters in 2011.
+The broad definition includes any gift or inheritance over \$10,000 reported in any wave, or a large parental gift in 2012.
+This allows for household fixed effects and a larger sample, but captures only large transfers and includes some non-parental ones.
+
+In a univariate regression, receiving a parental transfer is associated with a 2.9 percentage point higher probability of becoming a homeowner in 2013 among renters in 2011.
+With a full set of controls parental transfers are associated with a 1.2 percentage point higher probability of becoming a homeowner.
+This represents a substantial increase relative to the baseline rent-to-own transition rate of about 14\%, though the effect is not statistically significant.
+
+To increase statistical power, I now use the broader transfer definition, which includes all large transfers.
+This expands the sample size tenfold but reduces the transfer receipt rate to around 3\%, while the median transfer size rises from \$1,500 to \$24,200.
+The broader sample also allows me to follow renters over time and include household fixed effects.
+In a univariate regression, transfer receipt is associated with a 21.9 percentage point increase in the probability of becoming a homeowner.
+Even with a full set of controls, the effect remains large and statistically significant: 13.5 percentage points in the OLS specification ($p<0.001$) and 7.9 percentage points with household fixed effects ($p<0.1$).
+These magnitudes are substantial compared to the baseline rent-to-own transition rate of 18\%.
+
+\begin{table}
+	\center
+	\begin{threeparttable}
+		\caption{The Transition to Ownership}
+		\label{tab:newowners}
+		\small 
+				\input{../tabfig/regr/newowners.tex}
+
+		{\begin{footnotesize}\begin{flushleft}
+		\textit{Notes:} Standard errors in parentheses, clustered at the household level.
+\textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
+		\end{flushleft}\end{footnotesize}}		
+	\end{threeparttable}
+\end{table}
 
@@
 
-\begin{table}
-	\center
-	\begin{threeparttable}
-		\caption{The Transition to Ownership}
-		\label{tab:newowners}
-		\small 
-				\input{../tabfig/regr/newowners.tex}
-
-		{\begin{footnotesize}\begin{flushleft}
-		\textit{Notes:} Standard errors in parentheses, clustered at the household level.
-\textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
-		\end{flushleft}\end{footnotesize}}		
-	\end{threeparttable}
-\end{table}
-
 \begin{figure}\caption{Calibrated Income Process}\label{fig:calincome}

### 2025-05-01 docs/Homeownership_BankOfMomAndDad.tex
More minor text changes

@@
 \end{equation}
-where \(Own_{i,t}=0\) denotes a renter at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income and log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.
+where \(Own_{i,t}=0\) denotes renting at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income, log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.
 The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners.

### 2025-05-01 docs/Homeownership_BankOfMomAndDad.tex
Added new retention model results

@@
 
-I first present empirical evidence indicating that parental wealth helps not only in the transition to ownership, but also in maintaining ownership.
-\textcolor{red}{bring in new empirical results here.
-Highlight that the results on behind and unemployment are new}.
-I show that, controlling for a wide set of household and parental characteristics, households with wealthier parents are less likely to fall behind on mortgage payments after becoming homeowners, even though they buy more expensive homes.
-Finally, using a variation of the event study in \cite{Chetty2007}, I show that while most households downsize their housing during unemployment spells, those with wealthy parents retain their current homes.
+I first present empirical evidence showing that parental wealth helps not only with the transition into ownership, but also in retention of ownership.
+In addition to replicating previous findings that transfers increase the probability of becoming a homeowner and that parental wealth is associated with sustained ownership \citep[see e.g.,][]{Lee2018,bond2021role}, I provide novel results on the mechanism through which parental wealth aids retention.
+Controlling for a wide set of household and parental characteristics, I show that households with wealthier parents are less likely to fall behind on mortgage payments.
+Finally, using a variation of the event study in \cite{Chetty2007}, I show that while households with less wealthy parents downsize their housing during unemployment spells, those with wealthy parents retain their current homes.
 
@@
 I find that parental transfers account for 14 percentage points (29\%) of the homeownership rate among young households.
-
-I then use the model to evaluate policies aiming to increase homeownership.
-I find that such policies disproportionately benefit households with wealthy parents.
-Relaxing mortgage constraints---for example, by lowering minimum down payments---strengthens the link between parental wealth and children's housing outcomes, since more households with wealthy parents are on the margin to become homeowners.
+In one adaption of my model, I quantify the contribution of parental transfers to the Black-White homeownership gap of 50\% among young adults, following the  methodology of \cite{Ashman2020} and \cite{aliprantis2022dynamics}.
+I find that parental transfers account for 7 percentage points (14\%) of the Black-White homeownership gap.
+
+The interaction between altruistic transfers and illiquid housing generates surprising implications for several widely used housing policy levers, many of which disproportionately benefit households with wealthy parents.
+Relaxing borrowing constraints---for example, by lowering minimum down payments---strengthens the link between parental wealth and children's housing outcomes, since more households with wealthy parents are on the margin to become homeowners.
 Nonmortgage policies that lower barriers to entry, such as reducing prices or purchase costs, increase homeownership for all households, though disproportionately among those with wealthier parents.
@@
 
-I then study how illiquidity affects the strategic behavior arising in the absence of commitment.
+I then study how illiquidity affects the strategic behavior in the absence of commitment.
 I find that 27\% of young households prefer illiquid housing---due to sales costs---to liquid housing.
 These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth.'' Lower liquid wealth implies higher marginal propensities to consume (MPC) and, hence, more parental transfers.
-As an illustrative example, consider a child who buys a house and brings no liquid wealth to the next period.
-If the parent does not transfer in the next period, the child will either face very low consumption (with a correspondingly high marginal value of wealth) or be forced to liquidate the house and incur sales costs.
+As an illustration, consider a child who buys a house and brings no liquid wealth to the next period.
+If the parent does not transfer in the next period, the child will either face very low consumption---implying  a high marginal value of wealth---or be forced to liquidate the house and incur sales costs.
 The parent dislikes both outcomes and may find it optimal to transfer enough to keep the child in the home.
-This driver of preference for illiquidity differs from those studied in the literature, such as temptation or time inconsistency \citep{attanasio2024temptation,laibson1997golden}.
-Furthermore, parental wealth generating preference for illiquidity provides a novel explanation for the existence of high-MPC households \cite[see e.g.,][]{kaplan2022marginal}.
-That children with wealthy parents find illiquidity less problematic---or even prefer it---is consistent with the empirical results in \cite{choukhmane2023benefits}.
-They find that households with wealthy parents are more likely to invest in illiquid retirement accounts and less likely to make costly early withdrawals, even after controlling for a wide set of characteristics.
-
-Finally, I quantify the contribution of parental transfers to the Black-White homeownership gap of 50\% among young adults.
-I adapt the methodology of \cite{Ashman2020} and \cite{aliprantis2022dynamics}, who study the racial wealth gap, and recalibrate the income process to capture racial differences.
-I find that parental transfers account for 7 percentage points (14\%) of the Black-White homeownership gap.
-For context, the Black-White homeownership gap has remained roughly constant since the passing of the Fair Housing Act in 1968.
+
+That parental wealth generates a preference for illiquidity is not only a new theoretical mechanism---distinct from explanations like temptation or time inconsistency \citep[e.g.,][]{attanasio2024temptation,laibson1997golden}---but one that helps explain the positive link between parental wealth and successfull homeownership documented earlier.
+It also aligns with \citet{choukhmane2023benefits}, who find that households with wealthy parents are more likely to invest in illiquid retirement accounts and avoid costly early withdrawals, even after controlling for a wide set of characteristics.
+
+I then examine the implications of transfers and housing on marginal propensities to consume (MPCs).
+First, the presence of altruistic transfers reduces average MPCs by nearly 30\%, suggesting that standard life-cycle models without parent-child interactions may overstate MPCs.
+Second, the preference for illiquidity induced by transfers offers a novel explanation for the existence of high-MPC households \cite[e.g.,][]{kaplan2022marginal}; in fact, the entire decline in average MPCs is driven by lower MPCs among the wealthy hand-to-mouth.
+Third, a common explanation for high MPCs is heterogeneity in preferences---such as impatience \citep[e.g.,][]{aguiar2024hand}---but such behavior is observationally similar to that of households with wealthy, altruistic parents.
 
@@
 
-\subsubsection{Households with Wealthier Parents Retain Ownership}
+\subsubsection{Households with Wealthier Parents Retain Ownership}\label{sec:maintaining}
 While most research focuses on the transition into ownership, I now focus on the challenge of retaining it.
@@
 \end{table}
-\subsubsection{Nontargeted Moments}\label{sec:externalval}
+\subsection{Nontargeted Moments}\label{sec:externalval}
 To validate the model, I show that it also matches nontargeted moments.
@@
 
-\begin{figure}
+\begin{figure}[tb]
 	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates_model}
@@
 
-Taking stock, this section shows that the model can match targeted moments regarding homeownership, transfers, and the transition to homeownership with realistic preference parameter values.
-The model also matches a set of untargeted moments.
-Perhaps more importantly, the model replicates the household-level associations between parental wealth and downsizing during income losses.
+\subsubsection{Parental Wealth and Ownership Retention}
+Next, I repeat the exercise in Section~\ref{sec:maintaining} (see Figure~\ref{fig:owner_hazard}) for first-time homebuyers, where we saw that (a) parental wealth is associated with more retention, (b) more so than own wealth, and (c) that the role of parental wealth inreases over time horizons.
+In the model, I replicate the regression as closely as possible.
+For example, there are no year effects or controls for family size.
+The model captures these patterns well, though it overpredicts the role of parental wealth at shorter horizons.
+
+\begin{figure}[tb]
+	\caption{Homeownership Retention Among First-Time Owners}\label{fig:remain_model}
+	\includegraphics[width=0.5\textwidth]{../tabfig/model_remain_own.pdf}%
+	\includegraphics[width=0.5\textwidth]{../tabfig/model_remain_parent.pdf}
+
+	{\begin{footnotesize} \textit{Notes:} Lines denote means, and shaded areas indicate 95\% confidence intervals.
+The light orange area denotes the empirical estimates (Section \ref{sec:maintaining}), and the darker grey area denotes the model simulation.
+The sample consists of households first-time owners aged 25-45.
+\end{footnotesize}}
+\end{figure}
+
+\subsubsection{Taking Stock on Model Fit}
+The model matches targeted moments related to homeownership, transfers, and rent-to-own transitions using realistic preference parameters.
+It also performs well on untargeted aggregate moments.
+Perhaps more importantly, it replicates key non-targetted household-level patterns in the associations between parental wealth, downsizing behavior, and ownership retention.
 
@@
 Removing altruism reduces the gap to 45\%, implying that parental transfers account for 7 percentage points---or 14\%---of the gap.
+This is a sizeable effect, especially given that the Black–White homeownership gap has remained largely unchanged since the passage of the Fair Housing Act in 1968.
 While Black and White households receive transfers at similar rates in the model, transfers are less influential for Black households, who are further from the ownership threshold and receive smaller amounts.
@@
 These results have several implications for the MPC literature.
-First, model-implied MPCs may be overstated when informal intra-family transfers are omitted, particularly for households with wealthy parents.
+First, model-implied MPCs may be overstated when intra-family transfers are omitted, particularly for households with wealthy parents.
 Second, a common strategy to match high MPCs is to introduce households with low target wealth—e.g., through impatience \citep{aguiar2024hand}—but such behavior is observationally similar to that of households expecting altruistic support.

### 2025-05-02 docs/Homeownership_BankOfMomAndDad.tex
More polish

@@
 While homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}, households face different barriers to ownership.
-Between 2009 and 2016, around 30\% of American first-time homebuyers received direct parental assistance for their down payment, with such assistance averaging \$48,000 per receiving household\textcolor{red}{Check numbers}.\footnote{All data from the Survey of Household Economic Decisionmaking and the Panel Study of Income Dynamics (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
+Between 2009 and 2016, around 30\% of American first-time homebuyers received direct parental assistance for their down payment.\footnote{Source: Survey of Household Economic Decisionmaking (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
 
@@
 I made several simplifying assumptions to keep the model tractable.
-Adding richer housing dynamics and equilibrium effects would allow one to study other interesting questions, such as the extent to which transfers contribute to house price growth.
-My model omits a feedback loop, in which transfers raise demand and, thus, house prices, boosting the wealth of homeowning parents and, thus, higher transfers, and so on.
-Another potential extension could examine whether illiquidity may reduce the commitment problem in the family.
-Additionally, my preferences-for-illiquidity results have implications for the literature on high-MPC households \citep[see e.g.,][]{kaplan2022marginal}.
-In particular, we are likely overestimating the average MPC of liquidity-constrained households with wealthy parents, since these households use parental transfers to smooth shocks.
-For example, following a negative shock, consumption may fall less due to transfers, while for positive shocks, consumption may rise less as parents reduce transfers.
+Adding richer housing dynamics and equilibrium effects would allow analysis of other interesting questions, such as the extent to which transfers contribute to house price growth.
+My model omits a feedback loop, in which transfers raise housing demand, which leads to higher prices and thus increasing the wealth of homeowning parents and enabling further transfers.
+Another potential extension could examine whether illiquidity reduces the commitment problem in the family.
+Additionally, my results on preferences for illiquidity have implications for the literature on high-MPC households \citep[see e.g.,][]{kaplan2022marginal}.
+In particular, by ignoring transfers, existing work overestimates the average MPC of liquidity-constrained households, some of whom rely on parental transfers to smooth shocks.
 

### 2025-05-02 src/Julia/run_all.jl
Fix scoping issue

@@ -20,14 +20,12 @@ store = true # true to store results to .pdf/.tex
 #######################
 ## Structural Estimation
 #######################
-
+par_lims = par_lims_ini()
 for shrinkstep = 0:Nshrinks # Shrink the search space iteralatively
-    if shrinkstep == 0 # Initialize search space
-        par_lims = par_lims_ini()
-    else # Shrink search space, based on the last Nest iterations
+    if shrinkstep > 0 # shrink search space
         ilo = 1 + Nest*(shrinkstep-1)
         ihi = Nest*shrinkstep
-        par_lims = shrink_searchspace(ilo,ihi,nstate,nchoice,par_lims);
+        global par_lims = shrink_searchspace(ilo,ihi,nstate,nchoice,par_lims);
     end
     GlobalSearch(par_lims, nstate,nchoice,targ_moms;nsim = (Nest*(shrinkstep+1)),start=(1+Nest*(shrinkstep)))
 end

### 2025-05-02 docs/Homeownership_BankOfMomAndDad.tex
Followed Will feedback

@@
 
-I then study how illiquidity affects the strategic behavior in the absence of commitment.
-I find that 27\% of young households prefer illiquid housing---due to sales costs---to liquid housing.
-These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth.'' Lower liquid wealth implies higher marginal propensities to consume (MPC) and, hence, more parental transfers.
+I then study how illiquidity affects strategic behavior in the absence of commitment.
 As an illustration, consider a child who buys a house and brings no liquid wealth to the next period.
-If the parent does not transfer in the next period, the child will either face very low consumption---implying  a high marginal value of wealth---or be forced to liquidate the house and incur sales costs.
+If the parent does not transfer in the next period, the child will either face very low consumption—implying a high marginal value of wealth—or be forced to liquidate the house and incur sales costs.
 The parent dislikes both outcomes and may find it optimal to transfer enough to keep the child in the home.
-
-That parental wealth generates a preference for illiquidity is not only a new theoretical mechanism---distinct from explanations like temptation or time inconsistency \citep[e.g.,][]{attanasio2024temptation,laibson1997golden}---but one that helps explain the positive link between parental wealth and successfull homeownership documented earlier.
+More precisely, children with wealthy parents use the sales costs as a commitment device to maintain low liquid wealth in the next period, raising their next-period marginal utility of wealth and thereby strengthening the parent’s transfer motive.
+As a result, 27\% of young households prefer illiquid housing over liquid housing.
+That parental wealth generates a preference for illiquidity is a new theoretical mechanism---distinct from existing explanations such as temptation or time inconsistency \citep[e.g.,][]{attanasio2024temptation,laibson1997golden}.
+
+This lower cost of illiquidity for households with wealthy parents offers a natural explanation for the positive empirical relationship between parental wealth and successful homeownership I found in the PSID.
 It also aligns with \citet{choukhmane2023benefits}, who find that households with wealthy parents are more likely to invest in illiquid retirement accounts and avoid costly early withdrawals, even after controlling for a wide set of characteristics.
 
-I then examine the implications of transfers and housing on marginal propensities to consume (MPCs).
-First, the presence of altruistic transfers reduces average MPCs by nearly 30\%, suggesting that standard life-cycle models without parent-child interactions may overstate MPCs.
-Second, the preference for illiquidity induced by transfers offers a novel explanation for the existence of high-MPC households \cite[e.g.,][]{kaplan2022marginal}; in fact, the entire decline in average MPCs is driven by lower MPCs among the wealthy hand-to-mouth.
-Third, a common explanation for high MPCs is heterogeneity in preferences---such as impatience \citep[e.g.,][]{aguiar2024hand}---but such behavior is observationally similar to that of households with wealthy, altruistic parents.
+I then examine the interplay  of transfers and housing for marginal propensities to consume (MPC).
+A substantial literature has sought to reconcile structural models with the high MPCs observed in empirical data.
+In my model, the introduction of altruistic transfers lowers average MPCs by about 30\%, implying that standard life-cycle models may understate the discrepancy.
+Moreover, one common explanation for high MPCs is that some households are wealthy but liquidity constrained---such as the ``wealthy hand-to-mouth'' in \cite{kaplan2022marginal}---but I find that ignoring parental transfers overstates their MPCs by about 50\%.
+Another explanation for high MPCs is heterogeneity in preferences---such as impatience \citep[e.g.,][]{aguiar2024hand}---yet such behavior is observationally similar to that of households with wealthy, altruistic parents.
 

### 2025-05-02 src/Julia/export_estimation.jl

@@ -232,6 +232,7 @@ function PlotGlobalEstimation(dat,est,pars,targ_moms;store=false)
     ### Plots the effect of each parameter
     for (ipar,par) in enumerate(pars)
         ptmp = deepcopy(plot(p2[ipar,:]...,plots_obj[ipar],titlefontsize=9,layout=(2,3),size=(800,300).*0.8))
+        display(ptmp)
         if store == true
             Plots.pdf(ptmp,"tabfig/est/identification/"*_parnames_s[pars[ipar]])
         end

### 2025-05-02 src/Julia/run_all.jl

@@ -32,7 +32,7 @@ end
 
 ## Plot estimation
 df_boot, globest,dat = LoadEstimation(nstate,nchoice,targ_moms,pars);
-PlotGlobalEstimation(dat,globest,pars,targ_moms;store=true)
+PlotGlobalEstimation(dat,globest,pars,targ_moms;store=store)
 
 #######################
 ## Model fit etc

### 2025-05-07 docs/Homeownership_BankOfMomAndDad.tex
New results (still need to update text though!)

@@
 	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates}
-	\noindent\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both}\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both}
+	\noindent%
+	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both}%
+	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both}
 
@@
 		\caption{Homeownership Policies on the role of Parental Wealth}\label{tab:policy}
-
-ratio
 		\begin{tabular}{@{}llll@{}}
@@
 		\caption{Altruistic Transfers Decrease the MPC}
-		\input{../tabfig/postest/MPCs}
+		\input{../tabfig/postest/MPCs.tex}
 	\end{threeparttable}

### 2025-05-07 src/HomeownershipBankMomDad.jl

@@ -16,6 +16,9 @@ using Dates # Allows you append todays date to filenames
 using StableRNGs
 using PanelDataTools
 using BenchmarkTools
+using ShiftedArrays
+using GLM
+using CategoricalArrays
 
 ## Load files
 include("Julia/struct.jl")

### 2025-05-07 src/Julia/run_all.jl

@@ -53,6 +53,7 @@ end
 output_esttable(df_boot,pars,targ_moms,Mb,store=store)
 output_extramoments(Mb,nontarg_moms,store=store)
 pChetty = ChettyEventStudy(Mb,0.75,store=store)
+pRetain = maintaining_ownership_plots(Mb,store=store)
 #######################
 ## Post-Estimation Analaysis - Impact of Altruism on Homeownership
 #######################

### 2025-05-14 docs/Homeownership_BankOfMomAndDad.tex
Improve table titile
Ready for submission

@@
 Without parental transfers, the model simplifies to a standard life-cycle model with housing.
-I find that parental transfers account for 14 percentage points (29\%) of the homeownership rate among young households.
+I find that parental transfers account for 13 percentage points (27\%) of the homeownership rate among young households.
 In one adaption of my model, I quantify the contribution of parental transfers to the Black-White homeownership gap of 50\% among young adults, following the  methodology of \cite{Ashman2020} and \cite{aliprantis2022dynamics}.
-I find that parental transfers account for 7 percentage points (14\%) of the Black-White homeownership gap.
+I find that parental transfers account for 11 percentage points (20\%) of the Black-White homeownership gap.
 
@@
 More precisely, children with wealthy parents use the sales costs as a commitment device to maintain low liquid wealth in the next period, raising their next-period marginal utility of wealth and thereby strengthening the parent’s transfer motive.
-As a result, 27\% of young households prefer illiquid housing over liquid housing.
+As a result, 22\% of young households prefer illiquid housing over liquid housing.
 That parental wealth generates a preference for illiquidity is a new theoretical mechanism---distinct from existing explanations such as temptation or time inconsistency \citep[e.g.,][]{attanasio2024temptation,laibson1997golden}.
@@
 A substantial literature has sought to reconcile structural models with the high MPCs observed in empirical data.
-In my model, the introduction of altruistic transfers lowers average MPCs by about 30\%, implying that standard life-cycle models may understate the discrepancy.
-Moreover, one common explanation for high MPCs is that some households are wealthy but liquidity constrained---such as the ``wealthy hand-to-mouth'' in \cite{kaplan2022marginal}---but I find that ignoring parental transfers overstates their MPCs by about 50\%.
+In my model, the introduction of altruistic transfers lowers average MPCs by about 20\%, implying that standard life-cycle models may understate the discrepancy.
+Moreover, one common explanation for high MPCs is that some households are wealthy but liquidity constrained---such as the ``wealthy hand-to-mouth'' in \cite{kaplan2022marginal}---but I find that ignoring parental transfers overstates their MPCs by about 30\%.
 Another explanation for high MPCs is heterogeneity in preferences---such as impatience \citep[e.g.,][]{aguiar2024hand}---yet such behavior is observationally similar to that of households with wealthy, altruistic parents.
@@
 Second, the effect of current parental wealth grows with the time horizon.
-For instance, a 10\% increase in parental wealth raises the probability of retaining ownership after eight years by 0.45 percentage points among first-time owners.
+For instance, a 10\% increase in parental wealth raises the probability of retaining ownership after eight years by about 0.45 percentage points among first-time owners.
 I return to these results later as part of the model validation.
@@
 In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$) With the full set of controls, the effect increases slightly, though it is no longer statistically significant.
-These magnitudes are economically meaningful since 1.6\% of first-time owners are behind on their mortgage.In Columns 3 to 5, I expand the analysis to include all owners who are currently not behind, increasing the sample sizes and allowing us to include household fixed effects.
+These magnitudes are economically meaningful since 1.6\% of first-time owners are behind on their mortgage.
+In Columns 3 to 5, I expand the analysis to include all owners who are currently not behind, increasing the sample sizes and allowing us to include household fixed effects.
 In these specifications, a 10\% increase in parental wealth is associated with a 0.09 percentage point decrease in delinquency risk in the univariate OLS regression ($p<0.001$), a 0.02 percentage point decrease in the multivariate OLS regression (not significant), and a 0.07 percentage point decrease in the fixed effects regression ($p<0.1$).
@@
 Among households with non-wealthy parents, housing consumption falls by about 5\% during unemployment ($p<0.05$).
-In contrast, households with wealthy parents see a slight increase (not significant)in housing consumption during unemployment.\footnote{In Appendix \ref{app:eventstudy_details}, I also report the results of a formal event study using the same set of controls as in the previous regressions.}  I return to these results later as part of the model validation.
+In contrast, households with wealthy parents see a slight increase (not significant) in housing consumption during unemployment.
+As shown in Appendix~\ref{app:eventstudy_details}, the main finding---that households with wealthy parents do not downsize during unemployment---is also retained in a formal event study that includes the same set of controls as the previous regressions.
+I return to these results later as part of the model validation.
 
@@
 Similarly, I estimate the homeownership utility $\chi$, since transfers generate new ownership motives.
-Finally, I estimate the minimum size of owner-occupied housing, $h_0$, a key parameter in models of homeownership, because housing choices interact with transfer behavior.
+Finally, I estimate the relative size of owner-occupied housing, $h_0$, a key parameter in models of homeownership, because housing choices interact with transfer behavior.
 
@@
 First, with a calibrated discount factor of $\beta=0.92$, the model matches the wealth levels of both young and old households.
-It also replicates the correlation between parental wealth and children's homeownership: in the PSID, young homeowners' parents are 2.5 times wealthier than young renters' parents, compared to about 2 times in the model.
+It also replicates the correlation between parental wealth and children's homeownership: in the PSID, young homeowners' parents are 2.6 times wealthier than young renters' parents, compared to about 2 times in the model.
 This parental wealth gradient effectively captures the intergenerational link in homeownership.
@@
 Second, the model aligns well with other moments related to homeownership.
-It slightly overshoots the age of first homeownership and overpredicts the aggregate homeownership rate of 65\%, primarily due to an excess of homeowner parents stemming from their lack of income risk (see Appendix \ref{sec:robust_incomerisk}).
+It slightly overshoots the age of first homeownership and overpredicts the aggregate homeownership rate of 64\%, primarily due to an excess of homeowner parents stemming from their lack of income risk (see Appendix \ref{sec:robust_incomerisk}).
 Mortgages are slightly smaller than in the data, as households repay and re-borrow flexibly.
@@
 Third, the model also matches transfer-related moments well.
-It predicts average parental transfers of 2.3\% of parental wealth, close to the 2.6\% estimated from the 2013 Transfer Module (excluding outliers below 0\% or above 100\%).
-It accurately captures the share of households receiving transfers within two years of house purchases, suggesting proper synchronization of housing transactions and transfers.
+It correctly predicts average parental transfers of 2.6\% of parental wealth, as estimated from the 2013 Transfer Module (excluding outliers below 0\% or above 100\%).
+It roughly matches the share of households receiving transfers within two years of house purchases, suggesting proper synchronization of housing transactions and transfers.
 Renters also receive transfers more often than owners in both the model and the data, though the model slightly overstates this gap.
@@
 Table \ref{tab:modelfitnoa} reports how the moments change without altruism and transfers.
-Homeownership rates decline by 14 percentage points from 49\%, a decrease of 29\% for households aged 25 to 44.
+Homeownership rates decline by 13 percentage points from 48\%, a decrease of 27\% for households aged 25 to 44.
 This is a large decrease, since households endogenously choose to save more without transfers.
@@
 
-The reduction in homeownership cannot be attributed to an inability to afford buying: Without altruism, median wealth of the young increases (increasing homeownership, in isolation) to the average wealth level at purchase with altruism.
+The reduction in homeownership cannot be attributed to an inability to afford buying: Without altruism, median wealth of the young increases (increasing homeownership, in isolation).
 The main driver is that the ownership threshold shifts out in wealth without altruism, as we saw in the policy functions (Section \ref{sec:policyfuncs}).
-Without transfers then, the average wealth at purchase increases by about \$12,500 (about 33\%), the age of first ownership is delayed by about 2.5 years, and the LTV at purchase falls by 13 percentage points.
-Although homeownership of young households declines sharply, the overall homeownership rate of all households (ages 25 to 74) only falls by 6 percentage points (8\%).
+Without transfers then, the average wealth at purchase increases by about 25\%, the age of first ownership is delayed by about 2.5 years, and the LTV at purchase falls by 12 percentage points.
+Although homeownership of young households declines sharply, the overall homeownership rate of all households (ages 25 to 74) only falls by 6 percentage points (9\%).
 Hence, the main aggregate effect of parental transfers is that they induce earlier homeownership.
-Finally, without altruism, the parental wealth gradient in homeownership decreases from 1.94 to 1.14, with the remaining effect driven by the integenerational correlation in initial wealth and persistent labor productivity.
+Finally, without altruism, the parental wealth gradient in homeownership decreases from 2.0 to 1.15, with the remaining effect driven by the integenerational correlation in wealth and income.
+Hence, the positive correlation between parental wealth and housing outcomes is almost entirely driven by parental transfers.
 
@@
 		%			\footnotesize
-		\caption{Black-White Homeownership Rate}\label{tab:BW_table}
+		\caption{Altruism and the Black-White Homeownership Gap}\label{tab:BW_table}
 		\input{../tabfig/postest/BW_mostrecent.tex}
@@
 
-The observed Black-White homeownership gap is $\frac{0.54 - 0.27}{0.54} = 50\%$, compared to 52\% in the model with altruism.
-Removing altruism reduces the gap to 45\%, implying that parental transfers account for 7 percentage points---or 14\%---of the gap.
+The observed Black-White homeownership gap is $\frac{0.53 - 0.26}{0.53} = 51\%$, compared to 54\% in the model with altruism.
+Removing altruism reduces the gap to 43\%, implying that parental transfers account for 11 percentage points---or 20\%---of the gap.
 This is a sizeable effect, especially given that the Black–White homeownership gap has remained largely unchanged since the passage of the Fair Housing Act in 1968.
@@
 This is due to the small difference in homeownership rates (ages 25 to 74) with and without altruism.
-Prices decline by just 7\% even with a supply elasticity as low as 1.
+Prices decline by just 5\% even with a supply elasticity as low as 1.
 Thus, transfers account for between 10 and 13 percentage points of the homeownership rate, depending on the supply elasticity.
@@
 \textit{Relaxing Mortgaging Regulation:} To study how parental transfers interact with down payment requirements, I tighten the LTV limit by 3 percentage points.
-The tighter limit decreases homeownership among the young by 5 percentage points (10\%).
+The tighter limit decreases homeownership among the young by 7 percentage points (15\%).
 Homeownership declines more among households with parents in the middle of the wealth distribution, reducing the parental wealth gradient.
@@
 \textit{Lowering PMI:} To study how parental transfers interact with the PMI, which increases the interest rate on high LTV loans, I relax the PMI limit by 7 percentage points.
-The relaxed PMI limit increases homeownership by 2 percentage points (4\%).
+The relaxed PMI limit increases homeownership by 1 percentage points (2\%).
 Similar to the LTV regulation, homeownership among households with parents in the middle of the wealth distribution rises more, further widening the parental wealth gradient.
@@
 
-\textit{Lowering House Prices:} To study how parental transfers interact with housing affordability, I lower house prices by \$6,000 (7\%).
-As prices decline, homeownership increases for all households, though more so for households with the richest parents, slightly increasing the role of parental wealth.
-
-\textit{Reducing Purchase Costs:} To study how closing costs and first-time buyer assistance interact with parental transfers, I lower the purchase cost ($m_b$) by 2 percentage points---equivalent to a grant of about \$6,250.\footnote{Many U.S.
+\textit{Lowering House Prices:} To study how parental transfers interact with housing affordability, I lower house prices by 2\%.
+As prices decline, homeownership increases for all households, though more so for households with the richest parents, increasing the role of parental wealth.
+
+\textit{Reducing Purchase Costs:} To study how closing costs and first-time buyer assistance interact with parental transfers, I lower the purchase cost ($m_b$) by 2 percentage points---equivalent to a grant of about \$6,000.\footnote{Many U.S.
 states offer closing cost assistance to first-time buyers.
@@
 In contrast, for household with non-wealthy parents, lower purchase costs reduce the effective price—the main remaining barrier to ownership.
-This suggests that grants or credits for first-time buyers primarily benefit households of the wealthiest.
+This suggests that grants or credits for first-time buyers primarily benefit households with richer parents.
 
@@
 The earlier empirical results suggested that parental wealth helps offset the illiquidity of housing.
-To study this interaction, I lower the sales cost $m_s$ from \parms\% to 5.
-Strikingly, the homeownership rate decreases by 2 percentage points (4\%).
+To study this interaction, I lower the sales cost $m_s$ by 2 percentage points.
+Strikingly, the homeownership rate decreases by 3 percentage points (6\%).
 This counterintuitive result reflects selection into ownership.
-homeownership rises slightly among households with the wealthiest and poorest parents but drops sharply---by 9 percentage points---among households with moderately wealthy parents.
+Homeownership rises slightly for most households but drops sharply---by 6 percentage points---among households with moderately wealthy parents.
 For these households, lower sales costs weaken the ability to use housing as a commitment device to secure future transfers.
@@
 The key determinant is parental wealth: Parents of children who prefer sales costs are approximately three times wealthier.
-Children who prefer sales costs are also more likely to receive transfers (48\% vs.
-13\%) and, conditional on receiving transfers, receive three times larger transfers.
+Children who prefer sales costs are also more likely to receive transfers (51\% vs.
+15\%) and, conditional on receiving transfers, receive three times larger transfers.
 Notably, households that own their homes are more likely to prefer sales costs on their housing.
@@
 To explore why some households prefer illiquidity, Figure~\ref{fig:prefliq} breaks the sample by liquidity preference and age.
-Figure~\ref{fig:pref} shows that illiquidity preference declines with age, consistent with reduced reliance on parental transfers; around 50\% of the youngest households prefer illiquidity.
+Figure~\ref{fig:pref} shows that illiquidity preference declines with age, consistent with reduced reliance on parental transfers; around 40\% of the youngest households prefer illiquidity.
 The next panels plot homeownership, parental wealth, and child wealth by liquidity preference.
@@
 We begin by computing counterfactual MPCs for households in the stationary distribution, assuming no altruistic transfers.
-The average MPC of young households is 0.35.
-The MPC of liquidity constrained households (0.44) is about three times that of unconstrained households (0.16).
+The average MPC of young households is 0.32.
+The MPC of liquidity constrained households (0.43) is about three times that of unconstrained households (0.14).
 While these MPCs are lower than empirical estimates---typically ranging from 0.5 to 0.9---they are relatively standard in models not designed to generate high MPCs.
@@
 	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}
-		\textit{Notes:} Sample includes all households aged 25-44 in the stationary distribution.
+		\textit{Notes:} The sample includes all households aged 25-44 in the stationary distribution.
 Following \cite{Kaplan2014}, households are liquidity constrained if they have negative liquid wealth $(b_k)$ exceeding 16.5\% of annual earnings.
-The MPC is calculated from a \$5,000 positive shock.\end{flushleft}\end{footnotesize}}
+The MPC is calculated from a \$5,000 positive shock.
+To calculate the MPCs with future transfers only, we use the policy functions with altruism (baseline) but simulate households after unexpectedly confiscating any transfers.
+To calculate the MPCs with current transfers only, we use the parental policy function with altruism (baseline) and the children’s policy function without altruism.
+The children then receive the transfers but behave as if there are no future transfers.
+\end{flushleft}\end{footnotesize}}
 \end{table}
@@
 
-Using a counterfactual experiment without altruism (a standard life-cycle model), I find that transfers account for 14 percentage points (29\%) of the homeownership rate.
+Using a counterfactual experiment without altruism (a standard life-cycle model), I find that transfers account for 13 percentage points (27\%) of the homeownership rate.
 I show that policies that lower entry barriers to homeownership amplify the importance of parental wealth on housing outcomes, while those that lower the downsides of illiquid homeownership reduce the importance of parental wealth.
@@
 It also induces households to buy housing slightly later, with lower LTV's and higher wealth at purchase.
-However, the main quantitative finding remains: Parental transfers now account for 11pp (25\%) of the homeownership rate.
-I have also perform robustness checks where I allow for transitory idiosyncratic dividend/return shocks as in \cite{Chang2024} for homeowners, and find that, as expected, idiosyncratic price risk have even smaller quantitative effects.
+However, the main quantitative finding remains: Parental transfers now account for 10pp (23\%) of the homeownership rate.
+I have also perform robustness checks where I allow for transitory idiosyncratic dividend/return shocks as in \cite{Chang2024} for homeowners, and find that, as expected, idiosyncratic transitory price risk have even smaller effects.
 
@@
 The value and policy functions are linearly interpolated over child and parental wealth, conditional on the discrete choices.
-This solution algorithm satisfies both criteria established by \cite[p.30]{Barczyk2020}: ``algorithms should be able to deal with locally convex and even discontinuous value functions''.
-Naturally, the solution accuracy improves with denser grids.
+This solution algorithm satisfies both criteria established by \cite[p.30]{Barczyk2020}: ``algorithms should be able to deal with locally convex and even discontinuous value functions.'' Naturally, the solution accuracy improves with denser grids.
 I use {\parnstate} nodes in the state vectors and {\parnchoice} nodes in choice variable grids.
@@
 The cycle then repeats, with the child buying in one iteration and not buying in the next.
-When I solve the model without homeownership (reducing it to a standard altruistic consumption-savings model as in \cite{Barczyk2020a} and \cite{Chu2020}), the value function iterations consistently converge to a unique fixed point.
+When I solve the model without homeownership (reducing it to a standard altruistic consumption-savings model as in \cite{Barczyk2020a} and \cite{Chu2020}), the value functions always converge to a unique fixed point.
 Similarly, removing transaction costs, making housing more liquid and no longer a state variable, reduces the likelihood of cyclical fixed points.

### 2025-05-14 src/Julia/export_estimation.jl

@@ -4,7 +4,7 @@ function varnames()
     :medwealthold =>"Median Wealth (55-74)",
     :medwealthfyoung =>"Median Family Wealth (25-44)",
     :wealthatpurchase => "Wealth at Purchase (25-44)",
-    :transfrateyoung => "Transfer Rate (55-74)",
+    :transfrateyoung => "Transfer Receipt Rate (25-44)",
     :h2wyoung =>"House Value / Wealth (25-44)",
     :r2incyoung =>"Rent / Income (25-44)",
     :wealthatpurchaseyoung => "Wealth at Purchase (25-44)",
@@ -15,8 +15,8 @@ function varnames()
     :hand2mouthyoung  => "Hand to Mouth (25-44)",
     :mortg2incyoung => "Mortgage / Income (25-44)",
     :transfsizeyoung => "Transfer Size (55-74)",
-    :ownerall => "Owner (25-73)",
-    :ownerold => "Owner (55-73)",
+    :ownerall => "Owner (25-74)",
+    :ownerold => "Owner (55-74)",
     :price => "House Price",
     :α1 => "Supply Elasticity (\$\\alpha_1\$)",
     :α0 => "Supply El. Intercept",

### 2025-05-14 src/Julia/policyplots.jl

@@ -6,11 +6,11 @@ function plot_policyfuncs(Mb,Mn;store=false)
     bph = 400
     ibph = argmin(abs.(np.bp_grd.-bph))
 
-    bpl = 0
+    bpl = 100
     ibpl = argmin(abs.(np.bp_grd.-bpl))
 
     ihp = 1
-    ivk = 3
+    ivk = 2
     iok = 1
     is = 1
     iak = 1

### 2025-05-15 test/runtests.jl
Fix tests (numbers are new since income grid is new)

@@ -13,8 +13,8 @@ using Test
         Mb = Model(mp,np)
         solvemodel2nd!(Mb,np,mp)
 
-        @test Mb.moms[Mb.moms.type.=="Model",:owneryoung][1] == 0.413
-        @test Mb.simpan.xp[15] == 78.40585097145825
+        @test Mb.moms[Mb.moms.type.=="Model",:owneryoung][1] == 0.424
+        @test Mb.simpan.xp[15] == 76.30713470213524
         @test Mb.g.gp.t[1,10,3,2,2,1,14,1]  == 357.96712339010395
     end
 
@@ -28,6 +28,6 @@ using Test
 
         df_boot, globest,dat = LoadEstimation(nstate,nchoice,targ_moms,pars);
 
-        @test globest.smm_obj[1]== 0.7092005784698485 
+        @test globest.smm_obj[1]== 0.7634896057824219 
     end
 end

### 2025-05-15 docs/Homeownership_BankOfMomAndDad.tex
Post-QJE Revision (#11)
* Update data through 2021 (not, code not return, but model implications are tiny!)

* Show less subgroups for the summary statistics table

* Include more and better transfer/inheritance data

* Add new empirical results on ownership retention

* Added more non-targeted moments

* Updated text 
---------

Co-authored-by: Eirik Eylands Brandsaas <eirik.e.brandsaas@frb.gov>

@@
 In contrast, households with wealthy parents see a slight increase (not significant) in housing consumption during unemployment.
-As shown in Appendix~\ref{app:eventstudy_details}, the main finding---that households with wealthy parents do not downsize during unemployment---is also retained in a formal event study that includes the same set of controls as the previous regressions.
+Appendix~\ref{app:eventstudy_details} shows that these patterns are robust to an event-study specification that controls for household characteristics.
 I return to these results later as part of the model validation.

### 2025-05-15 src/HomeownershipBankMomDad.jl

@@ -16,6 +16,9 @@ using Dates # Allows you append todays date to filenames
 using StableRNGs
 using PanelDataTools
 using BenchmarkTools
+using ShiftedArrays
+using GLM
+using CategoricalArrays
 
 ## Load files
 include("Julia/struct.jl")
@@ -33,5 +36,6 @@ include("Julia/quant_endoprice.jl")
 include("Julia/quant_blackwhite.jl")
 include("Julia/quant_preferadjustment.jl")
 include("Julia/quant_robustness.jl")
+include("Julia/quant_MPC.jl")
 
 include("Julia/postest_plots.jl")

### 2025-05-15 src/Julia/estimation.jl

@@ -126,7 +126,7 @@ function shrink_searchspace(firstiter,lastiter,nstate,nchoice,par_lims)
     @printf("η : (%1.3f,%1.3f) vs (%1.3f,%1.3f). Size ratio of new over old %1.3f \n",par_lims_n.η[1][1],par_lims_n.η[1][2],par_lims.η[1][1],par_lims.η[1][2], ηr)
     @printf("ho: (%1.3f,%1.3f) vs (%1.3f,%1.3f). Size ratio of new over old %1.3f \n",par_lims_n.ho[1][1],par_lims_n.ho[1][2],par_lims.ho[1][1],par_lims.ho[1][2], hor)
     @printf("χ : (%1.3f,%1.3f) vs (%1.3f,%1.3f). Size ratio of new over old %1.3f \n",par_lims_n.χ[1][1],par_lims_n.χ[1][2],par_lims.χ[1][1],par_lims.χ[1][2],χr)
-    @printf("Total reduction in search area: %1.3f \n",ηr*hor*χr)
+    @printf("Total reduction in search area: %1.3f percent  \n",(1 - ηr*hor*χr)*100)
 
     return par_lims_n
 end

### 2025-05-15 src/Julia/export_estimation.jl

@@ -4,19 +4,19 @@ function varnames()
     :medwealthold =>"Median Wealth (55-74)",
     :medwealthfyoung =>"Median Family Wealth (25-44)",
     :wealthatpurchase => "Wealth at Purchase (25-44)",
-    :transfrateyoung => "Transfer Rate (55-74)",
+    :transfrateyoung => "Transfer Receipt Rate (25-44)",
     :h2wyoung =>"House Value / Wealth (25-44)",
     :r2incyoung =>"Rent / Income (25-44)",
     :wealthatpurchaseyoung => "Wealth at Purchase (25-44)",
     :first_ownyoung => "Age First Own (25-44)",
     :LTVatpurchaseyoung => "LTV at Purchase (25-44)",
-    :tp2wealthpyoung => "Transfers / Parental Wealth (55-74)",
+    :tp2wealthpyoung => "Transfers / Parental Wealth (55-74) (\\%)",
     :transferbuyersyoung => "Transfers Around Purchase (25-44)",
     :hand2mouthyoung  => "Hand to Mouth (25-44)",
     :mortg2incyoung => "Mortgage / Income (25-44)",
     :transfsizeyoung => "Transfer Size (55-74)",
-    :ownerall => "Owner (25-73)",
-    :ownerold => "Owner (55-73)",
+    :ownerall => "Owner (25-74)",
+    :ownerold => "Owner (55-74)",
     :price => "House Price",
     :α1 => "Supply Elasticity (\$\\alpha_1\$)",
     :α0 => "Supply El. Intercept",
@@ -36,6 +36,8 @@ function varnames()
     :owneryoung_ppoor => "Owner (25-44), Bottom 33\\%",
     :transfrateyoung_prich => "Transfers (25-44), Parent Top 50\\%",
     :transfrateyoung_ppoor => "Transfers (25-44), Parent Bot 50\\%",
+    :transfraterenteryoung => "Transfers (25-44), Renters",
+    :transfrateowneryoung => "Transfers (25-44), Owners",
     :smm_obj => "SMM Objective Function",
     )
 end
@@ -230,6 +232,7 @@ function PlotGlobalEstimation(dat,est,pars,targ_moms;store=false)
     ### Plots the effect of each parameter
     for (ipar,par) in enumerate(pars)
         ptmp = deepcopy(plot(p2[ipar,:]...,plots_obj[ipar],titlefontsize=9,layout=(2,3),size=(800,300).*0.8))
[…]

### 2025-05-15 src/Julia/policyplots.jl

@@ -6,18 +6,18 @@ function plot_policyfuncs(Mb,Mn;store=false)
     bph = 400
     ibph = argmin(abs.(np.bp_grd.-bph))
 
-    bpl = 0
+    bpl = 100
     ibpl = argmin(abs.(np.bp_grd.-bpl))
 
     ihp = 1
-    ivk = 3
+    ivk = 2
     iok = 1
     is = 1
     iak = 1
     ihk = 2
     _h2size = Dict(1=>L"h_r",2=>L"h_o")
 
-    plot(title="Child's Bond and Housing Choice $(L"b{\prime}_c,h{\prime}_c")",xlabel="Child Wealth $(L"x_c")")
+    plot(xlabel="Child Cash-on-Hand $(L"\tilde x_c")",ylabel="Child Bond Choice $(L"b'_c")")
 
     iown = findfirst(x->x==true,Mn.g.gk.disc[:,ibpl,ihp,ivk,iok,is,iak,ihk].==1)
     plot!(np.xk_grd[1:iown],Mn.g.gk.b[1:iown,ibpl,ihp,ivk,iok,is,iak,1] ,label="No Altruism",marker=:circle,color=:black,linewidth=2)
@@ -40,7 +40,7 @@ function plot_policyfuncs(Mb,Mn;store=false)
     if store == true
         Plots.pdf(p2,"tabfig/model/pol_bond")
         open("tabfig/model/pol_bond_state.tex", "w") do file
-            write(file, "$(L"h'_p")=$(_h2size[ihp]), $(L"y_c")=$(round(np.v_grd[iak,ivk];digits=1)), $(L"h_c")=$(_h2size[ihk]), and $(L"a_c")=$(round(Int,np.age_grd[iak])). The poor and rich parent have $(L"b_p")=$(round(Int,bpl)) and $(L"b_p")=$(round(Int,bph))" )
+            write(file, "$(L"h'_p")=$(_h2size[ihp]), $(L"y_c")=$(round(np.v_grd[iak,ivk];digits=1)), $(L"h_c")=$(_h2size[ihk]), and $(L"a_c")=$(round(Int,np.age_grd[iak])). The non-rich and rich parent have $(L"b'_p")=$(round(Int,bpl)) and $(L"b'_p")=$(round(Int,bph))" )
         end
 
     

### 2025-05-15 src/Julia/postest_plots.jl

@@ -77,3 +77,153 @@ function ChettyEventStudy(M,wealthypar_lim::Float64;store=false)
 
     return plot(p1,p2)
 end
+
+
+## Maintain ownership
+function find_firsttime_owner!(df)
+    # Find first-time owners
+    transform!(groupby(df, :id),
+    :own => (v -> begin
+        idx = findfirst(identity, v)     # first TRUE in that household
+        out  = falses(length(v))
+        if !isnothing(idx)               # mark only that one row
+            out[idx] = true
+        end
+        out
+    end) => :first_own)
+end
+
+function maintaining_ownership_plots(M;store=false)
+    
+    np = M.np
+    df = deepcopy(M.simpan)
+    @warn "Not obvious what is right here... Should it be: you count as owner the period you own (like I do now) or should it be the period you wake up as owner?"
+    df.own = df.ihk.==2
+
+    
+    df = df[:,[:id,:iak,:own, :xk, :xp, :ivk, :iop, :iok,:iν]]
+    find_firsttime_owner!(df)
+    df.wk .= NaN64
+    df.wp .= NaN64
+    for i = 1:nrow(df)
+        iak = df.iak[i]::Int
+        ivk = df.ivk[i]::Int
+        iν = df.iν[i]::Int
+        df.wk[i] = wk(iak,ivk,np)
+        df.wp[i] = wp(iak+np.nyoung,iν,np)
+    end
+
+    #
+    maxlead = 15
+    dff = DataFrame(Lead=Int[],first=Float64[],all=Float64[])
+    df.age=Int.(np.age_grd[df.iak])
+    for F in 0:maxlead
+        # 1.  Create a lead‑F version of :own within each household
+        leadcol = Symbol("Fown", F)
+        transform!(groupby(df, :id), :own => (v -> ShiftedArrays.lag(v, -F)) => leadcol)
+    end
+
+    agelo = 25
+    agehi = 44
+    for F = 0:maxlead
+        _mfirst = mean(skipmissing(df[(df.first_own.==true) .&& (agelo .<= df.age .<= agehi),Symbol(:Fown,F)]))
+        _mall = mean(skipmissing(df[(df.own.==true)  .&& (agelo .<= df.age .<= agehi),Symbol(:Fown,F)]))
[…]

### 2025-05-15 src/Julia/quant_MPC.jl

@@ -1,0 +1,184 @@
+## Find MPCs
+function find_choices(ivk,iok,iop,is,iak,xk,xp,np,mp,gdiscp,gtp,gcp,gdisck,gck)
+    ihp = Int(gdiscp[ivk,iok,iop,is,iak,1](xk,xp) <= gdiscp[ivk,iok,iop,is,iak,2](xk,xp) ) + 1
+    hp = np.h_grd[ihp]
+    ownp = isowner(ihp,np)
+    tp = gtp[ivk,iok,iop,is,iak,ihp](xk,xp)*(1.0 - mp.τt )
+    cp = gcp[ivk,iok,iop,is,iak,ihp](xk,xp)
+    bp = BCp_netshock(xp,cp,tp,hp,ownp,iak+np.nyoung,iop,np.p_grd[is],np,mp) 
+    cp, bp =  enforce_feasibleborrowing(cp,bp,ihp,is,iak,mp,np)
+    
+    xkcash = xk + tp
+    ihk = Int(gdisck[ihp,ivk,iok,is,iak,1](xkcash,bp) <gdisck[ihp,ivk,iok,is,iak,2](xkcash,bp)) + 1
+    hk = np.h_grd[ihk]
+    ownk = isowner(ihk,np)
+
+    ##
+    
+    
+    xkcash = xk + tp
+    ck = gck[ihp,ivk,iok,is,iak,ihk](xkcash,bp)
+    bk = BCk(xkcash,ck,0.0,hk,ownk,iak,iok,np.p_grd[is],np,mp,ivk)
+    ck,bk = enforce_feasibleborrowing(ck,bk,ihk,is,iak,mp,np)
+    
+    return ihp,hp,cp,bp,tp,ihk,hk,ck,bk
+end
+
+function find_mpcs(M,Δ)
+    np = M.np
+    mp = M.mp
+    df = deepcopy(M.simpan)
+
+    gtp,gcp,gdiscp = itp_gpage(np,M.g.gp.t,M.g.gp.c,M.g.gp.disc);
+    gck,gdisck = itp_gkage(np,M.g.gk.c,M.g.gk.disc);
+
+
+    mpc = fill(NaN64,nrow(df))
+    mpt = fill(NaN64,nrow(df))
+    
+    for i = 1:nrow(df)
+        dfr = df[i,:]
+        ivk = dfr.ivk
+        iok = dfr.iok
+        iop = dfr.iop
+        is = dfr.is
+        iak = dfr.iak
+
+        xk = dfr.xk
[…]

### 2025-05-15 src/Julia/run_all.jl

@@ -8,7 +8,7 @@ include("../HomeownershipBankMomDad.jl")
 ## Set "parameters"
 pars = [:η, :χ, :ho] # Estimated parameters
 targ_moms =  [:owneryoung, :r2incyoung, :wealthatpurchaseyoung,:transfrateyoung]
-nontarg_moms = [:medwealthyoung, :medwealthold, :medwealthpgradyoung, :first_ownyoung, :ownerall, :mortgyoung,:LTVatpurchaseyoung,:transferbuyersyoung,]
+nontarg_moms = [:medwealthyoung, :medwealthold, :medwealthpgradyoung, :first_ownyoung, :ownerall, :mortgyoung,:LTVatpurchaseyoung, :tp2wealthpyoung, :transferbuyersyoung,:transfraterenteryoung,:transfrateowneryoung]
 nstate = 65 # Grid points in state
 nchoice = 145 # Grid points in choice
 Nest = 1500 # How many iterations in each step of the estimation
@@ -20,21 +20,19 @@ store = true # true to store results to .pdf/.tex
 #######################
 ## Structural Estimation
 #######################
-
+par_lims = par_lims_ini()
 for shrinkstep = 0:Nshrinks # Shrink the search space iteralatively
-    if shrinkstep == 0 # Initialize search space
-        par_lims = par_lims_ini()
-    else # Shrink search space, based on the last Nest iterations
+    if shrinkstep > 0 # shrink search space
         ilo = 1 + Nest*(shrinkstep-1)
         ihi = Nest*shrinkstep
-        par_lims = shrink_searchspace(ilo,ihi,nstate,nchoice,par_lims);
+        global par_lims = shrink_searchspace(ilo,ihi,nstate,nchoice,par_lims);
     end
     GlobalSearch(par_lims, nstate,nchoice,targ_moms;nsim = (Nest*(shrinkstep+1)),start=(1+Nest*(shrinkstep)))
 end
 
 ## Plot estimation
 df_boot, globest,dat = LoadEstimation(nstate,nchoice,targ_moms,pars);
-PlotGlobalEstimation(dat,globest,pars,targ_moms;store=true)
+PlotGlobalEstimation(dat,globest,pars,targ_moms;store=store)
 
 #######################
 ## Model fit etc
@@ -55,12 +53,14 @@ end
 output_esttable(df_boot,pars,targ_moms,Mb,store=store)
[…]

### 2025-05-15 src/Julia/various_functions.jl

@@ -138,6 +138,22 @@ function find_moments(np::NumPar,mp::ModPar,pan,period::String)
     smom[!,:transfrateyoung_pmiddle] .= mean(dfy.tp[(dfy.middlep .== 1),:] .> 0)
     smom[!,:transfrateyoung_ppoor] .= mean(dfy.tp[(dfy.poorp .== 1),:] .> 0)
 
+    ## Calculate a wealth percentiles, ownership by income groups age 35-36, and transferrate by ownership
+    df35 = dfy[dfy.age.==35,:]
+    q = quantile(df35.xk, [0.10, 0.25, 0.5, 0.75, 0.90]) # Wealth percentiles
+    smom[!,:wealth_age3536_p10young] .= q[1]
+    smom[!,:wealth_age3536_p25young] .= q[2]
+    smom[!,:wealth_age3536_p50young] .= q[3]
+    smom[!,:wealth_age3536_p75young] .= q[4]
+    smom[!,:wealth_age3536_p90young] .= q[5]
+    # Ownership by income levels
+    smom[!,:owner_age3536_inc1young] .= mean(df35[df35.ivk.==1,:iok])-1  # subtract 1 since iok =1 if renting and =2 if owning
+    smom[!,:owner_age3536_inc2young] .= mean(df35[df35.ivk.==2,:iok])-1 
+    smom[!,:owner_age3536_inc3young] .= mean(df35[df35.ivk.==3,:iok])-1 
+    # Transfer rate by ownership
+    smom[!,:transfraterenteryoung] .= mean(dfy[dfy.iok.==1,:tp] .> 0)
+    smom[!,:transfrateowneryoung] .= mean(dfy[dfy.iok.==2,:tp] .> 0)
+
     df_owners = df_owners[!,[:iak, :id, :ownk, :first_own, :tp]]
     df_owners = df_lag(df_owners,:tp,:iak,:id,forward=true)
     df_owners = df_lag(df_owners,:tp,:iak,:id,forward=false)

### 2025-08-14 docs/Homeownership_BankOfMomAndDad.tex
Close #10

@@
 The MPC of liquidity constrained households (0.43) is about three times that of unconstrained households (0.14).
-While these MPCs are lower than empirical estimates---typically ranging from 0.5 to 0.9---they are relatively standard in models not designed to generate high MPCs.
+While these MPCs are lower than empirical estimates---typically ranging from 0.3 to 0.9---they are relatively standard in models not designed to generate high MPCs.
 

### 2025-08-14 docs/Homeownership_BankOfMomAndDad.tex
Close #12

@@
 My results highlight that policies increasing ownership through subsidies to first-time buyers---a common approach in many cities---will lead to increased housing inequality.
-My model-based results also complement some of the empirical research on the effect of parental resources and housing outcomes \citep[see e.g.,][ for recent work]{wold2024housing,daysal2023intergenerational,benetton2022dynastic} by quantifying the contribution of transfers to homeownership.
+My model-based results also complement some of the empirical research on the effect of parental resources and housing outcomes \citep[see e.g.,][ for recent work]{wold2024housing,daysal2023intergenerational,benetton2022dynastic,boileau2023drives} by quantifying the contribution of transfers to homeownership.
 

### 2025-08-14 docs/Homeownership_BankOfMomAndDad.tex
Close #13

@@
 
-\textit{Definition of Transfers:} To measure the transfer receipt rate, I rely on the transfer supplement which asked all parent households whether they gave money, gifts, or loans of \$100 or more to each child in 2012.
+\textit{Definition of Transfers:} To measure the transfer receipt rate, I rely on the transfer supplement which asked all parents whether they gave money, gifts, or loans of \$100 or more to each child in 2012.
 I follow the literature \citep[e.g.,][]{mcgarry2016dynamic} and treat all transfers as gifts.
-Since this paper focuses on transfers that (a) relate to housing and (b) are quantitatively meaningful, I ignore transfers below \$500.
-About 25\% of transfers are below this threshold, and ignoring them increases the conditional mean transfer from \$2,921 to \$3,960 and reduces the transfer receipt rate from 45\% to 22\%.\footnote{These transfer rates are relatively high and exceed those reported by \citealt{feiveson2019lifecycle} using the Survey of Consumer Finances (SCF).
-The PSID transfer supplement likely captures a broader range of transfers than the SCF, which deliberately focuses on substantial transfers.
-In addition, the SCF asks about transfers further back in time, increasing the risk of recall error and underreporting.}
+I ignore transfers below \$500 to focus on transfers that are quantitatively meaningful for housing.
+About 25\% of transfers are below this threshold, and ignoring them increases the conditional mean transfer from \$2,921 to \$3,960 and reduces the transfer receipt rate from 45\% to 22\%.
+It is unsurprising that PSID-based transfer rates exceed those in the SCF \citep[e.g.,][]{feiveson2019lifecycle,sabelhaus2024household}, which focuses on substantial transfers.
+In the United States, inter-vivos transfers of the magnitudes observed here are far below the federal gift tax limits and are therefore untaxed for both giver and recipient.
+In 2024, gifts of up to \$18{,}000 per recipient require no filing; gifts above that require a gift-tax return, but tax is owed only if cumulative lifetime gifts and bequests exceed \$13.61 million.
 

### 2025-08-15 docs/Homeownership_BankOfMomAndDad.tex
Ready for third submission!

@@
 
+\usepackage[skip=4pt]{caption}    % space between caption and body (default ~10pt)
+\setlength{\textfloatsep}{10pt}    % text↔float (above & below)  [default ~20pt]
+% \setlength{\intextsep}{12pt}       % inline floats (here/ht)     [default ~12–20pt]
 \newcommand{\E}{\mathbb{E}}
@@
 E-mail: eirik.e.brandsaas@frb.gov.
-I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts, Corina Boar, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobson, William Gamber, Claudia Sahm, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, Boston Fed, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, YES 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop.
+I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts, Corina Boar, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobson, William Gamber, Claudia Sahm, John Sabelhaus, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, Boston Fed, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, YES 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop.
 Code and replication instructions are available at the \href{https://github.com/eirikbrandsaas/HomeownershipBankMomDad.jl}{author's  website and Github}.\newline \textit{Disclaimer: }The views expressed in this paper are solely the responsibility of the author and should not be interpreted as reflecting the views of the Board of Governors of the Federal Reserve System or of anyone else associated with the Federal Reserve System.}}
@@
 I build and estimate a life-cycle overlapping generations model with housing, where adult children and parents interact without commitment.
-I find that parental transfers account for 14 percentage points (29\%) of young households' homeownership.
+I find that parental transfers account for 13 percentage points (27\%) of young households' homeownership.
 Transfers from wealthy parents not only help households overcome borrowing constraints, but also help sustain homeownership, mitigating the drawbacks of illiquidity.
@@
 While homeownership is seen as central to the ``American Dream'' \citep{goodman2018homeownership}, households face different barriers to ownership.
-Between 2009 and 2016, around 30\% of American first-time homebuyers received direct parental assistance for their down payment.\footnote{Source: Survey of Household Economic Decisionmaking (see Section \ref{sec:data}).} This assistance highlights disparities in access, as only those with sufficiently wealthy parents can expect substantial transfers.
+Between 2009 and 2016, about 30\% of first-time buyers received direct parental help with the down payment (Survey of Household Economics and Decisionmaking; see Section~\ref{sec:data}), highlighting disparities in access by parental wealth.
 
@@
 Parents and children interact without commitment through both inter-vivos transfers and end-of-life bequests.
-In models of altruism without housing (e.g., \citealp{Altonji1997a,Barczyk2014}), the transfer motive is to increase the child's consumption when they are borrowing constrained, thus having high marginal utility of wealth.
-The inclusion of a frictional housing market generates two new transfer motives.
+In altruism models without housing (e.g., \citealp{Altonji1997a,Barczyk2014}), transfers generally flow to borrowing-constrained children with high marginal utility of wealth.
+Introducing a frictional housing market adds two additional channels.
 First, transfers can directly alleviate credit constraints, enabling wealthier parents to help their children meet down-payment requirements.
@@
 Moreover, the illiquidity of housing serves as a commitment device enabling children to secure future transfers.
-These mechanisms differ from those examined by \citet{Barczyk2020a}, who study how elderly parents’ homeownership decisions shape nursing‑home entry, care arrangements, and intergenerational transfers.
-As in all transfer models, the expectation of future transfers reduces children's saving motives, potentially lowering the homeownership rate.
+These mechanisms differ from those examined by \citet{Barczyk2020a}, who study how elderly parents’ homeownership affects nursing-home entry, care arrangements, and transfers.
+As in all transfer models, anticipated transfers crowd out children’s saving, potentially lowering homeownership in isolation.
 
@@
 I find that parental transfers account for 13 percentage points (27\%) of the homeownership rate among young households.
-In one adaption of my model, I quantify the contribution of parental transfers to the Black-White homeownership gap of 50\% among young adults, following the  methodology of \cite{Ashman2020} and \cite{aliprantis2022dynamics}.
+In one application of my model, I quantify the contribution of parental transfers to the Black-White homeownership gap of 50\% among young adults, following the  methodology of \cite{Ashman2020} and \cite{aliprantis2022dynamics}.
 I find that parental transfers account for 11 percentage points (20\%) of the Black-White homeownership gap.
+On the housing side, the baseline is intentionally parsimonious---a simple rent-or-own choice with only two sizes, a fixed rent–price ratio, and prices held constant---to isolate the new parent–child channel.
 
@@
 
-This lower cost of illiquidity for households with wealthy parents offers a natural explanation for the positive empirical relationship between parental wealth and successful homeownership I found in the PSID.
+This lower cost of illiquidity for households with wealthy parents offers a natural explanation for the positive empirical relationship between parental wealth and successful homeownership I find in the PSID.
 It also aligns with \citet{choukhmane2023benefits}, who find that households with wealthy parents are more likely to invest in illiquid retirement accounts and avoid costly early withdrawals, even after controlling for a wide set of characteristics.
@@
 My results highlight that policies increasing ownership through subsidies to first-time buyers---a common approach in many cities---will lead to increased housing inequality.
-My model-based results also complement some of the empirical research on the effect of parental resources and housing outcomes \citep[see e.g.,][ for recent work]{wold2024housing,daysal2023intergenerational,benetton2022dynastic,boileau2023drives} by quantifying the contribution of transfers to homeownership.
+My model-based results also complement some of the empirical research on the effect of parental resources and housing outcomes \citep[see e.g.,][for recent work]{wold2024housing,daysal2023intergenerational,benetton2022dynastic,boileau2023drives} by quantifying the contribution of transfers to homeownership.
 
@@
 In these models, children have an incentive to undersave to increase parental transfers (e.g., \citealp{Altonji1997a,Boar2018,Barczyk2014,Chu2020}).
-Illiquid housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}, which children of wealthy parents strategically exploit to encourage future parental transfers.
+Illiquid housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}, which children of wealthy parents strategically use to encourage future parental transfers.
 The paper most closely related to mine is \cite{Barczyk2020a}, who study how the homeownership of retired parents affects the economic behavior of the elderly and their adult children.
@@
 My contribution is to study the complementary problem: how parental transfers affect the homeownership decisions of young adults.
-Finally, while \cite{kaplan2012moving} studies cohabitation with parents among individuals under the age of 25, my focus is on homeownership decisions for households from age 25 onwards.
+Finally, while \cite{kaplan2012moving} studies cohabitation with parents among individuals under the age of 25, my focus is on homeownership decisions for households from age 25 onward.
 
@@
 
-\subsection{Parental Transfers for Downpayments Over Time}\label{sec:overtime}
-I use the Survey of Household Economics and Decisionmaking (SHED) and American Housing Survey (AHS) datasets to show the time trends in parental housing transfers.
-A benefit of these surveys is that they ask questions directly relating transfers to downpayments, unlike the PSID.
+\subsection{Parental Transfers for Down Payments Over Time}\label{sec:overtime}
+I use the Survey of Household Economics and Decisionmaking (SHED) and American Housing Survey (AHS) to show the time trends in parental housing transfers.
+Unlike the PSID, these surveys ask questions directly relating transfers to down payments.
 
@@
 
-	 {\begin{footnotesize} \textit{Notes:} The left panel uses data from the SHED with the sample restricted to current owners who report this being their first home.
-The right panel uses data from the AHS with the sample restricted to all owner-occupied units that did not indicate the sale of a previous home as the main source of funding.
-Note that the left figure has year of purchase on the horizontal axis while the right figure has survey year.
+	\caption*{\begin{footnotesize} \textit{Notes:}  Left panel (SHED): sample is current owners reporting this is their first home; x-axis is purchase year; outcome is the share receiving any down-payment transfer.
+Right panel (AHS): sample is owner-occupied units that do not report proceeds from a prior home sale as the main funding source; x-axis is survey year; outcome is the share whose down payment was mostly funded by a transfer.
 Dashed lines denote 95\% confidence intervals.\end{footnotesize}}
@@
 
+Unfortunately, the Health and Retirement Study (HRS)---though it collects transfer data in many waves---does not include detailed information on child household wealth, a key determinant of homeownership.
+I therefore rely on the PSID, which includes detailed household-level data for both generations.
+
 \textit{Sample Selection:} Throughout this paper, the sample includes all households aged 25 to 84 in the PSID.
@@
 I can observe each household's reported transfers to and from parents and children.
-This leads to discrepancies, where the child and parent do not agree on the amount given from the parent to the child.
+This leads to discrepancies, where the parent and child do not agree on the amount given from the parent to the child.
 For each child, I use the parent's reported transfer to that child.
@@
 
-\textit{Definition of Transfers:} To measure the transfer receipt rate, I rely on the transfer supplement which asked all parents whether they gave money, gifts, or loans of \$100 or more to each child in 2012.
+\textit{Definition of Transfers:} To measure the transfer receipt rate, I rely on the transfer supplement, which asked all parents whether they gave money, gifts, or loans of \$100 or more to each child in 2012.
 I follow the literature \citep[e.g.,][]{mcgarry2016dynamic} and treat all transfers as gifts.
@@
 
-In all other waves, there is a general question of gifts and inhertitances from people outside the household.
-We do not know the source of these transfers (e.g., they could be from friends or grandparents) and hence these do not know the extent to which these captures parental assistance.
+In all other waves, there is a general question on gifts and inheritances from people outside the household.
+We do not know the source of these transfers (e.g., they could be from friends or grandparents), and hence we do not know the extent to which they capture parental assistance.
 Nevertheless, they still capture some of the impact of transfer receipt on household decisions.
@@
 
-Finally, why focus on inter-vivos transfers when most intra-family wealth transfers occur through bequests \citep[see e.g.,][]{feiveson2019lifecycle}?
+Finally, why focus on inter-vivos transfers when most intrafamily wealth transfers occur through bequests \citep[see e.g.,][]{feiveson2019lifecycle}?
 First, parental bequests are typically received after age 45, too late to affect young adults' transition into homeownership.
@@
 
-The first two columns of the top panel compare households by transfers receipt.
+The first two columns of the top panel compare households by transfer receipt.
 The mean transfer is relatively high at \$3,960, and 21\% of households received transfers in the last year.
@@
 Second, the transfers in the transfer supplement are much smaller than than those in the gift/inheritance question.
-Indeed, the median parental transfer is about \$1k, the mean about \$3k, and the 95th percentile is at \$10k, compared to the median gift/inheritance of \$25k, mean of \$60k, and a 95th percentile of about \$130k.
+Indeed, the median parental transfer is about \$1{,}000, the mean about \$3{,}000, and the 95th percentile is about \$10{,}000, compared to the median gift/inheritance of \$25{,}000, mean of \$60{,}000, and a 95th percentile of about \$130{,}000.
 
@@
 
-	 {\begin{footnotesize} \textit{Notes:} The left panel uses the transfers, gifts, and inheritance variable which is available in all waves of the PSID while the right panel uses the parental transfer data only available in the 2013 transfer supplement.
+	 \caption*{\begin{footnotesize} \textit{Notes:} The left panel uses the transfers, gifts, and inheritances variable, which is available in all waves of the PSID, while the right panel uses the parental-transfer data only available in the 2013 transfer supplement.
 See main text for a discussion.
@@
 Second, in the gift/inheritance question, the source of the transfer is unknown.
-Third, it is unclear whether intra-family loans are consistently reported or repaid.
+Third, it is unclear whether intrafamily loans are consistently reported or repaid.
 
@@
 
-I now utilize microdata from the PSID to examine the empirical association between parental financial resources on housing outcomes.
+I now utilize microdata from the PSID to examine the empirical association between parental financial resources and housing outcomes.
 The sample is limited to households aged 25 to 44 to focus on the housing choices of young adults.
 I use the sum of parental net worth and income (denoted as wealth) to measure parents' financial resources in all specifications for two reasons.
-First, this is consistent with models of altruism, as the one used later in the paper.
+First, this is consistent with models of altruism, such as the one used later in the paper.
 Second, since parents' wealth and income are highly correlated and I have few observations---fewer than 1,000 for many specifications---the standard errors are large when both are included.
+Note that neither transfer receipt nor parental wealth is treated as exogenous.
+For example, many transfers may be made precisely because the renter is buying a home, or parents may save in order to help their children enter the housing market.
 
 \subsubsection{Transfers and the Transition into Homeownership}
-Since many papers have already estimated the relationship between parental transfers and transitions into homeownership (e.g.\ \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Lee2018}), I replicate these regressions to highlight the association between transfer receipt and rent-to-own transitions, but relegate the full results to Appendix~\ref{app:rent_to_own}.
+Since many papers have already estimated the relationship between parental transfers and transitions into homeownership (e.g., \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Lee2018}), I replicate these regressions to highlight the association between transfer receipt and rent-to-own transitions, but relegate the full results to Appendix~\ref{app:rent_to_own}.
 Receiving a parental transfer above \$500 is associated with a 1.2 percentage point increase in the probability of becoming a homeowner (not statistically significant), after controlling for household characteristics.
@@
 Among first-time owners, only 80\% have retained ownership four years later.
-For non-first-time owners, there is still a marked decrease though it is more muted, after four years about 10\% of all owners will be renting.
-
-To understand which variables predicts retainment of ownership, I estimate
+For non-first-time owners, there is still a marked decrease, though it is more muted; after four years about 10\% of all owners will be renting.
+
+To understand which variables predicts retention of ownership, I estimate
 \begin{equation}	\label{eq:regmaintain}
@@
 \end{equation}
-where $Own_{i,t+\Delta}=1$ denote ownership at time $t+\Delta$, $\beta$ and $\zeta$ measure the effects of log household net worth and parental wealth, respectively.
+where $Own_{i,t+\Delta}=1$ denotes ownership at time $t+\Delta$, $\beta$, and $\zeta$ measure the effects of log household net worth and parental wealth, respectively.
 The vector of control variables  $X_{i,t}$ consists of log household income, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.
 The sample is limited to households aged 25 to 44.
-These regressions are motivated by \cite{bond2021role}, who estimate similar equations in the PSID and the Health and Retirement Study.
+These regressions are motivated by \cite{bond2021role}, who estimate similar equations using the PSID and the HRS.
 Figures \ref{fig:coef_allowner} and \ref{fig:coef_ftowner} plot the two coefficients of interest over different time horizons.
@@
 I first limit the sample to first-time buyers to focus on distress among new owners.
-In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$) With the full set of controls, the effect increases slightly, though it is no longer statistically significant.
+In a univariate regression, a 10\% increase in parental wealth is associated with a 0.07 percentage point decrease in the probability of being behind on mortgages two years later ($p<0.1$).
+With the full set of controls, the effect increases slightly, though it is no longer statistically significant.
 These magnitudes are economically meaningful since 1.6\% of first-time owners are behind on their mortgage.
 In Columns 3 to 5, I expand the analysis to include all owners who are currently not behind, increasing the sample sizes and allowing us to include household fixed effects.
-In these specifications, a 10\% increase in parental wealth is associated with a 0.09 percentage point decrease in delinquency risk in the univariate OLS regression ($p<0.001$), a 0.02 percentage point decrease in the multivariate OLS regression (not significant), and a 0.07 percentage point decrease in the fixed effects regression ($p<0.1$).
+In these specifications, a 10\% increase in parental wealth is associated with a 0.09 percentage point decrease in delinquency risk in the univariate OLS regression ($p<0.001$), a 0.02 percentage point decrease in the multivariate OLS regression (not significant), and a 0.07 percentage point decrease in the fixed-effects regression ($p<0.1$).
 
@@
 	{\begin{footnotesize}\begin{flushleft}
-		\textit{Notes:} The estimated association between parental wealth and the probability of behind on mortgage payments in the next period from regressions using specification \ref{eq:regbehind}.
+		\textit{Notes:} Estimated association between parental wealth and the probability of being behind on mortgage payments in the next period from regressions using specification \ref{eq:regbehind}.
 See Appendix Table~\ref{tab:hypo_long} for all results.
@@
 \subsubsection{Owners with Wealthy Parents Don't Downsize in Unemployment}\label{sec:eventstudy}
-Distressed and forced home sales are costly and often triggered by unemployment \citep{kermani2021racial,hsu2018unemployment}.
-To test whether parental wealth mitigates this risk, I perform a simple event study of how unemployment affects the likelihood of downsizing.
-
-The exercise follows \cite{Chetty2007} closely.
-The outcome of interest is the change in log housing consumption, which is set to zero for households who do not move.
-For movers, housing consumption is defined as yearly rent while renting and as imputed rent for owners, set to 5\% of the market value \citep{Davis2008}.
-The sample is limited to household heads who experience only one unemployement spell between ages 25 and 44.
-I divide the sample by whether a household's parents were in the top parental wealth quartile at the time of unemployment and compare the two groups' housing consumption growth at unemployment.
-The results are displayed in Figure \ref{fig:housinggrowthrates}.
+Because unemployment often triggers costly distressed sales \citep{kermani2021racial,hsu2018unemployment}, I test whether parental wealth mitigates this risk by estimating an event study of unemployment’s effect on downsizing.
+
+The exercise follows \citet{Chetty2007}.
+The outcome is the change in log housing consumption (winsorized at the 1st and 99th percentiles), set to zero for non-movers.
+For movers, housing consumption equals annual rent if renting and imputed rent if owning (5\% of market value; \citealp{Davis2008}).
+The sample includes household heads with exactly one unemployment spell between ages 25 and 44.
+I split the sample by whether parental wealth in the unemployment year lies in the top quartile and compare housing-consumption growth around unemployment.
+The results are displayed in Figure~\ref{fig:housinggrowthrates}.
 Among households with non-wealthy parents, housing consumption falls by about 5\% during unemployment ($p<0.05$).
-In contrast, households with wealthy parents see a slight increase (not significant) in housing consumption during unemployment.
-Appendix~\ref{app:eventstudy_details} shows that these patterns are robust to an event-study specification that controls for household characteristics.
-I return to these results later as part of the model validation.
+In contrast, households with wealthy parents exhibit a small, statistically insignificant increase.
+Appendix~\ref{app:eventstudy_details} gives additional details and shows that these patterns are robust to an event-study specification with household controls and to alternative cutoffs.
+I return to these results in the model validation.
 
@@
 
-	{\begin{footnotesize} \textit{Notes:} Solid lines denote means, and bars denote the 95\% confidence interval.
+	\caption*{\begin{footnotesize} \textit{Notes:} Solid lines denote means, and bars denote 95\% confidence intervals.
 The sample consists of households aged 25-45 with exactly one unemployment spell and without changes in head and/or spouse in the four years before and after unemployment.
@@
 \subsection{Taking Stock of the Data}
-Taking stock, we have seen that parental transfers have become more important for young homeowners over the past two decades.
+Taking stock, parental transfers have become more important for young homeowners over the past two decades.
 About 20\% of young households receive transfers from their parents in a given year, and recipients have wealthier parents.
-Recipients are more likely to transition from renting to owning, even after conditioning on a rich set of covariates.
+Recipients are more likely to transition from renting to owning, after conditioning on a rich set of covariates.
 Parental wealth is also positively associated with maintaining homeownership.
-Households with richer parents are less likely to fall behind on mortgages and less likely to downsize during unemployment spells.
-These results suggest that homeowners with wealthy parents find illiquidity of homeownership less problematic.
-
-\subsection{Transfers and Taxes}
-In the United States, transfers and bequests are subject to taxation due to the estate tax.
-Taxes are paid by the giver, not the recipient.
-In 2024, an individual can give \$13.61 million before they start owing gift taxes.
-However, individuals have to file if they give transfers above \$18,000 (in 2024) to any one individual within a calendar year.
-Thus, the gift tax is irrelevant for the vast majority of households, as it applies only once lifetime transfers---including bequests---exceed \$13.61 million.
+Households with richer parents are less likely to fall behind on mortgages and to downsize during unemployment spells.
+These results suggest that homeowners with wealthy parents find the illiquidity of homeownership less problematic.
+
+Throughout the empirical section, I do not treat transfer receipt or parental wealth as exogenous, and the estimates document associations rather than causal effects.
+For example, many transfers may be made precisely because the renter is buying a home.
+This motivates the structural model I describe in the next section, which provides a disciplined framework to interpret mechanisms and conduct counterfactuals.
 
 \section{A Quantitative Model of Parental Transfers and Homeownership}\label{sec:model}
-This section describes my life-cycle model of housing choices with overlapping generations, idiosyncratic earnings risk, and altruistic inter-vivos transfers from parents to their offspring (``children'').
-To answer how important parental transfers are for housing outcomes, I combine two models: altruism without commitment and illiquid housing.
+This section describes my life-cycle model of housing choices with overlapping generations, idiosyncratic earnings risk, and altruistic inter-vivos transfers from parents to their offspring.
+To answer how important parental transfers are for housing outcomes and how transfers interact with popular housing policies, I combine two models: altruism without commitment and housing.
 
@@
 The lack of commitment aligns with empirical transfer patterns and generates strategic behavior, as both children and parents internalize the effect of their choices on each other's future choices.
-The central theoretical prediction is that wealthy parents transfer to children with high marginal utility of wealth, typically poor or borrowing-constrained households \citep{Chu2020,Barczyk2020}.
+Although no commitment complicates computation, it is essential: under full commitment, transfer timing is indeterminate---unattractive due to down payment and borrowing constraints.
+The central theoretical prediction is, generally, that wealthy parents transfer to children with high marginal utility of wealth, typically poor or borrowing-constrained households \citep{Chu2020,Barczyk2020}.
 
@@
 Finally, while illiquid housing makes selling costly, wealthy parents can offer partial insurance, thus reducing the risk of costly liquidation or downsizing.
-
-These two housing frictions---borrowing constraints and illiquidity---are not only theoretically important, but also empirically relevant.
+The housing block is deliberately simple---constant prices and one unit size per tenure---but captures key barriers to homeownership: down payment constraints, adjustment costs and illiquidity, indivisibility, and market segmentation.
+
+Among these frictions, borrowing constraints and illiquidity are the focus of the analysis and are empirically important.
 In the SHED, 57\% of renters could not afford a down payment or did not qualify for a mortgage, while 26\% said that renting was more convenient and 23\% cited planning to move as reasons for renting.
 
-\subsection{Demographics, Preferences, and Technologies}
-Time is discrete and finite.
-Each period consists of two years.
+\subsection{Demographics, Preferences, and Technologies}\label{sec:dem}
+Time is discrete and finite.\footnote{I solve the model in discrete time, consistent with much of the life-cycle housing literature.
+While continuous-time approaches can simplify aspects of altruistic interactions \citep{Barczyk2014}, discrete-time models can be solved using global methods, including the brute-force grid search used here \citep{Barczyk2020}.
+See Appendix~\ref{sec:computational} for details.} Each period consists of two years.
 At the beginning of each period, a constant mass of new households enters and exits the economy.
@@
 Second, the child transitions to become a parent household.
-Third, the child of the child becomes economically active as a child household.
+Third, the child of the child becomes economically active as a child household, see Figure~\ref{fig:overview}.
 
@@
 	\input{modeloverview.tex}
+	{\begin{footnotesize} \textit{Notes:} An illustration of the dynastic and life cycle described in Section \ref{sec:dem}.\end{footnotesize}}
 \end{center}
@@
 \textit{Intergenerational Transfers:} In the last period before death, the parent can leave a bequest, which the child receives in the next period.
-In all other periods, the parent can give inter-vivos transfers of a nonnegative amount, $t_p$, that the child receives immediately.
+In all other periods, the parent can give inter-vivos nonnegative transfers, $t_p$, that the child receives immediately.
 
@@
 \end{cases}
-\end{equation} where $m_s$ and $m_b$ denote selling and buying costs, respectively. A household enters the period house $h_a$, while $h_{a+1}$ is the house chosen in this period.
+\end{equation} where $m_s$ and $m_b$ denote selling and buying costs, respectively. A household enters the period with housing $h_a$, while $h_{a+1}$ is the house chosen in this period.
 
@@
 \end{align}
-I assume that parents face no income risk for simplicity, consistent with the decrease in income risk with age \citep{Sanchez2020} and the fact that much of retirement income---such as Social Security and defined benefit pensions---is relatively certain.
-Furthermore, this paper focuses on the role of parental transfers for children' housing choices, where parental income risk is not a first-order concern.
-However, in Appendix \ref{sec:robust_incomerisk}, I show that parental income and health expenditure risks do not meaningfully change my results.
-
-\textit{Initial Conditions of the New Child:} When children enter the economy at age $a_c=25$, their initial net worth $x_{25}$ and productivity $y_{25}$ are stochastic but correlated with the parent's wealth $x_{53}$ and productivity $y_{53}$ according to $x_{25}, y_{25} \sim F(x_{53}, y_{53})$.
-The initial net worth is not funded by parental transfers, though children can receive transfers in the first period.
-While I abstract from parental investment in human capital—an important driver of intergenerational persistence \citep[see e.g.,][]{Daruich2018,Lee2019}—the function $F$ matches observed correlations in wealth and income, ensuring realistic transfer motives.
-All households begin as renters.
+I assume that parents face no income risk for simplicity, as income risk decreases with age \citep{Sanchez2020} and that much of retirement income---Social Security and defined benefit pensions---is relatively certain.
+Furthermore, this paper focuses on the role of parental transfers for children's housing choices, where parental income risk is not a first-order concern.
+Moreover, in Appendix \ref{sec:robust_incomerisk}, I show that parental income and health expenditure risks do not meaningfully change my results.
+
+\textit{Initial conditions of the entering child.} At age $a_c=25$, each child enters as a renter with net worth $x_{25}$ and productivity $y_{25}$ drawn from a reduced-form mapping $F(x_{53},y_{53})$ estimated from the PSID parent–child joint distribution.
+Initial wealth $x_{25}$ summarizes the child’s own savings and liabilities and the cumulative effects of pre-25 parental investments and support (e.g., tuition, co-residence, early gifts)---channels documented as important \citep[e.g.,][]{Daruich2018,Lee2019}.
+I do not model these pre-25 flows explicitly because the PSID lacks child-specific transfer histories.
+The reduced-form $F$ matches observed parent–child correlations in wealth and income, disciplining both selection into homeownership and the strength of transfer motives.
+The parent can give transfers once the initial draw is observed.
 
@@
 \textit{Decision Problems:} 
-I now show the decision problems for a dynasty at age $a_c < 53$. For simplicity, I assume the child is a renter choosing to become an owner, while the parent remains a renter.
-
-\textit{Child - Second Stage:} The child, conditional on becoming a new owner ($h_c=h_r,h'_c=h_o$), chooses consumption and bonds to maximize expected utility:
+I now show the decision problems for a dynasty at age $a_c < 53$ for a renting child choosing to become an owner, while the parent remains a renter.
+
+\textit{Child -- Second Stage:} The child, conditional on becoming a new owner ($h_c=h_r,h'_c=h_o$), chooses consumption and bonds to maximize expected utility:
 \begin{equation}\label{eq:Vk}
@@
 \end{equation}
-The first constraint is the budget constraint, which determines the net bond position as available resources (net worth, transfers, and income) minus expenditures (consumption, purchase price, and purchase cost).
-The second constraint is the law of motion for net worth, which is the sum of the bond position and housing wealth after maintenace costs.
+The first constraint is the budget constraint, defining the net bond position as available resources (net worth, transfers, and income) minus expenditures (consumption, purchase price, and purchase cost).
+The second constraint is the law of motion for net worth, which is the sum of the bond position and housing wealth after maintenance costs.
 The third constraint is the LTV constraint.
-We see that a transfer $t_p$ can move the child away from the LTV constraint by increasing the bond position $b_c'$, holding consumption and housing fixed.
+A transfer $t_p$ can move the child away from the LTV constraint by increasing the bond position $b_c'$, holding consumption and housing fixed.
 The expectation is conditional on the child's current productivity.
@@
 
-\textit{Parent - First Stage:} The parent, conditional on remaining a renter ($h_p=h'_p=h_r$), chooses consumption, transfers, and bonds to maximize expected utility:
+\textit{Parent -- First Stage:} The parent, conditional on remaining a renter ($h_p=h'_p=h_r$), chooses consumption, transfers, and bonds to maximize expected utility:
 \begin{equation}\label{eq:Vp}
 \begin{split}
-V_p(\mathbf{s}_p;\mathbf{g}_c)^{rent} = \max_{c_p,b_p',t_p} & u(c_p,h'_p) + {\eta} u\left (c_c^*({\mathbf{s}_c}),h_c'^*({\mathbf{s}_c})\right ) + \beta \E_{y_c} \left[V_{p}({\mathbf{s}_p';\mathbf{g}_c})\right] \\
+V_p(\mathbf{s}_p;\mathbf{g}_c)^{rent} = \max_{c_p,b_p',t_p} & u(c_p,h'_p) + {\eta} u\left(c_c^*({\mathbf{s}_c}),h_c'^*({\mathbf{s}_c})\right) + \beta \E_{y_c} \left[V_{p}(\mathbf{s}_p';\mathbf{g}_c)\right] \\
 \text{s.t.}\quad & b'_p = x_p + w_p - c_p - t_p - q p h_p'\\
@@
 \end{equation}
-The first constraint is the budget constraint, which determines the net bond position as available resources (net worth and income) minus expenditures (consumption, transfers, and rent).
+The first constraint is the budget constraint, defining the net bond position as net worth and income minus expenditures (consumption, transfers, and rent).
 For renters, next-period wealth is simply given by the bond position.
@@
 
-\subsection{Intuition - Transfers and Homeownership}\label{sec:modelintuition}
+\subsection{Intuition -- Transfers and Homeownership}\label{sec:modelintuition}
 Before turning to estimation, there are several points to highlight about how housing interacts with altruistic transfers.
@@
 Due to nonconvexities introduced by housing, the policy and value functions are not differentiable everywhere.
-To focus on the underlying intuition, I will sidestep the technical issues and refer to \cite{Barczyk2020} and \cite{Chu2020} for a formal derivation.
+To focus on the underlying intuition behind the classical altruistic transfer motive, I will sidestep the technical issues and refer to \cite{Barczyk2020} and \cite{Chu2020} for a formal derivation.
 
@@
 This is the classic first-order condition from \cite{Altonji1997a} and says that the parent transfers until their own marginal utility equals the altruism-weighted marginal utility of the child.
-This is the fundamental driver of the transfer motives in the model and, as we will see, interacts with housing choices in interesting ways.
-As discussed extensively in \cite{Barczyk2020a} and \cite{Chu2020}, other transfer motives exist, such as ``pushes to autarky'', though these do not appear in my parameterization.
+This transfer motive, as we will see, interacts with housing choices in interesting ways.
+As \citet{Barczyk2020a} and \citet{Chu2020} discuss, dynamic altruism models give rise to additional transfer motives---such as ``pushes to autarky''---that require the model to be solved globally (see Appendix~\ref{sec:computational}).
 
@@
 Third, transfers alleviate the LTV constraint in two ways.
-First, the parent give transfers to alleviate the borrowing constraint for a child who would otherwise be unable to buy.
-Second, the parent give transfers to a borrowing-constrained child, who can only afford very low consumption, shifting the ownership threshold down.
+First, the parent gives transfers to alleviate the borrowing constraint for a child who would otherwise be unable to buy.
+Second, the parent gives transfers to a borrowing-constrained child, who can only afford very low consumption, shifting the ownership threshold down.
 The first increases homeownership directly and the second indirectly.
@@
 On the right, where all households own, the child takes out a larger mortgage, holding wealth fixed.
-One the far right, the child is wealthy and behave as if there is no altruism.
+On the far right, the child is wealthy and behaves as if there is no altruism.
 Comparing the orange and red lines, we see that an increase in parental wealth decreases saving at all wealth levels.
@@
 	\end{center}
-	{\vspace{-0.2in}\begin{footnotesize}\textit{Notes:} The solid lines plots the child’s bond choice ($b'_c$) and the dashed lines the housing choice ($h'_c$) as a function of child cash-on-hand $\tilde x_c$.
+	{\vspace{-0.2in}\begin{footnotesize}\textit{Notes:} The solid lines plot the child’s bond choice ($b'_c$) and the dashed lines the housing choice ($h'_c$) as a function of child cash-on-hand $\tilde x_c$.
 The dashed vertical lines mark the wealth thresholds at which the child switches from renting to owning.
@@
 
-More generally, the child's savings choices show that with altruism, households are more likely to choose to be liquidity constrained, often called hand-to-mouth, as they become homeowners with less wealth and are closer to the borrowing constraints.
+With altruism, households are more likely to choose to be liquidity constrained, often called hand-to-mouth, as they become homeowners with less wealth and are closer to the borrowing constraints.
 Altruism reduces the downside of liquidity constraints---namely, the inability to smooth income shocks through partial insurance---while creating the advantage of receiving larger transfers.
@@
 
-To calibrate the stochastic income process $y_{i,a}$, I first set $N_y=3$ to create a three-state Markov Chain.
+To calibrate the stochastic income process $y_{i,a}$, I first set $N_y=3$ to create a three-state Markov chain.
 The sample is divided into age-specific income tertiles.
@@
 	\begin{threeparttable}
-		\caption{Summary of Externally and Independently Estimated Parameter}\label{tab:calpar}
+		\caption{Summary of Externally and Independently Estimated Parameters}\label{tab:calpar}
 		\small
@@
 		\footnotesize
-		\textit{Notes:} Superscript $^\dagger$ denotes parameters that are bi-annualized.
+		\textit{Notes:} Superscript $^\dagger$ denotes parameters expressed on a biennial basis.
 	\end{threeparttable}
@@
 Similarly, I estimate the homeownership utility $\chi$, since transfers generate new ownership motives.
-Finally, I estimate the relative size of owner-occupied housing, $h_0$, a key parameter in models of homeownership, because housing choices interact with transfer behavior.
+Finally, I estimate the relative size of owner-occupied housing, $h_o$, a key parameter in models of homeownership, because housing choices interact with transfer behavior.
 
 \textit{Identification and Moment Estimation:}
-I derive all empirical moments from the PSID sample discussed in Section \ref{sec:PSID}. I calculate moments by aggregating over all years, assigning equal weight to each year.
+I derive all empirical moments from the PSID sample discussed in Section \ref{sec:PSID}.
+Moments are calculated by aggregating across all years, with each year receiving equal weight.
+The parent-child transfer rate is estimated exclusively using the 2013 Transfer Supplement.
 
@@
 Mortgages are slightly smaller than in the data, as households repay and re-borrow flexibly.
-The model also matches the LTV ratio for buyers well.\footnote{The empirical LTV is measured at the first observation of ownership, not necessarily at origination, so it can reflect price appreciation, improvements, and mortgage paydown.}
+The model also matches the LTV ratio for buyers well.\footnote{The empirical LTV is measured at the first observation of ownership, not necessarily at origination, so it can reflect price appreciation, improvements, mortgage paydown, and cash-only buyers.}
 
@@
 
-	{\begin{footnotesize} \textit{Notes:} Lines denote means, and shaded areas indicate 95\% confidence intervals.
+	\caption*{\begin{footnotesize} \textit{Notes:} Lines denote means, and shaded areas indicate 95\% confidence intervals.
 The light orange area denotes the empirical estimates (Section \ref{sec:eventstudy}), and the darker grey area denotes the model simulation.
@@
 \subsubsection{Parental Wealth and Ownership Retention}
-Next, I repeat the exercise in Section~\ref{sec:maintaining} (see Figure~\ref{fig:owner_hazard}) for first-time homebuyers, where we saw that (a) parental wealth is associated with more retention, (b) more so than own wealth, and (c) that the role of parental wealth inreases over time horizons.
-In the model, I replicate the regression as closely as possible.
+Next, I repeat the exercise in Section~\ref{sec:maintaining} (see Figure~\ref{fig:owner_hazard}) for first-time homebuyers, where we saw that (a) parental wealth is associated with more retention, (b) more so than own wealth, and (c) that the role of parental wealth increases over time.
+In the model, I replicate the regression as closely as possible, see figure \ref{fig:remain_model}.
 For example, there are no year effects or controls for family size.
-The model captures these patterns well, though it overpredicts the role of parental wealth at shorter horizons.
+The model captures the three patterns well, though it overpredicts the role of parental wealth at shorter horizons.
 
@@
 
-	{\begin{footnotesize} \textit{Notes:} Lines denote means, and shaded areas indicate 95\% confidence intervals.
+	\caption*{\begin{footnotesize} \textit{Notes:} Lines denote means, and shaded areas indicate 95\% confidence intervals.
 The light orange area denotes the empirical estimates (Section \ref{sec:maintaining}), and the darker grey area denotes the model simulation.
-The sample consists of households first-time owners aged 25-45.
+The sample consists of first-time owners aged 25-45.
+The y-axis measures the increase in probability of ownership retention of a 1\% increase in own/parental wealth.
 \end{footnotesize}}
@@
 
-\subsubsection{Taking Stock on Model Fit}
+\subsubsection{Taking Stock of Model Fit}
 The model matches targeted moments related to homeownership, transfers, and rent-to-own transitions using realistic preference parameters.
 It also performs well on untargeted aggregate moments.
-Perhaps more importantly, it replicates key non-targetted household-level patterns in the associations between parental wealth, downsizing behavior, and ownership retention.
+Perhaps more importantly, it replicates key nontargetted household-level patterns in the associations between parental wealth, downsizing behavior, and ownership retention.
 
@@
 The main driver is that the ownership threshold shifts out in wealth without altruism, as we saw in the policy functions (Section \ref{sec:policyfuncs}).
-Without transfers then, the average wealth at purchase increases by about 25\%, the age of first ownership is delayed by about 2.5 years, and the LTV at purchase falls by 12 percentage points.
+Without transfers, the average wealth at purchase increases by about 25\%, the age of first ownership is delayed by about 2.5 years, and the LTV at purchase falls by 12 percentage points.
 Although homeownership of young households declines sharply, the overall homeownership rate of all households (ages 25 to 74) only falls by 6 percentage points (9\%).
 Hence, the main aggregate effect of parental transfers is that they induce earlier homeownership.
-Finally, without altruism, the parental wealth gradient in homeownership decreases from 2.0 to 1.15, with the remaining effect driven by the integenerational correlation in wealth and income.
+Finally, without altruism, the parental wealth gradient in homeownership decreases from 2.0 to 1.15, with the remaining effect driven by the intergenerational correlation in wealth and income.
 Hence, the positive correlation between parental wealth and housing outcomes is almost entirely driven by parental transfers.
@@
 While existing structural models of racial inequality focus primarily on income and bequests, they typically omit housing and inter-vivos transfers.
-In contrast, many empirical studies focus on racial difference in housing outcomes.\footnote{For example, \citet{Ashman2020} and \citet{aliprantis2022dynamics} use overlapping-generations models to study the racial wealth gap, attributing most of it to income differences.
+In contrast, many empirical studies focus on racial differences in housing outcomes.\footnote{For example, \citet{Ashman2020} and \citet{aliprantis2022dynamics} use overlapping-generations models to study the racial wealth gap, attributing most of it to income differences.
 Though the latter include pooled bequests, they do not model strategic behavior or direct transfers from parents to children.
-Many empirical studies highlight the role of parental wealth in racial housing gaps: \citet{charles2002transition} find that it explains 25\% of the mortgage application gap,  \citet{bond2021role} attribute 28\% of the gap in maintaining homeownership to parental wealth, and \cite{kermani2021racial} find a large gap in realized housing returns due to distressed sales.}
+Many empirical studies highlight the role of parental wealth in racial housing gaps: \citet{charles2002transition} find that it explains 25\% of the mortgage application gap,  \citet{bond2021role} attribute 28\% of the gap in maintaining homeownership to parental wealth, and \citet{kermani2021racial} find a large gap in realized housing returns due to distressed sales.}
 
@@
 Removing altruism reduces the gap to 43\%, implying that parental transfers account for 11 percentage points---or 20\%---of the gap.
-This is a sizeable effect, especially given that the Black–White homeownership gap has remained largely unchanged since the passage of the Fair Housing Act in 1968.
+This is a sizable effect, especially given that the Black–White homeownership gap has remained largely unchanged since the passage of the Fair Housing Act in 1968.
 While Black and White households receive transfers at similar rates in the model, transfers are less influential for Black households, who are further from the ownership threshold and receive smaller amounts.
@@
 Thus, transfers account for between 10 and 13 percentage points of the homeownership rate, depending on the supply elasticity.
+Nevertheless, this exercise holds the rent–price ratio fixed and abstracts from broader general-equilibrium channels (e.g., rent pass-through).
 
@@
 \subsection{Policies and the Importance of Parental Transfers}
-Many countries have policies aimed at increasing homeownership, and housing affordability and delayed homeownership is at the forefront of economic policy and research \citep[see e.g.,][]{Mabille2020}.
-I now evaluate how five different policy levers influence the role of parental wealth in housing outcomes.
-For each policy, I change the relevant parameter slightly and solve the model before simulating a new distribution.
-The idea is not to see what the effect of various policy proposals would be in general equilibrium, but how different policy levers interact with parental transfers.
-The results are reported in Table \ref{tab:policy}.
-
-\textit{Relaxing Mortgaging Regulation:} To study how parental transfers interact with down payment requirements, I tighten the LTV limit by 3 percentage points.
-The tighter limit decreases homeownership among the young by 7 percentage points (15\%).
-Homeownership declines more among households with parents in the middle of the wealth distribution, reducing the parental wealth gradient.
-While initially counterintuitive, these middle-wealth households rely heavily on small transfers to overcome borrowing constraints; poorer households remain constrained, and richer ones are unaffected.
-Thus, looser borrowing constraints not only boost homeownership overall but also amplify the importance of parental wealth.
-Consistent with this mechanism, \citet{wold2024housing} find that owners with poorer parents have lower LTV ratios at origination.
-
-\textit{Lowering PMI:} To study how parental transfers interact with the PMI, which increases the interest rate on high LTV loans, I relax the PMI limit by 7 percentage points.
-The relaxed PMI limit increases homeownership by 1 percentage points (2\%).
-Similar to the LTV regulation, homeownership among households with parents in the middle of the wealth distribution rises more, further widening the parental wealth gradient.
-This result suggests that relaxing the PMI limit increases not only homeownership for all households, but also the role of parental wealth.
-
-\textit{Lowering House Prices:} To study how parental transfers interact with housing affordability, I lower house prices by 2\%.
-As prices decline, homeownership increases for all households, though more so for households with the richest parents, increasing the role of parental wealth.
-
-\textit{Reducing Purchase Costs:} To study how closing costs and first-time buyer assistance interact with parental transfers, I lower the purchase cost ($m_b$) by 2 percentage points---equivalent to a grant of about \$6,000.\footnote{Many U.S.
-states offer closing cost assistance to first-time buyers.
-In England, stamp duty was reduced for first-time buyers in 2017.} While homeownership rises, the gains are concentrated among households with wealthy parents, amplifying the parental wealth gradient.
+Many countries pursue policies to raise homeownership, and concerns about affordability and delayed entry are central to current economic policy debates \citep[e.g.,][]{Mabille2020}.
+I evaluate how five policy levers affect the role of parental wealth in housing outcomes.
+For each policy, I modestly perturb the relevant parameter, re-solve the model, and simulate the new stationary distribution.
+The aim is not to assess general-equilibrium effects, but to study how popular policy levers interact with parental transfers.
+Results appear in Table~\ref{tab:policy}.
+
+\textit{Tighter LTVs:} LTV caps, a standard macroprudential tool, directly regulate down-payment requirements; as shown in Figure~\ref{fig:motivation}, reliance on down-payment assistance moves with changes in mortgage regulation.
+I tighten the LTV limit by 3 percentage points.
+Homeownership among the young falls by 7 percentage points (15\%), with larger declines for children of middle-wealth parents, flattening the parental wealth gradient.
+The mechanism is that these families rely on modest transfers to clear borrowing constraints; poorer households remain constrained, and richer ones are largely unaffected.
+Symmetrically, loosening borrowing constraints would raise homeownership and amplify the role of parental wealth.
+Indeed, \citet{wold2024housing} find that owners with poorer parents have lower LTV ratios at origination.
+
+\textit{Lowering PMI:} Several government-backed mortgages (e.g., VA, USDA) do not require PMI; to mimic easing on this margin, I relax the PMI limit by 7 percentage points.
+Homeownership rises by 1 percentage point (2\%).
+As with the LTV rule, gains are larger for households with parents in the middle of the wealth distribution, further widening the parental-wealth gradient.
+Thus, relaxing PMI increases both overall homeownership and the importance of parental wealth.
+
+\textit{Lowering House Prices:} Although there is no single “price lever,” policy shapes prices through supply regulation, taxes, interest rates, and subsidies.
+I capture these channels in reduced form by lowering house prices by 2\%.
+Homeownership increases for all households, but more so among those with the richest parents, thereby increasing the role of parental wealth.
+
+\textit{Reducing Purchase Costs:}
+Closing-cost programs and first-time-buyer relief---e.g., England’s 2017 stamp duty cut for first-time buyers---directly lower entry costs; to study this margin’s interaction with parental transfers, I lower the purchase cost ($m_b$) by 2 percentage points, equivalent to a grant of about \$6,000.
+While homeownership rises, the gains are concentrated among households with wealthy parents, amplifying the parental wealth gradient.
 Lowering purchase costs has little effect for renters with non-wealthy parents, who remain constrained by LTV limits and sales costs.
-In contrast, for household with non-wealthy parents, lower purchase costs reduce the effective price—the main remaining barrier to ownership.
+In contrast, for households with wealthy parents, lower purchase costs reduce the effective price—the main remaining barrier to ownership.
 This suggests that grants or credits for first-time buyers primarily benefit households with richer parents.
@@
 \textit{Reducing Sales Costs:} 
-The earlier empirical results suggested that parental wealth helps offset the illiquidity of housing.
-To study this interaction, I lower the sales cost $m_s$ by 2 percentage points.
+Sales costs consist of government taxes and fees and broker commissions.
+Following the 2024 National Association of Realtors settlement, commissions are expected to decline; to study how sales costs interact with parental transfers, I lower the sales cost $m_s$ by 2 percentage points.
 Strikingly, the homeownership rate decreases by 3 percentage points (6\%).
-This counterintuitive result reflects selection into ownership.
-Homeownership rises slightly for most households but drops sharply---by 6 percentage points---among households with moderately wealthy parents.
-For these households, lower sales costs weaken the ability to use housing as a commitment device to secure future transfers.
+This reflects selection into ownership: homeownership rises slightly for most households but drops by 6 percentage points among those with moderately wealthy parents.
+For these households, lower sales costs weaken the commitment value of housing used to secure future transfers.
 In contrast, children of the poorest parents (who rarely receive transfers) and the wealthiest parents (who expect transfers regardless) are less affected.
@@
 The previous section established that higher sales costs can increase ownership among households with wealthy parents.
-Here, I present a more surprising result: Households with wealthy parents prefer illiquidity.
-
-I solve the model both with and without sales costs ($m_s=0$).
-For each household in the stationary distribution with sales costs, I assess whether they would prefer that neither they nor future generations in their dynasty face sales costs.
-This hypothetical corresponds to a benevolent agent paying the sales costs for current and future generations in that dynasty only.
-The results are reported in Table \ref{tab:adjcost}.
+Here, I present a more surprising result: Households with wealthy parents prefer illiquidity.\footnote{Throughout, ``illiquidity'' means the inability to costlessly adjust housing, captured by adjustment costs.
+Because mortgages are modeled as one-period risk-free bonds, owners away from the loan-to-value (LTV) constraint can extract equity costlessly.
+Introducing refinancing frictions would make housing less liquid and likely strengthen the mechanism.}
+
+I solve the model in two versions: the baseline with sales costs and a counterfactual without ($m_s=0$).
+For each household in the baseline distribution, I ask whether it would prefer a dynasty-specific waiver of sales costs.
+Results are in Table \ref{tab:adjcost}.
 
@@
 	\center
-	\caption{Household Observables and Support for Keeping Adjustment Cost}\label{tab:adjcost}
+	\caption{Household Observables and Support for Keeping Adjustment Costs}\label{tab:adjcost}
 	\begin{threeparttable}
@@
 First, future transfers increase MPCs by decreasing precautionary savings motives.
-This effect is small, and only increases the average MPC by about 3 percentage points.
+This effect is small, and only increases the average MPC by about 4 percentage points.
 The increase is almost entirely driven by unconstrained households.
 Second, current transfers provide partial insurance and decrease MPCs.
-This effect is large, reducing the MPC by 11 percentage points.
+This effect is large, reducing the MPC by 8 percentage points.
 The decrease is almost entirely driven by liquidity constrained households.
-Finally, altruism---including current and future transers---reduces the MPC from 0.35 to 0.25.
+Finally, altruism---including current and future transers---reduces the MPC from 0.32 to 0.26.
 Altruistic transfers lower the MPC of liquidity constrained households but increases it for unconstrained households---the ratio of MPC for liquidity constrained to unconstrained decreases from 3 to 2.
@@
 Following \cite{Kaplan2014}, households are liquidity constrained if they have negative liquid wealth $(b_k)$ exceeding 16.5\% of annual earnings.
-The MPC is calculated from a \$5,000 positive shock.
-To calculate the MPCs with future transfers only, we use the policy functions with altruism (baseline) but simulate households after unexpectedly confiscating any transfers.
-To calculate the MPCs with current transfers only, we use the parental policy function with altruism (baseline) and the children’s policy function without altruism.
-The children then receive the transfers but behave as if there are no future transfers.
+MPCs are calculated from a \$5,000 positive shock.
+``Future transfers only'' MPCs use the baseline (altruistic) policy functions but set the contemporaneous transfer at the shock date to zero, while leaving expectations about future transfers unchanged.
+``Current transfers only'' MPCs combine the parental policy function from the baseline with children’s policy functions solved without altruism; children receive the contemporaneous transfer but behave as if there are no future transfers.
 \end{flushleft}\end{footnotesize}}
@@
 \section{Conclusion}
-In this paper, I demonstrate that parental transfers play a large role in households' homeownership decisions.
-I build and estimate a dynamic life-cycle model of homeownership where altruistic parents may transfer to their adult children in every period.
-I use the model to quantify the importance of parental transfers for young adults' homeownership and to understand how policies that lower barriers to homeownership interact with these transfers.
+This paper shows that parental transfers are central to households’ homeownership decisions.
+I build and estimate a dynamic life-cycle model in which altruistic parents can make inter-vivos transfers to adult children each period.
+The model is used both to quantify the contribution of transfers to young adults’ homeownership and to study how policies that lower entry barriers interact with these transfers.
 
 Using a counterfactual experiment without altruism (a standard life-cycle model), I find that transfers account for 13 percentage points (27\%) of the homeownership rate.
-I show that policies that lower entry barriers to homeownership amplify the importance of parental wealth on housing outcomes, while those that lower the downsides of illiquid homeownership reduce the importance of parental wealth.
-Finally, I study the role of housing illiquidity and parental transfers.
-My paper thus contributes to a better understanding of the determinants of homeownership and to the growing literature on the role of family economics in shaping macroeconomic outcomes (e.g., \cite{Doepke2016a,Daruich2018}).
-
-Another contribution of the paper is to include an illiquid asset in the altruism framework.
-With altruistic parents, transfers generally flow to borrowing-constrained households, since they have a large marginal utility of wealth \citep{Barczyk2020,Chu2020}.
-With a single asset, this framework implies that households who receive transfers are poor.
-In the data, around 20\% of all households are ``wealthy hand-to-mouth'' and have positive wealth but no liquid wealth \citep{Kaplan2014a,Attanasio2018}.
-My paper bridges these two strands of the literature.
-First, I show that some households with wealthy parents choose to be liquidity-constrained to receive larger transfers.
-While it is difficult to test this prediction, my empirical results are consistent with this behavior: parental wealth improves housing outcomes after purchase, including ownership retention, lower delinquency, and less downsizing in unemployment.
-Second, I show that the lack of commitment generates preferences for illiquidity, even when households are rational and without behavioral biases \citep[see e.g.,][]{attanasio2024temptation}.
-
-I made several simplifying assumptions to keep the model tractable.
-Adding richer housing dynamics and equilibrium effects would allow analysis of other interesting questions, such as the extent to which transfers contribute to house price growth.
-My model omits a feedback loop, in which transfers raise housing demand, which leads to higher prices and thus increasing the wealth of homeowning parents and enabling further transfers.
-Another potential extension could examine whether illiquidity reduces the commitment problem in the family.
-Additionally, my results on preferences for illiquidity have implications for the literature on high-MPC households \citep[see e.g.,][]{kaplan2022marginal}.
-In particular, by ignoring transfers, existing work overestimates the average MPC of liquidity-constrained households, some of whom rely on parental transfers to smooth shocks.
+I then analyze policies in place or under debate---LTV caps, PMI rules, first-time-buyer assistance, and broker commissions---and show that policies lowering entry barriers amplify the dependence of housing outcomes on parental wealth, whereas policies that reduce the downside of illiquid ownership attenuate it.
+I also study the interaction of housing illiquidity and parental transfers.
+These results clarify the determinants of homeownership and contribute to the literature on family economics and macroeconomic outcomes (e.g., \cite{Doepke2016a,Daruich2018}).
+
+A further contribution is to integrate an illiquid asset into the altruism framework.
+With altruistic parents, transfers typically flow to borrowing-constrained children with high marginal utility of wealth \citep{Barczyk2020,Chu2020}.
+In a single-asset setting this would imply that most recipients are poor, yet data show that roughly 20\% of households are “wealthy hand-to-mouth,” holding positive net worth but little liquid wealth \citep{Kaplan2014a,Attanasio2018}.
+The model reconciles these facts: some children with wealthy parents choose to remain liquidity-constrained to elicit larger transfers, and the lack of commitment generates preferences for illiquidity even under full rationality \citep[cf.][]{attanasio2024temptation}.
+While it is difficult to test this prediction, my empirical results are consistent with this behavior: parental wealth improves post-purchase outcomes---higher ownership retention, lower delinquency, and less downsizing during unemployment.
+
+Several simplifying assumptions keep the model tractable.
+The housing block could be enriched with multiple house sizes, refinancing frictions, and full general equilibrium to capture quality margins and price feedbacks.
+The dynasty is modeled as a single parent–child pair; in reality, resources are pooled across parents (including remarried/cohabiting partners) and allocated across multiple children.
+Pooling can raise transfer capacity, while sibling dilution spreads support, making the net strength of the transfer motive ambiguous and heterogeneous across families.
+Future work could also study whether illiquidity mitigates intrafamily commitment problems.
+Finally, the results on illiquidity preferences speak to the high-MPC literature \citep[e.g.,][]{kaplan2022marginal}: models that ignore parental transfers likely overstate average MPCs among liquidity-constrained households, some of whom rely on family support to smooth shocks.
 
@@
 \section{Transfers and the Rent-to-Own Transition}\label{app:rent_to_own}
-A large literature documents that parental transfers increase the probability that renters become homeowners (e.g., \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Guiso2002}; \citealp{Engelhardt1998}), I replicate this result to highlight the importance of transfers for housing outcomes.
+A large literature documents that parental transfers increase the probability that renters become homeowners (e.g., \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Guiso2002}; \citealp{Engelhardt1998}).
+I replicate this result to highlight the importance of transfers for housing outcomes.
 
@@
 \end{equation}
-where \(Own_{i,t}=0\) denotes renting at time \(t\), \(Transfer_{i,t}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income, log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.
+where \(Own_{i,t}=0\) denotes renting at time \(t\), \(Transfer_{i,t+2}=1\) if the household received a transfer between survey waves, and \(X_{i,t}\) includes log parental wealth, log household income, log household net worth, log household size, age, parental age, and dummies for  education, marital status, race, household composition changes, and state and year fixed effects.
 The sample is limited to renting households aged 25 to 44 who have never been observed as homeowners.
@@
 In a univariate regression, receiving a parental transfer is associated with a 2.9 percentage point higher probability of becoming a homeowner in 2013 among renters in 2011.
-With a full set of controls parental transfers are associated with a 1.2 percentage point higher probability of becoming a homeowner.
+With a full set of controls, parental transfers are associated with a 1.2 percentage point higher probability of becoming a homeowner.
 This represents a substantial increase relative to the baseline rent-to-own transition rate of about 14\%, though the effect is not statistically significant.
@@
 First, the continuation value is now the parent's value function at age 25 of the new child ($V_{p}({\mathbf{s}'_{p,25}})$).
-Second, the expectation is over the new child's initial productivity and net worth ($x_{c,25},v_{c,25}\sim F(x_{c,53},y_{c,53}$)), instead of the old child's productivity.
+Second, the expectation is over the new child's initial productivity and net worth $(x_{c,25},y_{c,25})\sim F(x_{c,53},y_{c,53})$, instead of the old child's productivity.
 Third, the law of motion now includes bequests from the old parent instead of a transfer.
 Fourth, it is necessary to map next-period variables for the old child into variables for the new parent.
-The problem, for a child aged 53 who choose to own, becomes:
+The problem for a child aged 53 who choose to own becomes:
 \begin{equation*}
 \begin{split}
-V_c(\mathbf{s}_{c,53;\mathbf{g}_p})^{own} = &\max_{c_{c,53},b'_{c,53},h'_{c,53}=h_o} u(c_{c,53},h'_{c,53}) + \beta \E_{y_c}\left[V_{p}({\mathbf{s}'_{p,25}};\mathbf{g_c}) \right] \\
+V_c(\mathbf{s}_{c,53};\mathbf{g}_p)^{own} = &\max_{c_{c,53},b'_{c,53},h'_{c,53}=h_o} u(c_{c,53},h'_{c,53}) + \beta \E_{y_c}\left[V_{p}({\mathbf{s}'_{p,25}};\mathbf{g_c}) \right] \\
 \text{s.t.}\quad & 	b'_{c,53} = \tilde x_{c,53} - c_{c,53} - p h'_{c,53}  - m_b p h'_{c,53} \\
@@
 \end{equation*} 
-where $c,53$ and $p,53$ denote the variables associated with old child and old parent in this period, and $c,25$ and $p,25$ denote the variables associated with the new child and new parent in the next period.
+where $c,53$ and $p,53$ denote the variables associated with the old child and old parent in this period, and $c,25$ and $p,25$ denote the variables associated with the new child and new parent in the next period.
 The last equation maps the old child's housing choice into the new parent's---the same household one period later---housing state.
@@
 	V_p(\mathbf{s}_{c,53};\mathbf{g}_c)^{own} = &\max_{c_{p,53},b'_{p,53},h'_{p,53}=h_r} u(c_{p,53},h'_{p,53}) + \eta u(c_{c,53},h'_{c,53}) + \beta \eta \E_{y_c}\left[V_{p}({\mathbf{s}'_{p,25}},\mathbf{g}_p) \right] \\
-	\text{s.t.}\quad & b'_{p,25} = x_{p,53} + w_{p,53} - c_{p,53} - t_{p,53} - q p h_{p,53}'\\
+	\text{s.t.}\quad & b'_{p,25} = x_{p,53} + w_{p,53} - c_{p,53} - t_{p,53} - q p h'_{p,53}\\
 	& x_{p,53}' = b'_{p,53}(1+r(b'_{p,53})) \\
-	& t_{p,53}\ge0, \\
+	& t_{p,53}\ge 0, \\
 	& b'_{p,53}\ge 0. \\
@@
 \subsection{Measures of Households}\label{sec:hhmeasures}
-The state-space of a parent is $S_p = X_c \times X_p \times Y_c \times H_c \times H_p \times A_c$, with $\mathbf{s}_p$ denoting generic elements therein and $\mc{S}_p$ the associated Borel-$\sigma$ algebra.
+The state space of a parent is $S_p = X_c \times X_p \times Y_c \times H_c \times H_p \times A_c$, with $\mathbf{s}_p$ denoting generic elements therein and $\mc{S}_p$ the associated Borel-$\sigma$ algebra.
 The state space of a child is $S_c= B_p \times H_p \times X_c \times Y_c \times H_c \times A_c$, where $B_p=\mathbb{R}$.
@@
 
-\textit{Law of Motion for Dynasties with Children Aged 25-51:} The mass of households in state $\mathbf{s}_p$ is the mass of families adopting housing and savings policies that lead them to this state, adjusted by the probability of experiencing a given income shock:
+\textit{Law of Motion for Dynasties with Children Aged 25-51:} The mass of households in state $\mathbf{s}'_p$ is the mass of families adopting housing and savings policies that lead them to this state, adjusted by the probability of experiencing a given income shock:
 \begin{equation}\label{eq:LoM}
@@
 \end{equation}
-Note that the child's state $\mathbf{s}_c(\mathbf{s}_p)$ in these expressions depend on the choices of the parent which depend on their state. For example, 
+Note that the child's state $\mathbf{s}_c(\mathbf{s}_p)$ in these expressions depends on the parent's choices which depend on their state. For example, 
 \begin{equation}
@@
 
-\textit{Law of Motion for Children Aged 53:} In this special case, the distribution will depend on the choices of the new parent (old child), the now deceased parent and the stochastic initial conditions of the new child.
+\textit{Law of Motion for Children Aged 53:} In this special case, the distribution will depend on the choices of the new parent (old child), the now-deceased parent and the stochastic initial conditions of the new child.
 \begin{equation}\label{eq:LoM55}
@@
 \end{equation}
-where we limit $s_p$ to only include the subset of the state-space where $a_c=53$.
-The initial wealth and productivity of the child depends on the wealth $x_c$ and productivity $y_c$ of the new parent at age 53.
+where we limit $s_p$ to only include the subset of the state space where $a_c=53$.
+The initial wealth and productivity of the child depend on the wealth $x_c$ and productivity $y_c$ of the new parent at age 53.
 Further, all children start out as renters, and the next-period wealth of the new parent is savings plus bequests.
@@
 \section{Estimation}\label{app:SMM}
-In the internal estimation, I first draw $N=\parNest$ 3-dimensional candidate parameter vectors $\theta$, using Sobol sequencing.
+In the internal estimation, I first draw $N=\parNest$ three-dimensional candidate parameter vectors $\theta$, using Sobol sequencing.
 I then rank the vectors by the objective function, and use the top 5\% to set the bounds of a new Sobol sequence, from which I draw another {\parNest} candidate parameter vectors.
@@
 
-\begin{figure}\caption{Identification of $\eta,h_0,\chi$.}\label{fig:idall}
+\begin{figure}\caption{Identification of $\eta,h_o,\chi$.}\label{fig:idall}
 	\begin{subfigure}{\textwidth}\caption{Identification of $h_o$}\label{fig:hoid}
@@
 The results are reported in Table \ref{tab:robust}.
-We see that the introduction of income risk for the old has neglible effects and leaves the main findings intact.
+We see that the introduction of income risk for the old has negligible effects and leaves the main findings intact.
 Income risk for parents lower their homeownership rate a touch while increasing their wealth a wee bit.
-Parental transfers still account for 14 percentange points of the homeownership rate of young households.
+Parental transfers still account for 13 percentange points of the homeownership rate of young households.
 
@@
 \end{equation}
-The aggregate house prices are set to be $(p_l,p_n,p_h)=(0.7,1.0,1.3)p_{bench}$, where $p_{bench}$ is set to be the estimated value from Table \ref{tab:esttable}.
+Aggregate house prices are set to be $(p_l,p_n,p_h)=(0.7,1.0,1.3)p_{bench}$, where $p_{bench}$ is set to be the estimated value from Table \ref{tab:esttable}.
 I then solve the model with price uncertainty, but simulate the economy when house prices are at the normal level.
@@
     \par}    {\begin{footnotesize} \textit{Notes:}
-        Dashed lines are the empirical age-medians, and solid lines are fitted second-order polynomials used in model calibration.
+        Dashed lines are the empirical medians, and solid lines are fitted second-order polynomials used in model calibration.
 The lines denote the value of $\nu_{i,a}$ in the first tertile (black), middle (gray), and top (blue) by age.
@@
 Due to the nonconvex nature of the decision problems---due to discrete housing choices, noncontinuous transfer choices, and occasionally binding constraints---I use grid search.
-That is, I define a grid of possible choices for each choice variable, and for each state I loop over all possible choices, and the optimal policy is the one that maximizes the utility.
+That is, I define a grid of possible choices for each choice variable and, for each state, I loop over all possible choices, and the optimal policy is the one that maximizes the utility.
 The value and policy functions are linearly interpolated over child and parental wealth, conditional on the discrete choices.
 This solution algorithm satisfies both criteria established by \cite[p.30]{Barczyk2020}: ``algorithms should be able to deal with locally convex and even discontinuous value functions.'' Naturally, the solution accuracy improves with denser grids.
-I use {\parnstate} nodes in the state vectors and {\parnchoice} nodes in choice variable grids.
-
-\textit{Solving Decisions Problems:} 
+I use {\parnstate} nodes for the state variables and {\parnchoice} nodes for choice variables.
+
+\subsection{Solving Decision Problems} 
 The parent's choice can be written as a three-stage problem, which increases computational speed dramatically.
@@
 
-\textit{Value Function Iteration Procedure:}
+\subsection{Value Function Iteration Procedure}
 I use the following procedure for value function iteration. The initial guess is set to zero for all functions, except for consumption, set to 1.0, and housing, set to renting.
 \begin{enumerate}
-    \item Next, I solve the decision problem of the child at age 53, that is the final period before he becomes a parent.
+    \item Solve the decision problem of the child at age 53, that is the final period before he becomes a parent.
 This yields $V_c(\cdot;a_c=53)$ and $g_c(\cdot;a_c=53)$.
@@
 The initial state of the new child is given by the joint distribution of initial productivity and wealth.
-    \item Next, I solve the parent's decision problem in the terminal age of 83, when the child is 53.
-This yields $V_p(\cdot;a_c=53)$ and $g_p(\cdot;a_c=53)$, which depend on the policy functions for the child at age 53 ($g_c(\cdot;a_c=53)$)---found in the previous step---and the new parent's value function $V_p(\cdot;a_c=25)$---which we still haven't found and is set to 0.
-    \item Next, I solve the child's decision problem at age 51.
-This yields $V_c(\cdot;a_c=51)$ and $g_c(\cdot;a_c=51)$, and depends on the value function we found in the first step ($V_c(\cdot;a_c=53)$) and the parent's policy function we found in the previous step ($g_p(\cdot;a_c=53)$).
-    \item Next, I solve, the parent's decision problem at age 81.
-This yields $V_p(\cdot,a_c=51)$ and $g_p(\cdot,a_c=51)$, and depends on the policy function for the child we found in the previous step ($g_c(\cdot;a_c=51)$) and the value function we found in the second step ($V_p(\cdot;a_p=83)$).
-    \item This is repeated backwards in age until the child is 25.
+    \item Solve the parent's decision problem in the terminal age of 83, when the child is 53.
+This yields $V_p(\cdot;a_c=53)$ and $g_p(\cdot;a_c=53)$, which depend on the policy functions for the child at age 53 ($g_c(\cdot;a_c=53)$)---found in the previous step---and the new parent's value function $V_p(\cdot;a_c=25)$---which we still have not found and is set to 0.
+    \item Solve the child's decision problem at age 51.
+This yields $V_c(\cdot;a_c=51)$ and $g_c(\cdot;a_c=51)$, and depends on the value function from the first step ($V_c(\cdot;a_c=53)$) and the parent's policy function we found in the previous step ($g_p(\cdot;a_c=53)$).
+    \item Solve the parent's decision problem at age 81.
+This yields $V_p(\cdot;a_c=51)$ and $g_p(\cdot;a_c=51)$, and depends on the policy function for the child from the previous step ($g_c(\cdot;a_c=51)$) and the value function we found in the second step ($V_p(\cdot;a_p=83)$).
+    \item Repeated backwards in age until the child is 25.
     \item Repeat these steps until the value functions converge.
@@
 
-\textit{Fixed Point Convergence:} 
+\subsection{Fixed Point Convergence}
 The parent-child interaction forms an infinitely repeated game within a dynasty, meaning the equilibrium may not be unique.
-In practice, value function iterations generally converge to a unique fixed poin, but when the altruism parameter is high, the equilibrium can become cyclical, with the solution alternating between iterations.
-For example, in iteration $n$ one solution appears, in $n+1$ another, and in $n+2$ the cycle returns.
-These behavioral differences are minor: the child homeownership rate varies by less than 0.3 percentage points, and average child wealth differs by less than 0.1 percent.
-While this cyclical behavior does not occur in the reported results, it sometimes occurs during the structural estimation, when the model is solved with higher altruism parameters $\eta$.
+In practice, value function iterations generally converge to a unique fixed point, but when the altruism parameter is high, the equilibrium can become cyclical, with the solution alternating between iterations.\footnote{Some factors likely contribute to the uniqueness of the equilibrium.
+First,  the biennial period length implies a relatively low discount factor $\beta$.
+Second, the estimated altruism parameter is modest.
+Third, although dynasties are infinite, each household lives finitely, curbing the strategic multiplicity noted by \citet[Footnote 9]{Barczyk2020a}.
+Fourth, the game is sequential.
+Fifth, idiosyncratic income shocks help smooth the value functions.} For example, in iteration $n$ one solution appears, in $n+1$ another, and in $n+2$ the cycle returns.
+These behavioral differences are minor: the child homeownership rate varies by less than 0.3 percentage points, and average child wealth differs by less than 0.1 percent between cycles.
+While this cyclical behavior does not occur in the reported results, it sometimes occurs during the structural estimation when the model is solved with higher altruism parameters $\eta$.
 In such cases, I simulate each dynasty once for every candidate solution.
+The fixed point does not change when I change the initial guess.
 
@@
 Similarly, removing transaction costs, making housing more liquid and no longer a state variable, reduces the likelihood of cyclical fixed points.
-Denser grids also decreases the likelihood of cyclical behavior.
+Denser grids also decrease the likelihood of cyclical behavior.
 In practice, the largest differences in policy functions within cycles are at the upper edges of the parental wealth grid.
@@
 
-\textit{Interpolation Details:} Because the policy functions for consumption and savings are discontinuous in both child and parental wealth (due to discrete housing choices and altruistic transfers), direct interpolation is unattractive.
+\subsection{Interpolation Details} Because the policy functions for consumption and savings are discontinuous in both child and parental wealth (due to discrete housing choices and altruistic transfers), direct interpolation is unattractive.
 However, conditional on discrete housing choices, the policy functions are smoother.
@@
 		\item If not, it takes a value on $(0,1)$.
-This happens only when the housing choice is not the same at all the four nearest nodes used to calculate the interpolation weights.
+This happens only when the housing choice is not the same at all four nearest nodes used to calculate the interpolation weights.
 Randomly assign the parent a housing choice.
@@
 
-\textit{Simulation of Households:} 
+\subsection{Simulation of Households} 
 I simulate N={\parNdyn} dynasties.
-The initial states of each dynasty (initial child and initial old) $(x_p,h_p,x_c,v_c,h_c,a_c=25)$ is drawn from a five-dimensional joint uniform distribution.
+The initial states of each dynasty (initial child and initial old) $(x_p,h_p,x_c,v_c,h_c,a_c=25)$ are drawn from a five-dimensional joint uniform distribution.
 I then simulate all dynasties for {\parNdyn} generations, since the distribution, as measured by average homeownership, wealth, and, productivity levels stabilizes after four generations.
@@
 
-\textit{Computational Packages:}
-The program to solve the model is written in Julia v1.10.4.
+\subsection{Computational Packages}
+The model is solved in Julia v1.10.4.
 I rely on the \texttt{interpolations.jl} v0.15.1 package for numerical interpolation routines.
-The Stata code used for the empirical analysis and the Julia code for the model is available through the \href{https://github.com/eirikbrandsaas/HomeownershipBankMomDad.jl}{author's website and Github}.
+The Stata code used for the empirical analysis and the Julia code for the model is available on my \href{https://github.com/eirikbrandsaas/HomeownershipBankMomDad.jl}{website and Github}.
 
 \section{Event Study Details}\label{app:eventstudy_details}
-In the end, the sample consists of 3,553 households with non-missing log growth rates of housing consumption from 948 households with one observed unemployment spells.
+In the end, the sample consists of 3,553 observations with non-missing log growth rates of housing consumption from 948 households with one observed unemployment spells.
 Means and standard errors are constructed using family weights.
 
+Housing consumption is measured at the household level.
+If a household takes on roommates or cohabits with non-parents, the unit remains a distinct household and housing expenses continue to be observed.
+We do not observe housing expenses (or employment status) if the household ceases to be defined as a PSID household, which may occur if it moves back home with parents.
+
+\subsection{Varying the Threshold}
+In the benchmark, I define ``high parental wealth’'' as above the 75th percentile.
+Figure~\ref{fig:housinggrowthrates_threshold} sweeps this cutoff from the 10th to the 90th percentile in 5-point steps.
+The patterns are essentially unchanged: households with less-wealthy parents downsize at unemployment, whereas those with wealthier parents do not.
+The profiles vary smoothly with the cutoff---though confidence bands widen in the extreme tails where cell sizes are small.
+Taken together, this suggests the mechanism operates broadly across the parental-wealth distribution and strengthens with parental wealth.
+
+\begin{figure}
+	\caption{Event Study at Various Cutoffs: Housing Consumption}\label{fig:housinggrowthrates_threshold} 
+	\noindent%
+	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoorall_both}%
+	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrichall_both}
+
+	{\begin{footnotesize} \textit{Notes:} Markers denote means, and bars denote the 95\% confidence interval.
+Lighter colors indicate lower thresholds for parents to be considered wealthy.
+The right panel plots households parents whose wealth is above the threshold.\end{footnotesize}}
+\end{figure}
+
+\subsection{Adding Control Variables}
 The main event study does not control for household characteristics for three reasons.
@@
 Third, for ease of interpretation.
-I now show that the results are robust to including household controls.
 
@@
 I then run linear regressions where I interact years relative to unemployment with a dummy for having wealthy parents at unemployment.
-The set of controls include dummies for children's wealth quintiles, a full set of age, year, and state dummies, and dummy variables for college, high-school, marriage, race, log income and log family size.
-The standard errors are clustered at the household level.
-The results are plotted in figure \ref{fig:housinggrowthrates_controls}.
-The main result is that households with wealthier parents do not significantly decrease housing consumption at unemployment while those with poor parents, on average, decrease housing consumption by 8 percent ($p<0.05$).
+The set of controls includes dummies for children's wealth quintiles, a full set of age, year, and state dummies, and dummy variables for college, high school, marriage, race, log income and log family size.
+Standard errors are clustered at the household level.
+The results are plotted in Figure \ref{fig:housinggrowthrates_controls}.
+The main result is that households with wealthier parents do not significantly decrease housing consumption at unemployment, while those with poor parents, on average, decrease housing consumption by 8 percent ($p<0.05$).
 However, we cannot reject the hypothesis that the effect of unemployment is the same for both groups.
 \begin{figure}
-	\caption{Event Study: Housing Consumption at Unemployment by Parental Wealth}\label{fig:housinggrowthrates_controls}
+	\caption{Event Study with Controls: Housing Consumption}\label{fig:housinggrowthrates_controls}
 	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both_controls}%
@@
 The reference (omitted) category is four years before unemployment.
+The right panel plots households whose parent's wealth is above the threshold.
 \end{footnotesize}}
@@
 		\textit{Transfer Related} \\ 
-		Received Transfer & ER67962 & 2/5years, gift/inherit \$10,000+ & Changing def. \\
-		Gave Transfer & RT13V125 & Loans/gifts to child in 2012 & 2013
-		prnt/chld file \\ 
+		Received Transfer & ER67962 & 2/5 years, gift/inherit \$10,000+ & Changing def. \\
+		Gave Transfer & RT13V125 & Loans/gifts to child in 2012 & 2013 prnt/chld file \\ 
 		Transfer Amount & RT13V125 & Amount given in 2012 & 2013 prnt/chld file \\
@@
 		Behind Mortgage & ER66062 & Behind on mortgage payments\\
-		Income & ER65349 & Total Household Income & \\
-		Employment & ER66164 & Working, Unemployed etc. \\
-		House Value & ER60031 & Reported Market Value & \\
-		Dollars Rent &	ER66090 & Monthly Rent\\
+		Income & ER65349 & Total household income & \\
+		Employment & ER66164 & Working, unemployed etc. \\
+		House Value & ER60031 & Reported market value & \\
+		Dollars Rent & ER66090 & Monthly rent\\
 		Family Weight &	ER71570 & Weight of family unit & \\
@@
 	\end{tabular}
-	{\textit{Notes:} This table lilsts the main variables I use from the PSID, as well as their variable code in the 2017 Main family level data set or the 2013 Transfer supplement.}
+	{\textit{Notes:} This table lists the main variables I use from the PSID, as well as their variable codes in the 2017 Main family-level data set or the 2013 Transfer Supplement.}
 	\end{threeparttable}
@@
 \begin{figure}\caption{Initial Distribution $F(x_{53},y_{53})$ by wealth $x_{53}$ and productivity $y_{53}$}\label{fig:inidistr}
-
 		\includegraphics[width=1\textwidth]{../tabfig/empirical/inidistr}
-		{\begin{footnotesize}\textit{Notes:} The vertial lines denote the first, second, and third productivity level for the child.
-Within each interval each point denotes a wealth quartile.
-For example, the first point on the black line in the left-most panel gives the probability that a new parent in the first wealth quartile (black lines) and the first income tertile (left panel), has a child in the lowest income tertile and lowest wealth tertile.
+		{\begin{footnotesize}\textit{Notes:} The vertical lines denote the first, second, and third productivity level for the child.
+Within each interval, each point denotes a wealth quartile.
+For example, the first point on the black line in the left-most panel gives the probability that a new parent in the first wealth quartile (black lines) and the first income tertile (left panel) has a child in the lowest income tertile and lowest wealth quartile.
 The second point gives the probability that the child is in the second wealth quartile but first income tertile.

### 2025-09-10 docs/Homeownership_BankOfMomAndDad.tex
added short abstract
some shortening

@@
 Finally, I show that children of wealthy parents strategically use the illiquidity of housing as a commitment device to encourage transfers, resulting in a preference for illiquidity.
+
+% <=100 word abstract This paper quantifies how parental transfers shape young adults’ homeownership.
+I estimate a life-cycle OLG model with housing in which parents and children interact without commitment.
+Parental transfers account for 13 percentage points (27\%) of young households’ homeownership.
+Transfers from wealthy parents both relax borrowing constraints and sustain ownership by mitigating housing illiquidity.
+Policy experiments show that easing entry barriers increases reliance on parental wealth, whereas increasing housing liquidity reduces it.
+Finally, I show that children of wealthy parents strategically use the illiquidity of housing as a commitment device to encourage transfers, resulting in a preference for illiquidity.
 \end{spacing}}
@@
 \section{Data on Transfers, Family, and Housing}\label{sec:data}
-I first present time trends in parental transfers for down payments, before describing the estimation sample taken from the PSID.
+I first present time trends in parental transfers for down payments, before describing the estimation sample.
 Finally, I use the estimation sample to show that parental wealth is positively associated with better housing outcomes, both before and after purchase.
@@
 Table \ref{tab:descrstats} contains the means of variables, by age, wealth, homeownership, and transfer receipt.
-Though limited to households with observed parents, the sample appears reasonably representative.
+Though limited to households with observed parents, the sample appears representative.
 Average income by age closely matches estimates from the SCF \cite[see e.g.,][]{kuhn20162013}, though average wealth is slightly lower.
 
-There are several main takeaways from the subgroup analysis.
 In 2012, 21\% of young households received a transfer, and transfers averaged \$3,960.
@@
 
-\section{A Quantitative Model of Parental Transfers and Homeownership}\label{sec:model}
+\section{A Model of Transfers and Homeownership}\label{sec:model}
 This section describes my life-cycle model of housing choices with overlapping generations, idiosyncratic earnings risk, and altruistic inter-vivos transfers from parents to their offspring.
@@
 
-\textit{Period Length:} Each period corresponds to two years, matching the PSID interview frequency.
-I report all parameters in biennial form (e.g., $\beta$ measures how much households discount two years).
+\textit{Period Length:} Each period corresponds to two years, matching the PSID interview frequency. I report all parameters in biennial form (e.g., $\beta$ discounts two years).
 
@@
 Within each age, I set $v_{i,a}$ as the ratio of the median income for that tertile to the overall median income (Figure \ref{fig:y_ia}).
-The empirical age-dependent transition matrices provide the transition matrix (Figure \ref{fig:Piv}).
+The age-dependent transition matrices are taken from the data (Figure \ref{fig:Piv}).
 
@@
 \textit{Initial Conditions of the Young $y_{c,25},x_{c,25}\sim F(x_{c,53},y_{c,53})$:} 
-The initial distribution is set to match the intergenerational correlation in wealth and income.
-I estimate it nonparametrically for transparency.
+The initial distribution is set to match the intergenerational correlation in wealth and income, estimated nonparametrically for transparency.
 First, I limit the sample to households aged 24 to 26 with parents 16 to 40 years older.
@@
 The PDF $F(\cdot)$ is then given by the empirical probability of each parent-child combination.
-The wealth level of a quartile is defined by its median value (Fig.
+The wealth level of a quartile is its median value (Fig.
 \ref{fig:inidistr})
@@
 I use a closed-form supply function because the model is too complex to introduce a construction sector.
-The rent-to-price ratio is unchanged, as in \cite{Kaplan2020}, where rental units are supplied by deep-pocketed landlords who convert rentals to owner-occupied units if the price deviates from the present value of rents.
+The rent-to-price ratio is unchanged, as in \cite{Kaplan2020}, where rental units are supplied by deep-pocketed landlords who convert them to owner-occupied units if the price deviates from the present value of rents.
 Letting $\alpha_1\to\infty$ yields a perfectly elastic supply function, leaving prices constant, as in the previous section.
-I first set the elasticity to 3, a typical estimated value for the United States \cite[see e.g.,][]{saiz2010geographic,aastveit2023changing}.
-Finally, I set the elasticity to 1 as a reasonable lower bound for the long-run aggregate elasticity.
+I experiment with setting the elasticity to 3, a typical estimated value for the United States \cite[see e.g.,][]{saiz2010geographic,aastveit2023changing}and to 1, a reasonable lower bound for the long-run aggregate elasticity.
 
@@
 
-\textit{Tighter LTVs:} LTV caps, a standard macroprudential tool, directly regulate down-payment requirements; as shown in Figure~\ref{fig:motivation}, reliance on down-payment assistance moves with changes in mortgage regulation.
+\textit{Tighter LTVs:} LTV regulation is a standard macroprudential tool and as shown in Figure~\ref{fig:motivation} reliance on down-payment assistance moves with changes in mortgage regulation.
 I tighten the LTV limit by 3 percentage points.
@@
 Homeownership rises by 1 percentage point (2\%).
-As with the LTV rule, gains are larger for households with parents in the middle of the wealth distribution, further widening the parental-wealth gradient.
+As with the LTV, gains are larger for households with parents in the middle of the wealth distribution, further widening the parental-wealth gradient.
 Thus, relaxing PMI increases both overall homeownership and the importance of parental wealth.
 
-\textit{Lowering House Prices:} Although there is no single “price lever,” policy shapes prices through supply regulation, taxes, interest rates, and subsidies.
+\textit{Lowering House Prices:} While policy does not set house prices, prices are affected by supply regulation, taxes, interest rates, and subsidies.
 I capture these channels in reduced form by lowering house prices by 2\%.
@@
 \textit{Reducing Purchase Costs:}
-Closing-cost programs and first-time-buyer relief---e.g., England’s 2017 stamp duty cut for first-time buyers---directly lower entry costs; to study this margin’s interaction with parental transfers, I lower the purchase cost ($m_b$) by 2 percentage points, equivalent to a grant of about \$6,000.
+Closing-cost programs and first-time-buyer relief---e.g., England’s 2017 stamp duty cut for first-time buyers---directly lower entry costs.
+I lower the purchase cost ($m_b$) by 2 percentage points, equivalent to a grant of about \$6,000.
 While homeownership rises, the gains are concentrated among households with wealthy parents, amplifying the parental wealth gradient.
@@
 
+% \cleardoublepage  % This ensures you start on a new page
+% \thispagestyle{empty}  % This removes page number and headers/footers
+% \vspace*{\fill}  % This pushes the text to the vertical center of the page
+% \begin{center}
+% \Large\textbf{Online Appendix}  % This creates the text, centered and bold
+% \end{center}
+% \vspace*{\fill}  % This completes the vertical centering
+% \clearpage  % This ends the current page
 \appendix
@@
 A large literature documents that parental transfers increase the probability that renters become homeowners (e.g., \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Guiso2002}; \citealp{Engelhardt1998}).
-I replicate this result to highlight the importance of transfers for housing outcomes.
-
-I estimate
+I replicate this result to highlight the importance of transfers for housing outcomes, by estimating
 \begin{equation}	\label{eq:transferrent}
@@
 	{\centering
-    \includegraphics[width=0.5\textwidth]{../tabfig/empirical/lifecycleproductivity_old_3}
+    \includegraphics[width=0.45\textwidth]{../tabfig/empirical/lifecycleproductivity_old_3}
     \par}    {\begin{footnotesize} \textit{Notes:}
@@
 	\noindent%
-	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoorall_both}%
-	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrichall_both}
+	\includegraphics[width=0.45\textwidth]{../tabfig/descr/PSID_housinggrowthpoorall_both}%
+	\includegraphics[width=0.45\textwidth]{../tabfig/descr/PSID_housinggrowthrichall_both}
 
@@
 	\caption{Event Study with Controls: Housing Consumption}\label{fig:housinggrowthrates_controls}
-	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both_controls}%
-	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both_controls}
+	\includegraphics[width=0.45\textwidth]{../tabfig/descr/PSID_housinggrowthpoor_both_controls}%
+	\includegraphics[width=0.45\textwidth]{../tabfig/descr/PSID_housinggrowthrich_both_controls}
 

### 2025-09-11 docs/test.tex
old latex is corruped, work on this one in the meantime

--- 2025-09-10 docs/Homeownership_BankOfMomAndDad.tex
@@
 Additionally, parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}.
-However, while these studies show that transfer receipt increases the probability of renters becoming homeowners, they do not address the overall importance of parental transfers.
-For example, households may anticipate future transfers and subsequently save less.
-In this case, even if transfers increase the likelihood of homeownership, all else being equal, they may not affect aggregate homeownership.
-Furthermore, this is the first paper quantifying the importance of parental transfers as a determinant of homeownership; the existing literature has primarily focused on family formation and credit constraints \citep[see e.g.,][]{Chang2024,Paz-Pardo2019,Mabille2020}.
+However, while these studies show that transfer receipt increases the probability of renters becoming homeowners, they do not establish the overall importance of parental transfers.
+Households may anticipate future transfers and reduce their own saving, so that the observed boost in homeownership conditional on receiving a transfer overstates the aggregate effect of transfers on homeownership.Furthermore, this is the first paper quantifying the importance of parental transfers as a determinant of homeownership; the existing literature has primarily focused on family formation and credit constraints \citep[see e.g.,][]{Chang2024,Paz-Pardo2019,Mabille2020}.
 
 I first present empirical evidence showing that parental wealth helps not only with the transition into ownership, but also in retention of ownership.
-In addition to replicating previous findings that transfers increase the probability of becoming a homeowner and that parental wealth is associated with sustained ownership \citep[see e.g.,][]{Lee2018,bond2021role}, I provide novel results on the mechanism through which parental wealth aids retention.
-Controlling for a wide set of household and parental characteristics, I show that households with wealthier parents are less likely to fall behind on mortgage payments.
+In addition to replicating previous findings that transfers increase the probability of becoming a homeowner and that parental wealth is associated with sustained ownership \citep[see e.g.,][]{Lee2018,bond2021role}, I provide novel evidence on one mechanism: households with wealthier parents are less likely to fall behind on mortgage payments, consistent with parental resources providing insurance against income shocks.
 Finally, using a variation of the event study in \cite{Chetty2007}, I show that while households with less wealthy parents downsize their housing during unemployment spells, those with wealthy parents retain their current homes.
@@
 
-I then examine the interplay  of transfers and housing for marginal propensities to consume (MPC).
+Finally, I examine the interplay  of transfers and housing for marginal propensities to consume (MPC).
 A substantial literature has sought to reconcile structural models with the high MPCs observed in empirical data.
@@
 
-	\caption*{\begin{footnotesize} \textit{Notes:}  Left panel (SHED): sample is current owners reporting this is their first home; x-axis is purchase year; outcome is the share receiving any down-payment transfer.
-Right panel (AHS): sample is owner-occupied units that do not report proceeds from a prior home sale as the main funding source; x-axis is survey year; outcome is the share whose down payment was mostly funded by a transfer.
+	\caption*{\begin{footnotesize} \textit{Notes:}  Left panel (SHED): first-time owners by purchase year; outcome is the share receiving a down-payment transfer.
+Right panel (AHS): owner-occupied units without prior-sale proceeds by survey year; outcome is the share whose down payment was primarily funded by a transfer.
+Higher levels in the SHED reflect its broader definition.
 Dashed lines denote 95\% confidence intervals.\end{footnotesize}}
@@
 In 2012, 21\% of young households received a transfer, and transfers averaged \$3,960.
-Receivers have significantly richer parents, have similar wealth and income as nonreceivers, and are less likely to own.
+Receivers have parents about 2.5 times richer, have similar wealth and income as nonreceivers, and are less likely to own.
 Receivers are more likely to transition from renting to owning, especially in the age groups where households are most likely to buy.
@@
 
-Next, I break the sample down by ownership.
-Homeowners are both wealthier and have wealthier parents than renters.
+Next, I split the sample by ownership.
+Homeowners are wealthier and have wealthier parents than renters.
 Renters and owners receive transfers at about the same rate and size.
-Receivers are more likely to switch from renting to owning: 21\% of receiving owners rented two years ago, compared to 14\% of nonreceiving owners.
+Among current owners, 21\% of receivers rented two years earlier, compared with 14\% of nonreceivers, suggesting that transfers ease the transition into ownership.
 
@@
 
-To understand which variables predicts retention of ownership, I estimate
+To understand which variables predict retention of ownership, I estimate
 \begin{equation}	\label{eq:regmaintain}
@@
 With the full set of controls, the effect increases slightly, though it is no longer statistically significant.
-These magnitudes are economically meaningful since 1.6\% of first-time owners are behind on their mortgage.
+These magnitudes are economically meaningful since 1.6\% of first-time owners are behind on their mortgage and parents of receivers are about 2.5 richer on average.
 In Columns 3 to 5, I expand the analysis to include all owners who are currently not behind, increasing the sample sizes and allowing us to include household fixed effects.

### 2025-09-11 docs/test.tex
minor changes
Small fix

@@
 \subsubsection{Households with Wealthier Parents Are Less Delinquent} 
-One potential roadblock to maintaining ownership is mortgage delinquency, as missed payments can lead to distressed or forced sales.
+One roadblock to maintaining ownership is mortgage delinquency, as missed payments can lead to distressed or forced sales.
 I show that households with wealthier parents are less likely to fall behind on their mortgages, despite purchasing more expensive homes.
@@
 h & \text{ if renting}, \\
-\chi h & \text{ if owning}.
+\chi \times h & \text{ if owning}.
 \end{cases}

### 2025-09-11 docs/test.tex
Feds feedback (#17)
* some shortening/text edits

* Make error bars thicker

---------

Co-authored-by: Eirik Eylands Brandsaas <eirik.e.brandsaas@frb.gov>

@@
 E-mail: eirik.e.brandsaas@frb.gov.
-I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts, Corina Boar, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobson, William Gamber, Claudia Sahm, John Sabelhaus, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, Boston Fed, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, YES 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop.
+I have benefited from the comments of Dean Corbae, Erwan Quintin, Ananth Seshadri, Corina Mommaerts, Corina Boar, Manuel Amador, Hamish Low, Patrick Moran, Dionissi Aliprantis, David Sturrock, Mike Eriksen, Aradhya Sood, Margaret Jacobson, William Gamber, Claudia Sahm, John Sabelhaus, Paul Lengermann, Raven Molloy, and Joel McMurry, as well as seminar participants Oxford Sa{\"i}d, Federal Reserve Board, Cal State Fullerton, Boston Fed, ECARES/Université libre de Bruxelles, Santa Clara, University of Lugano, NYU Furman, Norges Bank, BI Norwegian Business School, UW-Madison, and conference participants at NASEMS, AREUEA National, ReCapNet, Minneapolis Fed/OIGI, UEA EU, CFPB, CEF, SAET, Oxford NuCamp Macro Workshop, YES 2020 at UPenn, and Minnesota-Wisconsin Macro/International Workshop.
 Code and replication instructions are available at the \href{https://github.com/eirikbrandsaas/HomeownershipBankMomDad.jl}{author's  website and Github}.\newline \textit{Disclaimer: }The views expressed in this paper are solely the responsibility of the author and should not be interpreted as reflecting the views of the Board of Governors of the Federal Reserve System or of anyone else associated with the Federal Reserve System.}}

### 2025-09-15 docs/Homeownership_BankOfMomAndDad.tex
FEDSstuff

@@
 \begingroup\singlespacing
+\nocite{AHS,PSID,SHED}
 \bibliographystyle{ecta}

### 2025-10-16 src/Julia/run_all.jl
Taking on comments

@@ -170,7 +170,7 @@ open("tabfig/parameters.tex", "w") do f
     write(f,"\\newcommand{\\parmsperc}{$(round(Int,mp.ms[1]*100))}\n")
     write(f,"\\newcommand{\\parmb}{$(mp.mb[1])}\n")
     write(f,"\\newcommand{\\parmbperc}{$(round(Int,mp.mb[1]*100))}\n")
-    write(f,"\\newcommand{\\parprice}{$(round(mp.price;digits=2))\$\\times h_o\$}\n")
+    write(f,"\\newcommand{\\parprice}{$(round(mp.price;digits=2))}\n")
     write(f,"\\newcommand{\\parhr}{$(np.h_grd[1])}\n")    
 
     write(f,"\\newcommand{\\parNdyn}{$(np.Ndyn)}\n")    

### 2025-10-17 docs/Homeownership_BankOfMomAndDad.tex
Re stud (#20)
* some shortening

* old latex is corruped, work on this one in the meantime

* minor changes

* Make error bars thicker

* Better figures

* Small fix

* Update text

* FEDSstuff

* Taking on comments

---------

Co-authored-by: Eirik Eylands Brandsaas <eirik.e.brandsaas@frb.gov>

@@
 This paper quantifies the importance of parental transfers for young adults' homeownership and investigates how policies lowering homeownership barriers interact with parental transfers.
-A growing literature emphasizes that families matter for economic outcomes \citep{Doepke2016a}.
-In the context of homeownership, previous empirical studies highlight that parental transfer receipt is associated with an increase in the probability of becoming a homeowner, mainly by relaxing credit constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}.
-Additionally, parental transfers help households maintain homeownership, a mechanism that has received less attention \citep[see e.g.,][]{bond2021role}.
-However, while these studies show that transfer receipt increases the probability of renters becoming homeowners, they do not establish the overall importance of parental transfers.
-Households may anticipate future transfers and reduce their own saving, so that the observed boost in homeownership conditional on receiving a transfer overstates the aggregate effect of transfers on homeownership.Furthermore, this is the first paper quantifying the importance of parental transfers as a determinant of homeownership; the existing literature has primarily focused on family formation and credit constraints \citep[see e.g.,][]{Chang2024,Paz-Pardo2019,Mabille2020}.
+Previous empirical studies show that receiving a transfer increases the probability of becoming a homeowner, mainly by relaxing credit constraints \citep[see e.g.,][]{Lee2018,Blickle2019,wold2024housing}, and that transfers help households maintain ownership \citep[e.g.,][]{bond2021role}.
+Yet these estimates capture only the direct effect of receiving a transfer, holding expectations and savings behavior fixed.
+In reality, households anticipate future support and adjust their saving and tenure decisions accordingly.
+Observed correlations between transfers and homeownership may thus overstate the aggregate role of transfers.
+I address this limitation by building a structural life-cycle model in which expectations of parental transfers, savings, and homeownership are jointly determined.
+Finally, this is the first paper quantifying the importance of parental transfers as a determinant of homeownership, whereas the existing literature has primarily focused on family formation and credit constraints \citep[see e.g.,][]{Chang2024,Paz-Pardo2019,Mabille2020}.
 
@@
 As an illustration, consider a child who buys a house and brings no liquid wealth to the next period.
-If the parent does not transfer in the next period, the child will either face very low consumption—implying a high marginal value of wealth—or be forced to liquidate the house and incur sales costs.
+If the parent does not transfer in the next period, the child will either face very low consumption---implying a high marginal value of wealth---or be forced to liquidate the house and incur sales costs.
 The parent dislikes both outcomes and may find it optimal to transfer enough to keep the child in the home.
@@
 
-	 \caption*{\begin{footnotesize} \textit{Notes:} The left panel uses the transfers, gifts, and inheritances variable, which is available in all waves of the PSID, while the right panel uses the parental-transfer data only available in the 2013 transfer supplement.
+	 \caption*{\begin{footnotesize} \textit{Notes:} The left panel uses the parental-transfer data only available in the 2013 transfer supplement, while the right panel uses the transfers, gifts, and inheritances variable, which is available in all waves of the PSID.
 See main text for a discussion.
@@
 \subsubsection{Transfers and the Transition into Homeownership}
-Since many papers have already estimated the relationship between parental transfers and transitions into homeownership (e.g., \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Lee2018}), I replicate these regressions to highlight the association between transfer receipt and rent-to-own transitions, but relegate the full results to Appendix~\ref{app:rent_to_own}.
+Since many papers have already estimated the relationship between parental transfers and transitions into homeownership (e.g., \citealp{wold2024housing}; \citealp{Blickle2019}; \citealp{benetton2022dynastic}; \citealp{Lee2018}; \citealp{kurz2004labour}; \citealp{spilerman2012parental}), I replicate these regressions in my sample to highlight the association between transfer receipt and rent-to-own transitions, but relegate the full results to Appendix~\ref{app:rent_to_own}.
 Receiving a parental transfer above \$500 is associated with a 1.2 percentage point increase in the probability of becoming a homeowner (not statistically significant), after controlling for household characteristics.
@@
 Recipients are more likely to transition from renting to owning, after conditioning on a rich set of covariates.
-Parental wealth is also positively associated with maintaining homeownership.
-Households with richer parents are less likely to fall behind on mortgages and to downsize during unemployment spells.
-These results suggest that homeowners with wealthy parents find the illiquidity of homeownership less problematic.
-
-Throughout the empirical section, I do not treat transfer receipt or parental wealth as exogenous, and the estimates document associations rather than causal effects.
+Parental wealth is also positively associated with maintaining homeownership: Households with richer parents are less likely to fall behind on mortgages and to downsize during unemployment spells, suggesting that homeowners with wealthy parents find the illiquidity of homeownership less problematic.
+
+That said, the infrequency of homeownership transitions and small sample size limit statistical precision.
+Moreover, I do not treat transfer receipt or parental wealth as exogenous throughout the empirical section---the estimates document associations rather than causal effects.
 For example, many transfers may be made precisely because the renter is buying a home.
-This motivates the structural model I describe in the next section, which provides a disciplined framework to interpret mechanisms and conduct counterfactuals.
+Nonetheless, the patterns consistently point to strong associations between parental resources and both homeownership entry and persistence.
+This motivates the structural model I describe in the next section, which provides a disciplined framework to interpret mechanisms and conduct counterfactuals
 
@@
 I assume that prices are constant for tractability, but I model aggregate house price risk as in \cite{Corbae2015} in Appendix \ref{sec:robust_pricerisk} and find that the quantitative results are almost unchanged.
-Homeowners incur proportional maintenance and depreciation costs $\delta$.
+Homeowners must pay proportional maintenance and depreciation costs $\delta$.
 Illiquidity is captured by proportional adjustment and moving costs on owner-occupied housing, as in \cite{Yang2009}:
@@
 \textit{Housing Parameters:} I set the rental rate $q=0.10$ \citep{Davis2008}, housing depreciation $\delta=0.05$ \citep{Harding2007}, and sales cost $m_s=0.075$ and purchase cost $m_b=0.02$ \citep{Yang2009}.
-I set the price level to $p=\$232,918/h_o$ to match the average market value of owner-occupied houses for households aged 25 to 44 in my sample.
+I set the per-unit price level to $p=\$232,918/h_o$ to match the average market value of owner-occupied houses for households aged 25 to 44 in my sample.
 The size $h_o$ is estimated internally.
@@
 
-While the model and the PSID have the same two-year frequency, households were asked about transfers in the last calendar year.
+While the model and the PSID have the same two-year frequency, the PSID asks about transfers in the last calendar year.
 In my cleaned sample, the annual transfer rate is 24\%.
-Thus, the two-year transfer rate could range between 24\% and 48\%.
-To be very conservative, I assume that the two-year transfer rate equals the one-year transfer rate.
+Thus, the two-year transfer rate could range anywhere between 24\% and 48\%.
+To be conservative, I assume that the two-year transfer rate equals the one-year transfer rate.
 My estimate of altruism $\eta$ should be viewed as a lower bound.\footnote{I have experimented with targeted higher transfer rates.
 This requires stronger altruism, which, in turn, increases homeownership motives, lowering the estimated utility premium $\chi$.
-Overall, the results are qualitatively similar to my conservative estimate, which understates the role of altruistic transfers.}
+Overall, the results are qualitatively similar to my conservative estimate.}
 
@@
 Otherwise, I replicate the exercise exactly.
-Figure \ref{fig:housinggrowthrates_model} shows that the model matches both the qualitative and quantitative patterns from the empirical data relatively well: Housing consumption falls for households without wealthy parents during ``unemployment,'' while it does not decline for those with rich parents.
+Figure \ref{fig:housinggrowthrates_model} shows that the model matches both the qualitative and quantitative patterns from the empirical data relatively well: The level of housing consumption falls for households without wealthy parents during ``unemployment,'' while it does not decline for those with rich parents.
 
@@
 Lowering purchase costs has little effect for renters with non-wealthy parents, who remain constrained by LTV limits and sales costs.
-In contrast, for households with wealthy parents, lower purchase costs reduce the effective price—the main remaining barrier to ownership.
+In contrast, for households with wealthy parents, lower purchase costs reduce the effective price---the main remaining barrier to ownership.
 This suggests that grants or credits for first-time buyers primarily benefit households with richer parents.
@@
 	{\begin{footnotesize}\begin{flushleft}\vspace{-0.1in}
-		\textit{Notes:} The sample includes all households aged 25-44 in the stationary distribution.
-Following \cite{Kaplan2014}, households are liquidity constrained if they have negative liquid wealth $(b_k)$ exceeding 16.5\% of annual earnings.
-MPCs are calculated from a \$5,000 positive shock.
+		\textit{Notes:} Following \cite{Kaplan2014}, households are liquidity constrained if they have negative liquid wealth $(b_k)$ exceeding 16.5\% of annual earnings.
+MPCs are calculated from a \$5,000 positive shock to households aged 25-44 in the baseline (with altruism) stationary distribution.
 ``Future transfers only'' MPCs use the baseline (altruistic) policy functions but set the contemporaneous transfer at the shock date to zero, while leaving expectations about future transfers unchanged.
@@
 First, model-implied MPCs may be overstated when intra-family transfers are omitted, particularly for households with wealthy parents.
-Second, a common strategy to match high MPCs is to introduce households with low target wealth—e.g., through impatience \citep{aguiar2024hand}—but such behavior is observationally similar to that of households expecting altruistic support.
+Second, a common strategy to match high MPCs is to introduce households with low target wealth---e.g., through impatience \citep{aguiar2024hand}.
+Altruistic expectations can generate qualitatively similar behavior, as anticipated parental support reduces savings motives.
 Third, models that generate high MPCs by introducing illiquid assets with large return premia \citep{kaplan2022marginal} may overlook the fact that homeowners with wealthy parents have access to informal insurance, which lowers their effective MPC.
@@
 Pooling can raise transfer capacity, while sibling dilution spreads support, making the net strength of the transfer motive ambiguous and heterogeneous across families.
+Moreover, parents can also provide non-financial support, such as co-signing mortgages, allowing children to live at home rent-free, or offering informal childcare.
+These non-financial transfers could substitute for direct transfers.
 Future work could also study whether illiquidity mitigates intrafamily commitment problems.

### 2025-10-17 docs/outside_parameters.tex
Re stud (#20)
* some shortening

* old latex is corruped, work on this one in the meantime

* minor changes

* Make error bars thicker

* Better figures

* Small fix

* Update text

* FEDSstuff

* Taking on comments

---------

Co-authored-by: Eirik Eylands Brandsaas <eirik.e.brandsaas@frb.gov>

@@
     Deprecation$^\dagger$ 		& $\delta$			& \pardelta		& \cite{Harding2007} \\			
-    Risk-free Rate$^\dagger$		& $r^f$ 			& \parrf 	    & Literautre (see text) \\
+    Risk-free Rate$^\dagger$		& $r^f$ 			& \parrf 	    & Literature (see text) \\
     Mortgage premium$^\dagger$	& $r^m$ 			& \parrm 	    & Literature (see text) \\
@@
     PMI limit           & $\overline{PMI}$   & \parPMIlim     & \cite{goodman2017sixty} \\ 
-    House Price          & $p$		& \parprice      & Data (see text) \\
+    House Price Level   & $p$		& \parprice      & Data (see text) \\
     Max Loan-to-Value   & LTV 			& \parLTV  & Literature/data (see text) \\

### 2025-10-17 src/Julia/export_estimation.jl
close #19

@@ -177,7 +177,7 @@ function PlotGlobalEstimation(dat,est,pars,targ_moms;store=false)
     for (ipar,par) in enumerate(pars)
         for (imom,mom) in enumerate(targ_moms)
 
-            plot(dfq[!,par],[dfp[!,Symbol(par,"lo_",mom)] dfp[!,Symbol(par,"hi_",mom)]],color=:gray,linestyle=:dash)
+            plot(dfq[!,par],[dfp[!,Symbol(par,"lo_",mom)] dfp[!,Symbol(par,"hi_",mom)]],color=:gray,linestyle=:dash,xrot=30)
             plot!(dfq[!,par],[dfp[!,Symbol(par,"med_",mom)]],color=:black,)
             if mom != :smm_obj
                 hline!([dat[!,mom]],color="orange",style=:dot)
@@ -192,7 +192,7 @@ function PlotGlobalEstimation(dat,est,pars,targ_moms;store=false)
     for (ipar,par) in enumerate(pars)
         mom = :smm_obj
 
-            plot(dfq[!,par],[dfp[!,Symbol(par,"lo_",mom)] dfp[!,Symbol(par,"hi_",mom)]],color=:gray,linestyle=:dash)
+            plot(dfq[!,par],[dfp[!,Symbol(par,"lo_",mom)] dfp[!,Symbol(par,"hi_",mom)]],color=:gray,linestyle=:dash,xrot=30)
             plot!(dfq[!,par],[dfp[!,Symbol(par,"med_",mom)]],color=:black,)
             plot!(dfq[!,par],[dfp[!,Symbol(par,"mnm_",mom)]],color=:red,)
             ptemp  = plot!(legend=false,ylims=pylims[1,mom],title=_varnames[mom],yscale=:log)            
@@ -223,10 +223,10 @@ function PlotGlobalEstimation(dat,est,pars,targ_moms;store=false)
     Plots.pdf(plot(plots_obj...),"tabfig/est/identification/smmobj")
     ### Plots each moment for all parameter
     for (imom,mom) in enumerate(targ_moms)
-    ptmp = deepcopy(plot(p1[:,imom]...))
-    if store == true
-        Plots.pdf(ptmp,"tabfig/est/identification/"*String(mom))
-    end
+        ptmp = deepcopy(plot(p1[:,imom]...))
+        if store == true
+            Plots.pdf(ptmp,"tabfig/est/identification/"*String(mom))
+        end
     end
 
     ### Plots the effect of each parameter
